---
title: "bosc23_metab_fba"
author: "Riccardo Frizzo"
date: "2025-10-13"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

#Init
## Packages
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
require(cobrar); require(tidyverse); require(data.table); require(foreach); library(parallel); require(magrittr); require(ggpubr); require(vegan); require(caret);library(ggrepel);require(DT);require(rstatix)
```
## Functions
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
plot_pca<-function(pca, top_n=30,
                   plot_title = "",
                   size_var="mean_mag_abund",
                   fill_var="Phylum",
                   arrow_scale=8,
                   metadata=gem_analysis,
                   limits=4
                   ){
OTUs<-rownames(pca$x)
pca_res<-pca
scores <- as.data.frame(pca_res$x)
loadings <- as.data.frame(pca_res$rotation)
metadata<-merge(gem_analysis, metadata, by="OTU")

# --- Identify top contributors to PC1 & PC2 ---
loadings$feature <- rownames(loadings)
loadings <- loadings %>%
  mutate(contrib = abs(PC1) + abs(PC2)) %>%
  arrange(desc(contrib))

top_loadings <- loadings[1:top_n, ]

# --- Scale arrows for plotting ---
  # adjust this factor to control arrow length
top_loadings <- top_loadings %>%
  mutate(PC1 = PC1 * arrow_scale,
         PC2 = PC2 * arrow_scale)

# Additional plotting variables
if(!is.na(size_var)){scores$size=metadata[OTU%in% OTUs,get(size_var)]}else{scores$size=NULL}
if(!is.na(fill_var)){scores$fill=metadata[OTU%in% OTUs,get(fill_var)]}else{scores$fill=NULL}


# --- Plot ---
ggplot(scores, aes(x=PC1, y=PC2)) +
  geom_point(alpha = 0.8,
             shape = 21,
             color="black",
             aes(fill=fill,
                 size=size,
                 )) +
  #scale_color_manual(values=c("black", "green")) +
  scale_fill_manual(values=tree_colors) +
  geom_segment(
    data = top_loadings %>%
               mutate(PC1=PC1*0.9,
                      PC2=PC2*0.9),
    aes(x = 0, y = 0, xend = PC1, yend = PC2),
    arrow = arrow(length = unit(0.25, "cm")),
    color = "black",
    linewidth = 0.1
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none") +
  labs(
    title = plot_title,
    x = paste0("PC1 (", round(100 * (pca_res$sdev[1]^2 / sum(pca_res$sdev^2)), 1), "%)"),
    y = paste0("PC2 (", round(100 * (pca_res$sdev[2]^2 / sum(pca_res$sdev^2)), 1), "%)")
  ) + 
  geom_text_repel(data = top_loadings, aes(x = PC1, y = PC2, label = feature),
          hjust = 1.1, vjust = 1.1, color = "black", size = 2.5, fontface = "bold", force = 1.5 ) +
  geom_point(data = top_loadings,
             aes(x = PC1, y = PC2, label = feature)) +
    
  xlim(-limits,limits) + ylim(-limits,limits)->p;p
#plotly::ggplotly(p) -> pi;pi
}
rm(scores)

##############################

plot_metab_PCA<-
  function(
  pca_data=pca_exhanges_aero,
  met="Trehalose",
  pca_exchanges=fba_ex){
scores<-pca_data$x %>% data.frame %>% rownames_to_column("OTU") %>% data.table
scores[,hl:=OTU %in% pca_exchanges[name %in% met, OTU]]
scores$flux=0
scores[OTU %in% fba_ex[name %in% met, OTU]]$flux = 
  abs(fba_ex[name %in% met, fba_flux]) %>% as.numeric
scores$abund=mag_analysis[OTU %in% scores$OTU]$mean_mag_abund

ggplot(scores, aes(x=PC1, y=PC2, fill=flux)) +
  geom_point(shape=21,aes(size=abund), alpha=0.8) + 
  scale_fill_gradient(high = "red", low = "white") +
  theme_minimal() + 
  theme(legend.position = "none") +
  ggtitle(paste(met,collapse = ";"))
  }
################################################
find_metab_react<-function(met_id, reactions){
df<-reactions
df$OTU %>% unique %>% length->ngems
df[grepl(met_id, compound_ids)]-> selected_react
print(ngems)

# Classify metabolite utilisation
selected_react %>%
  mutate(
    utilisation=
      case_when(
      ((grepl(paste0(met_id,".*="), equation) & fba_flux>0)|
      (grepl(paste0("=.*", met_id), equation) & fba_flux<0)) &
      !(grepl(paste0(met_id, ".*=.*", met_id), equation)) ~ "consumed",
      
      ((grepl(paste0(met_id,".*="), equation) & fba_flux<0)|
      (grepl(paste0("=.*", met_id), equation) & fba_flux>0)) &
      !(grepl(paste0(met_id, ".*=.*", met_id), equation)) ~ "produced",
      
      (grepl(paste0(met_id, ".*=.*", met_id), equation)) ~ "transported"
      )
    ) %>%
  dplyr::select(OTU,react_id, react_name, utilisation, definition,
         fba_flux, red_costs) -> selected_react1

selected_react1 %>%
  group_by(utilisation, react_id, react_name, definition) %>% 
  summarise(
    median_c_flux=median(fba_flux[utilisation=="consumed"]) %>% round(3),
    median_p_flux=median(fba_flux[utilisation=="produced"]) %>% round(3),
    median_t_flux=median(fba_flux[utilisation=="transported"]) %>% round(3),
    prop_gems=round((n()/ngems)*100,3) ,
    n_gems=n()
  ) %>% add_column(metabolite=key_metabolites[id==met_id, name],
                   .before="utilisation") %>%
  left_join(sbs_react)
}

met_report<-
  function(selected_mets=NA,
         exchange_subset=NA,
         react_subset=NA,
         utilisation_type=c("consumed", "produced", "transported"),
         ko_data=NA,
         fisher_use_type)
    {
    # For each metabolite, extract linked reactions and related info
    # from the provided rection dataset
    foreach(met_id = selected_mets$id)%do%{
      #exchange_subset[id==met_id]->sel
      find_metab_react(met_id=met_id,reactions = react_subset) %>% 
        arrange(prop_gems) %>% 
        filter(utilisation %in% utilisation_type) %>%
        ungroup
    } %>% rbindlist()  %>% arrange(n_gems) -> react_summary
    
    react_summary %>%
      left_join(sbs_react) -> react_summary


  
  
  react_subset$OTU %>% unique %>% length->ngems  
  ko_data %>%
    dplyr::rename(react_id=ko_metab) %>%
    mutate(react_id2=str_remove_all(react_id, "EX_|_c0|_e0")) %>%
    filter(react_id2 %in% selected_mets$id) %>%
    group_by(name, react_id) %>%
    summarise(
    median_biomass1=median(biomass_free) %>% round(3),
    mean_biomass_decrease=mean(delta_prop) %>% round(3),
    n_ko_gems=n()
  ) %>% arrange(ngems)-> ko_summary
  
  left_join(ko_summary, select(fisher_use_type, -p_value))-> ko_fish_summary
  
  
  left_join(
    selected_mets[,c("react_id", "name","median","n", "prop")],
    ko_fish_summary
        ) %>%
    mutate(prop_ko_gems=(100*(n_ko_gems/n)) %>% round(2),
           p_adj = formatC(p_adj, format = "e", digits = 1)) %>%
    
    dplyr::select(name, median, n, prop, prop_ko_gems,
                  mean_biomass_decrease, subs_prop, prod_prop) %>%
    
    dplyr::rename(
           "Metabolite name"= name,
           "Median FBA flux"=median,
           "Used by GEMs"=n,
           "Used by GEMs (%)"=prop,
           "GEMs limited (%)"= prop_ko_gems,
           "Mean biomass decrease"=mean_biomass_decrease,
           "Used as substrate (%)" = subs_prop,
           "Used as product (%)" = prod_prop
           )-> ko_fish_summary2
  list("react_summary"=react_summary,
       "ko_fish_summary"=ko_fish_summary2
       )->> res
  list("react_summary"=react_summary,
       "ko_fish_summary"=ko_fish_summary2
       ) %>% lapply(datatable)
  }

########################
react_check<-function(met="H2S"){
tidy_reactions[
  grepl(key_metabolites[grepl(met, name), id], compound_ids)] %>%
  group_by(Order, react_name) %>%
  summarise(median_flux=median(fba_flux) %>% round(5),
            n=n()) %>% data.table %>%
    arrange(median_flux)
  }

```
## Variables
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
qprobs=seq(0,1,0.01)
load("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/data/tax/tree_colors_14apr25.RData")
tree_colors<-c(tree_colors, "gray10")
names(tree_colors)[15]="Thermoproteota" #Add the archaea :)
```
#---     
# Metabolites vocabulary
## Prepare
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=FALSE}
require(rcdk)
key_metabolites<-fread("/home/riccardo/rstudio-server/storage16TB/riccardo/tools/gapseq_feb25/gapseq/dat/seed_metabolites_edited.tsv") %>%
  mutate(has_C=str_detect(formula, "C[:digit:]|C[:upper:]"),
         has_N=str_detect(formula, "N"),
         has_S=str_detect(formula, "S"), 
         has_H=str_detect(formula, "H"),
         has_O=str_detect(formula, "O"),
         ) 
## Define chemical classes
cl<-makeCluster(25)
doParallel::registerDoParallel(cl)
foreach(met_row=1:nrow(key_metabolites[,]), .errorhandling = "remove")%dopar%{
  print(paste0(met_row, "/", nrow(key_metabolites)))
  mols <- rcdk::parse.smiles(as.character(key_metabolites[met_row,"smiles"]))
  if(!is.null(unlist(mols))){
    tibble::add_column(key_metabolites[met_row,],
               is_COH= rcdk::matches(query ="([CX4][OX2H])", mols),
               is_COOH= rcdk::matches(query ="C(=O)O", mols),
               carbonyl = rcdk::matches(query ="[CX3]=O", mols)
               #ketone =rcdk::matches(query = "[#6][CX3](=O)[#6]", mols),
               #aldehyde = rcdk::matches(query ="[CX3H1](=O)[#6]", mols),
               #carboxylic_acid = rcdk::matches(query ="[CX3](=O)[OX2H1]", mols),
               #carboxylate = rcdk::matches(query ="[CX3](=O)[O-]", mols)
               )
                    }else{
                      tibble::add_column(key_metabolites[met_row,],
                        is_COH=NA,
                        is_COOH=NA,
                        carbonyl = NA,
                        #ketone = NA,
                        #aldehyde = NA,
                        #carboxylic_acid = NA,
                        #carboxylate = NA
                                  
                        )
                    }
} %>% rbindlist -> key_metabolites
stopCluster(cl)

fwrite(key_metabolites, "/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/key_metabolites.tsv")
```
## Load
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
# Import classes
fread("/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/key_metabolites.tsv") %>%
  mutate(has_H=str_detect(formula, "H"),
         has_O=str_detect(formula, "O"),
         has_P=str_detect(formula, "P"),
        simple_category=case_when(
            has_S ~ "Sulfur compounds",
            has_C & !has_S & !has_N & carbonyl & as.numeric(mass)<100 ~ "Carbonyl compounds" ,
            has_C & !has_S & !has_N & is_COH & as.numeric(mass)>100 ~ "Carbohydrates" ,
            has_C & !has_S & has_N & carbonyl ~ "N-Carbonyl compounds",
            has_C & !has_S & has_N ~ "Nitrogen compounds",
            has_C & !has_S & has_P ~ "Phosporous compounds",
            !has_C | (has_C & !has_H) ~ "Inorganic compounds", .default = NA)) -> key_metabolites

carbonyl_cpp_str<-"Fumarate|Succinate|2-Oxoglutarate|Citrate|Itaconate|4-Hydroxybenzoate|Acetoacetate|Chorismate|Malate"
aa_cpp_str<-"Alanine|Glycine|Isoleucine|Leucine|Proline|Valine|Phenylalanine|Tryptophan|Aspartate|Glutamate|Arginine|Histidine|Lysine|Serine|Threonine|Cysteine|Methionine|Asparagine|Glutamine|Tyrosine|Ala-Ala"

key_metabolites %<>%
  mutate(simple_category=
           case_when(grepl(carbonyl_cpp_str, name) ~ "Carbonyl compounds",
                     grepl("Methanol|Ethanol|Glycerol|1,2-Propanediol", name) ~ "Alcohols",
                     grepl("Xylan|Inulin|starch", name) ~ "Polysaccharides",
                     grepl("Carbon monoxide", name) ~ "Inorganic compounds",
                     grepl("Dodecanoic acid|lipid", name) ~ "Fatty acids",
                     simple_category=="Carbohydrates" & grepl("ol", name) ~ "Sugar Alcohols",
                     grepl(aa_cpp_str, name) ~ "Amino acids",
                     grepl("chloroethene", name) ~ "Chlorocarbons",
                     grepl("Zn2+|Fe3+|Fe2+|Mn|Mg|K+|Cu2+|Zn2+|Ca2+|Cobalt", name) ~ "Metals",
                     grepl("Melibiose", name) ~ "Carbohydrates",
                     .default = simple_category
                     ))

fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/data/analysis/fba_summary_ox_curated_30dec24.tsv") %>%
  dplyr::select(id, subclass)-> hmdb_metabolite_classes
```

# Reactions vocabulary
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
#fread("/home/riccardo/rstudio-server/storage16TB/riccardo/tools/gapseq/dat/seed_reactions_corrected.tsv")-> key_seed_reactions
fread("/home/riccardo/rstudio-server/storage16TB/riccardo/tools/gapseq_feb25/gapseq/dat/seed_reactions_corrected.tsv")-> key_seed_reactions

# Watch out!! A single reaction name, matches several IDs
(count(key_seed_reactions, name)$n>1) %>% table
(count(key_seed_reactions, name)$n>5) %>% table

# IDs are unique
(count(key_seed_reactions, id)$n>1) %>% table

# Simplify reactions DB
names(key_seed_reactions)

key_seed_reactions %>% 
  dplyr::select(id, name, compound_ids,
                 definition, equation 
                ) %>%
  dplyr::rename(react_id2=id,
                react_name=name) -> key_reactions_simple
```

# Pathways vocabulary
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
# Import pathway hierarchy
gapseq_pwy_hierarchy<-
  rbind(fread("/home/riccardo/rstudio-server/storage16TB/riccardo/tools/gapseq_feb25/gapseq/dat/meta_pwy.tbl"),
        fread("/home/riccardo/rstudio-server/storage16TB/riccardo/tools/gapseq_feb25/gapseq/dat/custom_pwy.tbl"))
gapseq_pwy_hierarchy <- gapseq_pwy_hierarchy[!duplicated(id)]
gapseq_pwy_hierarchy %<>%
  mutate(sbs=str_remove_all(id, "\\|")) %>% 
  dplyr::select(sbs, name, hierarchy) %>% 
  dplyr::rename(Name=name) %>%
  unique

# read and prepare pathway DB
pwyDB <- rbind(fread("/home/riccardo/rstudio-server/storage16TB/riccardo/tools/gapseq_feb25/gapseq/dat/meta_pwy.tbl"),
               fread("/home/riccardo/rstudio-server/storage16TB/riccardo/tools/gapseq_feb25/gapseq/dat/custom_pwy.tbl"))

pwyDB <- pwyDB[!duplicated(id)]
pwyDB <- pwyDB[, .(id, name, spont, reaId)]
pwyDB[, reaNr := length(unlist(strsplit(reaId, ","))), by = id] # Count n° of reactions
pwyDB[, spontNr := length(unlist(strsplit(spont, ","))), by = id] # Count n° of spont reactions
pwyDB %<>% mutate(id=str_remove_all(id, "\\|"))

pwyDB %>% dplyr::select(id, name) %>% 
  dplyr::rename(sbs = id) %>% mutate(sbs=str_remove_all(sbs, "\\|")) -> key_pwy
rm(pwyDB)

gapseq_pwy_hierarchy$hierarchy %>%
  str_extract("\\|Pathways\\|[:graph:]+") %>%
  str_remove_all("\\|") %>%
  str_split(pattern = ",", simplify = T) %>% data.table -> sbs_hier
names(sbs_hier)=paste0("h", seq(1,8))

cbind(gapseq_pwy_hierarchy, sbs_hier) %>%
  dplyr::rename(sbs_ids=sbs) %>%
  select(sbs_ids, Name, h2, h3,h4,h5) -> sbs_hier1

```



#----
# Prepare metaB data
## Import metaB IDs and assign a modelSEED id
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
# Import assigned metabolites 
fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaB/LCGC_metabolite_ids_subclasses_curated_classified_hmdb_24dic24.csv") %>%
  mutate(Subclass = str_remove(Subclass, "\t"))-> metab_ids_curated

# Use KEGG IDs and HMDB IDs to find the cpd id in gapseq (modelSEED) db
foreach(i = 1:nrow(metab_ids_curated))%do%{
  
  metab_ids_curated[i,]$HMDB->hmdbid
  metab_ids_curated[i,]$KEGG->keggid
  metab_ids_curated[i,]$name_orig->name_orig
  metab_ids_curated[i,]$Subclass->subclass
  
  #Find modelSEED ids
  key_metabolites %>% 
    filter(
        grepl(keggid, keggID) | 
        grepl(hmdbid, hmdbID)) %>% 
    dplyr::select(id, name, keggID, hmdbID) %>%
    add_column(name_orig=name_orig,
               subclass=subclass)->hits
  if(nrow(hits)==0){
    data.frame(id=NA,
               name=NA,
               keggID=NA,
               hmdbID=NA,
               name_orig=name_orig,
               subclass=subclass) # Create empty line in case of nohit
    }else(hits)
  } %>% rbindlist -> metab_ids_modelseed

metab_ids_modelseed %>% 
  filter(is.na(id)) -> not_matched

# Manual curation of non matches (go to ModelSEED Compunds and look at synonyms)
not_matched[1,]$id="cpd24232";not_matched[2,]$name="eicosane";
not_matched[3,]$id="cpd35300";not_matched[3,]$name="gamma-Glu-Leu";
not_matched[8,]$id="cpd33424";not_matched[8,]$name="Glu-Thr";
not_matched[9,]$id="cpd25630";not_matched[9,]$name="N-acetyl-(L)-arginine";
not_matched[12,]$id="cpd33751";not_matched[12,]$name="2-octanone";
not_matched[13,]$id="cpd33751";not_matched[13,]$name="2-nonanone";
not_matched[15,]$id="cpd15618";not_matched[15,]$name="nonanal";
not_matched[16,]$id="cpd15644";not_matched[16,]$name="undecanal";
not_matched[18,]$id="cpd02498";not_matched[18,]$name="2,3-Dihydroxy-isovalerate";
not_matched[19,]$id="cpd32753";not_matched[19,]$name="3-methylinosine";
not_matched[20,]$id="cpd32380";not_matched[20,]$name="5-methylcytidine";
not_matched[21,]$id="cpd00093";not_matched[21,]$name="Anthranilate";
not_matched[22,]$id="cpd10271";not_matched[22,]$name="2-Ethyltoluene";

not_matched %>% 
  filter(!is.na(id)) -> rematched

metab_ids_modelseed %>% 
  filter(!is.na(id)) -> matched

rbind(matched, rematched) %>% add_column(orig="metab")-> metab_react_matched
not_matched->metab_react_not_matched


rm(rematched, matched, not_matched, hits, metab_ids_modelseed, metab_ids_curated)
```

## SANITY CHECK: Metabolite merging
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
## Check ambiguos matches
metab_react_matched %>%
  group_by(name_orig) %>% count %>% 
  filter(n>1) %>%
  pull(name_orig) -> ambiguos

metab_react_matched[name_orig %in% ambiguos] 

# Check duplicates
metab_react_matched %>%
  group_by(id) %>% count(id) %>%
  filter(n>1) %>% pull(id) -> dups
metab_react_matched[id %in% dups] %>% arrange(id)

# Check unmatched
metab_react_not_matched %>% 
  filter(is.na(id)) 

# Final dataset
metab_react_matched %>% data.table->metab_react_matched_cpds
rm(metab_react_matched, ambiguos, dups)

fwrite(metab_react_matched_cpds, "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/metab_react_matched_cpds.csv")
```
## Add average relative signal intensity
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
GC_area<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaB/GC_for_metaboAn_27may25.csv")
GC_area %>% select(-group) %>%
  column_to_rownames("sample") %>%
  reshape2::melt()%>%
  group_by(variable) %>%
  summarise(mean=mean(value)) ->gc_avg

gc_avg %<>%
  mutate(prop=round(mean/sum(gc_avg$mean)*100,3))%>%
  dplyr::rename(name=variable) %>% filter(!grepl("NA_", name)) %>%
  arrange(name)

LC_area<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaB/LCMS_ALL_MA_26may25.csv")
LC_area %>% select(-group) %>%
  column_to_rownames("sample") %>%
  reshape2::melt() %>%
  group_by(variable) %>%
  summarise(mean=mean(value))->lc_avg

lc_avg %<>%
  mutate(prop=(mean/sum(lc_avg$mean)*100) %>% round(3)) %>%
  dplyr::rename(name=variable) %>% arrange(name)  %>% filter(!grepl("U_", name))

rbind(lc_avg, gc_avg)

metab_react_matched_cpds3<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/metab_react_matched_cpds_rel_avg_abund_v2.csv")
```

# Prepare growth medium
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
min_medium<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaG/coassemby_07feb24/mags_c70c10a99/gapseq/mean_min_medium_6sep24.csv")
metabolites<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/metab_react_matched_cpds.csv")

metabolites %>%
  add_column(maxFlux=10) %>%
  dplyr::rename(compounds=id) %>%
  select(compounds, name,  maxFlux) -> metab_medium

metab_medium[!compounds %in% min_medium$compounds]
min_medium[compounds %in% metab_medium$compounds]

rbind(min_medium[!compounds %in% metab_medium$compounds],
      metab_medium) %>% unique->min_metab_medium

fwrite(min_metab_medium, "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/min_metab_medium_v4_all_metab10.csv")
```
Gapfill

#-
# Import MAG analysis and prepare GEM analysis
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
# Import MAG analysis for taxonomy and other parameters
mag_analysis<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/data/analysis/mag_analysis_co_de_24sep24.tsv") %>%
  rowwise %>%
  mutate(mean_mag_abund=
           mean(mean_mag_abund_Art, mean_mag_abund_Nat), .before = "mean_mag_abund_Art") %>%
  data.table %>%
  mutate(Phylum=str_remove(Phylum, "p__"),
         Order=str_remove(Order, "o__")
         )

mag_analysis$Order[mag_analysis$Order==""]="Unknown"
mag_analysis %>% pull(OTU)->mag_ids

# Make gem_analysis
gem_analysis <- 
  mag_analysis[,c("OTU", "Phylum", "Class", "Order", "Family","mean_mag_abund", "virus_presence")]
```

#Import models
```{r eval=FALSE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
#job::job({
mag_fba_metab <- function(m) {
  model<-cobrar::readSBMLmod(
    paste0("/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaG/coassemby_07feb24/mags_c70c10a99/gapseq_metab_27oct25_all_metab10/", m, ".xml"))
  return(model)
}

#mag_fba_min <- function(m) {
#  model<-cobrar::readSBMLmod(
#    paste0("/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaG/coassemby_07feb24/mags_c70c10a99/gapseq/mean_min_medium_models_nocore_reac/all_models_xml/", m, ".xml"))
#  return(model)
#}

#cl<-makeCluster(30)
#parLapply(cl, mag_ids, mag_fba_min)->models_min
#stopCluster(cl);gc()
#names(models_min)=mag_ids

cl<-makeCluster(30)
parLapply(cl, mag_ids, mag_fba_metab)->models_metab
stopCluster(cl);gc()
names(models_metab)=mag_ids
#})


#cl<-makeCluster(30)
#parLapply(cl, mag_ids, mag_fba_min)->models_min
#stopCluster(cl);gc()
#names(models_metab)=mag_ids
#})

# GEM stats
models_metab[[1]]


```

#Prepare sbs react (do)
```{r eval=FALSE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
foreach(mod=models_metab)%do%{
  sbs<-mod@subSys %>% as.matrix
  rownames(sbs)=mod@react_id
  sbs %>% reshape2::melt() %>% as.data.frame
} %>% rbindlist() %>% filter(value) %>%
  select(Var1, Var2) %>% unique %>%
  dplyr::rename(react_id=Var1, sbs=Var2)-> sbs_react

sbs_react %>% mutate(sbs=str_remove(sbs, "subsys_") %>%
                       str_replace("_", "-")) -> sbs_react1

sbs_react1 %>%
  left_join(key_pwy) -> sbs_react

fwrite(sbs_react, "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/sbs_react.csv")
```
#Prepare sbs react (load)
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
sbs_react<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/sbs_react.csv") %>%
  group_by(react_id) %>%
  summarise(
    sbs_ids=paste(sbs, collapse = "|"),
    sbs_names=paste(name, collapse = "|")) %>%
  mutate(react_id2=str_remove_all(react_id, "_c0"))
```


#Prepare react-gene df (do)
```{r eval=FALSE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
foreach(mod=models_metab)%do%{
tibble(
  OTU=mod@mod_id,
  react_id=mod@react_id,
  react_name=mod@react_name,
  gene=mod@genes) %>% rowwise() %>% 
    mutate(gene=paste(unlist(gene), collapse = ";"))
} %>% rbindlist -> gem_react_genes
fwrite(gem_react_genes, "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/gem_react_genes.csv")
```

#Prepare react-gene df (load)
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
gem_react_genes<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/gem_react_genes.csv")
```
#FBA: Get default reactions and exchanges fluxes
## Simulate growth (do)
```{r eval=FALSE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
# FBA METAB ----
cl<-makeCluster(20)
doParallel::registerDoParallel(cl)

foreach(mod = models_metab) %dopar% {
  mod <- cobrar::changeBounds(model = mod, react = "EX_cpd00007_e0", lb = -0.01, ub=0)  #O2
  # Simulate aerobic growth
  cobrar::fba(mod) -> sol
  tibble::add_column(cobrar::fva(mod), 
                     OTU=mod@mod_id,
                     sim_O2=T, .before = "react") -> fva
  
  data.table::data.table(
    OTU = mod@mod_id,
    sim_O2=T,
    react_id=mod@react_id,
    fba_flux=sol@fluxes,
    fva_min=fva$min.flux,
    fva_max=fva$max.flux,
    red_costs=sol@redCosts
  )-> fba_a

  
  # Simulate anaerobic growth by setting O2 lower reaction bound to 0
  mod <- cobrar::changeBounds(model = mod, react = "EX_cpd00007_e0", lb = 0)  #O2
  cobrar::fba(mod) -> sol
  tibble::add_column(cobrar::fva(mod), 
                     OTU=mod@mod_id,
                     sim_O2=F, .before = "react") -> fva
  
  data.table::data.table(
    OTU = mod@mod_id,
    sim_O2=F,
    react_id=mod@react_id,
    fba_flux=sol@fluxes,
    fva_min=fva$min.flux,
    fva_max=fva$max.flux,
    red_costs=sol@redCosts
  )-> fba_an
  
  rbind(fba_a, fba_an)
} -> fba_comp
fba_comp %>% rbindlist() -> fba_comp_metab

#fwrite(fba_comp_metab, "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/fba_fva_min_gems_20nov25.csv", nThread = 20)
fwrite(fba_comp_metab, "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/fba_fva_metab_gems_2feb26_02-0.01.csv", nThread = 20)
stopCluster(cl)
gc()
```

## Simulate growth (load)
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
fba_comp_metab<- fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/fba_fva_metab_gems_27oct25.csv")

#fba_comp_metab<- fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/pfba_fva_metab_gems_02feb25.csv")

```

#Import subsystem/reaction table
```{r eval=FALSE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
foreach(mod=models_metab)%do%{
  sbs<-mod@subSys %>% as.matrix
  rownames(sbs)=mod@react_id
  sbs %>% reshape2::melt() %>% as.data.frame
} %>% rbindlist() %>% filter(value) %>%
  select(Var1, Var2) %>% unique %>%
  dplyr::rename(react_id=Var1, sbs=Var2)-> sbs_react

sbs_react %>% mutate(sbs=str_remove(sbs, "subsys_") %>%
                       str_replace("_", "-")) -> sbs_react1
```

#---
# Analysis I

## Number of MAGs by phylum
```{r eval=FALSE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
gem_analysis %>%
  group_by(Phylum) %>%
  summarise(abund=sum(mean_mag_abund),
            n=n()) %>%
  arrange(desc(abund)) -> phylum_nmags
```
## Number of MAGs by order
```{r eval=T, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
gem_analysis %>%
  group_by(Order) %>%
  summarise(abund=sum(mean_mag_abund),
            n=n()) %>%
  arrange(desc(abund)) -> order_nmags


order_nmags %>%
  filter(n>=5) %>% pull(Order) %>% length

order_nmags %>%
  filter(n>=5) %>% pull(n) %>% sum
```

## Number of MAGs by family
```{r eval=T, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
gem_analysis %>%
  group_by(Family) %>%
  summarise(abund=sum(mean_mag_abund),
            n=n()) %>%
  arrange(desc(abund)) -> family_nmags


family_nmags %>%
  filter(n>=5) %>% pull(Family) %>% length

family_nmags %>%
  filter(n>=5) %>% pull(n) %>% sum -> testable_fam
family_nmags %>%
  filter(n>=5) %>% pull(Family) -> testable_fam
```
## Define who is growing with O2 and without (for analysis)
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
# Check distribution of biomass vector
fba_comp_metab[react_id=="EX_cpd11416_c0"] %>%
  filter(sim_O2) %>% pull(fba_flux) %>% sort-> growth_vec_aero

fba_comp_metab[react_id=="EX_cpd11416_c0"] %>%
  filter(!sim_O2) %>% pull(fba_flux) %>% sort-> growth_vec_anaero

qprobs<-seq(0,1,0.01)
quantile(growth_vec_aero, qprobs)%>% round(10)

qprobs<-seq(0,1,0.01)
quantile(growth_vec_anaero, qprobs) %>% round(10)

# Look for a big drop (several orders of magnitude) in the quantile distribution
# 0.1 is a suitable threshold

fba_comp_metab[react_id=="EX_cpd11416_c0"] %>%
  filter(sim_O2) %>% filter(fba_flux>0.1) -> OTU_aero_df
OTU_aero_df %>% pull(OTU) -> OTU_aero

gem_analysis[OTU_aero_df[,c("OTU", "fba_flux")],
             growth_aero := i.fba_flux, on = "OTU"]

fba_comp_metab[react_id=="EX_cpd11416_c0"] %>%
  filter(!sim_O2) %>% filter(fba_flux>0.1)-> OTU_anaero_df
OTU_anaero_df %>% pull(OTU) -> OTU_anaero

gem_analysis[OTU_anaero_df[,c("OTU", "fba_flux")],
             growth_anaero := i.fba_flux, on = "OTU"]

# Define anaerobes
gem_analysis[,anaerobic:=OTU %in% OTU_anaero]
gem_analysis %<>%
  mutate(anaerobic_random=sample(anaerobic))


# Create a common growth value
gem_analysis
gem_analysis[, growth_value := fcase(growth_aero >= 1.11, growth_aero,
                               growth_aero < 1.11, growth_aero,
                               default = NA)]
gem_analysis %<>%
  mutate(growth_value=case_when(anaerobic==T ~ growth_anaero, .default = growth_value ))


```
## Effect of O2 blocking in biomass
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}

gem_analysis %<>%
mutate(diff_O2=round(100*(growth_anaero-growth_aero)/growth_aero),2)

gem_analysis[!is.na(diff_O2), diff_O2] %>% sort
gem_analysis[!is.na(diff_O2) & diff_O2<0, diff_O2] %>% mean
gem_analysis[!is.na(diff_O2), diff_O2] %>% sd

```
# Define who is growing fast and slow
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
growth_vec<-gem_analysis$growth_value%>% sort
growth_vec %>% quantile()->growth_quant;growth_quant

#Define threshold at 50%
growth_threshold<-growth_quant[["50%"]]


# Define growth categories
gem_analysis[, growth_category := fcase(growth_value > growth_threshold, "High",
                                        growth_value <= growth_threshold, "Low",
                                        default = NA)]
 
```

#---
# Tidy raw datasets
## Exchange fluxes
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
# Tidy, add info
# Filter low fluxes 
fba_comp_metab %>%
  filter(
    grepl("EX_", react_id),
    !grepl("DM|bio|EX_cpd11416_c0", react_id)) %>%
  mutate(id=str_remove_all(react_id, "EX_|_c0|_e0")) %>%
  left_join(key_metabolites[,
                            c("id", "name", "formula",
                              "has_C", "has_N","has_S",
                              "is_COH", "is_COOH","mass","simple_category")]) %>%
  left_join(
    gem_analysis[, 
                 c("OTU","Phylum", "Class", "Order","Family",
                   "growth_category", "anaerobic", "anaerobic_random")]) %>%
  select(OTU, sim_O2, react_id,
         fba_flux, fva_min, fva_max, red_costs,
         id, name, formula,
         has_C, has_N, has_S, , is_COH, is_COOH, mass, simple_category,
         Phylum, Class, Order, Family,
         growth_category, anaerobic, anaerobic_random) -> fba_comp_metab1
  #filter(abs(fba_flux)>1e-3)-> fba_comp_metab1 #HERE FLUX FILTERING 

# Fix metaB categories
fba_comp_metab1[,c("name", "simple_category")] %>% unique -> metab_cat
metab_cat$simple_category %>% unique
metab_cat[is.na(simple_category)]

carbonyl_cpp_str<-"Fumarate|Succinate|2-Oxoglutarate|Citrate|Itaconate|4-Hydroxybenzoate|Acetoacetate|Chorismate|Malate"
aa_cpp_str<-"Alanine|Glycine|Isoleucine|Leucine|Proline|Valine|Phenylalanine|Tryptophan|Aspartate|Glutamate|Arginine|Histidine|Lysine|Serine|Threonine|Cysteine|Methionine|Asparagine|Glutamine"
lipid<-"Decanoate|Myristic acid|tetradecenoate|Palmitate|tetradecenoate|Dodecanoic acid|lipid"

fba_comp_metab1 %<>%
  mutate(simple_category=
           case_when(grepl(carbonyl_cpp_str, name) ~ "Carbonyl compounds",
                     grepl("Methanol|Ethanol|Glycerol|1,2-Propanediol|1,3-Propanediol", name) ~ "Alcohols",
                     grepl("Xylan|Inulin|starch", name) ~ "Polysaccharides",
                     grepl("Carbon monoxide|H2S|Sulfate|Sulfite", name) ~ "Inorganic compounds",
                     grepl(lipid, name) ~ "Fatty acids",
                     simple_category=="Carbohydrates" & grepl("ol", name) ~ "Sugar Alcohols",
                     grepl(aa_cpp_str, name) ~ "Amino acids",
                     grepl("chloroethene", name) ~ "Chlorocarbons",
                     grepl("Zn2+|Fe3+|Fe2+|Mn|Mg|K+|Cu2+|Zn2+|Ca2+|Cobalt", name) ~ "Metals",
                     .default = simple_category
                     ))


# Separate aero and anaero

tidy_exhanges<-
  rbind(
  fba_comp_metab1[sim_O2==T & anaerobic==F],
  fba_comp_metab1[sim_O2==F & anaerobic==T])

## Filter substrates
fba_comp_metab1 %>%
  filter(
    sim_O2,
    fba_flux<0,
    OTU%in% gem_analysis[anaerobic==F, OTU]
    ) %>% mutate(fba_flux=-(fba_flux)) -> fba_substrates_aero

fba_comp_metab1 %>%
  filter(
    !sim_O2,
    fba_flux<0,
    OTU%in% gem_analysis[anaerobic==T, OTU]
    )%>% mutate(fba_flux=-(fba_flux)) -> fba_substrates_anaero

## Filter products
fba_comp_metab1 %>%
  filter(
    sim_O2,
    fba_flux>0,
    OTU%in% gem_analysis[anaerobic==F, OTU]
    ) -> fba_products_aero

fba_comp_metab1 %>%
  filter(
    !sim_O2,
    fba_flux>0,
    OTU%in% gem_analysis[anaerobic==T, OTU]
    ) -> fba_products_anaero
```



## Reaction fluxes
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
fba_comp_metab[grepl("rxn", react_id) &
               !grepl("DM|bio|EX_", react_id)] %>%
  mutate(react_id2=str_remove(react_id,"_c0|_e0")) %>%
   left_join(key_reactions_simple) %>%
  left_join(gem_analysis[,
                         c("OTU","growth_category", "anaerobic")]) %>%
  left_join(gem_analysis[, c("OTU","Phylum", "Class", "Order")])->fba_reactions
fba_reactions$fba_flux %>% abs %>% quantile(seq(0,1,0.01)) %>%
  round(5)
# 1e-3 is a reasonable threshold 

fba_reactions[abs(fba_flux)>1e-3] %>%
  left_join(gem_analysis[, c("OTU","Phylum", "Class", "Order")]) -> fba_reactions_f
index_reactions<-fba_reactions_f[,c("react_name")] %>% unique

tidy_reactions_f<-
  rbind(
  fba_reactions_f[sim_O2==T & anaerobic==F],
  fba_reactions_f[sim_O2==F & anaerobic==T])

tidy_reactions<-
  rbind(
  fba_reactions[sim_O2==T & anaerobic==F],
  fba_reactions[sim_O2==F & anaerobic==T])
```

#----
# INTEGRATION WITH VIR
## Import vocabularies 
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
vfdb_vocabulary<-fread("~/rstudio-server/storage16TB/riccardo/tools/phold/vfdb_description_output.csv")
defencefinder_vocabulary<-fread("~/rstudio-server/storage16TB/riccardo/tools/phold/defensefinder_plddt_over_70_metadata.tsv")
card_vocabulary<-fread("~/rstudio-server/storage16TB/riccardo/tools/phold/card_plddt_over_70_metadata.tsv")
```

## Import and tidy data
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
vir_analysis<-fread("/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/data/analysis/vir_analysis_03jan25.csv")
mag_analysis<-fread("/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/data/analysis/mag_analsis_fba_sbs_16nov24.csv")
phold_annot<-fread("/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/vir/pipeline3_29aug24/pharokka_phold/phold/phold_per_cds_predictions.tsv") %>%
  filter(contig_id %in% vir_analysis$votu)
names(phold_annot)[1]="votu"

phold_annot %>%
  dplyr::rename(func=`function`) -> phold_annot1

phold_annot1 %>%
  mutate(func= 
           case_when(
             func %in% c("head and packaging", "tail", "connector") ~ "Structural components",
             func == "moron, auxiliary metabolic gene and host takeover" ~ "Auxiliary genes and host takeover",
             func == "other" ~ "Other", .default = func)) -> phold_annot2


phold_annot2 %>% dplyr::select(cds_id, func, product) -> phold_annot_simple

blastp_out<-list.files(
  "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaG/coassemby_07feb24/mags_c70c10a99/blastp_mags_phold_morons_07jan25/",
  pattern = ".faa", full.names = T)

foreach(i = blastp_out[])%do%{
fread(i) %>%
    add_column(OTU=str_extract(i, "[:alpha:]+_SemiBin_[:digit:]+")) -> hits
  names(hits) = c("qseqid", "sseqid", "pident", "length", "mismatch", "gapopen", 
             "qstart", "qend", "sstart", "send", "evalue", "bitscore", 
             "qlen", "slen", "OTU")
  hits
  } %>% rbindlist -> blastp

columns <- c("qseqid", "sseqid", "pident", "length", "mismatch", "gapopen", 
             "qstart", "qend", "sstart", "send", "evalue", "bitscore", 
             "qlen", "slen")
react_sbs_metab_info<-fread(
   "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/final/supplementary/gem_reactions_17mar25.csv")
```

# Expand VFDB annotations
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
phold_annot2 %>% 
  filter(func == "Auxiliary genes and host takeover") -> aux
aux %>% 
  dplyr::select(votu, cds_id, product, tophit_protein) %>%
  filter(product == "VFDB virulence factor protein") %>%
  dplyr::rename(prot_id=tophit_protein) %>% merge(vfdb_vocabulary, by="prot_id") -> aux_vfdb

aux_vfdb %>% 
  group_by(description) %>% 
  count(sort = T) -> aux_vfdb_count
names(aux_vfdb_count) = c("Gene", "Count")
aux_vfdb_count %<>% add_column(Category = "VFDB", .before = "Gene")

```

## BLASTn: All votu cds vs MAGs proteins
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
list.files("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/vir/pipeline3_29aug24/OTU_morons2/blastn_aux_oth_mags/", full.names = T)->files_paths
list.files("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/vir/pipeline3_29aug24/OTU_morons2/blastn_aux_oth_mags/", full.names = F)->files_names

foreach(i=files_paths)%do%{
  fread(i)
    } %>% rbindlist->blastn
names(blastn) = c("cds_id", "sseqid", "pident", "length", "mismatch", "gapopen", 
             "qstart", "qend", "sstart", "send", "evalue", "bitscore", 
             "qlen", "slen")
blastn -> blastn_f

left_join(
  blastn_f,phold_annot_simple) %>%
  mutate(
  qperc_align=(length/qlen)*100,
  sperc_align=(length/slen)*100,
  ) %>%
  filter(pident>95 & qperc_align>95)-> blastn_all_votu_genes
blastn_all_votu_genes$product %>%
table %>% sort

blastn_all_votu_genes %>%
  group_by(func, product) %>% count %>% view
  
```
## Moron and auxiliary, and others
### Prepare
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval = FALSE}
# Select and summarise OTHER annot
phold_annot2 %>% 
  filter(func == "Auxiliary genes and host takeover") -> aux

aux %>%
  count(product) %>% 
  arrange(desc(n)) -> aux_count
names(aux_count) = c("Gene", "Count")
aux_count %<>% add_column(Category = "AUX", .before = "Gene")


# Expand VFDB annotations
aux %>% 
  dplyr::select(votu, cds_id, product, tophit_protein) %>%
  filter(product == "VFDB virulence factor protein") %>%
  dplyr::rename(prot_id=tophit_protein) %>% merge(vfdb_vocabulary, by="prot_id") -> aux_vfdb

aux_vfdb %>% 
  group_by(description) %>% 
  count(sort = T) -> aux_vfdb_count
names(aux_vfdb_count) = c("Gene", "Count")
aux_vfdb_count %<>% add_column(Category = "VFDB", .before = "Gene")


# Expand DefenceFinder annotations
aux %>% 
  dplyr::select(votu, cds_id, product, tophit_protein) %>%
  filter(product == "DefenseFinder protein") %>%
  mutate(id = parse_number(str_extract(tophit_protein, "_[:digit:]+$"))) %>% 
  merge(defencefinder_vocabulary, by="id") -> aux_defencefinder


aux_defencefinder %>% 
  group_by(gene_name) %>% 
  count(sort = T) -> aux_defencefinder_count
names(aux_defencefinder_count) = c("Gene", "Count")
aux_defencefinder_count %<>% add_column(Category = "DF", .before = "Gene")

# Expand CARD annotations
aux %>% 
  dplyr::select(votu, cds_id, product, tophit_protein) %>%
  filter(product == "CARD resistance protein") %>% 
  dplyr::rename("Protein Accession" = tophit_protein) %>%
  merge(card_vocabulary, by="Protein Accession") -> aux_card


# Select and summarise OTHER annot
phold_annot2 %>% 
  filter(func == "Other") -> oth

oth %>%
  count(product) %>% 
  arrange(desc(n)) -> oth_count
names(oth_count) = c("Gene", "Count")
oth_count %<>% add_column(Category = "OTH", .before = "Gene")


rbindlist(list(aux_count, aux_defencefinder_count, aux_vfdb_count, oth_count)) %>%
  arrange(desc(Count)) -> aux_oth_gene_count 
 
 fwrite(
    aux_oth_gene_count,
    file = "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/final/supplementary/vir_aux_other_counts_11mar25.csv")

```

## BLASTp: Morons and other vs MAGs proteins
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
# Check that all the MAGs have been processed
length(unique(blastp$OTU))

blastp %>%
  mutate(qseqid_gene = str_split(qseqid, "\\:", simplify = T)[,2]) %>%
  mutate(
  qperc_align=(length/qlen)*100,
  sperc_align=(length/slen)*100,
  )-> blastp1

blastp1[qperc_align>130]


phold_annot$`function` %>% table
unique(blastp1$qseqid) %>% length
unique(blastp1$sseqid) %>% length
mag_analysis$Total_Coding_Sequences %>% mean

blastp1 %>%
  dplyr::select(sseqid, OTU)%>%
  unique %>%
  group_by(OTU) %>% count()

# Select mag_analysis MAGs only
blastp1 %>%
  filter(OTU %in% mag_analysis$OTU) -> blastp2

#blastp2 %>% 
#  filter(pident>=75,
#         qperc_align>=95,
#         evalue<1e-10
#         )-> blastp_f

blastp2 %>% 
  filter(pident>=30,
         qperc_align>=70,
         evalue<1e-5
         )-> blastp_f

blastp_f %>%
  dplyr::select(OTU, qseqid, qseqid_gene,sseqid, qperc_align, sperc_align, evalue, length) %>%
  mutate(qseqid = str_split(qseqid, ":", simplify = T)[,1]) %>%
  dplyr::rename(cds_id = qseqid_gene, votu=qseqid, mag_gene=sseqid) %>%
  left_join(phold_annot_simple, by = "cds_id") %>% data.table -> blast_phold

blast_phold %>% 
  group_by(func, product) %>%
  summarise(n=n())-> blast_phold_summary

phold_annot2[votu %in%
blast_phold[grepl("S-adenosylmethionine decarboxylase", product), votu]] 

mag_analysis[OTU %in% blast_phold$OTU]
vir_analysis[votu %in% blast_phold$qseqid] 

blast_phold %>%
  group_by(func, product) %>%
  summarise(n_mags_hits=length(unique(OTU)),
            v_otus_hits=length(unique(votu))
            )  -> blastp_count

phold_annot2 %>%
  group_by(func,product) %>%
  summarise(n_votu=length(unique(votu))) %>%
  right_join(blastp_count, by=c("func","product")) %>% 
  arrange(desc(v_otus_hits)) %>%
  mutate(prop_mags=round(100*(n_mags_hits/311),2)) 
phold_annot2
```

# Subset CDS
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
require(Biostrings)
readDNAStringSet("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/vir/pipeline3_29aug24/OTU_morons2/pharokka_phold/prodigal-gv.ffn")-> aux_cds.fa
names(aux_cds.fa) %>% str_extract(".*_[:digit:]+") -> cds_ids

selected_cds<-names(aux_cds.fa)[cds_ids%in%selected_aux[,cds_id]]
selected_sequences<-aux_cds.fa[selected_cds]
names(selected_sequences) %<>% 
  str_extract(".*_[:digit:]+") #%>% str_replace("\\|", "_")

names(selected_sequences)

writeXStringSet(
  selected_sequences,
  "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/vir/pipeline3_29aug24/contigs_versus_selected_AUX/selected_aux_cds.fa",format = "fasta")
```

# Check blastn hits on contigs
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
rbind(
fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/vir/pipeline3_29aug24/contigs_versus_selected_AUX/aux_cds_art_contigs.tsv"),
fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/vir/pipeline3_29aug24/contigs_versus_selected_AUX/aux_cds_nat_contigs.tsv"))-> blastn_aux_contigs

names(blastn_aux_contigs) = c("cds_id", "sseqid", "pident", "length", "mismatch", "gapopen", 
             "qstart", "qend", "sstart", "send", "evalue", "bitscore", 
             "qlen", "slen")

left_join(
  blastn_aux_contigs,phold_annot_simple) %>%
  mutate(
  qperc_align=(length/qlen)*100,
  sperc_align=(length/slen)*100
  ) %>%
  filter(pident>95 & qperc_align>95)-> blastn_all_votu_genes
```

# BlastP results
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
blast_phold %>%
  left_join(aux_vfdb[,c("cds_id", "description")], by="cds_id") %>%
  mutate(
  product=case_when(cds_id %in% aux_vfdb$cds_id ~ description, .default=product)) -> blast_phold
  
blast_phold %>%
  filter(grepl(aux_oth_string, product, ignore.case=T)) %>%
  left_join(gem_analysis[,c("OTU", "Phylum", "Order")]) -> blast_phold_aux

blast_phold_aux %>% group_by(Order) %>% count(sort = T)

blast_phold_aux %>%
  group_by(product) %>% count %>% 
  fwrite("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/blasp_aux_product_annot.csv")

vir_analysis[votu %in% blast_phold_aux$votu, mean_vir_all] %>% sum
```
# Host prediction
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
iphop_mags<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/vir/pipeline3_29aug24/iphop_mags_vOTUs/Host_prediction_to_genome_m90.csv") %>% janitor::clean_names()

iphop_mags$host_taxonomy %>%
  str_split(';', simplify = T) %>%
  as.data.frame -> iphop_expanded_tax
names(iphop_expanded_tax) = c('Domain', 'Phylum', 'Class', 'Order', 'Family', 'Genus', 'Species')
iphop_expanded_tax$Order %>% table %>% sort

cbind(iphop_mags, iphop_expanded_tax) %>% data.table -> iphop_mags2

# How many virus per order (no mags)
iphop_mags2[!grepl("SemiBin", host_genome)] %>%
  group_by(Order) %>%
  summarise(n_votu=length(unique(virus))) %>%
  arrange(desc(n_votu)) -> votu_order_count

# How many virus per order (only mags)
iphop_mags2[grepl("SemiBin", host_genome)] %>%
  group_by(Order) %>%
  summarise(n_votu=length(unique(virus))) %>%
  arrange(desc(n_votu))

iphop_mags2[grepl("SemiBin", host_genome)] %>%
  summarise(sum=length(unique(virus)))

iphop_mags2[grepl("SemiBin", host_genome)] %>%
  summarise(sum=length(unique(host_genome)))
  
mag_analysis$Order %>% table %>% sort

iphop_mags2[grepl("SemiBin", host_genome)] %>% 
  fwrite("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/iphop_vOTU-MAGs.csv")

# How many virus per order (all)
iphop_mags2 %>%
  group_by(Order) %>%
  summarise(n_votu=length(unique(virus))) %>%
  arrange(desc(n_votu)) %>%
  mutate(Order=str_remove(Order, "o__"))-> votu_order_count_all
```

# AUX: Find overlapping annotations between blastp_hits and gems genes (1)
```{r eval=FALSE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
mag_prot<-list.files("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaG/coassemby_07feb24/mags_c70c10a99/checkm2_mags/protein_files/", pattern = ".faa", full.names = T)

foreach(i=mag_prot[-1])%do%{
  data.table(
    OTU=str_extract(i, "//[:graph:]+.faa") %>% str_remove_all("//|.faa"),
    header=readAAStringSet(i) %>% names
  )
} %>% rbindlist -> mag_prot_regions

mag_prot_regions %>%
  mutate(mag_gene=str_extract(header, "megahit_[:digit:]+_[:digit:]+"),
         contig=str_remove(mag_gene, "_[:digit:]+$"),
         region_span=str_extract(header, "# [:digit:]+ # [:digit:]+ "),
         range=str_replace(region_span, " # ","_") %>% str_remove("# ") %>% str_remove_all(" "),
         gene=paste0(contig,"_", range),
         ) %>%
  select(OTU,mag_gene, gene) -> mag_prot_regions1   # load and tidy all CDS

mag_prot_regions1[mag_gene %in% blast_phold$mag_gene] %>% #select CDS that matched viral functions (blastp)
  rowwise %>%
  mutate(
         contig=str_remove(mag_gene, "_[:digit:]+$"),
         id=paste0(OTU,"-", contig) %>% str_remove_all(" "),
         range=str_extract(gene, "[:digit:]+_[:digit:]+$"),
          start=sort(parse_integer(unlist(strsplit(range,"_"))))[1],
          end=sort(parse_integer(unlist(strsplit(range,"_"))))[2]
             ) %>%
  select(id,mag_gene, contig, start, end) %>% unique -> blastp_mag_contig_regions

gem_react_genes %>%
  group_by(OTU, gene, react_id, react_name) %>%
  reframe(gene=unlist(strsplit(gene,";"))) -> gem_react_genes_l

gem_react_genes_l %>%
    mutate(
         contig=str_remove(gene, "_[:digit:]+_[:digit:]+$"),
         id=paste0(OTU,"-", contig) %>% str_remove_all(" "),
         range=str_extract(gene, "[:digit:]+_[:digit:]+$")) %>%
  filter(contig %in% blastp_mag_contig_regions$contig) %>%
  rowwise %>%
  mutate(
    start=sort(parse_integer(unlist(strsplit(range,"_"))))[1],
    end=sort(parse_integer(unlist(strsplit(range,"_"))))[2]
         ) %>%
  select(id, gene, start, end) -> gem_react_contig_regions

fwrite(blastp_mag_contig_regions %>% select(-c(mag_gene, contig)) %>% unique, "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/blastp_mag_contig_regions.bed",
col.names = F,
sep = "\t", dec = ",")

fwrite(gem_react_contig_regions %>% select(-gene) %>% unique, "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/gem_react_contig_regions.bed", col.names = F,
sep = "\t", dec = ",")

# do bedtools
"bedtools intersect -a blastp_mag_contig_regions.bed -b gem_react_contig_regions.bed -wa -wb > aux_gems"

fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/aux_gems") -> aux_gems
names(aux_gems)=c("aux_contig", "aux_start", "aux_end", "gem_contig", "gem_start", "gem_end")
names(gem_react_contig_regions)=c("gem_contig", "gem_gene", "gem_start", "gem_end")
blastp_mag_contig_regions %<>% select(-contig)
names(blastp_mag_contig_regions)=c("aux_contig", "aux_gene", "aux_start", "aux_end")

aux_gems %>%
  left_join(blastp_mag_contig_regions) %>%
  left_join(gem_react_contig_regions) %>%
  #select(aux_contig, aux_gene, gem_gene) %>%
  unique %>%
  left_join(
    blast_phold[,c("mag_gene","func", "product")] %>%
      dplyr::rename(aux_gene=mag_gene)
      ) %>%
  left_join(gem_react_genes_l[,c("gene", "react_id", "react_name")] %>% unique %>%
              dplyr::rename(gem_gene=gene)) %>% unique -> blastp_aux_gem_react




fwrite(blastp_aux_gem_react, "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/blastp_aux_gem_react.csv")


  
```

# Analyse AUX functions in GEMS
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
blastp_aux_gem_react<- fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/blastp_aux_gem_react.csv")

# select genes aux genes that matched in GEMs
blastp_aux_gem_react %>%
  filter(grepl("Other|Auxiliary", func)) %>%
    mutate(gem=str_split(gem_contig, "-", simplify = T)[,1],
           mag_gene=paste0(gem, "-", aux_gene)) %>%
  select(mag_gene, product, react_name, react_id) %>% 
  unique-> blastp_aux_gem_react_aux

# Link votu and MAGs to genes containing AUX
blast_phold %>%
  filter(grepl("Other|Auxiliary", func)) %>%
  mutate(mag_gene=paste0(OTU, "-", mag_gene)) %>%
  filter(mag_gene %in% blastp_aux_gem_react_aux$mag_gene) %>%
  select(mag_gene, OTU, votu) %>% unique-> mag_gene_votu

# Match gene annotation with MAG and vOTU identity
full_join(blastp_aux_gem_react_aux, mag_gene_votu, by = "mag_gene") %>% 
  unique %>%filter(!is.na(votu)|!is.na(OTU)) ->  gem_votu_matches

gem_votu_matches %>%
  select(mag_gene, product, react_name, react_id, OTU, votu) %>% unique -> gem_votu_matches2

gem_votu_matches2 %>%
  group_by(product, react_name, react_id) %>%
  filter(!grepl("transport|resistance|repressor", product),
         !grepl("ABC|transport", react_name)
         ) %>%
  summarise(
    n=n(),
    nmag=length(unique(OTU)),
    nvotu=length(unique(votu)),
    gems_id=paste(unique(OTU), collapse = "|"),
            votu_id=paste(unique(votu), collapse = "|")
            ) %>% 
  ungroup %>%
  group_by(product) %>%
  slice_max(order_by = n, n = 5, with_ties = FALSE)%>%
  left_join(sbs_react[, c("react_id", "sbs_names")]) -> top_aux_gem_react

fwrite(top_aux_gem_react, "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/putative_avg_in_gems_annot_09feb26_v2.csv", sep="\t", dec=",")

```

# Describe selection of putative AVGs
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}

top_aux_gem_react_classes<-
  fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/putative_avg_in_gems_annot_16feb26_v6.tsv") %>%
  unique

# How many initial genes is aux and oth?
phold_annot_simple %>%
  filter(grepl("Aux|Oth", func)) %>% pull(product) %>% unique %>% length

gem_votu_matches2 %>%
  group_by(product, react_name, react_id) %>%
  filter(grepl("transport|resistance|repressor", product),
         !grepl("ABC|transport", react_name)
         ) %>% pull(product) %>% unique

# Final selection of putative AVGs
top_aux_gem_react_classes %>% 
  filter(`Unlikely AMG`!="T") -> top_aux_gem_react_classes_f

top_aux_gem_react_classes_f$product
top_aux_gem_react_classes_f$react_name

top_aux_gem_react %>%
  select(product, react_name, votu_id, gems_id) %>% 
  right_join(top_aux_gem_react_classes_f, by = c("product", "react_name", "votu_id", "gems_id")) %>% 
  filter(!is.na(react_name)) %>% unique %>% arrange(nvotu) %>% data.table -> top_aux_gem_react_summary2

foreach(i=1:nrow(top_aux_gem_react_summary2))%do%{
  top_aux_gem_react_summary2[i,gems_id] %>% strsplit("\\|") %>% unlist %>% unique -> gem_ids
  top_aux_gem_react_summary2[i,react_name] -> react_nam
  
  gem_analysis[OTU %in% gem_ids, Order] %>% unique %>% paste(collapse = "|") -> order_vec
  tidy_reactions_f[OTU %in% gem_ids & react_name %in% react_nam, fba_flux] -> fba_flux
  
  top_aux_gem_react_summary2[i,] %>% 
    add_column(median_flux=(median(abs(fba_flux))),
               gem_orders=order_vec
               )
} %>% rbindlist-> top_aux_gem_react_summary3

top_aux_gem_react_summary3 %>%
  select(product, react_name, Class, votu_id, gems_id) %>%
  fwrite("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/aux_in_gems.csv", sep = "\t", dec = ".")


top_aux_gem_react_summary3 %>%
  select(Class, react_name, median_flux, nvotu, sbs_names, nmag, react_id) %>%
  group_by(Class, react_name,react_id, sbs_names) %>%
  summarise(median_flux=mean(median_flux),
            nvotu=sum(nvotu),
            nmag=sum(nmag),
            ) -> top_aux_gem_react_summary4

top_aux_gem_react_summary4 %>%
  group_by(Class) %>% count(name = "ngenes", sort = T) -> ngenes
ngenes$ngenes %>% sum
ngenes %>%
  mutate(prop=100*(ngenes/52))

top_aux_gem_react_summary4 %>%
  group_by(Class) %>% summarise(nvotu=sum(nvotu)) -> nvotus

full_join(ngenes, nvotus) %>% arrange(nvotu)->n_genes_votus

n_genes_votus$Class=factor(n_genes_votus$Class, levels = n_genes_votus$Class)

top_aux_gem_react_summary4$Class=
  factor(top_aux_gem_react_summary4$Class, levels = n_genes_votus$Class)
top_aux_gem_react_summary4$Class %>% unique

  ggplot() + 
  geom_bar(data= top_aux_gem_react_summary4,
           aes(y=Class, x=nvotu, fill=median_flux),
           stat="identity", color="black", size=0.7) +
  geom_text(data = n_genes_votus,
            aes(x = nvotu+1, y=Class, label = paste0("", ngenes))) +
  theme_bw() +
    xlab("Number of vOTUs carrying a putative AVG") +
    ylab("Microbial metabolism") +
  scale_fill_gradientn(na.value = "white",colours = 
                      c(
                      low = "cadetblue",
                      mid="orange",
                      high = "red")) +
    theme(legend.position = "bottom",axis.text.y = element_text(size = 12)) +
    labs(fill = expression(atop("Median reaction flux", mmol~g[DW]^{-1}~h^{-1}))) -> p;p
  
ggsave(plot = p,
    "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/barplot_putative_avg_in_gems.png",
    width = 10, height = 8, units = "in", dpi = 600)

file.show("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/barplot_putative_avg_in_gems.png")
```

# Verify react flux values and pathways
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
top_aux_gem_react_summary3 %>%
  filter(grepl("glutamine", react_name, ignore.case=T)) %>%
  pull(react_id) -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("glutamine", react_name)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags


tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )

"D-Glucosamine phosphate"
key_metabolites[name=="D-Glucosamine phosphate", c("id", "name")]
find_metab_react("cpd00288", tidy_reactions_f[]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable

####################

top_aux_gem_react_summary3 %>%
  filter(grepl("D-glycero-beta-d-manno-heptose", react_name, ignore.case=T)) %>%
  pull(react_id) -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("D-glycero-beta-d-manno-heptose", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags, definition] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )

####################

top_aux_gem_react_summary3 %>%
  filter(grepl("arabinose", react_name, ignore.case=T)) %>%
  pull(react_id) -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("arabinose", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags, definition] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )

key_metabolites[name=="D-Arabinose5-phosphate", c("id", "name")]
find_metab_react("cpd00817", tidy_reactions_f[]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable

####################

top_aux_gem_react_summary3 %>%
  filter(grepl("nicotin", react_name, ignore.case=T)) %>%
  pull(react_id) -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("nicotin", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

gem_analysis[OTU %in% mags]

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags, definition] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )

####################

top_aux_gem_react_summary3 %>%
  filter(grepl("cobal", react_name, ignore.case=T)) %>%
  pull(react_id) -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("cobal", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

gem_analysis[OTU %in% mags] %>% arrange(Order)

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags, definition] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )

####################

top_aux_gem_react_summary3 %>%
  filter(grepl("this", react_name, ignore.case=T)) %>%
  pull(react_id) -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("this", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

gem_analysis[OTU %in% mags] %>% arrange(Order)

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags, definition] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )

####################

top_aux_gem_react_summary3 %>%
  filter(grepl("S-adeno", react_name, ignore.case=T)) %>%
  pull(react_id) -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("S-adeno", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

gem_analysis[OTU %in% mags] %>% arrange(Order)

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags, definition] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )
####################

top_aux_gem_react_summary3 %>%
  filter(grepl("oxo-L", react_name, ignore.case=T)) %>%
  pull(react_id) -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("oxo-L", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

gem_analysis[OTU %in% mags] %>% arrange(Order)

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags, definition] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )

####################

top_aux_gem_react_summary3 %>%
  filter(grepl("glutamate", react_name, ignore.case=T)) %>%
  pull(react_id) -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("glutamate", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

gem_analysis[OTU %in% mags] %>% arrange(Order)

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags, definition] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )

####################

top_aux_gem_react_summary3 %>%
  filter(grepl("aspartate", react_name, ignore.case=T)) %>%
  pull(react_id) -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("aspartate", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

gem_analysis[OTU %in% mags] %>% arrange(Order)

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags, definition] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )
####################

top_aux_gem_react_summary3 %>%
  filter(grepl("photo", react_name, ignore.case=T)) %>%
  pull(react_id) -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("photo", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

gem_analysis[OTU %in% mags] %>% arrange(Order)

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags, c("react_name", "definition")] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )

####################

top_aux_gem_react_summary3 %>%
  filter(grepl("sulf", react_name, ignore.case=T)) %>%
  pull(react_id) %>% unique -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("sulf", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

gem_analysis[OTU %in% mags] %>% arrange(Order)

tidy_reactions[react_id %in% react_ids[] & OTU %in% mags, definition] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )

####################

top_aux_gem_react_summary3 %>%
  filter(grepl("ferre", react_name, ignore.case=T)) %>%
  pull(react_id) %>% unique -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("ferre", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

gem_analysis[OTU %in% mags] %>% arrange(Order)

tidy_reactions[react_id %in% react_ids[] & OTU %in% mags, definition] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )

####################

top_aux_gem_react_summary3 %>%
  filter(grepl("complex", react_name, ignore.case=T)) %>%
  pull(react_id) %>% unique -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("ferre", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

gem_analysis[OTU %in% mags] %>% arrange(Order)

tidy_reactions[react_id %in% react_ids[] & OTU %in% mags, definition] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )

####################

top_aux_gem_react_summary3 %>%
  filter(grepl("urea", react_name, ignore.case=T)) %>%
  pull(react_id) %>% unique -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("urea", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

gem_analysis[OTU %in% mags] %>% arrange(Order)

tidy_reactions[react_id %in% react_ids[] & OTU %in% mags, definition] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )

####################

top_aux_gem_react_summary3 %>%
  filter(grepl("acetate", react_name, ignore.case=T)) %>%
  pull(react_id) %>% unique -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("acetate", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

gem_analysis[OTU %in% mags] %>% arrange(Order)

tidy_reactions[react_id %in% react_ids[] & OTU %in% mags, definition] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )
####################

top_aux_gem_react_summary3 %>%
  filter(grepl(":pyruvate,phosphate phosphotransferase", react_name, ignore.case=T)) %>%
  pull(react_id) %>% unique -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl(":pyruvate,phosphate phosphotransferase", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

gem_analysis[OTU %in% mags] %>% arrange(Order)

tidy_reactions[react_id %in% react_ids[] & OTU %in% mags, definition] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )

key_metabolites[name=="Phosphoenolpyruvate"]
find_metab_react("cpd00061", tidy_reactions_f[OTU %in% mags]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable

####################

top_aux_gem_react_summary3 %>%
  filter(grepl("hexadecanal", react_name, ignore.case=T)) %>%
  pull(react_id) %>% unique -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("hexadecanal", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

gem_analysis[OTU %in% mags] %>% arrange(Order)

tidy_reactions[react_id %in% react_ids[] & OTU %in% mags, definition] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )

#################
key_metabolites[grepl("phosphoadenyl",name)]
find_metab_react("cpd00044", tidy_reactions[]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable

```

# Verify gene location in genome
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
top_aux_gem_react_summary3 %>%
  #filter(grepl("5-oxo", react_name)) %>% 
  filter(Class == "Carbonyl-compound") %>%
  pull(votu_id) %>% strsplit("\\|") %>% unlist %>% unique -> votu_selection

{
vir_func_cat<-phold_annot2$func %>% unique
vir_func_cat_colors<-hcl.colors(n=9, palette = "Zissou 1")
names(vir_func_cat_colors)<-vir_func_cat
vir_func_cat_colors["unknown function"]="white"
vir_func_cat_colors["Other"]="orange"
vir_func_cat_colors["Auxiliary genes and host takeover"]="red"
vir_func_cat_colors["Structural components"]="lightblue"
vir_func_cat_colors["integration and excision"]="#C05763"
vir_func_cat_colors["lysis"]="#FDF7DC"
vir_func_cat_colors["transcription regulation"]="#4C9168"
vir_func_cat_colors["DNA, RNA and nucleotide metabolism"]="#B7D886"
}

phold_annot2[votu %in% votu_selection] %>%
  group_by(votu) %>%
  summarise(max(end)) %>% arrange(votu)


vir_analysis[votu %in% votu_selection]
phold_annot2[votu %in% votu_selection] %>%
  mutate(annot=
           case_when(func == "Structural components" ~ product,
                     func == "Auxiliary genes and host takeover" ~ product,
                     func == "Other" ~ product,
                     ))->toplot
toplot$votu %<>% factor(levels=votu_selection)
require(gggenes)
ggplot(toplot,
       aes(xmin = start, xmax = end, y = votu, fill = func)) +
    geom_gene_arrow() +
  geom_gene_label(aes(label=product),min.size = 0) +
  scale_fill_manual(values=vir_func_cat_colors, name = "Virus gene\nfunctional category") +
  theme_genes() +
  facet_wrap(~ votu, scales = "free", ncol = 1) +
  theme(legend.position = "bottom") -> p;p

ggsave(plot = p,
       filename= "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/select_aux_votu_test.png",
       width=20, height = 30, units = "in", dpi = 600)
file.show("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/select_aux_votu_test.png")

```
# How many glycosyltransferases?
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
phold_annot2 %>%
  filter(grepl("Aux|Oth", func)) %>%
  group_by(product) %>%
  summarise(nvotu=length(unique(votu))) %>%
  arrange(desc(nvotu)) %>% print(n=50)
```

# How many vOTUs with the selected genes?
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
unique(unlist(strsplit(top_aux_gem_react_summary3$votu_id, "\\|"))) -> aux_votus

100*(length(aux_votus)/nrow(vir_analysis))

vir_analysis[votu %in% aux_votus, mean_vir_all] %>% sum

phold_annot2[grepl("Aux|Oth", func)]

```


# Analyse AUX functions in GEMS
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}

"Galactose-3-O-Sulfotransferase"


phold_annot2%>%
  filter(grepl("Galactose-3-O-Sulfotransferase", product, ignore.case=T)) %>% pull(votu) %>% unique-> votus

votus %>% length
vir_analysis[votu %in% votus] %>% view

blast_phold[grepl("Galactose-3-O-sulfotransferase", product), OTU] %>% unique %>% length

blastp_aux_gem_react_count %>%
  filter(grepl("Galactose", react_name, ignore.case=T))-> matches;
matches %>% datatable

tidy_reactions[react_id %in% matches$react_id] %>% 
  group_by(react_id, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25)

##################################################
"bifunctional heptose"

blast_phold[grepl("bifunctional heptose", product), OTU] %>% unique %>% length

blastp_aux_gem_react_count %>%
  filter(grepl("bifunctional heptose", product, ignore.case=T))-> matches;matches

tidy_reactions_f[react_id %in%  c("rxn03916_c0", "rxn03918_c0")] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") %>%
  datatable

gapseq_pwy_hierarchy[sbs=="PWY0-1241"]

##################################################
"phosphoheptose" #no clear match

blast_phold[grepl("phosphoheptose", product), OTU] %>% unique %>% length

blastp_aux_gem_react_count %>%
  filter(grepl("phosphoheptose", product, ignore.case=T))-> matches;matches 

tidy_reactions[react_id %in%  c("rxn03916_c0", "rxn03918_c0")] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
  res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

#################################################
"glutamine-D-fructose"
phold_annot2%>%
  filter(grepl("glutamine-D-fructose", product, ignore.case=T)) %>% pull(votu) %>% unique-> votus

votus %>% length
vir_analysis[votu %in% votus]

blastp_aux_gem_react_count %>%
  filter(grepl("glutamine-D-fructose", product, ignore.case=T))-> matches;matches

tidy_reactions_f[react_id %in% c("rxn00790_c0", "rxn00555_c0", "rxn15064_c0")] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions_f[react_id %in% c("rxn00790_c0", "rxn00555_c0", "rxn15064_c0")] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view
#################################################
"gamma-glutamyl cyclotransferase" # matches several oxido reductases acting on alcohols

blastp_aux_gem_react_count %>%
  filter(grepl("gamma-glutamyl cyclotransferase", product, ignore.case=T))-> matches;matches

tidy_reactions[react_id %in% matches$react_id] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions_f[react_id %in% matches$react_id] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("glutamyl",definition)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable


#################################################
"cysteine dioxygenase" # matches several different reactions, no reference to cysteine or alanine

blastp_aux_gem_react_count %>%
  filter(grepl("cysteine dioxygenase", product, ignore.case=T))-> matches;matches

tidy_reactions[react_id %in% matches$react_id] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions_f[react_id %in% matches$react_id] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("cysteine",react_name)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable


#################################################
"2OG" # no matches in GEMs
blast_phold%>%
  filter(grepl("2OG", product, ignore.case=T))

blastp_aux_gem_react_count %>%
  filter(grepl("2OG", product, ignore.case=T))-> matches;matches

tidy_reactions[react_id %in% matches$react_id] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions_f[react_id %in% matches$react_id] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("2-Oxoglutarate",react_name)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable


#################################################
"Porphyrin"
blast_phold%>%
  filter(grepl("Porphyrin", product, ignore.case=T))

blastp_aux_gem_react_count %>%
  filter(grepl("Porphyrin", product, ignore.case=T))-> matches;matches

tidy_reactions[react_id %in% matches$react_id] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(mean_fba=mean(abs(fba_flux)),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions[react_id %in% matches$react_id] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view


gem_analysis[OTU %in% 
               unique(tidy_reactions[react_id %in% matches$react_id, OTU])] %>%
  group_by(Order) %>% count(sort = T)

#################################################
"arabinose" 
phold_annot2%>%
  filter(grepl("arabinose", product, ignore.case=T)) %>% pull(votu) %>% unique-> votus
votus %>% length()
vir_analysis[votu %in% votus]


blastp_aux_gem_react_count %>%
  filter(grepl("arabinose", product, ignore.case=T))-> matches;matches

tidy_reactions[react_id %in% "rxn01117_c0"] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(mean_fba=mean(abs(fba_flux)),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions_f[react_id %in% "rxn01117_c0"] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("",react_name)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

#################################################
"S-adenosyl" 
blast_phold%>%
  filter(grepl("S-adenosyl", product, ignore.case=T))

blastp_aux_gem_react_count %>%
  filter(grepl("S-adenosyl", product, ignore.case=T))-> matches;matches

tidy_reactions[react_id %in% matches$react_id] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(mean_fba=median((fba_flux)),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions_f[react_id %in% matches$react_id] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("",react_name)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gem_analysis[OTU %in% 
               unique(tidy_reactions[react_id %in% matches$react_id, OTU])] %>%
  group_by(Order) %>% count(sort = T)

# Methanol
key_metabolites[name=="Putrescine"]
find_metab_react("cpd00118", tidy_reactions[]) %>% datatable

#################################################
"legionaminate" 
blast_phold%>%
  filter(grepl("legiona", product, ignore.case=T))

blastp_aux_gem_react_count %>%
  filter(grepl("legiona", product, ignore.case=T))-> matches;matches

tidy_reactions_f[react_id %in% matches$react_id] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(mean_fba=mean(abs(fba_flux)),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions[react_id %in% matches$react_id] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") %>%
  filter(grepl("phosphoenolpyruvate", react_name))  -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("",react_name)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

#################################################
"ferredoxin" 
phold_annot2%>%
  filter(grepl("ferredoxin", product, ignore.case=T)) %>% pull(votu) %>% unique-> votus

votus %>% length
vir_analysis[votu %in% votus] 


blastp_aux_gem_react_count %>%
  filter(grepl("ferredoxin", product, ignore.case=T))-> matches;matches

tidy_reactions_f[react_id %in% "rxn05937_c0"] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions_f[react_id %in% "rxn05937_c0"] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

mag_analysis[OTU %in% 
               unique(tidy_reactions_f[react_id %in% matches$react_id, OTU])] %>%
  group_by(Order,Family) %>% count(sort = T)

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("",react_name)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

#################################################
"sulfurylase"
blast_phold%>%
  filter(grepl("sulfurylase", product, ignore.case=T))

blastp_aux_gem_react_count %>%
  filter(grepl("sulfurylase", product, ignore.case=T))-> matches;matches

tidy_reactions[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions_f[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("",react_name)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

#################################################
"glutamine amidotransferase" # no matches in GEMs
blast_phold%>%
  filter(grepl("glutamine amidotransferase", product, ignore.case=T)) 

blastp_aux_gem_react_count %>%
  filter(grepl("glutamine amidotransferase", product, ignore.case=T))-> matches;matches

votus %>% length()
vir_analysis[votu %in% votus]


tidy_reactions[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("",react_name)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

#################################################
"enoyl-coa" # no matches in GEMs
blast_phold%>%
  filter(grepl("enoyl-coa", product, ignore.case=T))

blastp_aux_gem_react_count %>%
  filter(grepl("enoyl-coa", product, ignore.case=T))-> matches;matches

tidy_reactions[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("",react_name)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

#################################################
"pnuc" # no matches in GEMs
blast_phold%>%
  filter(grepl("pnuc", product, ignore.case=T))

blastp_aux_gem_react_count %>%
  filter(grepl("pnuc", product, ignore.case=T))-> matches;matches

tidy_reactions[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("",react_name)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable
#######################################
"methylamine" # no matches in GEMs
blast_phold%>%
  filter(grepl("methylamine", product, ignore.case=T))

blastp_aux_gem_react_count %>%
  filter(grepl("methylamine", product, ignore.case=T))-> matches;matches

tidy_reactions[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("",react_name)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

####################################################
"phosphofructokinase" # no matches in GEMs
blast_phold%>%
  filter(grepl("phosphofructokinase", product, ignore.case=T))

blastp_aux_gem_react_count %>%
  filter(grepl("phosphofructokinase", product, ignore.case=T))-> matches;matches

tidy_reactions[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("",react_name)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

####################################################
"cobalto" # no matches in GEMs
blast_phold%>%
  filter(grepl("cobalto", product, ignore.case=T))

blastp_aux_gem_react_count %>%
  filter(grepl("cobalto", product, ignore.case=T))-> matches;matches

tidy_reactions[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("",react_name)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

####################################################
"Fe-S" # no matches in GEMs
blast_phold%>%
  filter(grepl("Fe-S", product, ignore.case=T))

blastp_aux_gem_react_count %>%
  filter(grepl("Fe-S", product, ignore.case=T))-> matches;matches

tidy_reactions[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("",react_name)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

####################################################
"phosphoadenosine p"

phold_annot2%>%
  filter(grepl("phosphoadenosine", product, ignore.case=T)) %>% pull(votu) %>% unique-> votus
votus %>% length()
vir_analysis[votu %in% votus]

blastp_aux_gem_react_count %>%
  filter(grepl("phosphoadenosine", product, ignore.case=T))-> matches;matches

tidy_reactions_f[react_id %in% matches$react_id] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25)

tidy_reactions_f[react_id=="rxn00379_c0"] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

mag_analysis[OTU %in% 
               unique(tidy_reactions_f[react_id %in% matches$react_id, OTU])] %>%
  group_by(Order,Family) %>% count(sort = T)

#################################################
"nicotinamide-nucleo" # seems to have a fairly different function
blast_phold%>%
  filter(grepl("nicotinamide-nucleo", product, ignore.case=T))

blastp_aux_gem_react_count %>%
  filter(grepl("nicotinamide-nucleo", product, ignore.case=T))-> matches;matches

tidy_reactions[react_id %in% c("rxn03372_c0", "rxn00361_c0")] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("",react_name)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

####################################################
"nicotinamide phospho" # seems to have a fairly different function
blast_phold%>%
  filter(grepl("nicotinamide phosp", product, ignore.case=T))

blastp_aux_gem_react_count %>%
  filter(grepl("nicotinamide phosp", product, ignore.case=T))-> matches;matches

tidy_reactions_f[react_id %in% matches$react_id] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux) %>% round(4),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions[react_id %in% matches$react_id] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("",react_name)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

#################################################

```



#----
# GATHER RESULTS

##AUX: Merge blastp aux hits with gems genes
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
gem_react_genes %>% 
  group_by(OTU, react_id, react_name) %>%
  reframe(gene=strsplit(gene, ";") %>% unlist) -> gem_react_genes_l
```

# AUX genes co-occurring
```{r, echo = TRUE, warning=FALSE, message=FALSE, results='show', eval=F}
# Genes together
selected_aux1 %>%
  group_by(votu) %>%
  summarise(
    n_aux=n(),
    aux_genes=paste(product, collapse = ";")
  ) 

selected_aux1 %>%
  select(votu, product) %>%
  unique %>%
  add_column(p=1) %>%
  pivot_wider(names_from = product, values_from = p, values_fill = 0) %>%
  column_to_rownames("votu") %>% as.matrix -> aux_votu_mat
  aux_votu_mat[rowSums(aux_votu_mat)>1,]-> aux_votu_mat1
  
  Heatmap(aux_votu_mat1,
          col = c("white", "orange"),
          rect_gp = gpar(col = "gray20", lwd = 0.4),
          row_names_gp = gpar(fontsize=0),
          column_names_gp = gpar(fontsize=8),
          column_names_rot = 45,
          heatmap_width = unit(10, "in"),
          heatmap_height = unit(10, "in")
          ) 
```

# AUX: Distribution of AVG across predicted vOTUs
```{r, echo = TRUE, warning=FALSE, message=FALSE, results='show', eval=F}
iphop_mags2[] %>%
  select(votu=virus, Order) %>% 
  unique %>%
  left_join(phold_annot3) %>%
  select(votu, Order, func, product) %>% unique %>%
    filter(grepl("Aux|Oth", func)) %>%
  ungroup %>%
  group_by(Order) %>%
  count(product) %>%
  view
```

# DEF: Classify metab use by order
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
foreach(metabolite=unique(tidy_exhanges$id), .errorhandling = "remove")%do%{
  tidy_exhanges[id==metabolite] -> dt
  metab_name<-unique(dt$name)
  dt %>%
    group_by(Order) %>%
    summarise(
              median_imp=median(fba_flux[fba_flux<0]),
              median_exp=median(fba_flux[fba_flux>0]),
              n_imp=sum(fba_flux<0),
              n_exp=sum(fba_flux>0),
              n_use=n()
              ) %>%
    left_join(order_nmags) %>%
    filter(n>=5) %>%
    rowwise() %>%
    mutate(
      fisher = list(
        fisher.test(
          matrix(c(n_imp, n - n_imp,
                   n_exp, n - n_exp),
                 nrow = 2, byrow = TRUE)
        )
      ),
      p_value   = fisher$p.value,
      odds_ratio = fisher$estimate) %>%
    ungroup() %>%
    mutate(p_adj = p.adjust(p_value, method = "fdr")) %>%
    filter(p_value<=0.05) %>%
    add_column(id=metabolite, name=metab_name) %>%
    select(id, name, Order, median_imp, median_exp, n_imp, n_exp, n_use, n, odds_ratio, p_adj)
  } %>% rbindlist -> order_metab_fisher

order_metab_fisher %>%
  filter(odds_ratio>1) -> order_metab_fisher_imported
order_metab_fisher %>%
  filter(odds_ratio<1) -> order_metab_fisher_exported
  

# Number of orders classifiables
order_metab_fisher %>%
  pull(Order) %>% unique %>% length #20

# Average number of aminoacids classified
order_metab_fisher %>% 
  left_join(key_metabolites %>%
  select(name, simple_category) %>% unique)->order_metab_fisher1


# Check frequent substrates
order_metab_fisher1 %>%
  filter(odds_ratio>1) %>%
  group_by(simple_category, name) %>% 
  summarise(prop=100*(length(unique(Order))/20)) %>%
  filter(prop>=90) -> frequent_metab

frequent_metab %>%
  group_by(simple_category) %>% count(sort = T)


# how many frequent metabolites are detect?
frequent_metab %>%
  filter(!simple_category %in% c("Metals", "Inorganic compounds")) %>%
  pull(name) %>% length

intersect(
metab_react_matched_cpds3$name,
frequent_metab$name) %>% length

13/22

# What is their median fluxes
order_metab_fisher1[name %in% frequent_metab$name] %>%
    filter(odds_ratio>1) %>%
  group_by(name) %>%
    summarise(median=median(median_imp )) -> freq_median_f

table(freq_median_f$median<(-1))

```

# Top producers by Order
```{r eval=T, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
order_nmags %>%
  filter(n>=5) %>% pull(Order) -> teastable

foreach(i=unique(tidy_exhanges$name), .errorhandling = "remove")%do%{
  tidy_exhanges %>%
  #filter(abs(fba_flux) > 0.00) %>%
  filter(Order %in% teastable) %>%
  filter(name == i) -> dt
  g <- unique(dt$Order)
  do.call(rbind, lapply(g, function(gr) {
    x <- dt$fba_flux[dt$Order == gr]
    y <- dt$fba_flux[dt$Order != gr]
    data.frame(
      group = gr,
      n_group = length(x),
      n_rest = length(y),
      median_group = median(x, na.rm = TRUE),
      median_rest = median(y, na.rm = TRUE),
      p_value = wilcox.test(x, y, alternative = "greater", exact = FALSE)$p.value
    )
      }))  %>% 
    add_column(name=i) -> df
  df %>%
      add_column(p_adj=
        df$p_value %>% p.adjust(method = "fdr"))
  } %>% rbindlist() %>%
  filter(p_value<0.05 & n_group>=5) %>% filter(median_group>0) %>%
  mutate(p_adj=round(p_adj, 5)) %>%
    mutate(FC=round((median_rest-median_group)/median_group, 3))-> top_prod

```

# Top consumers by Order
```{r eval=T, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
order_nmags %>%
  filter(n>=5) %>% pull(Order) -> teastable

foreach(i=unique(tidy_exhanges$name), .errorhandling = "remove")%do%{
  tidy_exhanges %>%
  filter(Order %in% teastable) %>%
  #filter(fba_flux < 0.00) %>%
  filter(name == i) -> dt
  g <- unique(dt$Order)
  do.call(rbind, lapply(g, function(gr) {
    x <- dt$fba_flux[dt$Order == gr]
    y <- dt$fba_flux[dt$Order != gr]
    data.frame(
      group = gr,
      n_group = length(x),
      n_rest = length(y),
      median_group = median(x, na.rm = TRUE),
      median_rest = median(y, na.rm = TRUE),
      p_value = wilcox.test(x, y, alternative = "less", exact = FALSE)$p.value
    ) 
  })) %>% 
    add_column(name=i) -> df
  df %>%
      add_column(p_adj=
        df$p_value %>% p.adjust(method = "BH"))
  } %>% rbindlist() %>%
  filter(p_value<0.05) %>% filter(median_group<0) %>% 
  mutate(p_adj=round(p_adj, 5))%>%
    mutate(FC=round((median_rest-median_group)/median_group, 3))-> top_sub
```

# Produced plots
```{r eval=F, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
top_prod1<-rbind(top_prod[p_adj<=0.05],top_sub[p_adj<=0.05])

iphop_mags2 %>% 
  group_by(Phylum, Order) %>%
  summarise(sum=length(unique(virus))) %>%
    arrange(sum) %>% 
  filter(sum>=5 & 
         Order %in% mag_analysis$Order) %>% unique %>%
  pull(Order) %>% str_remove_all("o__")-> sorted_votus2

order_nmags %>%
  filter(n>=5) %>% pull(Order)-> testable

tidy_exhanges %>%
  filter(Order %in% testable) %>%
  filter(name=="Acetate") %>%
  mutate(votu=case_when(Order %in% top_prod1[name=="Acetate", group]~ T))-> dt

dt$fba_flux %>% median->med_p1
dt %>% group_by(Order) %>% 
    summarise(median=median(fba_flux)) %>%
  arrange((median)) -> sorted

ggplot(dt %>%
        mutate(Order=factor(Order, levels=sorted$Order)),
       aes(x = fba_flux, y = Order, fill = votu)) +
  geom_boxplot() +
  geom_vline(aes(xintercept=med_p1), color="red", linetype = "dashed", size=1)+
  theme_bw() +
  ggtitle("Acetate") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
  ylab("") +
  xlab("") +
  scale_fill_manual(
    values = c("lightblue"), na.value = "white")-> p1


#####################
tidy_exhanges %>%
  filter(Order %in% testable) %>%
  filter(name=="Succinate") %>%
  mutate(votu=case_when(Order %in% top_prod1[name=="Succinate", group] ~ T))-> dt

dt$fba_flux %>% median->med_p2
dt %>% group_by(Order) %>% 
    summarise(median=median(fba_flux)) %>%
  arrange((median)) -> sorted
ggplot(dt %>%
        mutate(Order=factor(Order, levels=sorted$Order)),
       aes(x = fba_flux, y = Order, fill = votu)) +
  geom_boxplot() +
  geom_vline(aes(xintercept=med_p2), color="red", linetype = "dashed", size=1)+
  theme_bw() +
  ggtitle("Succinate") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
  ylab("") +
  xlab("")+
  scale_fill_manual(
    values = c("lightblue"), na.value = "white")->p2

#####################
tidy_exhanges %>%
  filter(Order %in% testable) %>%
  filter(name=="Ammonium")%>%
  mutate(votu=case_when(Order%in%top_prod1[name=="Ammonium", group] ~ T))-> dt

dt$fba_flux %>% median->med_p3
dt %>% group_by(Order) %>% 
    summarise(median=median(fba_flux)) %>%
  arrange((median)) -> sorted
ggplot(dt %>%
        mutate(Order=factor(Order, levels=sorted$Order)),
       aes(x = fba_flux, y = Order, fill = votu)) +
  geom_boxplot() +
  geom_vline(aes(xintercept=med_p3), color="red", linetype = "dashed", size=1)+
  theme_bw() +
  ggtitle("Ammonium") +
  
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
  ylab("") +
  xlab(expression(atop("Exchange flux", mmol~g[DW]^{-1}~h^{-1}))) +
  scale_fill_manual(
    values = c("lightblue"), na.value = "white")->p3

#####################
tidy_exhanges %>%
  filter(Order %in% testable) %>%
  filter(name=="CO2")%>%
  mutate(votu=case_when(Order%in%top_prod1[name=="CO2", group] ~ T))-> dt

dt$fba_flux %>% median->med_p4

dt %>% group_by(Order) %>% 
    summarise(median=median(fba_flux)) %>%
  arrange((median)) -> sorted
ggplot(dt %>%
        mutate(Order=factor(Order, levels=sorted$Order)),
       aes(x = fba_flux, y = Order, fill = votu)) +
  geom_boxplot() +
  geom_vline(aes(xintercept=med_p4), color="red", linetype = "dashed", size=1)+
  theme_bw() +
  ggtitle("Carbon dioxide") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
  ylab("") +
  xlab(expression(atop("Exchange flux", mmol~g[DW]^{-1}~h^{-1}))) +
  scale_fill_manual(
    values = c("lightblue"), na.value = "white")->p4


require(patchwork)
(p1+p2) / (p3+p4)->pf

ggsave(plot = pf, filename = "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/boxplot_products_vir.png",
       height = 8,width = 10, units = "in", dpi = 600
       )
file.show("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/boxplot_products_vir.png")


ggsave(plot = pf, filename = "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/boxplot_products_vir.pdf",
       height = 8,width = 10, units = "in", dpi = 600
       )
file.show("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/boxplot_products_vir.pdf")


#####################
tidy_exhanges %>%
  filter(Order %in% testable) %>%
  filter(name=="Ethanol")%>%
  mutate(votu=case_when(Order%in%top_prod1[name=="Ethanol", group] ~ T))-> dt

dt$fba_flux %>% median->med_p4

dt %>% group_by(Order) %>% 
    summarise(median=median(fba_flux)) %>%
  arrange((median)) -> sorted
ggplot(dt %>%
        mutate(Order=factor(Order, levels=sorted$Order)),
       aes(x = fba_flux, y = Order, fill = votu)) +
  geom_boxplot() +
  geom_vline(aes(xintercept=med_p4), color="red", linetype = "dashed", size=1)+
  theme_bw() +
  ggtitle("Methanethiol") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
  ylab("") +
  xlab("Exchange flux [mmol/gDWh]") +
  scale_fill_manual(
    values = c("lightblue"), na.value = "white")
```

```{r eval=FALSE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}

#####
tidy_exhanges %>%
  filter(Order %in% testable) %>%
  filter(name=="Succinate")%>%
  mutate(votu=case_when(Order%in%top_prod[p_adj<=0.05 & name=="Succinate", group]  ~ "Produced",
                        Order%in%top_sub[p_adj<=0.05 & name=="Succinate", group] ~ "Consumed")
         )-> dt

dt$fba_flux %>% median->med_p1

dt %>% group_by(Order) %>% 
    summarise(median=median(fba_flux)) %>%
  arrange((median)) -> sorted
ggplot(dt %>%
        mutate(Order=factor(Order, levels=sorted$Order)),
       aes(x = fba_flux, y = Order, fill = votu)) +
  geom_boxplot() +
  geom_vline(aes(xintercept=med_p1), color="red", linetype = "dashed", size=1)+
  theme_bw() +
  ggtitle("Succinate") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
  ylab("") +
  xlab("") +
  scale_fill_manual(
    values = c(Produced="red3",Consumed="lightblue"), na.value = "white")-> p1
#####


tidy_exhanges %>%
  filter(Order %in% testable) %>%
  filter(name=="Acetate")%>%
  mutate(votu=case_when(Order%in%top_prod[p_adj<=0.05 & name=="Acetate", group]  ~ "Produced",
                        Order%in%top_sub[p_adj<=0.05 & name=="Acetate", group] ~ "Consumed")
         )-> dt

dt$fba_flux %>% median->med_p2

dt %>% group_by(Order) %>% 
    summarise(median=median(fba_flux)) %>%
  arrange((median)) -> sorted
ggplot(dt %>%
        mutate(Order=factor(Order, levels=sorted$Order)),
       aes(x = fba_flux, y = Order, fill = votu)) +
  geom_boxplot() +
  geom_vline(aes(xintercept=med_p2), color="red", linetype = "dashed", size=1)+
  theme_bw() +
  ggtitle("Acetate") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
  ylab("") +
  xlab("") +
  scale_fill_manual(
    values =  c(Produced="red3",Consumed="lightblue"), na.value = "white")-> p2

####

tidy_exhanges %>%
  filter(Order %in% testable) %>%
  filter(name=="Ammonium")%>%
  mutate(votu=case_when(Order%in%top_prod[p_adj<=0.05 & name=="Ammonium", group]  ~ "Produced",
                        Order%in%top_sub[p_adj<=0.05 & name=="Ammonium", group] ~ "Consumed")
         )-> dt

dt$fba_flux %>% median->med_p3

dt %>% group_by(Order) %>% 
    summarise(median=median(fba_flux)) %>%
  arrange((median)) -> sorted
ggplot(dt %>%
        mutate(Order=factor(Order, levels=sorted$Order)),
       aes(x = fba_flux, y = Order, fill = votu)) +
  geom_boxplot() +
  geom_vline(aes(xintercept=med_p3), color="red", linetype = "dashed", size=1)+
  theme_bw() +
  ggtitle("Ammonium") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
  ylab("") +
  xlab("Exchange flux [mmol/gDWh]") +
  scale_fill_manual(
    values =  c(Produced="red3",Consumed="lightblue"), na.value = "white")-> p3

####

tidy_exhanges %>%
  filter(Order %in% testable) %>%
  filter(name=="CO2")%>%
  mutate(votu=case_when(Order%in%top_prod[p_adj<=0.05 & name=="CO2", group]  ~ "Produced",
                        Order%in%top_sub[p_adj<=0.05 & name=="CO2", group] ~ "Consumed")
         )-> dt

dt$fba_flux %>% median->med_p4

dt %>% group_by(Order) %>% 
    summarise(median=median(fba_flux)) %>%
  arrange((median)) -> sorted
ggplot(dt %>%
        mutate(Order=factor(Order, levels=sorted$Order)),
       aes(x = fba_flux, y = Order, fill = votu)) +
  geom_boxplot() +
  geom_vline(aes(xintercept=med_p4), color="red", linetype = "dashed", size=1)+
  theme_bw() +
  ggtitle("Carbon dioxide") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
  ylab("") +
  xlab("Exchange flux [mmol/gDWh]") +
  scale_fill_manual(
    values =  c(Produced="red3",Consumed="lightblue"), na.value = "white") -> p4





(p1+p2) / (p3+p4)->pf

ggsave(plot = pf, filename = "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/boxplot_products_vir.png",
       height = 8,width = 10, units = "in", dpi = 600
       )
file.show("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/boxplot_products_vir.png")

```

#---
#FIGURES
## Metabolites ranked by flux
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
tidy_exhanges %>%
  group_by(name,simple_category) %>%
  summarise(median_pos=median(fba_flux[fba_flux>0], na.rm = T),
            median_neg=median(fba_flux[fba_flux<0], na.rm = T),
            prop_pos=round(100*(sum(fba_flux[fba_flux>0], na.rm = T)/311),2),
            prop_neg=round(100*(sum(fba_flux[fba_flux<0], na.rm = T)/311),2)
            ) %>%
  arrange(desc(median_pos)) -> exchanges_summary
```

##Heatmap: MetaB utilisation by order v2
```{r eval=FALSE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
#####  HM O2
require(ComplexHeatmap)

# vOTU abundance by host order
votu_host_order_abund<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/votu_host_order_mean.csv")
votu_host_order_abund$Order = str_remove(votu_host_order_abund$Order,"o__")
votu_host_order_abund_vec<-votu_host_order_abund$mean
names(votu_host_order_abund_vec) = votu_host_order_abund$Order

# Select thos with at least two votus
votu_order_count_all %>%
  filter(n_votu>=2) -> order_min2_votu

# Select which categories to plot
unique(tidy_exhanges$simple_category)
selected_classes<-c(
    "Polysaccharides", "Carbohydrates", "Sugar Alcohols",
    "Alcohols", "Carbonyl compounds",
    "Sulfur compounds",
    "N-Carbonyl compounds","Amino acids", "Nitrogen compounds", "Inorganic compounds")

tax_lev="Order"

tidy_exhanges[Order %in% order_min2_votu$Order] %>%
  filter(abs(fba_flux)>0.001) %>%
  group_by(name, Order, simple_category) %>%
  summarise(n_flux=n()) %>%
  left_join(order_nmags[,c("Order", "n")]) %>%
  mutate(prop=n_flux/n) %>%
  filter(prop>0.2) -> freq_exc

tidy_exhanges[Order %in% order_min2_votu$Order] %>%
  filter(abs(fba_flux)>0.001) %>%
  select(name, simple_category, !!sym(tax_lev), fba_flux) %>%
  filter(simple_category %in% selected_classes) %>%
  group_by(name, simple_category, !!sym(tax_lev)) %>%
  summarise(median=median(fba_flux)) %>%
  mutate(
    tax=as.factor(!!sym(tax_lev)),
    name=factor(name, levels=unique(tidy_exhanges$name)))-> medians

semi_join(medians, freq_exc, by=c("name", "Order")) %>%
  select(name, simple_category, Order, median, tax)-> medians

quant<-medians$median %>% quantile(seq(0,1,0.05));quant

medians %>%
  ungroup %>%
  mutate(f_cat=case_when(
    median>0 & median<=quant[["85%"]] ~ 0.5,
    median>quant[["85%"]] & median<=quant[["100%"]] ~ 1,
    median<0 & median>=quant[["35%"]] ~ -0.5,
    median>=quant[["0%"]] & median<quant[["35%"]] ~ -1
  )) %>% 
  dplyr::select(tax, name, f_cat) %>% data.table ->medians1 
  medians1$name %>% unique
  medians1[CJ(tax = unique(tax), name = levels(medians1$name)), on = .(tax, name)]->medians2
  medians2$name %>% unique
  medians2[is.na(f_cat), f_cat := 0]
  
  medians2 %>%
  pivot_wider(names_from = name, values_from = f_cat, values_fill = 0) %>%
  column_to_rownames("tax") %>%
  as.matrix  -> order_mat

tidy_exhanges[name %in% medians$name,c("name", "simple_category")] %>%
    filter(simple_category %in% selected_classes) %>%
  unique %>% filter(name %in% colnames(order_mat)) %>%
  mutate(simple_category=
           factor(
           simple_category,
         levels = 
          selected_classes)) %>% arrange(simple_category) -> metab_cat

metab_class_vec=metab_cat$simple_category
names(metab_class_vec)=metab_cat$name


# Define splitting vectors
gem_analysis$Phylum %>% as.factor->phy_vec
names(phy_vec)=gem_analysis %>% pull(!!sym(tax_lev))

# Define HM body colors
flux_colors=c(alpha("cadetblue",1),
              alpha("cadetblue",0.3),
              "white",
              alpha("red",0.4),
              alpha("red",1)
              )

# Transpose matrix
reorder_mat<-order_mat %>% t
reorder_mat<-reorder_mat[names(metab_class_vec),]

# Bottom annot

# Number of predicted host
votu_order_count_all %>%
  filter(Order %in% colnames(reorder_mat)) -> votu_hits
votu_vec<-votu_hits$n_votu
names(votu_vec) = votu_hits$Order


# Relative abundance by Order
abund_order_df<-gem_analysis %>%
  group_by(Order) %>%
  summarise(sum=sum(mean_mag_abund))
abund_order_df$sum->abund_order_df_vec
names(abund_order_df_vec)=abund_order_df$Order
abund_order_df_vec %>% sort

# nGEMs by order
ngems_order_df<-gem_analysis %>%
  group_by(Order) %>%
  summarise(sum=length(OTU))
ngems_order_df$sum->ngems_order_df_vec
names(ngems_order_df_vec)=ngems_order_df$Order
ngems_order_df_vec %>% sort


# Predicted biomass growth
biomass_order_df<-gem_analysis %>%
  group_by(Order) %>%
  summarise(mean=mean(growth_value))
biomass_order_df$mean->biomass_order_df_vec
names(biomass_order_df_vec)=biomass_order_df$Order


bottom_annot<-
columnAnnotation(
  
  
   "Number of vOTUs\n linked to order" = 
     anno_barplot(x = votu_vec[colnames(reorder_mat)],
                 gp=gpar(fill="gray80"),
                 axis_param = list(labels_rot = 0,
                                   gp = gpar(fontsize = 6),
                                   side = "right"
                                   #direction="reverse"
                                   ),
     ),
  

  "Avg. rel. abundance of vOTUs\n linked to order" = 
     anno_barplot(x = votu_host_order_abund_vec[colnames(reorder_mat)],
                 gp=gpar(fill="gray80"),
                 axis_param = list(labels_rot = 0,
                                   gp = gpar(fontsize = 6),
                                   side = "right"
                                   #direction="reverse"
                                   ),
     ),
  
    "MAG rel. abundance\n by order (%)" = 
    anno_barplot(x = abund_order_df_vec[colnames(reorder_mat)],
                 gp=gpar(fill="gray80"),
                 axis_param = list(labels_rot = 0,
                                   gp = gpar(fontsize = 6),
                                   side = "right"

                                   #direction="reverse"
                                   ),
                 ),
  
    spacer=anno_empty(border = F, height = unit(2, "mm")),
  show_legend = F,
  annotation_name_rot = 0,
  annotation_name_side = "right",
  annotation_name_gp = gpar(fontsize = 10),
  annotation_name_offset = unit(0.6, "cm"),
  gap=unit(2,"mm")
  )

# Top annot
HeatmapAnnotation(
  "MAG phylum" = phy_vec[colnames(reorder_mat)],
  which = "column",
  col = list("MAG phylum"=tree_colors),
  border = T,  show_legend = F,
  show_annotation_name = F,
    annotation_name_side = "right"
)-> top_annot

# Right annot
rownames(reorder_mat)%in%
metab_react_matched_cpds3[!is.na(avg_rel_abund), name] %>% as.character->metaB_presence_vec
names(metaB_presence_vec)=rownames(reorder_mat)


right_annot<-
rowAnnotation(
  #spacer=anno_empty(border = F, height = unit(0.5, "mm")),
  "Detected metabolites"=metaB_presence_vec[rownames(reorder_mat)],
  annotation_name_side = "bottom",
  col=list("Detected metabolites"=c("FALSE"="white","TRUE"="gray80")),
  border=T,
  show_legend = F)

# left annot
fba_median_pos<-exchanges_summary$median_pos
names(fba_median_pos)<-exchanges_summary$name

fba_median_neg<-exchanges_summary$median_neg %>% abs
names(fba_median_neg)<-exchanges_summary$name

fba_median_neg[fba_median_neg>30]
fba_median_neg[fba_median_neg>30] = 30

left_annot<-rowAnnotation(
  "Median export flux\n[mmol/gDWh]" = 
    anno_barplot(x = fba_median_pos[rownames(reorder_mat)],
                 gp=gpar(fill="gray80"),
                 axis_param = list(labels_rot = 90,
                                   gp = gpar(fontsize = 6),
                                   direction="reverse"
                                   )),
    "Median import flux\n[mmol/gDWh]" = 
    anno_barplot(x = fba_median_neg[rownames(reorder_mat)],
                gp=gpar(fill="gray80"),
                axis_param = list(labels_rot = 90,
                                  gp = gpar(fontsize = 6),
                                  direction="reverse"
                                  )),
  spacer=anno_empty(border = F, width = unit(1, "mm")),

  show_legend = F,
  annotation_name_rot = 90,
  annotation_name_side = "top",
  annotation_name_gp = gpar(fontsize = 10),
  annotation_name_offset = unit(0.1, "cm"),
  gap=unit(2,"mm")
  )


#reorder_mat<-order_mat[,(metab_cat$name)]
dim(order_mat[, names(metab_class_vec)])
Heatmap(
  reorder_mat,
  col=flux_colors, na_col = "white",
  rect_gp = gpar(col = "gray20", lwd = 0.3),
  border_gp = gpar(col = "black", lwd = 0.7),
  row_gap= unit(3, "mm"),
  column_gap= unit(3, "mm"),
  row_split = metab_class_vec,
  column_title_rot = 45,
  column_title_gp = gpar(fontsize=13),
  column_names_rot = 45,
  column_names_gp  = gpar(fontsize=13),
  column_title_side = "bottom",
  column_split = phy_vec[colnames(reorder_mat)],
  column_names_side = "top",
  row_title_rot=0,
  row_names_gp = gpar(fontsize=10),
  row_title_gp  = gpar(fontsize=13),
  border = T,
  cluster_rows = T,
  cluster_columns = T,
  cluster_column_slices = F,
  cluster_row_slices = F,
  show_column_dend = F,
  show_row_dend = F,
  heatmap_height = unit(19, "in"),
  #heatmap_width = unit(14, "in"),
  show_heatmap_legend = F,
  right_annotation = right_annot,
  top_annotation = bottom_annot,
  bottom_annotation = top_annot,
  left_annotation = left_annot
) ->hm
png(filename = 
      paste0(
      "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/hm_exchanges_test_v2.png"),
    width = 17, height = 20, units = "in", res = 300)
draw(hm, heatmap_legend_side = "bottom", annotation_legend_side = "bottom")
dev.off()

file.show("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/hm_exchanges_test_v2.png")

```

## Distribution of complete pws for key categories
```{r eval=FALSE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
# Import pwy data

files<-list.files(path = "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaG/coassemby_07feb24/mags_c70c10a99/gapseq/",
                  pattern ="-all-Pathways.tbl", full.names = T)
foreach(i = files)%do%{
fread(i) %>%
  filter(Prediction) %>% 
  mutate(sbs_ids = str_remove_all(ID, "\\|")) %>%
  left_join(sbs_hier1) %>%
    add_column(OTU=str_extract(i, "[:alpha:]{3}_SemiBin_[:digit:]+"))
  } %>% rbindlist -> pws

pws %>%
  select(OTU, Name, h2, h3, h4, h5, ReactionsFound) %>%
  left_join(gem_analysis[,c("OTU", "Order")]) %>%
  rowwise %>%
  mutate(n_react=length(unlist(strsplit(ReactionsFound, " ")))) %>% data.table -> pws1


# vOTU abundance by host order
votu_host_order_abund<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/votu_host_order_mean.csv")
votu_host_order_abund$Order = str_remove(votu_host_order_abund$Order,"o__")
votu_host_order_abund_vec<-votu_host_order_abund$mean
names(votu_host_order_abund_vec) = votu_host_order_abund$Order

# Select thos with at least two votus
votu_order_count_all %>%
  filter(n_votu>=2) %>% arrange(n_votu) -> order_min2_votu


# Select relevant pwys and summarise
pws1$h4 %>% unique
pws1[h3 == "CARBOXYLATES-DEG", Name] %>% unique %>% view
pws1[grepl("Xylan|Inulin|Pectin|Cellulose|Starch", Name, ignore.case = T) &
     !grepl("Biosynthesis", Name, ignore.case = T)] %>% add_column(type="Polysaccharide \ndegradation") -> df1

pws1[grepl("Calvin|Wood|Reductive TCA|3-hydroxypropanoate cycle", Name)] %>% add_column(type="CO2 fixation") -> df2

pws1[grepl("lactate ferment|pyruvate fermentation to acetate|ethanol|fermentation to butanoate|	
propanoate fermentation|acetate conversion", Name, ignore.case = F) & 
       h4 == c("Fermentation-to-Acids")] %>% add_column(type="Fermentation") -> df3

pws1[grepl("glyoxylate cycle", Name, ignore.case = F)] %>% add_column(type="Energy metabolism") -> df4

pws1[grepl("Tyrosine|Arginine|histidine", Name, ignore.case = T) & 
     !grepl("arginine decarboxylase", Name, ignore.case = T) &
     h4 %in% c("Proteinogenic-Amino-Acids-Degradation") | 
     h4 == "4-Aminobutyraye-Degradation"] %>% 
  add_column(type="Proteinogenic \namino acid \ndegradation") -> df5

pws1[h4 %in% c("Purine-Degradation", "Pyrimidine-Degradation") &
     !grepl("L-arginine degradation IV", Name, ignore.case = T)] %>% add_column(type="Nucleotide degradation")-> df6

pws1[grepl("trehalose|xylose|fucose|melibiose|lactose|cellobiose", Name, ignore.case = T) & 
      h4 %in% c("Sugars-And-Polysaccharides-Degradation") &
     !grepl("periplasmic|cytosolic|trehalose degradation IV", Name, ignore.case = T)
     ] %>% 
  add_column(type="Carbohydrate degradation")-> df7

rbind(df1, df2) %>%
  rbind(df3) %>%
  rbind(df4) %>%
  rbind(df5) %>%
  rbind(df6) %>%
  rbind(df7) %>%
    filter(Order %in% order_min2_votu$Order)-> df

df$type %>% unique

df %>%
  group_by(Order, Name, type) %>% 
  summarise(nmag = length(unique(OTU))) %>% 
  left_join(order_nmags, by= "Order") %>%
  mutate(p=100*(nmag/n)) %>% select(Order, Name, type) %>%
  unique %>% 
  mutate(p=case_when(type=="Polysaccharide \ndegradation"~ 1,
                     type=="CO2 fixation"~ 2,
                     type=="Fermentation"~ 3,
                     type=="Energy metabolism"~4,
                     type=="Nucleotide degradation"~ 5,
                     type=="Proteinogenic \namino acid \ndegradation"~ 6,
                     type=="Carbohydrate degradation" ~ 7
                     )) %>% select(Order, Name, p) -> pwy_count
  

pwy_count %>% pivot_wider(names_from = Order, values_from = p, values_fill = 0) %>% 
  column_to_rownames("Name")  %>% as.matrix -> mat


# Define splitting vectors columns
gem_analysis$Phylum %>% as.factor->phy_vec
names(phy_vec)=gem_analysis %>% pull(Order)

# Define row splits
df %>% select(Name, type) %>% unique -> rsplit
rsplit$type %<>% factor(levels = c(
                         "Polysaccharide \ndegradation",
                         "Carbohydrate degradation",
                         "CO2 fixation",
                         "Fermentation",
                         "Energy metabolism",
                         "Nucleotide degradation",
                         "Proteinogenic \namino acid \ndegradation"
                       ))
rsplit$type -> rsplit_vec
names(rsplit_vec)=rsplit %>% pull(Name)

# Top annot
HeatmapAnnotation(
  "MAG phylum" = phy_vec[colnames(mat)],
  which = "column",
  col = list("MAG phylum"=tree_colors),
  border = T,  show_legend = F,
  show_annotation_name = F,
    annotation_name_side = "right"
)-> top_annot



require(ComplexHeatmap)

library(circlize)
col_fun <- colorRamp2(
  breaks = c(0, 1, 2, 3, 4, 5, 6, 7),
  colors = c(
    "white",                  
    "green4",
    "gray50",
    "orange",
    "lightgoldenrod1",
    "lightblue",
    "lightcoral",
    "olivedrab3"
  ),transparency = 0.3
)

Heatmap(
  mat[rsplit$Name,],
  name="xxx",
  col = col_fun,
  rect_gp = gpar(col = "gray20", lwd = 0.5),
  border_gp = gpar(col = "black", lwd = 0.9),
  row_gap= unit(3, "mm"),
  column_gap= unit(3, "mm"),
  row_split = rsplit_vec[],
  column_title_rot = 45,
  column_names_rot = 45,
  #column_names_gp  = gpar(fontsize=2),
  column_split = phy_vec[colnames(mat)],
  column_title_side = "bottom",
  #column_title_gp = gpar(fontsize=1)
  #row_split = aux_gem_annot_vec[rownames(mag_vir_hm)],
  row_title_rot=0, 
  #row_title_gp = gpar(fontsize=1),
  row_names_gp = gpar(fontsize=9),
  column_names_side = "top",
  border = T,
  cluster_rows = T,
  cluster_columns = T,
  cluster_column_slices = F,
  cluster_row_slices = F,
  show_column_dend = F,
  show_row_dend = F,
  #heatmap_height = unit(3, "in"),
  heatmap_width = unit(11, "in"),
  show_heatmap_legend = F,
  top_annotation = top_annot
) -> hm;hm

png(filename = 
      paste0(
      "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/hm_pws.png"),
    width = 16, height = 12, units = "in", res = 600)
draw(hm)
dev.off()

file.show("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/hm_pws.png")

cairo_pdf(file = 
      paste0(
      "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/hm_pws.pdf"),
    width = 16, height = 12)
draw(hm)
dev.off()

file.show("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/hm_pws.pdf")
```


#---- METABOLITE UTILISATION ----
# SULFUR
## Examine substrates
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
selected_mets<-sorted_substrates_aero[simple_category=="Sulfur compounds"];selected_mets
met_report(selected_mets,
           fba_substrates_aero,
           fba_reactions_f[sim_O2==T & anaerobic==F],
           c("consumed","produced", "transported"),
           ko_f_aero, fisher_aero)

selected_mets<-sorted_substrates_anaero[simple_category=="Sulfur compounds"];selected_mets
met_report(selected_mets,
           fba_substrates_anaero,
           fba_reactions_f[sim_O2==F & anaerobic==T],
           "consumed",
           ko_f_anaero, fisher_anaero)

# Sulfate
key_metabolites[name=="Sulfate"]
find_metab_react("cpd00048", tidy_reactions[]) %>% datatable
tidy_reactions[react_name=="ATP:sulfate adenylyltransferase"] %>%
  group_by(Order) %>% count %>% arrange(desc(n)) %>% print(n=50)


key_metabolites[name=="APS"]
find_metab_react("cpd00193", tidy_reactions[]) %>% datatable
fba_reactions[react_id=="rxn00379_c0"] %>%
  group_by(react_name) %>% count

key_metabolites[name=="3-phosphoadenylylsulfate"]
find_metab_react("cpd00044", tidy_reactions[]) %>% datatable

# CHECK PAPS reductases
tidy_reactions[react_id=="rxn05239_c0"] %>%
  select(OTU, react_name, Phylum, Order) %>% unique-> sel_react

sel_react$OTU %>% unique %>% length
sel_react %>% group_by(Phylum) %>% 
  count(sort = T) %>%
left_join(phy_count) %>% mutate(100*(n/n_phyla))

sel_react %>% group_by(Order) %>% 
  count(sort = T) %>%
left_join(order_count) %>% mutate(prop=100*(n/n_order))

tidy_reactions[react_id=="rxn09069_c0"] %>% 
  select(OTU, react_name, Phylum, Order) %>% unique-> sel_react

sel_react$OTU %>% unique %>% length
sel_react %>% group_by(Phylum) %>% 
  count(sort = T) %>%
left_join(phy_count) %>% mutate(100*(n/n_phyla))

sel_react %>% group_by(Order) %>% 
  count(sort = T) %>%
left_join(order_count) %>% mutate(prop=100*(n/n_order))


key_metabolites[name=="Taurine"]
find_metab_react("cpd00210", tidy_reactions) %>% datatable
#Is taurine linked to O2?
tidy_exhanges[name=="Taurine", ]

key_metabolites[name=="Sulfoacetaldehyde"]
find_metab_react("cpd00460", tidy_reactions) %>% datatable

key_metabolites[name=="Sulfoacetate"]
find_metab_react("cpd09878", tidy_reactions) %>% datatable

# Sulfide
key_metabolites[name=="L-Methionine"]
find_metab_react("cpd00060", tidy_reactions) %>% datatable


key_metabolites[name=="S-Adenosyl-L-methionine"]
find_metab_react("cpd00017", tidy_reactions) %>% datatable

# How produces more S-Methilthioribose?
key_metabolites[grep("5-Methylthio-D-ribose", name)]
tidy_exhanges[react_id=="EX_cpd01981_e0", Order] %>% table %>% sort

tidy_reactions_f[react_id=="rxn05909_c0"] %>% 
  select(OTU, react_name, Phylum, Order) %>% unique-> sel_react

sel_react$OTU %>% unique %>% length
sel_react %>% group_by(Phylum) %>% 
  count(sort = T) %>%
left_join(phy_count) %>% mutate(100*(n/n_phyla))


key_metabolites[name=="L-Cysteine"]
find_metab_react("cpd00084", tidy_reactions) %>% datatable

key_metabolites[grepl("Cyste",name)]
find_metab_react("cpd00084", tidy_reactions) %>% datatable
# Cysteine conversion to Serine and Pyruvate
tidy_reactions_f[react_id=="rxn05909_c0"] %>% 
  select(OTU, react_name, Phylum, Order) %>% unique-> sel_react

sel_react$OTU %>% unique %>% length
sel_react %>% group_by(Phylum) %>% 
  count(sort = T) %>%
left_join(phy_count) %>% mutate(100*(n/n_phyla))

sel_react %>% group_by(Order) %>% 
  count(sort = T) %>%
left_join(order_count) %>% mutate(prop=100*(n/n_order))


tidy_reactions_f[react_id=="rxn00566_c0"] %>% 
  select(OTU, react_name, Phylum, Order) %>% unique-> sel_react

sel_react$OTU %>% unique %>% length
sel_react %>% group_by(Phylum) %>% 
  count(sort = T) %>%
left_join(phy_count) %>% mutate(100*(n/n_phyla))

sel_react %>% group_by(Order) %>% 
  count(sort = T) %>%
left_join(order_count) %>% mutate(prop=100*(n/n_order)) %>% print(n=50)


# Virus
#Sulfate
blast_phold[grepl("sulfate",product)]
gem_react_genes[OTU=="nat_SemiBin_881" & grepl("megahit_391107", gene)] %>%
  left_join(gem_analysis[,c("OTU", "Order")])

phold_annot2[grepl("phosphoadenosine phosphosulfate reductase", product)]

phold_annot2[grepl("Galactose-3-O-sulfotransferase", product), votu] %>%
  unique->votus
length(votus)
phold_annot2[votu%in% votus] %>% group_by(product) %>% 
  count(sort = T) %>% print(n=200)
vir_analysis[votu%in%votus, c("Class", "size_kb", "host_taxonomy")] 

phold_annot2[grepl("phosphoadenosine phosphosulfate reductase", product), votu] %>% unique->votus
length(votus)
phold_annot2[votu%in% votus] %>% 
  group_by(product) %>% 
  count(sort = T) %>% print(n=200)
vir_analysis[votu%in%votus, c("Class", "size_kb", "host_taxonomy")]

phold_annot2[grepl("S-adenosyl", product), votu] %>% unique->votus
length(votus)
phold_annot2[votu%in% votus] %>% 
  group_by(product) %>% 
  count(sort = T) %>% print(n=200)
vir_analysis[votu%in%votus, c("Class", "size_kb", "host_taxonomy")]

phold_annot2[grepl("cyst", product), votu] %>% unique->votus
length(votus)
phold_annot2[votu%in% votus] %>% 
  group_by(product) %>% 
  count(sort = T) %>% print(n=200)
vir_analysis[votu%in%votus, c("Class", "size_kb", "host_taxonomy")]
```

## Examine products
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
selected_mets<-sorted_products_aero[simple_category=="Sulfur compounds"];selected_mets
met_report(selected_mets,
           fba_products_aero,
           fba_reactions_f[sim_O2==T & anaerobic==F],
           "produced",
           ko_f_aero, fisher_aero)

selected_mets<-sorted_products_anaero[simple_category=="Sulfur compounds"];selected_mets
met_report(selected_mets,
           fba_products_anaero,
           fba_reactions_f[sim_O2==F & anaerobic==T],
           "produced",
           ko_f_anaero, fisher_anaero)

find_metab_react("cpd00239", tidy_reactions) %>% datatable

key_metabolites[name=="5-Methylthio-D-ribose"]
find_metab_react("cpd01981", tidy_reactions) %>% datatable
```

## Examine mixed
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
selected_mets<-sorted_mixed_aero[simple_category=="Sulfur compounds"];selected_mets
met_report(selected_mets,
           fba_comp_metab1[sim_O2==T & anaerobic==F],
           fba_reactions_f[sim_O2==T & anaerobic==F],
           c("consumed", "produced", "transported"),
           ko_f_aero, fisher_aero)

selected_mets<-sorted_mixed_anaero[simple_category=="Sulfur compounds"];selected_mets
met_report(selected_mets,
           fba_comp_metab1[sim_O2==F & anaerobic==T],
           fba_reactions_f[sim_O2==F & anaerobic==T],
           c("consumed", "produced", "transported"),
           ko_f_anaero, fisher_anaero)
```



#-
# CARBOHYDRATES
## Examine substrates
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
selected_mets<-sorted_substrates_aero[simple_category=="Polysaccharides"];selected_mets
selected_mets<-sorted_substrates_aero[simple_category=="Carbohydrates"];selected_mets
selected_mets<-sorted_substrates_aero[simple_category=="Sugar Alcohols"];selected_mets

met_report(selected_mets,
           fba_substrates_aero,
           fba_reactions_f[sim_O2==T & anaerobic==F],
           "consumed",
           ko_f_aero, fisher_aero)


selected_mets<-sorted_substrates_anaero[simple_category=="Polysaccharides"];selected_mets
selected_mets<-sorted_substrates_anaero[simple_category=="Carbohydrates"];selected_mets
selected_mets<-sorted_substrates_anaero[simple_category=="Sugar Alcohols"];selected_mets


met_report(selected_mets,
           fba_substrates_anaero,
           fba_reactions_f[sim_O2==F & anaerobic==T],
           "consumed",
           ko_f_anaero, fisher_anaero)

# POlysaccharides
tidy_reactions_f[grepl("starch", definition), react_name] %>% table
tidy_reactions_f[grepl("starch", definition), Order] %>% table %>% sort

tidy_reactions_f[grepl("Xyla", definition), react_name] %>% table
tidy_reactions_f[grepl("Xyla", definition), Order] %>% table %>% sort

tidy_reactions_f[grepl("Inulin", definition), react_name] %>% table
tidy_reactions_f[grepl("Inulin", definition), Order] %>% table %>% sort

#Monosaccharides
key_metabolites[name=="D-Glucose"]
find_metab_react("cpd00027", tidy_reactions) %>% datatable

key_metabolites[name=="Galactose"]
find_metab_react("cpd00108", tidy_reactions) %>% datatable

key_metabolites[name=="Dulcose"]
find_metab_react("cpd01171", tidy_reactions) %>% datatable

key_metabolites[name=="D-Fructose", id, name]
find_metab_react("cpd00082", tidy_reactions) %>% datatable


key_metabolites[name=="Sorbitol", id, name]
find_metab_react("cpd00588", tidy_reactions) %>% datatable


key_metabolites[name=="D-fructose-6-phosphate", id, name]
find_metab_react("cpd00072", tidy_reactions) %>% datatable

tidy_reactions[grepl("ATP:D-fructose 1-phosphotransferase", react_name)]

tidy_reactions[grepl("gamma-glutamylaminopeptidase", react_name)]
key_metabolites[name=="Cys-Gly", id, name]
find_metab_react("cpd01017", tidy_reactions) %>% datatable


key_metabolites[grepl("Oxoproline",name), id, name]
find_metab_react("cpd01293", tidy_reactions) %>% datatable

```

## Examine products
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
selected_mets<-sorted_products_aero[simple_category=="Polysaccharides"];selected_mets
selected_mets<-sorted_products_aero[simple_category=="Carbohydrates"];selected_mets
selected_mets<-sorted_products_aero[simple_category=="Sugar Alcohols"];selected_mets

met_report(selected_mets,
           fba_products_aero,
           fba_reactions_f[sim_O2==T & anaerobic==F],
           "produced",
           ko_f_aero, fisher_aero)

selected_mets<-sorted_products_anaero[eval(exp)];selected_mets
met_report(selected_mets,
           fba_products_anaero,
           fba_reactions_f[sim_O2==F & anaerobic==T],
           "produced",
           ko_f_anaero, fisher_anaero)
```

## Examine mixed
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
selected_mets<-sorted_mixed_aero[simple_category=="Polysaccharides"];selected_mets
met_report(selected_mets,
           fba_comp_metab1[sim_O2==T & anaerobic==F],
           fba_reactions_f[sim_O2==T & anaerobic==F],
           c("consumed", "produced", "transported"),
           ko_f_aero, fisher_aero)

selected_mets<-sorted_mixed_anaero[eval(exp)];selected_mets
met_report(selected_mets,
           fba_comp_metab1[sim_O2==F & anaerobic==T],
           fba_reactions_f[sim_O2==F & anaerobic==T],
           c("consumed", "produced", "transported"),
           ko_f_aero, fisher_anaero)
```

#-
# ALCOHLS
## Examine substrates
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
selected_mets<-sorted_substrates_aero[simple_category=="Alcohols"];selected_mets
met_report(selected_mets,
           fba_substrates_aero,
           fba_reactions_f[sim_O2==T & anaerobic==F],
           "consumed",
           ko_f_aero, fisher_aero)


selected_mets<-sorted_substrates_anaero[simple_category=="Alcohols"];selected_mets
met_report(selected_mets,
           fba_substrates_anaero,
           fba_reactions_f[sim_O2==F & anaerobic==T],
           "consumed",
           ko_f_anaero, fisher_anaero)

# Methanol
key_metabolites[name=="Methanol"]
find_metab_react("cpd00116", tidy_reactions[]) %>% datatable

#Converted all to formaldehyde
tidy_exhanges[name=="Formaldehyde", ] 

key_metabolites[name=="Formaldehyde"]
find_metab_react("cpd00055", tidy_reactions[]) -> res; res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$sbs_ids , ";"))] %>% view

```

## Examine products
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
selected_mets<-sorted_products_aero[simple_category=="Alcohols"];selected_mets
met_report(selected_mets,
           fba_products_aero,
           fba_reactions_f[sim_O2==T & anaerobic==F],
           "produced",
            ko_f_aero, fisher_aero)


selected_mets<-sorted_products_anaero[eval(exp),];selected_mets
met_report(selected_mets,
           fba_products_anaero,
           fba_reactions_f[sim_O2==F & anaerobic==T],
           "produced",
           ko_f_aero, fisher_aero)
react_hm(react_df=fba_reactions_f[sim_O2==T & anaerobic==F])
```

## Examine mixed
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
selected_mets<-sorted_mixed_aero[eval(exp)];selected_mets
met_report(selected_mets,
           fba_comp_metab1[sim_O2==T & anaerobic==F],
           fba_reactions_f[sim_O2==T & anaerobic==F],
           c("consumed", "produced", "transported"),
           ko_f_aero, fisher_aero)

selected_mets<-sorted_mixed_anaero[eval(exp)];selected_mets
met_report(selected_mets,
           fba_comp_metab1[sim_O2==F & anaerobic==T],
           fba_reactions_f[sim_O2==F & anaerobic==T],
           c("consumed", "produced", "transported"),
           ko_f_aero, fisher_anaero)
```
#-
# CARBOXYLS
## Examine substrates
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
selected_mets<-sorted_substrates_aero[simple_category=="Carbonyl compounds",];selected_mets
met_report(selected_mets,
           fba_substrates_aero,
           tidy_reactions_f,
           "consumed",
           ko_f_aero, fisher_aero)

 selected_mets<-sorted_substrates_anaero[simple_category=="Carbonyl compounds"];selected_mets
met_report(selected_mets,
           fba_substrates_anaero,
           fba_reactions_f[sim_O2==F & anaerobic==T],
           "consumed",
           ko_f_anaero, fisher_anaero)


substrate_fba_median
############# Formate ################
key_metabolites[name=="Formate"]
find_metab_react("cpd00047", tidy_reactions[]) -> res; res %>% datatable
gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$sbs_ids , ";"))] %>% view

############# Lactate ################
key_metabolites[name=="L-Lactate"]
find_metab_react("cpd00159", tidy_reactions[]) -> res; res %>% datatable
gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$sbs_ids , ";"))] %>% view

############# Malate ################
key_metabolites[name=="L-Malate"]
find_metab_react("cpd00130", tidy_reactions_f[]) -> res; res %>% datatable
gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$sbs_ids , ";"))] %>% view

key_metabolites[name=="Oxaloacetate"]
find_metab_react("cpd00032", tidy_reactions[]) -> res; res %>% datatable

key_metabolites[name=="Malyl-CoA"]
find_metab_react("cpd02662", tidy_reactions[]) -> res; res %>% datatable

key_metabolites[name=="Glyoxalate"]
find_metab_react("cpd00040", tidy_reactions[]) -> res; res %>% datatable


key_metabolites[name=="Sinapoyl malate"]
find_metab_react("cpd01849", tidy_reactions[]) -> res; res %>% datatable

############# 2-Oxoglutarate ################
key_metabolites[name=="2-Oxoglutarate"]
find_metab_react("cpd00024", tidy_reactions_f[]) -> res; res %>% datatable
gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$sbs_ids , ";"))] %>% view

```

## Examine products
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
selected_mets<-sorted_products_aero[simple_category=="Carbonyl compounds",];selected_mets
met_report(selected_mets,
           fba_products_aero,
           fba_reactions_f[sim_O2==T & anaerobic==F],
           "produced",
            ko_f_aero, fisher_aero)


selected_mets<-sorted_products_anaero[simple_category=="Carbonyl compounds",];selected_mets
met_report(selected_mets,
           fba_products_anaero,
           fba_reactions_f[sim_O2==F & anaerobic==T],
           "produced",
           ko_f_anaero, fisher_anaero)
react_hm(react_df=fba_reactions_f[sim_O2==T & anaerobic==F])
react_hm(react_df=fba_reactions_f[sim_O2==F & anaerobic==T])

```

## Examine mixed
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
selected_mets<-sorted_mixed_aero[simple_category=="Carbonyl compounds"];selected_mets
met_report(selected_mets,
           fba_comp_metab1[sim_O2==T & anaerobic==F],
           fba_reactions_f[sim_O2==T & anaerobic==F],
           c("consumed", "produced", "transported"),
           ko_f_aero, fisher_aero)

selected_mets<-sorted_mixed_anaero[simple_category=="Carbonyl compounds"];selected_mets
met_report(selected_mets,
           fba_comp_metab1[sim_O2==F & anaerobic==T],
           fba_reactions_f[sim_O2==F & anaerobic==T],
           c("consumed", "produced", "transported"),
           ko_f_aero, fisher_anaero)
```
#-
# Metals
## Examine substrates
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
selected_mets<-sorted_substrates_aero[simple_category=="Metals",
                                      ];selected_mets

met_report(selected_mets,
           fba_substrates_aero,
           tidy_reactions,
           c("consumed", "produced", "transported"),
           ko_f_aero, fisher_aero)

met_report(selected_mets,
           fba_substrates_aero,
           fba_reactions[sim_O2==F & anaerobic==T],
           c("consumed", "produced", "transported"),
           ko_f_aero, fisher_aero)

# Who uses cobalt for biosynth?
#Cobaltochelatase
tidy_reactions_f[react_id=="rxn04045_c0"]%>%
  group_by(Order) %>% count(sort = T)

#hydrogenobyrinic-acid-a,c-diamide:cobalt cobalt-ligase (ADP-forming)
tidy_reactions_f[react_id=="rxn03542_c0"] %>%
  group_by(Order) %>% count(sort = T)

# Fe3+ oxidation

tidy_reactions_f[react_id=="rxn00068_c0"]%>%
  group_by(Order) %>% count(sort = T)

# Mg in chlorophyllide
tidy_reactions_f[react_id=="rxn02733_c0"]%>%
  group_by(Order) %>% count(sort = T) #nope
```
#-
# N-carbonyls
## Examine substrates
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
selected_mets<-sorted_substrates_aero[simple_category=="N-Carbonyl compounds"];selected_mets

met_report(selected_mets,
           fba_substrates_aero,
           fba_reactions_f[sim_O2==T & anaerobic==F],
           "consumed",
           ko_f_aero, fisher_aero)


selected_mets<-sorted_substrates_anaero[simple_category=="N-Carbonyl compounds"];selected_mets
met_report(selected_mets,
           fba_substrates_anaero,
           fba_reactions_f[sim_O2==F & anaerobic==T],
           "consumed",
           ko_f_anaero, fisher_anaero)


# Follow betaine degradation
res$react_summary[metabolite=="Betaine"] %>% datatable

key_metabolites[grepl("Betaine",name, ignore.case=T)]
find_metab_react(met_id="cpd00540",
                 reactions =  tidy_reactions_f) %>% datatable()

key_metabolites[grepl("Betaine aldehyde",name, ignore.case=T)]
find_metab_react(met_id="cpd00447",
                 reactions =  tidy_reactions) %>% datatable()

key_metabolites[grepl("Choline",name, ignore.case=T)]
find_metab_react(met_id="cpd00098",
                 reactions =  tidy_reactions) -> res;res%>%datatable()

sbs_hier1[sbs %in% unlist(strsplit(res$sbs_ids , ";")), ]

key_metabolites[grepl("Methylamine",name, ignore.case=T)]
find_metab_react(met_id="cpd00441",
                 reactions =  tidy_reactions) %>% datatable()

key_metabolites[grepl("Dimethylglycine",name, ignore.case=T)]
find_metab_react(met_id="cpd00756",
                 reactions =  tidy_reactions) %>% datatable()


# Follow S-Adenosyl-L-methionine
fba_reactions[grepl("S-Adenosyl-L-methionine",react_name)] %>% 
  group_by(react_name,definition) %>% count 


phold_annot2[votu %in%
blast_phold[grepl("S-adenosylmethionine decarboxylase", product), votu]] 


fba_reactions[grepl("methioninamine",react_name)] %>% 
  group_by(react_name, definition) %>% count 

key_metabolites[grepl("Niacin",name, ignore.case=T)]
find_metab_react(met_id="cpd00218",
                 reactions =  fba_reactions_f[sim_O2==T & anaerobic==F]) %>% datatable()

fba_reactions[grepl("nicotinamide",react_name)] %>% 
  group_by(react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) 

fba_reactions[grepl("methylamine",react_name)] %>% 
  group_by(react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) 

```

## Examine products 
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
selected_mets<-sorted_products_aero[simple_category=="N-Carbonyl compounds"];selected_mets
met_report(selected_mets,
           fba_products_aero,
           fba_reactions_f[sim_O2==T & anaerobic==F],
           "produced",
           ko_f_aero, fisher_aero)

selected_mets<-sorted_products_anaero[simple_category=="N-Carbonyl compounds"];selected_mets
met_report(selected_mets,
           fba_products_anaero,
           fba_reactions_f[sim_O2==F & anaerobic==T],
           "produced",
           ko_f_anaero, fisher_anaero) 
react_hm(react_df=fba_reactions_f[sim_O2==T & anaerobic==F])

key_metabolites[grepl("Urea",name, ignore.case=T)]
find_metab_react(met_id="cpd00073",
                 reactions =  tidy_reactions) -> res;res%>%datatable()
key_metabolites[grepl("Ornithine",name, ignore.case=T)]
find_metab_react(met_id="cpd00064",
                 reactions =  tidy_reactions) -> res;res%>%datatable()
```

## Examine mixed
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
selected_mets<-sorted_mixed_aero[simple_category=="N-Carbonyl compounds"];selected_mets
met_report(selected_mets,
           fba_comp_metab1[sim_O2==T & anaerobic==F],
           fba_reactions_f[sim_O2==T & anaerobic==F],
           c("consumed", "produced", "transported"),
           ko_f_aero, fisher_aero)

selected_mets<-sorted_mixed_anaero[simple_category=="N-Carbonyl compounds"];selected_mets
met_report(selected_mets,
           fba_comp_metab1[sim_O2==F & anaerobic==T],
           fba_reactions_f[sim_O2==F & anaerobic==T],
           c("consumed", "produced", "transported"),
           ko_f_aero, fisher_anaero)
```
# -
# Amino acids
## Examine substrates
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
selected_mets<-sorted_substrates_aero[simple_category=="Amino acids"];selected_mets

met_report(selected_mets,
           fba_substrates_aero,
           fba_reactions_f[sim_O2==T & anaerobic==F],
           "consumed",
           ko_f_aero, fisher_aero)

res$react_summary[metabolite=="L-Methionine"] %>% datatable
res$react_summary[metabolite=="L-Cysteine"] %>% datatable


selected_mets<-sorted_substrates_anaero[simple_category=="Amino acids"];selected_mets
met_report(selected_mets,
           fba_substrates_anaero,
           fba_reactions_f[sim_O2==F & anaerobic==T],
           "consumed",
           ko_f_anaero, fisher_anaero)

# Follow methionine degradation
key_metabolites[grepl("L-Methionine",name, ignore.case=T)]
find_metab_react(met_id="cpd00060",
                 reactions =  tidy_reactions_f) -> res;res%>%datatable()
sbs_hier1[sbs %in% unlist(strsplit(res$sbs_ids , ";")), ]

key_metabolites[grepl("2-Aminobut-2-enoate",name, ignore.case=T)]
find_metab_react(met_id="cpd17424",
                 reactions =  fba_reactions_f[sim_O2==T & anaerobic==F]) %>% datatable()

key_metabolites[grepl("2-Oxobutyrate",name, ignore.case=T)]
find_metab_react(met_id="cpd00094",
                 reactions =  fba_reactions_f[sim_O2==T & anaerobic==F]) %>% datatable()

key_metabolites[grepl("S-Adenosyl-L-methionine",name, ignore.case=T)]
find_metab_react(met_id="cpd00017",
                 reactions =  fba_reactions_f[sim_O2==T & anaerobic==F]) -> res;res%>%datatable()


sbs_hier1[sbs %in% unlist(strsplit(res$sbs_ids , ";")), ]

######################################################
# Follow threonine degradation
key_metabolites[grepl("L-threonine",name, ignore.case=T)]
find_metab_react(met_id="cpd00161",
                 reactions =  fba_reactions_f[sim_O2==T & anaerobic==F]) -> res;res%>%datatable()
sbs_hier1[sbs %in% unlist(strsplit(res$sbs_ids , ";")), ]


key_metabolites[grepl("Amino-acetoacetate",name, ignore.case=T)]
find_metab_react(met_id="cpd02211",
                 reactions =  fba_reactions_f[sim_O2==T & anaerobic==F]) -> res;res%>%datatable()
sbs_hier1[sbs %in% unlist(strsplit(res$sbs_ids , ";")), ]

######################################################

key_metabolites[grepl("L-Histidine",name, ignore.case=T)]
find_metab_react(met_id="cpd00119",
                 reactions =  tidy_reactions_f) -> res;res%>%datatable()
sbs_hier1[sbs %in% unlist(strsplit(res$sbs_ids , ";")), ]


######################################################
key_metabolites[grepl("L-Alanine",name, ignore.case=T)]
find_metab_react(met_id="cpd00035",
                 reactions =  tidy_reactions_f) -> res;res%>%datatable()
sbs_hier1[sbs %in% unlist(strsplit(res$sbs_ids , ";")), ]
######################################################
key_metabolites[grepl("L-Serine",name, ignore.case=T)]
find_metab_react(met_id="cpd00054",
                 reactions =  tidy_reactions_f) -> res;res%>%datatable()
sbs_hier1[sbs %in% unlist(strsplit(res$sbs_ids , ";")), ]

######################################################
key_metabolites[grepl("L-Glutamate",name, ignore.case=T)]
find_metab_react(met_id="cpd00023",
                 reactions =  tidy_reactions_f) -> res;res%>%datatable()
sbs_hier1[sbs %in% unlist(strsplit(res$sbs_ids , ";")), h4] %>% table %>% sort
######################################################

key_metabolites[grepl("Cysteate",name, ignore.case=T)]
find_metab_react(met_id="cpd00051",
                 reactions =  fba_reactions_f[sim_O2==T & anaerobic==F]) %>% datatable()

fba_reactions[grepl("S-Adenosyl-L-methionine",react_name)] %>% 
  group_by(react_name,definition) %>% count 


phold_annot2[votu %in%
blast_phold[grepl("S-adenosylmethionine decarboxylase", product), votu]] 


fba_reactions[grepl("methioninamine",react_name)] %>% 
  group_by(react_name, definition) %>% count 

key_metabolites[grepl("Niacin",name, ignore.case=T)]
find_metab_react(met_id="cpd00218",
                 reactions =  fba_reactions_f[sim_O2==T & anaerobic==F]) %>% datatable()

fba_reactions[grepl("nicotinamide",react_name)] %>% 
  group_by(react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% view

fba_reactions[grepl("methylamine",react_name)] %>% 
  group_by(react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% view

```

## Examine products (NO products)
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
selected_mets<-sorted_products_aero[simple_category=="N-Carbonyl compounds"];selected_mets
met_report(selected_mets,
           fba_products_aero,
           fba_reactions_f[sim_O2==T & anaerobic==F],
           "produced",
           ko_f_aero, fisher_aero)

selected_mets<-sorted_products_anaero[simple_category=="N-Carbonyl compounds"];selected_mets
met_report(selected_mets,
           fba_products_anaero,
           fba_reactions_f[sim_O2==F & anaerobic==T],
           "produced",
           ko_f_anaero, fisher_anaero) 
react_hm(react_df=fba_reactions_f[sim_O2==T & anaerobic==F])
```

## Examine mixed
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
selected_mets<-sorted_mixed_aero[simple_category=="N-Carbonyl compounds"];selected_mets
met_report(selected_mets,
           fba_comp_metab1[sim_O2==T & anaerobic==F],
           fba_reactions_f[sim_O2==T & anaerobic==F],
           c("consumed", "produced", "transported"),
           ko_f_aero, fisher_aero)

selected_mets<-sorted_mixed_anaero[simple_category=="N-Carbonyl compounds"];selected_mets
met_report(selected_mets,
           fba_comp_metab1[sim_O2==F & anaerobic==T],
           fba_reactions_f[sim_O2==F & anaerobic==T],
           c("consumed", "produced", "transported"),
           ko_f_aero, fisher_anaero)
```
#----

# ORDER - SPECIFIC REPORTS
#################################################

## Bacteroidales
```{r eval=FALSE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
# Acetate production is high and enhriched in Chromatiales, why?
key_metabolites[name=="Acetate"]
find_metab_react("cpd00029", tidy_reactions_f[Order == "Bacteroidales"]) %>%
  filter(utilisation == "produced")%>% arrange(desc(prop_gems))-> res; res %>% datatable
res$react_name[1:5]

tidy_reactions_f[react_id2 == 'rxn00172'] %>%
  group_by(Order) %>% count %>% arrange(desc(n))

key_metabolites[name=="Acetyl-CoA", c("id", "name")]
find_metab_react("cpd00022", tidy_reactions_f[Order == "Bacteroidales"]) %>%
  filter(utilisation == "produced") %>% arrange(desc(prop_gems))-> res; res %>% datatable
res$react_name[1:5]


key_metabolites[name=="Acetylphosphate", c("id", "name")]
find_metab_react("cpd00196", tidy_reactions_f[Order == "Bacteroidales"]) %>%
  filter(utilisation == "produced") %>% arrange(desc(prop_gems))-> res; res %>% datatable
res$react_name[1:5]


key_metabolites[name=="Pyruvate", c("id", "name")]
find_metab_react("cpd00020", tidy_reactions_f[Order == "Bacteroidales"]) %>%
  filter(utilisation == "produced") %>% arrange(desc(prop_gems))-> res; res %>% datatable


key_metabolites[name=="L-Lactate", c("id", "name")]
find_metab_react("cpd00159", tidy_reactions[Order == "Bacteroidales"]) %>%
  filter(utilisation == "produced") %>% arrange(desc(prop_gems))-> res; res %>% datatable
Lactate
```

## Chromatiales
```{r eval=FALSE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
# Acetate production is high and enhriched in Chromatiales, why?
tidy_exhanges[name=="CO2" & fba_flux<0, c("OTU", "Order")]

key_metabolites[name=="Acetate"]
find_metab_react("cpd00029", tidy_reactions_f[Order == "Chromatiales"]) %>%
  filter(utilisation == "produced")%>% arrange(desc(prop_gems))-> res; res %>% datatable
res$react_name[1:5]

key_metabolites[name=="Acetyl-CoA", c("id", "name")]
find_metab_react("cpd00022", tidy_reactions_f[Order == "Chromatiales"]) %>%
  filter(utilisation == "produced") %>% arrange(desc(prop_gems))-> res; res %>% datatable
res$react_name[1:5]

key_metabolites[name=="Pyruvate", c("id", "name")]
find_metab_react("cpd00020", tidy_reactions_f[Order == "Chromatiales" & abs(fba_flux)>1]) %>%
  filter(utilisation == "produced") %>% arrange(desc(prop_gems))-> res; res %>% datatable

key_metabolites[name=="CO2"]
find_metab_react("cpd00011", tidy_reactions_f[Order == "Chromatiales"]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable

tidy_reactions_f[react_id=="rxn07189_c0"] %>% count(Order, sort = T)
#################################################

key_metabolites[name=="Oxaloacetate", c("id", "name")]
find_metab_react("cpd00032", tidy_reactions_f[Order == "Chromatiales"]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable

key_metabolites[name=="Succinate"]
find_metab_react("cpd00036", tidy_reactions_f[Order == "Chromatiales"]) %>%
  filter(utilisation == "produced")-> res; res %>% datatable 

key_metabolites[name=="Ethanol"]
find_metab_react("cpd00363", tidy_reactions_f[Order == "Chromatiales"]) %>%
  filter(utilisation == "produced")-> res; res %>% datatable 

key_metabolites[name=="Acetaldehyde"]
find_metab_react("cpd00071", tidy_reactions_f[Order == "Chromatiales"]) %>%
  filter(utilisation == "produced")-> res; res %>% datatable 


key_metabolites[name=="Fumarate"]
find_metab_react("cpd00106", tidy_reactions_f[Order == "Chromatiales"]) -> res; res %>% datatable

key_metabolites[name=="CO2"]
find_metab_react("cpd00011", tidy_reactions_f[Order == "Chromatiales"]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable

key_metabolites[name=="Oxaloacetate"]
find_metab_react("cpd00032", tidy_reactions_f[Order == "Chromatiales"]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable

key_metabolites[name=="Hydrogen"]
find_metab_react("cpd11640", tidy_reactions_f[Order == "Chromatiales"]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable

key_metabolites[name=="Hydrogen"]
find_metab_react("cpd11640", tidy_reactions_f[Order == "Chromatiales"]) %>%
  filter(utilisation == "produced")-> res; res %>% datatable
```

## Caldilineales
```{r eval=FALSE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
key_metabolites[name=="CO2", c("id", "name")]
find_metab_react("cpd00011", tidy_reactions_f[Order == "Caldilineales"]) %>%
  filter(utilisation == "produced" , prop_gems>50) -> res; res %>% datatable

key_metabolites[name=="Formate", c("id", "name")]
find_metab_react("cpd00047", tidy_reactions[Order == "Caldilineales"]) %>%
  filter(utilisation == "produced")-> res; res %>% datatable
```

## Pseudomonadales
```{r eval=FALSE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
key_metabolites[name=="Acetate"]
find_metab_react("cpd00029", tidy_reactions_f[Order == "Pseudomonadales"]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable

key_metabolites[name=="Acetyl-CoA"]
find_metab_react("cpd00022", tidy_reactions[Order == "Pseudomonadales"]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable

key_metabolites[name=="Succinate"]
find_metab_react("cpd00036", tidy_reactions_f[Order == "Pseudomonadales"]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable
gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$sbs_ids , ";"))] %>% view

key_metabolites[name=="Succinyl-CoA"]
find_metab_react("cpd00078", tidy_reactions_f[Order == "Pseudomonadales"]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable

key_metabolites[name=="indol"]
find_metab_react("cpd00281", tidy_reactions_f[Order == "Pseudomonadales"]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable

key_metabolites[name=="GABA"]
find_metab_react("cpd00281", tidy_reactions_f[Order == "Pseudomonadales"]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable

key_metabolites[name=="L-Lactate"]
find_metab_react("cpd00159", tidy_reactions_f[Order == "Pseudomonadales"]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable

key_metabolites[name=="Butyrate"]
find_metab_react("cpd00211", tidy_reactions[Order == "Pseudomonadales"]) %>%
  filter(utilisation == "produced")-> res; res %>% datatable

key_metabolites[name=="Butyryl-CoA"]
find_metab_react("cpd00120", tidy_reactions[Order == "Pseudomonadales"])-> res; res %>% datatable

key_metabolites[name=="Crotonyl-CoA"]
find_metab_react("cpd00650", tidy_reactions[Order == "Pseudomonadales"])-> res; res %>% datatable

key_metabolites[name=="(S)-3-Hydroxybutyryl-CoA"]
find_metab_react("cpd00842", tidy_reactions[Order == "Pseudomonadales"])-> res; res %>% datatable

key_metabolites[name=="Oxalate"]
find_metab_react("cpd00180", tidy_reactions_f[Order == "Pseudomonadales"]) %>%
  filter(utilisation == "produced")-> res; res %>% datatable

key_metabolites[name=="Ammonium"]
find_metab_react("cpd00013", tidy_reactions_f[Order == "Pseudomonadales"]) %>%
  filter(utilisation == "produced")-> res; res %>% datatable


```

## Desulfobacterales
```{r eval=FALSE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
key_metabolites[name=="Acetate"]
find_metab_react("cpd00029", tidy_reactions_f[Order == "Desulfobacterales"]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable

key_metabolites[name=="Acetyl-CoA"]
find_metab_react("cpd00022", tidy_reactions_f[Order == "Desulfobacterales"]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable


key_metabolites[name=="2-Oxoglutarate"]
find_metab_react("cpd00024", tidy_reactions_f[Order == "Desulfobacterales"]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable

key_metabolites[name=="L-Lactate"]
find_metab_react("cpd00159", tidy_reactions_f[Order == "Desulfobacterales"]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable

key_metabolites[name=="Acetate"]
find_metab_react("cpd00029", tidy_reactions_f[Order == "Desulfobacterales"]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable

key_metabolites[name=="GABA"]
find_metab_react("cpd00281", tidy_reactions_f[Order == "Desulfobacterales"]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable

key_metabolites[name=="Butyrate"]
find_metab_react("cpd00211", tidy_reactions_f[Order == "Desulfobacterales"]) %>%
  filter(utilisation == "produced")-> res; res %>% datatable
```

## Rhizobiales
```{r eval=FALSE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
key_metabolites[name=="Ethanol"]
find_metab_react("cpd00363", tidy_reactions_f[Order == "Rhizobiales"]) %>%
  filter(utilisation == "produced")-> res; res %>% datatable

key_metabolites[name=="Acetaldehyde"]
find_metab_react("cpd00071", tidy_reactions_f[Order == "Rhizobiales"]) %>%
  filter(utilisation == "produced")-> res; res %>% datatable

key_metabolites[name=="Methanol"]
find_metab_react("cpd00116", tidy_reactions_f[Order == "Rhizobiales"]) %>%
  filter(utilisation == "produced")-> res; res %>% datatable

key_metabolites[name=="Propionate"]
find_metab_react("cpd00141", tidy_reactions_f[Order == "Rhizobiales"]) %>%
  filter(utilisation == "produced")-> res; res %>% datatable


key_metabolites[name=="indol"]
find_metab_react("cpd00359", tidy_reactions_f[Order == "Rhizobiales"]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable

key_metabolites[name=="Thiamin", c("id", "name")]
find_metab_react("cpd00305", tidy_reactions_f[Order == "Rhizobiales"]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable

key_metabolites[name=="Hydrogen"]
find_metab_react("cpd11640", tidy_reactions[Order == "Rhizobiales"]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable
```


## Desulfobulbales
```{r eval=FALSE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
key_metabolites[name=="Propionate"]
find_metab_react("cpd00141", tidy_reactions_f[Order == "Desulfobulbales"]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable

key_metabolites[name=="CO2"]
find_metab_react("cpd00011", tidy_reactions_f[Order == "Desulfobulbales"]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable
tidy_exhanges[Order == "Desulfobulbales" & fba_flux<0 & name == "Hydrogen"]

key_metabolites[name=="indol"]
find_metab_react("cpd00359", tidy_reactions_f[Order == "Desulfobulbales"]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable

key_metabolites[name=="Hydrogen"]
find_metab_react("cpd11640", tidy_reactions[Order == "Desulfobulbales"]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable

key_metabolites[name=="Butyrate"]
find_metab_react("cpd00211", tidy_reactions_f[Order == "Desulfobulbales"]) %>%
  filter(utilisation == "produced")-> res; res %>% datatable

key_metabolites[name=="Butyryl-CoA"]
find_metab_react("cpd00120", tidy_reactions_f[Order == "Desulfobulbales"]) %>%
  filter(utilisation == "produced")-> res; res %>% datatable

key_metabolites[name=="Methanethiol"]
find_metab_react("cpd00324", tidy_reactions_f[Order == "Desulfobulbales"]) %>%
  filter(utilisation == "produced")-> res; res %>% datatable
Urea
key_metabolites[name=="Urea", c("id", "name")]
find_metab_react("cpd00073", tidy_reactions[Order == "Desulfobulbales"]) %>%
  filter(utilisation == "produced")-> res; res %>% datatable
```

## Bacteroidales 2
```{r eval=FALSE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
tidy_exhanges[Order == "Bacteroidales" &
                simple_category == "Carbonyl compounds" &
                  fba_flux>0
                ] %>% 
  group_by(name) %>%
  summarise(n=n(),
            median=median(fba_flux))  %>%
  arrange(n, median)

key_metabolites[name=="Xylose"]
find_metab_react("cpd00154", tidy_reactions_f[Order == "Bacteroidales"]) %>%
  filter(utilisation == "produced")-> res; res %>% datatable

key_metabolites[name=="Glucuronate"]
find_metab_react("cpd00164", tidy_reactions[Order == "Bacteroidales"]) %>%
  filter(utilisation == "produced")-> res; res %>% datatable

key_metabolites[name=="L-Fucose", c("name", "id")]
find_metab_react("cpd00751", tidy_reactions[Order == "Bacteroidales"]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable


key_metabolites[name=="Acetate"]
find_metab_react("cpd00029", tidy_reactions_f[Order == "Bacteroidales"]) %>%
  filter(utilisation == "produced")-> res; res %>% datatable

key_metabolites[name=="Galactose"]
find_metab_react("cpd00108", tidy_reactions_f[Order == "Bacteroidales"]) %>%
  filter(utilisation == "produced")-> res; res %>% datatable

key_metabolites[name=="Galactose"]
find_metab_react("cpd00108", tidy_reactions_f[Order == "Bacteroidales"]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable

key_metabolites[name=="L-Fucose"]
find_metab_react("cpd00751", tidy_reactions_f[Order == "Bacteroidales"]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable

key_metabolites[name=="L-Fuculose"]
find_metab_react("cpd01186", tidy_reactions_f[Order == "Bacteroidales"]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable

key_metabolites[name=="Maltose"]
find_metab_react("cpd00179", tidy_reactions_f[Order == "Bacteroidales"]) %>%
  filter(utilisation == "produced")-> res; res %>% datatable

```

## Woeseiales
```{r eval=FALSE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
key_metabolites[name=="Cytidine"]
find_metab_react("cpd00367", tidy_reactions_f[Order == "Woeseiales"]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable
```

