---
title: "Abiotic measures Boschettona 23"
author: "Riccardo Frizzo"
date: "2026-06-26"
output:
  pdf_document: default
  html_document: default
editor_options: 
  chunk_output_type: console
---
#Init
```{r setup, include=FALSE}
libs<-c("tidyverse", "foreach","data.table", "readxl",
        "magrittr", "ggstatsplot", "colorspace", "rstatix",
        "knitr", "kableExtra", "ComplexHeatmap"); lapply(libs,require,character.only=T)
knitr::opts_chunk$set(echo = TRUE, fig.width = 12, fig.height = 6)
hcl.colors(n=3, palette ="Zissou 1",alpha = 0.9)
#hcl.colors(n=10, palette ="Zissou 1",alpha = 0.9) %>% colorspace::swatchplot(cvd = TRUE)
zy=c("#3B99B1E6","#FADB2BE6","#A5191CE6")
#zy=c("#A5191CE6","#FADB2BE6","#65B59AE6")%>% colorspace::swatchplot(cvd = TRUE)
zy=c("#A5191CE6","#FADB2BE6","#65B59AE6")

zy=c("#A5191CE6","white","#65B59AE6")
scales::show_col(zy[c(3,1,2)],ncol = 3)
saltmarsh_colors = c(Constructed=zy[1], Natural=zy[3], ns = zy[2])
saltmarsh_colors2 = c(Restored=zy[1], Natural=zy[3], ns = zy[2])
```

# Granulometry
```{r eval=TRUE, tidy = TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
libs<-c("tidyverse", "foreach","data.table", "readxl", "magrittr", "ggstatsplot", "colorspace", "rstatix","knitr", "kableExtra"); lapply(libs,require,character.only=T)
setwd("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/granulometry/")
files<-list.files(pattern = ".xlsx", path = "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/granulometry",
                  full.names = T)

# Import data ----
foreach(i = files, .combine = "rbind")%do%{
  read_excel(i,sheet = "Data") %>% add_column(Sample = str_remove(i, ".xlsx"), .before ="Particle Size (µm)")
} %>% dplyr::filter(!is.na(`Detector Number`)) %>% dplyr::select(Sample, `Particle Size (µm)`, matches("Vol")) -> gran

range(gran$`Particle Size (µm)`)

# Group data on the Wentworth scale
gran %<>% mutate(Wentworth=
                  case_when(
                    `Particle Size (µm)`> 0.95e-3 & `Particle Size (µm)`< 0.97 ~ "Colloid",
                    `Particle Size (µm)`> 0.98 & `Particle Size (µm)`< 3.9 ~ "Clay",
                    `Particle Size (µm)`> 3.9 & `Particle Size (µm)`< 62.5 ~ "Silt",
                    `Particle Size (µm)`> 62.5 & `Particle Size (µm)`< 125 ~ "Very fine sand",
                    `Particle Size (µm)`> 125 & `Particle Size (µm)`< 250 ~ "Fine sand",
                    `Particle Size (µm)`> 250 & `Particle Size (µm)`< 500 ~ "Medium sand",
                    `Particle Size (µm)`> 500 & `Particle Size (µm)`< 1000 ~ "Coarse sand",
                    `Particle Size (µm)`> 1200 & `Particle Size (µm)`< 2000 ~ "Very coarse sand")) %>% filter(!is.na(Wentworth))

#Add group information 
gran %<>% mutate(Group = factor(str_extract(Sample, "A|N"), levels = c("A", "N"), labels = c("Constructed", "Natural")),
                 Wentworth= factor(Wentworth, levels=c("Colloid", "Clay", "Silt",
                                                       "Very fine sand", "Fine sand","Medium sand","Coarse sand", "Very coarse sand")))

#Calculate mean of the technical replicas
gran_l<-gran %>% pivot_longer(cols = matches("Vol"), names_to = "Run", values_to = "perc_vol")
gran_mean<-gran_l %>% group_by(Sample, Group, Wentworth, `Particle Size (µm)`) %>% 
  summarise(
    mean_perc_vol=mean(perc_vol),
    sd_perc_vol=sd(perc_vol))
```

```{r, echo = FALSE}
# Plot %vol per Particle size
ggplot(gran_mean %>% filter(Wentworth != "Very coarse sand"),
       aes(y=mean_perc_vol, x=`Particle Size (µm)`, color = Group)) + geom_point() + 
  theme_minimal() + ylab("Volume Proportion (%)") + 
  scale_color_brewer(type = "qual", palette = "Set2") + 
  ggtitle("Particle size distribution of all samples, colored by Wentworth classes")
```

### Test distribution within experimental groups and Wentworth classes

H1: Granulometry differs between Constructed and Natural in at least one Wentworth class

Null hypothesis: no significantly different Wentworth classes between Constructed and Natural

### Test distribution: Shapiro test

```{r, echo = FALSE, warning=F, message=FALSE, tidy=TRUE}
# Group particle size into Wentworth classes and sum
gran_group_sum <- gran_mean%>% group_by(Sample, Group, Wentworth) %>% summarise(sample_sum = sum(mean_perc_vol))

# Test distribution
gran_group_sum %>% group_by(Group, Wentworth) %>% summarise(
  shapiro_statistic=shapiro.test(sample_sum)$statistic,
  shapiro_pval=shapiro.test(sample_sum)$p.value) %>% kable()

```

### Gran mean and SD

```{r, echo = FALSE, warning=F, message=FALSE, tidy=TRUE}
gran_group_sum %>% group_by(Wentworth)%>%
  summarise(
  mean=mean(sample_sum),
  sd=sd(sample_sum)) %>%
  arrange(mean) -> gran_mean
```

Data is normally distributed within experimental groups and Wentworth classes

### Test significance: T-Test and Wilcox
```{r, echo = FALSE, warning=F, message=FALSE}
# T-test
gran_group_sum %>% group_by(Wentworth)%>%
  summarise(
  t_statistic=t.test(sample_sum ~ Group)$statistic,
  t_pval=t.test(sample_sum ~ Group)$p.value) %>% kable()

# Wilcox test
gran_group_sum %>% group_by(Wentworth)%>%
  summarise(
  t_statistic=wilcox.test(sample_sum ~ Group)$statistic,
  t_pval=t.test(sample_sum ~ Group)$p.value) %>% kable()

```

## Visual summary: boxplots
```{r, eval=T, echo = FALSE, warning=F, message=FALSE}
require(ggbreak)
ggplot(gran_group_sum, aes(y=(sample_sum), x= Wentworth)) +
  geom_boxplot(linewidth = 0.5, outliers = F) + 
  geom_jitter(position = position_dodge2(0.8),
               alpha = 0.5, size=0.3) +
  theme_bw() + 
  ylab("Volume proportion (%)") + 
  xlab("") + 
  ggtitle("Granulometry") +
    ylim(c(0,80)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 14),
          axis.text.y = element_text(angle = 0, hjust = 0, size = 15),
          axis.title.y = element_text(size=15),
          legend.position = "none",
          plot.title = element_text(size = 18, face = "bold")
          )-> gg_gran;gg_gran

ggsave(plot = gg_gran, 
       "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/figures/granulometry_wentworth_11may26.png",
       width = 9,
       height = 7,
       units = "cm",
       dpi = 600,bg = "white"
       )

```

No significant differences were found, accepting the null hypothesis for all Wentworth classes

#---
# CHNS
## Test significance
```{r, echo = FALSE, warning=F, message=FALSE}
TOC<-
  read_excel("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/chns/TOC Boschettona.XLS") %>%
  janitor::clean_names() %>% 
  dplyr::select(number, toc) %>%
  dplyr::rename(FILENAME=number, TOC=toc) %>%
  filter(!is.na(FILENAME)) %>%
  mutate(FILENAME = str_replace(FILENAME, "_", "\\."))

CNH<-read_excel("/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/chns/chns_02aug24.xls") %>% 
  dplyr::select(FILENAME, WEIGHT, NITROGEN, CARBON, HYDROGEN, SULPHUR) %>%
  mutate(FILENAME=str_remove(FILENAME, "!")) %>% 
  filter(!grepl("BBOT", FILENAME) & !grepl("Sulf|SULF", FILENAME) & !grepl("Blank", FILENAME) & !grepl("#", FILENAME)) %>%
  mutate(FILENAME=str_replace(str_remove(FILENAME, " "), "-", "\\."),
         Group=factor(str_extract(FILENAME, "A|N"), levels=c("A", "N"), labels = c("Constructed", "Natural")))
CNH_TOC<-merge(TOC, CNH)

CNH_TOC %>% 
  pivot_longer(
    cols=c(NITROGEN, CARBON, HYDROGEN, SULPHUR, TOC),
    names_to = "Element",
    values_to = "Proportion") -> CHN_l

CHN_l %<>% 
  mutate(
    Element = 
      factor(Element, levels = c("CARBON", "TOC", "NITROGEN", "HYDROGEN", "SULPHUR"),
              labels = c("TC", "TOC", "TN", "TH", "TS")))

# Check distribution
CHN_l %>% group_by(Group, Element) %>% summarise(
  shapiro_statistic=shapiro.test(Proportion)$statistic,
  shapiro_pval=shapiro.test(Proportion)$p.value) %>% kable()

# Check 
CHN_l %>% group_by(Element)%>%
  summarise(
  t_statistic=t.test(Proportion ~ Group)$statistic,
  t_pval=t.test(Proportion ~ Group)$p.value) %>% kable()

```

## Visual summary: boxplots
```{r, echo = FALSE, warning=F, message=FALSE}
CHN_l %>% group_by(Element, Group) %>% count
levels(CHN_l$Element)

CHN_l %>%
  filter(Element != "TS") %>%
  ggplot(aes(x=Element, y=Proportion)) +
  geom_boxplot(linewidth = 0.5, outliers = F) + 
  geom_jitter(position = position_dodge2(0.8),
              size= 1) +
    theme_bw() + 
    scale_color_manual(values = saltmarsh_colors) +
    ylab(" Mass composition (%)") + 
    xlab("") +
    ggtitle("CHN & TOC") +
      theme(axis.text.x = element_text(angle = 0, hjust = 1, size = 15),
          axis.text.y = element_text(angle = 0, hjust = 0, size = 15),
          axis.title.y = element_text(hjust = 0.5, vjust =  2, size = 15),
          legend.position = "none",
          plot.margin = unit(c(1, 1, 1, 1), "cm"),
          plot.title = element_text(size=18, face = "bold")
          )->gg_chn;gg_chn

ggsave(plot = gg_chn, 
       "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/figures/chns_toc.png",
       width = 9,
       height = 7,
       units = "cm",
       dpi = 600,
       bg = "white"
       )

ggsave(plot = gg_chn, 
       "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/figures/presentations/chns_toc.png",
       width = 8,
       height = 4,
       units = "cm",
       dpi = 600,
       bg = "white"
       )
```

## Summary statistics
```{r, echo = FALSE, warning=F, message=FALSE}
CHN_l %>% group_by(Element)%>%
  summarise(mean=mean(Proportion),
            sd=sd(Proportion)
            ) -> chn_mean
chn_mean %>% kable

mult_factor<-1/chn_mean$mean[3]

# Calculate ratios
chn_mean %>%
  mutate(prop=mean*mult_factor)

# Calculate organic carbon prop
chn_mean$mean[2]/chn_mean$mean[1]
```

# ----
#Vegetation coverage
```{r, echo = FALSE, warning=F, message=FALSE}
require(kableExtra)
list.files(path = "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/copernicus",
           pattern ="Sentinel-2.*NDVI",full.names = T) -> ndvi_files
foreach(i = ndvi_files, .combine = "rbind")%do%{
fread(i) %>% add_column(
  sample=str_extract(i, "_[:alpha:]+[:digit:].csv") %>% str_remove_all("_|.csv"),
  group=str_extract(sample, "Con|Nat"), .before = "C0/date")
  } %>%  
  mutate(group=factor(group, levels = c("Con", "Nat"), labels = c("Constructed", "Natural"))) -> nvdi

nvdi %>% filter(`C0/cloudCoveragePercent` == 0) -> nvdi_nc

nvdi_nc %>%
  group_by(sample, group) %>% 
  summarise(ndvi=median(`C0/mean`)) %>% ungroup -> ndvi_median

ggstatsplot::ggbetweenstats(ndvi_median, x = group, y=ndvi)

ndvi_median %>% 
  group_by(group) %>% 
  summarise(mean=mean(ndvi),
            sd=sd(ndvi)
            )

ndvi_median %>%
    summarise(
  t_statistic=t.test(ndvi ~ group)$statistic,
  t_pval=t.test(ndvi ~ group)$p.value) %>% kable()


ggplot(ndvi_median, aes(x=group, y=ndvi, color = group)) +
  geom_boxplot(linewidth = 0.7, outliers = F) + geom_jitter() +
    theme_minimal() +
    scale_color_manual(values = saltmarsh_colors) +
    ylab("") + xlab("Normalized difference vegetation index") +
  theme(legend.position = "none",
        axis.text.y = element_text(angle = 0, size = 13),
        axis.text.x = element_text(angle = 0, size = 13),
        axis.title.y = element_text(hjust = 0.5, vjust =  2, size = 12))-> gg_ndvi;gg_ndvi

ggsave(plot = gg_ndvi, 
       "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/figures/presentations/ndvi_boxplot.png",
       width = 4,
       height = 4,
       dpi = 600,
       bg = "white"
       )
  
  
nvdi_nc %>%
    summarise(
  t_statistic=t.test(`C0/mean` ~ group)$statistic,
  t_pval=t.test(`C0/mean` ~ group)$p.value) #%>% kable()

nvdi_nc %>% group_by(group) %>% summarise(
  shapiro_statistic=shapiro.test(`C0/mean`)$statistic,
  shapiro_pval=shapiro.test(`C0/mean`)$p.value) #%>% kable()

nvdi_nc %>%
    summarise(
  t_statistic=wilcox.test(`C0/mean` ~ group)$statistic,
  t_pval=wilcox.test(`C0/mean` ~ group)$p.value) #%>% kable()

ggplot(nvdi_nc, aes(x=group, y=`C0/mean`)) +
  geom_boxplot() + geom_jitter()

```
#---
# MetaB
## Assigned signals: preprocess
```{r, echo = FALSE, warning=F, message=FALSE}
require(readxl);
#Import LC data
read_excel("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaB/Features_Identified_Riccardo2_bysample.xlsx", sheet = 1)%>% 
  dplyr::select(`Metabolite name`, matches("ART|NAT"))-> LC_metab_areas
 colnames(LC_metab_areas) %<>% str_replace("ART", "CON") #fix sample names
 
# Import LC assigned metabolites metadata
read_excel("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaB/Features_Identified_Riccardo2.xlsx", sheet = 1) %>% add_column(data_type="LC")-> LC_metab_ids



# Remove ID missing in the matrix
#LC_metab_ids %>% filter(`Metabolite name` != "Tryptophan betaine") -> LC_metab_ids

# Import GC metadata
read_excel("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaB/BoschettonaGC_riccardo_21jan25.xlsx", sheet = 1) %>% add_column(data_type="GC")-> GC_metab_ids

# Merge LC and GC data, use metaboanalyst to tidy names and assign db ids
read_excel("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaB/LCGC_metabolite_ids_curated_20dec24.xlsx")-> metab_db_ids

```

## Merge and tidy dataset of db annotation
```{r echo=FALSE, message=FALSE, warning=FALSE}
metab_db_ids %>% dplyr::select(Query, Match, HMDB, KEGG, `HMDB&KEGG`) %>%
  dplyr::rename(Metabolite=Match)-> metab_info2
```

## Assigned signals: Classes
```{r, echo = FALSE, warning=F, message=FALSE}
fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaB/LCGC_metabolite_superclasses.csv")-> metab_superclasses
fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaB/LCGC_metabolite_mainclasses.csv")-> metab_mainclasses
fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaB/LCGC_metabolite_subclasses.csv")-> metab_subclasses

# Tidy subclasses
foreach(i = metab_subclasses$Group)%do%{
  
  metab_subclasses %>% filter(Group == i) %>%
    pull(Members) %>% str_split_1(";") %>% str_remove("^ ")-> metab
  
  data.frame(Subclass=i,
             Metabolite=metab)
  } %>% rbindlist -> metab_subclasses_2

# Merge with metab_info
full_join(metab_info2, metab_subclasses_2, by= "Metabolite")-> metab_info3

metab_info3 %>% 
  mutate(HMDB_link=paste0("https://hmdb.ca/metabolites/", HMDB))-> metab_info4

# Export for manual curation of subclasses
fwrite(metab_info4, "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaB/LCGC_metabolite_ids_subclasses_curated_24dic24.csv")
```

## Check uncategorised
```{r, echo = FALSE, warning=F, message=FALSE}
metab_info3$Subclass %>% is.na() %>% table #35 cmpds uncategorised
table(metab_info3$HMDB == "NA") #6 cmpds wothoud hmdb hit
```

## Import curated data and tidy
```{r, echo = FALSE, warning=F, message=FALSE}
fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaB/LCGC_metabolite_ids_subclasses_curated_classified_hmdb_24dic24.csv") %>%
  mutate(Subclass = str_remove(Subclass, "\t"))-> metab_ids_curated

metab_ids_curated %>% dplyr::select(name_orig, Metabolite, Subclass)->
  metab_ids_curated_names

# Fix duplicated values
metab_ids_curated_names[!duplicated(metab_ids_curated_names$Metabolite),] -> metab_ids_curated_names

# Make df
metab_ids_curated_names[,-1] %>% column_to_rownames("Metabolite")-> metab_ids_curated_names_df


metab_classes<-
  data.table(
  Subclass=metab_ids_curated_names_df$Subclass %>% unique,
  Class=coarse_class <- c(
    "Fatty Acyls",
    "Other nitrogen/sulfur compounds",
    "Alcohols & Polyols",
    "Fatty Acyls",
    "Amino acids & Peptides",
    "Organic acids",
    "Carbohydrates",
    "Organic acids",
    "Organic acids",
    "Fatty Acyls",
    "Fatty Acyls",
    "Organic acids",
    "Other nitrogen/sulfur compounds",
    "Nucleosides & Nucleotides",
    "Nucleosides & Nucleotides",
    "Nucleosides & Nucleotides",
    "Nucleosides & Nucleotides",
    "Nucleosides & Nucleotides",
    "Alkaloids",
    "Organic acids",
    "Organic acids",
    "Organic acids",
    "Organic acids",
    "Aromatic compounds",
    "Aromatic compounds",
    "Other nitrogen/sulfur compounds",
    "Aromatic compounds",
    "Fatty Acyls",
    "Aromatic compounds",
    "Other nitrogen/sulfur compounds"
  ))

metab_ids_curated_names %>%
  left_join(metab_classes) -> metab_ids_curated_names1

metab_ids_curated_names1 %>% 
  fwrite("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/metab_ids_curated_names_03giu26.csv")
```

# LC-MS
##Tidy bysample dataset and prepare matrix (Remove sample NAT3.1)
```{r, echo = FALSE, warning=F, message=FALSE}
# Tidy names
LC_metab_areas %>% dplyr::rename(name_orig=`Metabolite name`) %>%
  merge(metab_ids_curated_names[,c(1,2)], by="name_orig") ->LC_metab_areas2

# Remove sample NAT3.1, as signal is very low across the board
LC_metab_areas2<-LC_metab_areas2[,-which(colnames(LC_metab_areas2)=="NAT3.1")]
colnames(LC_metab_areas2)

# Fix duplicated values, remove original IDs
rownames(LC_metab_areas2)=NULL
  
LC_metab_areas2[,-1] %>%
  column_to_rownames('Metabolite') %>% as.matrix -> LC_metab_areas_mat

# Format data for metaboanalyst, export
LC_metab_areas_mat %>% t %>%
  as.data.frame %>%
  rownames_to_column("sample") %>%
  mutate(
    group=factor(levels=c("CON", "NAT"), labels= c("Constructed", "Natural"),
                 str_extract(sample, "CON|NAT")), .after = "sample") -> LC_metab_areas_df

LC_metab_areas_df %>%
  fwrite("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaB/for_metaboAn.csv")
```

## LC: Normalise and scale with MetaboAnalyst
```{r, echo = FALSE, warning=F, message=FALSE}
setwd("/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaB/metaboanalyst/")
require(MetaboAnalystR)
mSet<-InitDataObjects("msspec", "stat", FALSE);
mSet<-Read.TextData(mSet, "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaB/for_metaboAn.csv", "rowu", "disc");
mSet<-SanityCheckData(mSet);
mSet<-ReplaceMin(mSet);
mSet<-PreparePrenormData(mSet);
mSet<-Normalization(mSet,
                    rowNorm = "SumNorm",
                    transNorm ="NULL",
                    scaleNorm =  "AutoNorm");
mSet$dataSet$norm %>% t -> LC_mat_norm
fwrite(x=LC_mat_norm %>% as.data.frame %>% rownames_to_column("metabolite"),
       file="~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/data/analysis/LC_mat_norm_23apr25.csv")

```

## Summarise abundance by categories
```{r, echo = FALSE, warning=F, message=FALSE}
# Count and sort classes
metab_ids_curated_names1 %>%
  filter(
    name_orig %in% metab_ids_curated_names$name_orig & 
      name_orig %in% LC_metab_areas$`Metabolite name`) %>%
  group_by(Subclass) %>% count %>% arrange(desc(n)) -> sorted_metabolite_subclass

metab_ids_curated_names1 %>%
  filter(
    name_orig %in% metab_ids_curated_names$name_orig & 
      name_orig %in% LC_metab_areas$`Metabolite name`) %>%
  group_by(Class) %>% count %>% arrange(desc(n)) -> sorted_metabolite_class

# Summarise signal area by metabolite
LC_metab_ids %>% 
  dplyr::select(`Metabolite name`, `Area media`) %>%
  dplyr::rename(name_orig=`Metabolite name`) -> LC_metab_avg_area

## Merge with annotation and filter out GC annotations
merge(metab_ids_curated_names, LC_metab_avg_area, by="name_orig") %>%
  filter(!is.na(`Area media`))-> LC_metab_avg_area_1

LC_metab_avg_area_1 %>%
  group_by(Subclass) %>%
  summarise(sum=sum(`Area media`)) %>%
  ggplot(aes(y=reorder(Subclass, sum, decreasing = F), x=(sum))) +
  geom_bar(stat="identity", color = "black", fill='yellow3') +
  theme_bw() +
  ylab('HMDB Subclass') +
  xlab('Average signal area')

```

## Visual summary: heatmap by sample
```{r, echo = FALSE, warning=F, message=FALSE, eval = F}
min_fontsize=4
# Set molecular classes as row annotation, sort by size of the category
lev<-metab_ids_curated_names_df[rownames(LC_metab_areas_mat),"Subclass"] %>%
  factor(levels=sorted_metabolite_class$Subclass)

# Set colors for molecular classes
set.seed(984)
row_col<-sample(hcl.colors(length(unique(lev)), palette = "Tropic"))
names(row_col) = sorted_metabolite_class$Subclass
subclass_colors<-row_col
save(subclass_colors,
     file="~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/data/analysis/subclass_colors.Rdata")

# Reset colors also considering FBA metabolites
load(file="~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/data/analysis/all_metabolite_subclasses_colors.Rdata")

row_col<-all_metabolite_subclasses_colors[sorted_metabolite_class$Subclass]
# Convert the full Subclass vector into a color vector
LC_metab_avg_area_1 %>% 
  mutate(color=factor(Subclass,
                      labels = row_col,
                      levels = names(row_col)
                      )) -> bar_fill_cols

bfc=as.character(bar_fill_cols$color)
names(bfc)=bar_fill_cols$Metabolite

# Calculate intensity mean across sample and log transform
rowm<-rowMeans(log(LC_metab_areas_mat))

# Set left annot as colored bars
left_annot<-
  rowAnnotation(
  "Molecular classification" = lev,
  show_legend = F,
  col = list("Molecular classification"= row_col),
  annotation_name_rot = 60,
  annotation_name_gp = gpar(fontsize = min_fontsize +2),
  width = unit(0, "cm")
  )

# Set right annot as barplots of the mean
right_annot<-
rowAnnotation(
  "Average\n log(S.I.)" = anno_barplot(
    x = rowm,
    gp = gpar(fill= bfc[names(rowm)]),
    axis_param = list(gp = gpar(fontsize = min_fontsize+1))
    ), 
  show_legend = F,
  annotation_name_rot = 0,
  annotation_name_side = "bottom",
  annotation_name_gp = gpar(fontsize = min_fontsize +2),
  width = unit(1, "cm")
  )

# Plot

LC_mat_norm[,LC_metab_areas_df$sample] %>%
Heatmap(
          col=hcl.colors(10, palette = "PuBuGn", rev = T),
          cluster_rows = F,
          cluster_columns = F,
          show_row_names = T,
          row_names_side = "right", 
          row_names_gp = gpar(fontsize=min_fontsize),
          #top_annotation = col_annot,
          left_annotation = NULL,
          right_annotation = right_annot,
          column_split = LC_metab_areas_df$group,
          rect_gp = gpar(col = "white", lwd = 0),
          column_title_gp = gpar(fontsize = min_fontsize + 5,
                                 fontface="bold"
                                 ),
          column_title_rot = 20, 
          row_split = lev,
          gap = unit(0.5, "mm"),  #Size of gaps
          row_title_gp = gpar(fontsize = min_fontsize + 2),
          row_title_side = "left",
          row_title_rot = 0,
          column_names_rot=60,
          column_names_gp = gpar(fontsize=min_fontsize),
          show_heatmap_legend = T,
          heatmap_legend_param = list(
            title = "Normalised S.I.",  # Set legend title
            legend_direction = "horizontal",    # Set legend position
            legend_width = unit(2, "cm"),
            title_gp = gpar(fontsize = min_fontsize+2, fontface = "bold"),
            labels_gp = gpar(fontsize = min_fontsize)
            ),       # Set legend size if horizontal
          height = unit(13.5, "cm"),  # Adjust heatmap height
          width = unit(2.05, "cm")   # Adjust heatmap width
          
  )->hm_LC_bysample

png(filename = "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/final/figures/heatmap_lc_bysample.png", width = 11.55, height = 17, units = "cm", res = 600)
draw(hm_LC_bysample,
     heatmap_legend_side = "bottom"
     )
dev.off()

```

## LC-MS heatmap by mean
```{r, eval = F, echo = FALSE, warning=F, message=FALSE}
# Load diff information
fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/final/supplementary/Supplementary file 1- GC- LC- ICP-MS summary tables.xlsx - LC-MS summary table.csv", skip = 3) %>% 
  select("Database Annotation", "Significantly more abundant in") %>%
  rename(diff="Significantly more abundant in",
         Metabolite = "Database Annotation"
         ) %>% filter(diff != "") -> LC_diff

min_fontsize=4
# Set molecular classes as row annotation, sort by size of the category
lev<-metab_ids_curated_names_df[rownames(LC_metab_areas_mat),"Subclass"] %>%
  factor(levels=sorted_metabolite_class$Subclass)

# Reset colors also considering FBA metabolites
load(file="~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/data/analysis/all_metabolite_subclasses_colors.Rdata")


row_col<-all_metabolite_subclasses_colors[sorted_metabolite_class$Subclass]
# Convert the full Subclass vector into a color vector
LC_metab_avg_area_1 %>% 
  mutate(color=factor(Subclass,
                      labels = row_col,
                      levels = names(row_col)
                      )) -> bar_fill_cols

bfc=as.character(bar_fill_cols$color)
names(bfc)=bar_fill_cols$Metabolite

#Include diff data
left_join(bar_fill_cols, LC_diff) %>% select(Metabolite, diff) %>% data.table -> LC_diff2
LC_diff2[is.na(diff),"diff"] = "ns"
LC_diff2 %>% column_to_rownames("Metabolite")->LC_diff3

# Calculate intensity mean across sample and log transform
rowm<-rowMeans(log(LC_metab_areas_mat))

# Sort each category by rowmean
foreach(i = levels(lev))%do%{
  metab_ids_curated_names_df %>% filter(Subclass == i) %>% rownames() -> metab
  rowm[metab] %>% sort(decreasing = T)
} %>% unlist -> ordered_sorted_rowmeans

sorted_lev<-metab_ids_curated_names_df[names(ordered_sorted_rowmeans),]
sorted_lev<-factor(sorted_lev, levels=levels(lev))

# Set left annot as colored bars
left_annot<-
  rowAnnotation(
  "Molecular classification" = sorted_lev,
  show_legend = F,
  col = list("Molecular classification"= row_col),
  annotation_name_rot = 60,
  annotation_name_gp = gpar(fontsize = min_fontsize +2),
  width = unit(0, "cm")
  )

# Set right annot as barplots of the mean
right_annot<-
rowAnnotation(
 "Salt marsh\n zone" = anno_simple(
    x = LC_diff3[names(ordered_sorted_rowmeans),"diff"],
    col = saltmarsh_colors2,
    border = T,
    width = unit(2.5, "mm")
   ),
  "Average\n log(SI)" = anno_barplot(
    x = ordered_sorted_rowmeans,
    gp = gpar(fill= bfc[names(ordered_sorted_rowmeans)]),
    axis_param = list(gp = gpar(fontsize = min_fontsize+1))
    ), 
  show_legend = F,
  annotation_name_rot = c(45, 0),
  annotation_name_side = "bottom",
  annotation_name_gp = gpar(fontsize = min_fontsize +2),
  annotation_name_offset = unit(c(2, 5), "mm") 
  )


# Calculate mean by zone
cbind(
  LC_mat_norm[,grepl("NAT",colnames(LC_mat_norm))] %>% rowMeans,
  LC_mat_norm[,grepl("CON",colnames(LC_mat_norm))] %>% rowMeans
) %>% as.matrix -> mat
colnames(mat) = c("Natural", "Restored")

# Plot
ordered_sorted_rowmeans
mat[names(ordered_sorted_rowmeans),] %>%
Heatmap(
          col=hcl.colors(10, palette = "PuBuGn", rev = T),
          cluster_rows = F,
          cluster_columns = F,
          show_row_names = T,
          row_names_side = "right", 
          row_names_gp = gpar(fontsize=min_fontsize),
          #top_annotation = col_annot,
          left_annotation = NULL,
          right_annotation = right_annot,
          column_split = colnames(mat),
          rect_gp = gpar(col = "white", lwd = 0),
          column_title_gp = gpar(fontsize = min_fontsize + 2,
                                 fontface="bold"
                                 ),
          column_title_rot = 20, 
          row_split = sorted_lev,
          gap = unit(0.5, "mm"),  #Size of gaps
          row_title_gp = gpar(fontsize = min_fontsize + 2),
          row_title_side = "left",
          row_title_rot = 0,
          column_names_rot=60,
          column_names_gp = gpar(fontsize=0),
          show_heatmap_legend = T,
          heatmap_legend_param = list(
            title = "Normalised S.I.",  # Set legend title
            legend_direction = "horizontal",    # Set legend position
            legend_width = unit(2, "cm"),
            title_gp = gpar(fontsize = min_fontsize+2, fontface = "bold"),
            labels_gp = gpar(fontsize = min_fontsize)
            ),       # Set legend size if horizontal
          height = unit(16.5, "cm"),  # Adjust heatmap height
          width = unit(0.8, "cm")   # Adjust heatmap width
          
  )->hm_LC_mean

png(filename = "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/final/figures/heatmap_lc_mean.png", width = 11.7, height = 20, units = "cm", res = 600)
draw(hm_LC_mean,
     heatmap_legend_side = "top"
     )
dev.off()

```
## export LC_mean
```{r, echo = FALSE, warning=F, message=FALSE}

LC_metab_areas %>%
  column_to_rownames("Metabolite name") %>% colSums()

#Remove NAT3.1, much lower signal

LC_metab_areas %>%
  pivot_longer(cols = matches("NAT|CON")) %>%
  filter(name != "NAT3.1") %>%
  group_by(name_orig=`Metabolite name`) %>%
  summarise(
    mean=mean(value),
    sd=sd(value)
  ) %>% arrange(desc(mean)) %>% add_column(type="LC") -> lc_mean

fwrite(lc_mean, 
      "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/LC_means.csv")

```

## LC-MS stats (Table 1)
```{r, echo = FALSE, warning=F, message=FALSE}
#Metabolites stats
# Percentile by intensity
LC_metab_avg_area_1 %>%
  mutate(percentile_group = ntile(`Area media`, 100),
         area_prop = (`Area media`/sum(LC_metab_avg_area_1$`Area media`))*100
         ) -> percentile
percentile %>% filter(percentile_group>=90) %>% arrange(desc(`Area media`)) %>% pull(Metabolite) %>% cat

# Categories stats
## Number of metabolites per category
LC_metab_avg_area_1 %>% 
  group_by(Subclass) %>% count %>%
  arrange(desc(n)) %>%
  mutate(prop=n/nrow(LC_metab_avg_area_1)*100)

## Cumulative signal per category
LC_metab_avg_area_1 %>% 
  group_by(Subclass) %>% 
  summarise(sum=sum(`Area media`),
            "Number of compounds"=n()
            ) %>%
  arrange(desc(sum)) %>%
  mutate("Average signal proportion (%)"=sum/sum(LC_metab_avg_area_1$`Area media`)*100) -> LC_area_count_nice
LC_area_count_nice[,c(1,3,4)] %>% fwrite(file = "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/final/tables/LC_MS_subclass_area_count.tsv", sep = "\t", dec = ",")

```


#---
# GC-MS
##Import and tidy bysample dataset
```{r, echo = FALSE, warning=F, message=FALSE}
read_excel("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaB/BoschettonaGC_riccardo_21jan25.xlsx", sheet = 2) %>%
  dplyr::select("Metabolite name", matches("ART|NAT"))-> GC_metab_areas
colnames(GC_metab_areas) %<>% str_replace("ART", "CON") %>% str_replace(" ", "") %>% str_remove("G") #fix sample names

# Tidy names
GC_metab_areas %>% dplyr::rename(name_orig=`Metabolitename`) %>%
  merge(metab_ids_curated_names[,c(1,2)], by="name_orig") -> GC_metab_areas2

# Fix duplicated values, remove original IDs
rownames(GC_metab_areas2)=NULL
  
GC_metab_areas2[,-1] %>%
  column_to_rownames('Metabolite') %>% as.matrix -> GC_metab_areas_mat

# Format data for metaboanalyst, export
GC_metab_areas_mat %>% t %>%
  as.data.frame %>%
  rownames_to_column("sample") %>%
  mutate(
    group=factor(levels=c("CON", "NAT"), labels= c("Constructed", "Natural"),
                 str_extract(sample, "CON|NAT")), .after = "sample") %>% as.data.table -> GC_metab_areas_df

# Table 1
GC_metab_areas_df %>%
  fwrite("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaB/GC_for_metaboAn.csv")
```

## GC: Normalise and scale with MetaboAnalyst
```{r, echo = FALSE, warning=F, message=FALSE}
setwd("/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaB/metaboanalyst/")
require(MetaboAnalystR)
mSet<-InitDataObjects("msspec", "stat", FALSE);
mSet<-Read.TextData(mSet, "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaB/GC_for_metaboAn.csv", "rowu", "disc");
mSet<-SanityCheckData(mSet);
mSet<-ReplaceMin(mSet);
mSet<-PreparePrenormData(mSet);
mSet<-Normalization(mSet,
                    rowNorm = "SumNorm",
                    transNorm ="NULL",
                    scaleNorm =  "AutoNorm");
mSet$dataSet$norm %>% t -> GC_mat_norm
rm(mSet)
```

## Summarise abundance by categories
```{r, echo = FALSE, warning=F, message=FALSE}
# Count and sort classes
metab_ids_curated_names %>%
  filter(
    name_orig %in% metab_ids_curated_names$name_orig & 
      name_orig %in% GC_metab_areas$`Metabolitename`) %>%
  group_by(Subclass) %>% count %>% arrange(desc(n)) -> GC_sorted_metabolite_class

# Summarise signal area by class
GC_metab_areas_mat %>% rowMeans->rmeans
data.frame(Metabolite = names(rmeans),
           mean_area = rmeans)-> GC_metab_avg_area

## Merge with annotation and filter out GC annotations
merge(metab_ids_curated_names, GC_metab_avg_area, by="Metabolite") %>%
  filter(!is.na(mean_area)) -> GC_metab_avg_area_1

```

## GC_mean
```{r, echo = FALSE, warning=F, message=FALSE}
GC_metab_areas %>%
  pivot_longer(cols = matches("NAT|CON")) %>%
  group_by(Metabolitename) %>%
  summarise(
    mean=mean(value),
    sd=sd(value)
  ) %>% arrange(desc(mean)) -> gc_mean
```

## Visual summary: heatmap by sample
```{r, echo = FALSE, warning=F, message=FALSE, eval = F}
min_fontsize=4
# Set molecular classes as row annotation, sort by size of the category
lev<-metab_ids_curated_names_df[rownames(GC_metab_areas_mat),"Subclass"] %>%
  factor(levels=GC_sorted_metabolite_class$Subclass)

# Reset colors also considering FBA metabolites
load(file="~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/data/analysis/all_metabolite_subclasses_colors.Rdata")

# Assign colors to already colored classes
ass<-intersect(GC_sorted_metabolite_class$Subclass, names(all_metabolite_subclasses_colors))
col_colors_assigned<-all_metabolite_subclasses_colors[ass]

# Find GC categories missing colors and assign new colors
diff<-setdiff(GC_sorted_metabolite_class$Subclass, names(all_metabolite_subclasses_colors))
col_colors_not_assigned<-sample(hcl.colors(length(diff), palette = "Berlin"))
names(col_colors_not_assigned) = diff

col_colors<-c(col_colors_assigned, col_colors_not_assigned)

# Convert the full Subclass vector into a color vector
GC_metab_avg_area_1 %>% 
  mutate(color=factor(Subclass,
                      labels = col_colors,
                      levels = names(col_colors)
                      )) -> bar_fill_colors

bfc=as.character(bar_fill_colors$color)
names(bfc)=bar_fill_colors$Metabolite

# Calculate intensity mean across sample and log transform
rowm<-rowMeans(log(GC_metab_areas_mat))

# Set right annot as barplots of the mean
top_annot<-
columnAnnotation(
  "Average\n log(SI)" = anno_barplot(
    x = rowm,
    gp = gpar(fill= bfc[names(rowm)]),
    axis_param = list(gp = gpar(fontsize = min_fontsize+1))
    ), 
  show_legend = F,
  annotation_name_rot = 0,
  annotation_name_side = "left",
  annotation_name_gp = gpar(fontsize = min_fontsize +5)
  )

# log transform the whole matrix and plot
sample_ord<-(colnames(GC_mat_norm))
t(GC_mat_norm)[sample_ord,] %>%
Heatmap(
          col=hcl.colors(10, palette = "PuBuGn", rev = T),
          cluster_rows = F,
          cluster_columns = F,
          show_row_names = T,
          row_names_side = "right", 
          row_names_gp = gpar(fontsize=min_fontsize+1),
          top_annotation = top_annot,
          left_annotation = NULL,
          right_annotation = NULL,
          column_split = lev,
          rect_gp = gpar(col = "white", lwd = 0),
          column_title_gp = gpar(fontsize = min_fontsize + 2
                                 ),
          column_title_rot = 20,
          row_split = GC_metab_areas_df[match(sample,sample_ord)]$group,
          gap = unit(0.5, "mm"),  #Size of gaps
          row_title_gp = gpar(fontsize = min_fontsize + 5, fontface="bold"),
          row_title_side = "left",
          row_title_rot = 0,
          column_names_rot=60,
          column_names_gp = gpar(fontsize=min_fontsize+2),
          show_heatmap_legend = T,
          heatmap_legend_param = list(
            title = "Normalised SI",  # Set legend title
            legend_direction = "vertical",    # Set legend position
            legend_width = unit(2, "cm"),
            title_gp = gpar(fontsize = min_fontsize+2, fontface = "bold"),
            labels_gp = gpar(fontsize = min_fontsize)
            ),       # Set legend size if horizontal
          height = unit(2, "cm"),  # Adjust heatmap height
          width = unit(18, "cm")   # Adjust heatmap width
          
  )->hm_GC_bysample

png(filename = "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/final/figures/heatmap_gc_bysample.png", width = 25, height = 7, units = "cm", res = 600)
draw(hm_GC_bysample,
     heatmap_legend_side = "left"
     )
dev.off()

```

## GC-MS heatmap mean
```{r, echo = FALSE, warning=F, message=FALSE, eval=F}
min_fontsize=4
# Set molecular classes as row annotation, sort by size of the category
lev<-metab_ids_curated_names_df[rownames(GC_metab_areas_mat),"Subclass"] %>%
  factor(levels=GC_sorted_metabolite_class$Subclass)

# Reset colors also considering FBA metabolites
load(file="~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/data/analysis/all_metabolite_subclasses_colors.Rdata")

# Assign colors to already colored classes
ass<-intersect(GC_sorted_metabolite_class$Subclass, names(all_metabolite_subclasses_colors))
col_colors_assigned<-all_metabolite_subclasses_colors[ass]

# Find GC categories missing colors and assign new colors
diff<-setdiff(GC_sorted_metabolite_class$Subclass, names(all_metabolite_subclasses_colors))
col_colors_not_assigned<-sample(hcl.colors(length(diff), palette = "Berlin"))
names(col_colors_not_assigned) = diff

col_colors<-c(col_colors_assigned, col_colors_not_assigned)

# Convert the full Subclass vector into a color vector
GC_metab_avg_area_1 %>% 
  mutate(color=factor(Subclass,
                      labels = col_colors,
                      levels = names(col_colors)
                      )) -> bar_fill_colors

bfc=as.character(bar_fill_colors$color)
names(bfc)=bar_fill_colors$Metabolite

# Calculate intensity mean across sample and log transform
rowm<-rowMeans(log(GC_metab_areas_mat))

# Sort each category by rowmean
foreach(i = levels(lev))%do%{
  metab_ids_curated_names_df %>% filter(Subclass == i) %>% rownames() -> metab
  rowm[metab] %>% sort(decreasing = T)
} %>% unlist -> ordered_sorted_rowmeans

sorted_lev<-metab_ids_curated_names_df[names(ordered_sorted_rowmeans),]
sorted_lev<-factor(sorted_lev, levels=levels(lev))
# Set left annot as colored bars
left_annot<-
  rowAnnotation(
  "Molecular classification" = sorted_lev,
  show_legend = F,
  col = list("Molecular classification"= row_col),
  annotation_name_rot = -90,
  annotation_name_offset = unit(1, "cm"),
  annotation_name_gp = gpar(fontsize = min_fontsize + 2),
  width = unit(0, "cm")
  )

# Set right annot as barplots of the mean
right_annot<-
rowAnnotation(
  "Average\n log(SI)" = anno_barplot(
    x = ordered_sorted_rowmeans,
    gp = gpar(fill= bfc[names(ordered_sorted_rowmeans)]),
    axis_param = list(gp = gpar(fontsize = min_fontsize+1),
                      labels_rot = 0
                      )
    ), 
  show_legend = F,
  annotation_name_rot = -90,
  annotation_name_side = "bottom",
  annotation_name_gp = gpar(fontsize = min_fontsize + 2),
  width = unit(1, "cm")
  )

# Calculate mean by zone
cbind(
  GC_mat_norm[,grepl("NAT",colnames(GC_mat_norm))] %>% rowMeans,
  GC_mat_norm[,grepl("CON",colnames(GC_mat_norm))] %>% rowMeans
) %>% as.matrix -> mat
colnames(mat) = c("Natural", "Restored")

ordered_sorted_rowmeans
mat[names(ordered_sorted_rowmeans),] %>%
Heatmap(
          col=hcl.colors(10, palette = "PuBuGn", rev = T),
          cluster_rows = F,
          cluster_columns = F,
          show_row_names = T,
          row_names_side = "right", 
          row_names_gp = gpar(fontsize=min_fontsize),
          row_names_rot = -20,
          #top_annotation = col_annot,
          left_annotation = NULL,
          right_annotation = right_annot,
          column_split = colnames(mat),
          column_title_gp = gpar(fontsize = min_fontsize + 1,
                                 fontface="bold"
                                 ),
          column_title_rot = 20, 
          row_split = sorted_lev,
          gap = unit(0.5, "mm"),  #Size of gaps
          row_title_gp = gpar(fontsize = min_fontsize + 2),
          row_title_side = "left",
          row_title_rot = -90, 
          column_names_gp = gpar(fontsize=0),
          show_heatmap_legend = T,
          heatmap_legend_param = list(
            title = "Normalised SI",  # Set legend title
            legend_direction = "horizontal",    # Set legend position
            legend_width = unit(2, "cm"),
            title_gp = gpar(fontsize = min_fontsize+2, fontface = "bold"),
            labels_gp = gpar(fontsize = min_fontsize)
            ),       # Set legend size if horizontal
          height = unit(5, "cm"),  # Adjust heatmap height
          width = unit(0.8, "cm")  # Adjust heatmap width
  )->hm_GC_mean

png(filename = "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/final/figures/heatmap_gc_mean.png",
    width = 11.55, height = 10, units = "cm", res = 600, bg = "transparent")
draw(hm_GC_mean,
     heatmap_legend_side = "top"
     )
dev.off()

```

## GC-MS stats
```{r, echo = FALSE, warning=F, message=FALSE}
#Metabolites stats
# Percentile by intensity
GC_metab_avg_area_1 %>%
  mutate(percentile_group = ntile(mean_area, 10),
         area_prop = (mean_area/sum(GC_metab_avg_area_1$mean_area))*100
         ) -> percentile
percentile %>% filter(percentile_group>=9) %>% arrange(desc(mean_area)) %>% pull(Metabolite) %>% cat

# Categories stats
## Number of metabolites per category
GC_metab_avg_area_1 %>% 
  group_by(Subclass) %>% count %>%
  arrange(desc(n)) %>%
  mutate(prop=n/nrow(GC_metab_avg_area_1)*100)

## Cumulative signal per category
GC_metab_avg_area_1 %>% 
  group_by(Subclass) %>% 
  summarise(sum=sum(mean_area),
            n=n()
            ) %>%
  arrange(desc(sum)) %>%
  mutate(prop=sum/sum(GC_metab_avg_area_1$mean_area)*100) %>% view

```

## export GC_mean
```{r, echo = FALSE, warning=F, message=FALSE}

GC_metab_areas %>%
  pivot_longer(cols = matches("NAT|CON")) %>%
  group_by(name_orig=Metabolitename) %>%
  summarise(
    mean=mean(value),
    sd=sd(value)
  ) %>% arrange(desc(mean)) %>% add_column(type="GC") -> gc_mean

fwrite(gc_mean, 
      "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/GC_means.csv")
```


#---
# ICP-MS
## Tidy dataset
```{r, echo = FALSE, warning=F, message=FALSE}
read_excel("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaB/ICPMetalli_Boschettona_Riccardo.xlsx") -> icp
names(icp)[1]="Element"
colnames(icp) %<>% str_replace("ART", "CON") %>% 
      str_replace("_", "") #fix sample names

elements <- c(
  "Li", "Be", "B", "Na", "Mg", "Al", "P", "S", "K", "Ca",
  "Ti", "V", "Cr", "Mn", "Fe", "Co", "Ni", "Cu", "Zn", "Ga", "As",
  "Se", "Rb", "Sr", "Zr", "Mo", "Pd", "Ba", "La", "Ce", "Pr", "Nd",
  "Sm", "Gd", "Tb", "Pb", "Th", "U"
)

elements <- c(
  "Na", "Mg", "Al", "P", "S", "K", "Ca",
  "Mn", "Fe", "Cu", "Zn", "Tb"
)


icp %>% pivot_longer(
  cols=matches("CON|NAT"),
  names_to = "Sample",
  values_to = "ppm"
  ) %>%
  mutate(
    Group=
      str_extract(Sample, "CON|NAT") %>% 
      factor(levels = c("CON", "NAT"), labels= c("Constructed", "Natural"))) %>%
  filter(Element %in% elements)-> icp_l
```
## ICP mean
```{r, echo = FALSE, warning=F, message=FALSE}
icp_l %>%
  group_by(Element) %>%
  summarise(mean=mean(ppm),
            sd=sd(ppm)
            ) %>% arrange(desc(mean)) -> icp_mean
```

## Check distribution and T-test
```{r, echo = FALSE, warning=F, message=FALSE}
icp_l %>% 
  group_by(Element) %>%
  mutate(
    shapiro = shapiro.test(ppm)$p.value) %>%
  summarise(
    shapiro = unique(shapiro),
    t_test = broom::tidy(t.test(ppm ~ Group))$p.value,
    wilcox_test = broom::tidy(wilcox.test(ppm ~ Group))$p.value,
    .groups = 'drop') %>% 
  mutate(
    test_pval=
     case_when(
     shapiro<0.05 ~ wilcox_test,
     shapiro>0.05 ~ t_test)) -> icp_univariate

icp_l %>%
    group_by(Element, Group) %>%
    summarise(mean=mean(ppm)) %>%
    pivot_wider(names_from = Group, values_from = mean) -> icp_mean
  merge(icp_univariate, icp_mean) %>% dplyr::select(Element, test_pval, Constructed, Natural) -> icp_summary
  
icp_summary %<>% 
  mutate(
    sig_sym=case_when(test_pval < 0.05 ~ "*", .default = ""),
    mean_ppm=(Constructed + Natural)/2,
    log2FC=log(Constructed/Natural,base = 2)
    )

```

## Summary statistics
```{r, echo = FALSE, warning=F, message=FALSE}
icp_summary %>% arrange(desc(mean_ppm)) %>% view
icp_summary %>% filter(mean_ppm>=1 & sig_sym == "*") %>%
  arrange(desc(mean_ppm))-> icp_summary_sig

icp_univariate %>% view

#Average fold change
abs(icp_summary_sig$log2FC) %>% mean
icp_summary %>% view
```

## ICP: normalise and scale with MetaboAnalyst
```{r, echo = FALSE, warning=F, message=FALSE}
setwd("/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaB/metaboanalyst/")
icp_l %>%
  pivot_wider(names_from = "Element", values_from = "ppm") %>%
  fwrite("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaB/ICP_for_metaboAn.csv")

require(MetaboAnalystR)
mSet<-InitDataObjects("msspec", "stat", FALSE);
mSet<-Read.TextData(mSet, "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaB/ICP_for_metaboAn.csv", "rowu", "disc");
mSet<-SanityCheckData(mSet);
mSet<-ReplaceMin(mSet);
mSet<-PreparePrenormData(mSet);
mSet<-Normalization(mSet,
                    rowNorm = "SumNorm",
                    transNorm ="NULL",
                    scaleNorm =  "AutoNorm");
mSet$dataSet$norm %>% t-> ICP_mat_norm
```

# ICP: Boxplots
```{r, eval=T, echo = FALSE, warning=F, message=FALSE}
icp_l %>%
  group_by(Element) %>%
  summarise(mean=mean(ppm)) %>%
  arrange(desc(mean)) %>% pull(Element)-> sorted

element_group <- c(
  # Alkali metals
  "Na" = "Alkali metals",       "K"  = "Alkali metals",
  "Li" = "Alkali metals",       "Rb" = "Alkali metals",
  "Cs" = "Alkali metals",

  # Alkaline earth metals
  "Ca" = "Alkaline earth metals", "Mg" = "Alkaline earth metals",
  "Sr" = "Alkaline earth metals", "Ba" = "Alkaline earth metals",
  "Be" = "Alkaline earth metals",

  # Transition metals
  "Ti" = "Transition metals", "Zr" = "Transition metals", "Hf" = "Transition metals",
  "V"  = "Transition metals", "Nb" = "Transition metals", "Ta" = "Transition metals",
  "Cr" = "Transition metals", "Mo" = "Transition metals", "W"  = "Transition metals",
  "Mn" = "Transition metals", "Re" = "Transition metals",
  "Fe" = "Transition metals", "Ru" = "Transition metals",
  "Co" = "Transition metals", "Rh" = "Transition metals", "Ir" = "Transition metals",
  "Ni" = "Transition metals", "Pd" = "Transition metals", "Pt" = "Transition metals",
  "Cu" = "Transition metals", "Ag" = "Transition metals", "Au" = "Transition metals",
  "Zn" = "Transition metals", "Cd" = "Transition metals",

  # Metalloids & post-transition metals (Boron–Carbon–Pnictogen groups)
  "B"  = "Metalloids",       "Si" = "Metalloids",
  "As" = "Metalloids",       "Te" = "Metalloids",
  "Sb" = "Metalloids",
  "Al" = "Post-transition metals", "Ga" = "Post-transition metals",
  "Tl" = "Post-transition metals", "Sn" = "Post-transition metals",
  "Pb" = "Post-transition metals", "Bi" = "Post-transition metals",

  # Nonmetals
  "P"  = "Nonmetals",  "S"  = "Nonmetals",  "Se" = "Nonmetals",

  # Halogens
  "I"  = "Halogens",   "Br" = "Halogens",

  # Lanthanides
  "La" = "Lanthanides", "Ce" = "Lanthanides", "Pr" = "Lanthanides",
  "Nd" = "Lanthanides", "Sm" = "Lanthanides", "Eu" = "Lanthanides",
  "Gd" = "Lanthanides", "Tb" = "Lanthanides", "Dy" = "Lanthanides",
  "Ho" = "Lanthanides", "Er" = "Lanthanides", "Tm" = "Lanthanides",
  "Yb" = "Lanthanides", "Lu" = "Lanthanides",

  # Actinides
  "U"  = "Actinides"
)


icp_l %>%
  left_join(
  data.table(
    Element=names(element_group),
    Element_Group = element_group
    )) %>%
    mutate(Element = factor(Element, levels = sorted)) %>% data.table-> icp_l1
```

# --
## Collect plots
## Visual summary: boxplots
```{r, echo = FALSE, warning=F, message=FALSE}
library(ggh4x)
options(scipen =-1) # strongly prefer scientific notation
fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/metabolite_classes.csv") %>% dplyr::rename(Class=hmdb_class, 
       Metabolite=metabolite) %>%
  mutate(Class = str_remove(Class, "carbohydrate |, peptides,"))-> metab_ids_curated_names2

metab_ids_curated_names2$Class %>% unique()

c("Amino acids & analogues","Nucleosides & nucleotides",
  "Carbohydrates & conjugates","Carboxylic acids",
  "Lipids & Fatty Acyls", "Alkaloids & N-heterocycles",
  "Alcohols & polyols", "Organosulfur compounds"
  )->sorted_m_cls


metab_class_cols<-hcl.colors(
  n=length(sorted_m_cls),
  palette = "Set3",rev = T, alpha = 0.5)
names(metab_class_cols)=sorted_m_cls
strip <- strip_themed(background_x = elem_list_rect(fill = alpha("darkgreen", 0.3)))


c("Amino acids & analogues","Nucleosides & nucleotides"
  )->sorted_m_cls1

c(
  "Carbohydrates & conjugates","Carboxylic acids",
  "Lipids & Fatty Acyls", "Alkaloids & N-heterocycles",
  "Alcohols & polyols", "Organosulfur compounds"
  )->sorted_m_cls2

# L-valine cutoff value
lc_mean %>% filter(mean>9037545) %>%
  left_join(select(LC_metab_areas2, name_orig, Metabolite)) %>%
  pull(Metabolite) -> above_valine

lc_mean %>% filter(mean<=9037545) %>%
  left_join(select(LC_metab_areas2, name_orig, Metabolite)) %>%
  pull(Metabolite) -> below_valine

lc_mean %>% filter(mean>3324084.4) %>%
  left_join(select(LC_metab_areas2, name_orig, Metabolite)) %>%
  pull(Metabolite) -> above_inosine

lc_mean %>% filter(mean<=3324084.4) %>%
  left_join(select(LC_metab_areas2, name_orig, Metabolite)) %>%
  pull(Metabolite) -> below_inosine

plot_subclasses<-c(
                    "Amino acids & analogues 1",
                    "Nucleosides & nucleotides 1",
                    "Amino acids & analogues 2",
                    "Nucleosides & nucleotides 2"
                      )
corrected_classes<-c(
    "Carbohydrates & conj.","Carboxylic acids",
    "Lipids & Fatty Acyls", "Alkaloids & heterocycles",
    "Alcohols & polyols", "Organosulphurs"
                     )


LC_metab_areas2 %>%
  left_join(metab_ids_curated_names2) %>%
  mutate(Metabolite = 
           case_when(Metabolite == "Uridine diphosphate-N-acetylgalactosamine" ~  "UDP-N-acetylgalactosamine", .default = Metabolite),
         Class=case_when(
           Metabolite %in% above_valine & 
             Class == sorted_m_cls1[1] ~ "Amino acids & analogues 1",
         .default = Class),
         
         Class=case_when(
           Metabolite %in% below_valine & 
             Class == sorted_m_cls1[1] ~ "Amino acids & analogues 2",
         .default = Class),
         
         Class=case_when(
           Metabolite %in% above_inosine & 
             Class == sorted_m_cls1[2] ~ "Nucleosides & nucleotides 1",
         .default = Class),
         
         Class=case_when(
           Metabolite %in% below_inosine & 
             Class == sorted_m_cls1[2] ~ "Nucleosides & nucleotides 2",
         .default = Class)
         ) %>%
  pivot_longer(cols=contains(c("NAT","CON"))) %>%
  mutate(Class=factor(Class, level=plot_subclasses)) %>%
  filter(Class %in% plot_subclasses) %>%
  ggplot(aes(x=value,y=reorder(Metabolite, value))) +
  geom_boxplot(outliers = F) +
  geom_jitter(size=0.5) +
  facet_wrap2(~Class, scales = "free", ncol = 2, strip=strip) +
  #facet_wrap(~Class, scales = "free", ncol = 2) +
  #force_panelsizes(rows = c(2, 1, 1, 0.2)) +
  force_panelsizes(rows = c(2, 2, 2, 2)) +
  theme_bw() +
  theme(axis.text.y=element_text(size = 15),
        axis.text.x=element_text(angle = 45, hjust = 1),
        strip.text =element_text(face = "bold", size = 18)) +
  ylab("") + xlab("") -> p1

LC_metab_areas2 %>%
  left_join(metab_ids_curated_names2) %>%
  mutate(
         
         Class=case_when(
           Class == sorted_m_cls2[1] ~ "Carbohydrates & conj.",
         .default = Class),
         
         Class=case_when(
           Class == sorted_m_cls2[4] ~ "Alkaloids & heterocycles",
         .default = Class),
         
         Class=case_when(
           Class == sorted_m_cls2[6] ~ "Organosulphurs",
         .default = Class)
         
         
         ) %>%
  pivot_longer(cols=contains(c("NAT","CON"))) %>%
  
  mutate(Class=factor(Class,
                      level=c(corrected_classes))) %>%
  filter(Class %in% c(corrected_classes)) %>%
  ggplot(aes(x=value,y=reorder(Metabolite, value))) +
  geom_boxplot(outliers = F) +
  geom_jitter(size=0.5) +
  facet_wrap2(~Class, scales = "free", ncol = 2, strip=strip) +
  force_panelsizes(rows = c(2, 1, 1, 0.2)) +
  theme_bw() +
  theme(axis.text.y=element_text(size = 15),
        axis.text.x=element_text(angle = 45, hjust = 1),
        strip.text =element_text(face = "bold", size = 18)) +
    ylab("") + xlab("Peak Area") +
  scale_x_continuous(
    labels = function(x) formatC(x, format = "e", digits = 1))-> p2

metab_ids_curated_names2$Class %>% unique()

c("Lipids & Fatty Acyls","Aromatic compounds",
  "Hydrocarbons", "Organosulphurs")->sorted_m_cls3

strip <- strip_themed(background_x = elem_list_rect(fill = alpha("sandybrown", 0.3)))
GC_metab_areas2 %>%
  left_join(metab_ids_curated_names2) %>%
  
    mutate(
         
         Class=case_when(
           Class == sorted_m_cls2[6] ~ "Organosulphurs",
         .default = Class)
         
         ) %>%
  
  pivot_longer(cols=contains(c("NAT","CON"))) %>%
  mutate(Class=factor(Class, level=sorted_m_cls3)) %>%
  ggplot(aes(x=value,y=reorder(Metabolite, value))) +
  geom_boxplot(outliers = F) +
  geom_jitter(size=1) +
  facet_wrap2(~Class, scales = "free", ncol = 4, strip=strip) +
  force_panelsizes(rows = c(0.5)) +
  theme_bw() +
  theme(axis.text.y=element_text(size = 15),
        axis.text.x=element_text(angle = 45, hjust = 1),
        strip.text =element_text(face = "bold", size = 18),
        axis.title.x = element_text(size=15)
        ) +
      ylab("") + xlab("Peak Area") +
  scale_x_continuous(
    labels = function(x) formatC(x, format = "e", digits = 1))-> p3


# ICP
icp_l1$Element_Group %>% unique->element_groups

hcl.colors(n = length(element_groups), palette = "Earth") ->unified_colors
qualitative_hcl(length(element_groups))->unified_colors

unified_colors <- c(
  "Alkali metals"         = "#F8766D",
  "Alkaline earth metals" = "#7CAE00",
  "Lanthanides"           = "#CD9600",
  "Nonmetals"             = "#00BFC4",
  "Post-transition metals"= "#00A9FF",
  "Transition metals"     = "#C77CFF"
)
names(unified_colors) = element_groups


icp_l1$Element_Group1 = factor(icp_l1$Element_Group, levels=element_groups)


ggplot(icp_l1[Element %in% c("Na", "Ca", "Mg","Fe","S", "Al" )], aes(x=Element, y=ppm)) + 
  geom_boxplot(outliers = F) +
  geom_jitter(size=1, aes(color = Element_Group1)) + 
    theme_bw() +
  scale_color_manual(values = unified_colors, drop = F)+
  ylab("mg/Kg")+ xlab("")+
  ggtitle("ICP-MS") +
  theme(legend.position = "none",
        axis.text.x = element_text(size=15),
        axis.title.y = element_text(size=15),
        plot.title = element_text(size=18, face = "bold")
        ) -> p4

ggplot(icp_l1[Element %in% c("K", "P", "Tb","Mn","Zn")], aes(x=Element, y=ppm)) + 
  geom_boxplot(outliers = F) +
  geom_jitter(size=1, aes(color = Element_Group1)) + 
    theme_bw() +
  scale_color_manual(values = unified_colors, drop = F) +
  ylab("") + xlab("")+
    ggtitle("ICP-MS") +
  theme(legend.position = "none",
        axis.text.x = element_text(size=15),
        plot.title = element_text(size=18, face = "bold")
        ) +
  scale_y_continuous(
    labels = function(x) formatC(x, format = "e", digits = 1)) + 
  ggtitle("ICP-MS")-> p5



legend_plot <- ggplot(icp_l1, aes(x = Element, y = ppm, color = Element_Group1)) +
  geom_point() +
  scale_color_manual(values = unified_colors, name = "Element Group") +
  theme_bw() +
  ylab("") +
  theme(legend.position = "bottom",
        text=element_text(size=15),
        legend.title = element_text(face = "bold")) +
  guides(color = guide_legend(
    ncol = 3,
    nrow = 2,
    override.aes = list(size = 10)  # increase point size in legend
  ))

# Extract just the legend
custom_legend <- get_legend(legend_plot)


wrap_elements(custom_legend) + ggplot() + theme_void() + 
  p4 + p5 + gg_chn + gg_gran +
  plot_layout(widths = c(0.5, 0.3, 2, 2, 2, 2), nrow = 1) -> icp;icp

(p4 + p5) / (wrap_elements(custom_legend)) + 
  plot_layout(widths = c(1,1,1), heights = c(1,1,0.5)) -> icp

gg_gran + gg_chn + icp +
   plot_layout(widths = c(1,1,1), heights = c(1,1,2))


require(patchwork)

gg_fill<-ggplot() + theme_void()

((gg_gran + gg_chn + p4 + p5) +
  plot_layout(nrow = 1)) /
  (gg_fill  + (wrap_elements(custom_legend))) +
  plot_layout(widths = c(1, 1), heights = c(1, 0.1)) +
  theme(plot.margin = margin(t = 0, b = -0, l = 2, r = 2)) -> first_row

((gg_gran + gg_chn + p4 + p5) +
   plot_layout(nrow = 1)) / 
  wrap_elements(custom_legend) +
   plot_layout(heights = c(10, 1)) -> first_row;first_row

 #lc_ms<-wrap_elements(p1/p1.1) +
#  theme(plot.margin = margin(t = -10, b = -10, l = -10, r = -10))

p1 + p2  + plot_layout(ncol = 2)-> lc_ms1

  (
  first_row
    /
  lc_ms1
    / 
   p3 
  ) +
  plot_layout(heights = c(0.5, 0.2, 1.8, 0.4),#c(2, 0.5, 0.5, 0.01),
              ) -> p


ggsave(p,
       filename = "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/boxplot_lcms.png", dpi = 600, width = 32, height = 25, units = "in", scale = 0.85
       )
file.show("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/boxplot_lcms.png")
  

```



## Visual summary: heatmap
```{r, eval=F, echo = FALSE, warning=F, message=FALSE}
require(ComplexHeatmap)
min_fontsize=4
icp %>% column_to_rownames("Element") %>% as.matrix -> icp_m

data.frame(
  Sample=colnames(icp_m),
  Group=str_extract(colnames(icp_m), "CON|NAT") %>% 
      factor(levels = c("CON", "NAT"), labels= c("Constructed", "Natural")))-> metadata_icp
metadata_icp[c(7:12, 1:6),]

rowSums(icp_m) %>% sort(decreasing = T) %>% names -> sorted_elements

col_ha = columnAnnotation(
  "Average \nlog(ppm)" = anno_barplot(log(rowMeans(icp_m)[sorted_elements]),
                                      axis_param = list(gp = gpar(fontsize = min_fontsize+1))
                                      ),
                         annotation_name_rot = 0,
  annotation_name_side = "right",
  annotation_name_gp = gpar(fontsize = min_fontsize+2)
  )
                   
colnames(ICP_mat_norm) %<>% str_replace("art_", "CON") %>% str_replace("nat_", "NAT")
ICP_mat_norm[sorted_elements,metadata_icp[c(7:12, 1:6),]$Sample] %>%
  t %>%
        Heatmap(
          col=hcl.colors(10, palette = "PuBuGn", rev = T),
          show_row_names = T, 
          row_names_side = "left",
          show_row_dend = FALSE,
          top_annotation = col_ha,
          cluster_rows = F,
          cluster_columns = F,
          #column_title_gp = gpar(fontsize = min_fontsize + 3, fontface="bold"),
          column_title_rot = 90,
          column_names_centered = T,
          column_title_side = "top",
          row_split = metadata_icp[c(7:12, 1:6),]$Group,
          rect_gp = gpar(col = "white", lwd = 1),
          gap = unit(1, "mm"),  
          row_title_gp = gpar(fontsize = min_fontsize + 5, fontface="bold"),
          row_title_side = "right",
          row_title_rot = 0,
          row_names_rot = 0,
          row_names_gp = gpar(fontsize= min_fontsize + 1),
          column_names_rot = 60,
          column_names_gp = gpar(fontsize = min_fontsize + 2),
          column_names_side = "bottom",
          show_heatmap_legend = T,
          heatmap_legend_param = list(
            title = "Normalised ppm values",  # Set legend title
            legend_direction = "horizontal",    # Set legend position
            legend_width = unit(2, "cm"),
            labels_gp = gpar(fontsize = min_fontsize),
            title_gp = gpar(fontsize = min_fontsize + 2, fontface = "bold")
            ),       # Set legend size if horizontal
           width= unit(13.5, "cm"),  # Adjust heatmap height
          height = unit(3, "cm")
  ) -> hm_icp;hm_icp

png(filename = "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/final/figures/heatmap_icp_bysample.png", height = 6, width = 17, units = "cm", res = 600)
draw(hm_icp, heatmap_legend_side = "bottom")
dev.off()
```

## Rank metabolites by abundance (quantile)
```{r, echo = FALSE, warning=F, message=FALSE}
lc_mean %>%
  mutate(quant = ntile(mean, 4),
         perc = ntile(mean, 100)
         ) -> lc_mean_quant

```
# ----
## SUPPLEMENTARY 3: Export GC-, LC-, ICP- MS data
```{r, echo = FALSE, warning=F, message=FALSE}
require(openxlsx)
rbind(lc_mean, gc_mean) %>% rename(name_orig = `Metabolite name`) -> ms_mean

rbind(
LC_metab_ids %>% 
  dplyr::select(`Metabolite name`, `Area media`, data_type) %>% 
  dplyr::rename(media = `Area media`),
GC_metab_ids %>% dplyr::select(`Metabolite name`, media, data_type),
by= "Metabolite name") -> LC_GC_names_area



LC_GC_names_area %<>% dplyr::rename(name_orig=`Metabolite name`)

merge(metab_ids_curated, ms_mean) %>%
  dplyr::select(name_orig, Metabolite, HMDB, KEGG, Subclass, mean, sd, type) %>%
  dplyr::rename(
    "Annotation" = name_orig,
    "Database Annotation" = Metabolite,
    "HMDB ID" = HMDB,
    "KEGG ID" = KEGG,
    "Molecular Subclass" = Subclass,
    "Average Signal Intensity"= mean,
    "Standard deviation"= sd
  ) -> LC_GC_names_area2

LC_GC_names_area2 %>% filter(type == "LC") -> LC_names_area2
LC_GC_names_area2 %>% filter(type == "GC") -> GC_names_area2


wb <- createWorkbook()
addWorksheet(wb, "Granulometry")
writeData(wb, "Granulometry", x = gran_mean)

addWorksheet(wb, "CHNS")
writeData(wb, "CHNS", x = chn_mean)

addWorksheet(wb, "LC-MS summary table")
writeData(wb, "LC-MS summary table", x = LC_names_area2)

addWorksheet(wb, "GC-MS summary table")
writeData(wb, "GC-MS summary table", GC_names_area2)

addWorksheet(wb, "ICP-MS summary table")
writeData(wb, "ICP-MS summary table", icp_mean)

saveWorkbook(wb,
             "/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/final/supplementary/Supplementary File 2 - Abiotic parameters.xlsx",
             overwrite = TRUE)
```
