---
title: "bosc23_network_analysis"
author: "Riccardo Frizzo"
date: "2024-08-12"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

#Init
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
libs<-c("data.table", "tidyverse", "foreach", "magrittr","ggtree", "ape")
invisible(lapply(libs, require, character.only = TRUE))

hcl.colors(n=3, palette ="Zissou 1",alpha = 0.9)
zy=c("#A5191CE6","#FADB2BE6","#65B59AE6")
#scales::show_col(zy[c(3,1,2)],ncol = 3)
saltmarsh_colors = c(Constructed=zy[1], Natural=zy[3], ns = zy[2])

# Load Phylum colors
load(file = "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/data/tax/tree_colors_14apr25.RData")
tree_colors<-c(tree_colors, "firebrick4")
names(tree_colors)[length(tree_colors)]="Thermoproteota" #Add the archaea :)


```

## Preproc MAGs (make mag_analysis)
### MAG analysis starts here!
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
# Import taxonomy, relative abundance, differential abundance data
mag_taxonomy<-data.table::fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/data/tax/mag_taxonomy_mags99c70c10a99.csv")
names(mag_taxonomy)[1]="OTU"

# Extrapolate phylogenetic distance
mag_tree <- ggtree::read.tree("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaG/coassemby_07feb24/mags_c70c10a99/gtdb220/classify/gtdbtk.backbone.bac120.classify.tree")
tree_dist_matrix <- ape::cophenetic.phylo(mag_tree)
mag_dist_mat<-tree_dist_matrix[grepl("art|nat", colnames(tree_dist_matrix)),grepl("art|nat", colnames(tree_dist_matrix))]
mag_dist_vector <- as.vector(mag_dist_mat[upper.tri(mag_dist_mat)])
dist(mag_dist_mat) %>% hclust-> tax_hclust
data.frame(mag=tax_hclust$labels,
           phylo_dist=tax_hclust$order) %>% rename(OTU=mag)->phylogenetic_distance_hclust

#Import quality data
mag_qual<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaG/coassemby_07feb24/mags_c70c10a99/checkm2_mags/quality_report.tsv")
names(mag_qual)[1]="OTU"

#Import differential abundance data
mag_diff<-data.table::fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/data/analysis/mag_diff_mags99c70c10a99.csv")

#Import MPSE analysis
load("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/data/analysis/mag_MPSE_norm_mags99c70c10a99.RData")

# Make wide abundance dataset for association analysis
mag_MPSE_norm %>% dplyr::select(OTU, Sample, Group, RelhellingerBySample) %>%  #RelhellingerBySample or Abundance
  pivot_wider(names_from = OTU,
              values_from = RelhellingerBySample #RelhellingerBySample or Abundance
              ) %>% 
  dplyr::select(!Group) %>% 
  column_to_rownames("Sample")->mag_abund_w

# Make mag_analysis

mag_taxonomy %>%
  merge(mag_diff) %>%
  merge(mag_qual) -> mag_analysis

# Fill empty values in diff analysis
mag_analysis %<>%
  mutate(condition_higher_abund = case_when(condition_higher_abund == "" ~ "ns", .default = condition_higher_abund))

# Import phylo_dist

mag_analysis %<>%
  full_join(phylogenetic_distance_hclust, by = "OTU")

# Export mag_analysis
fwrite(mag_analysis, "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/data/analysis/mag_analysis_17oct24.tsv", sep = "\t", dec = ".")



```

### Export MAG analysis for publishing
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
names(mag_analysis)

# Prepare a tidy version for printing
mag_analysis %>% 
  mutate(
    mean_mag_abund_Art=round(mean_mag_abund_Art,3),
    mean_mag_abund_Nat=round(mean_mag_abund_Nat,3),
    average= round((mean_mag_abund_Nat + mean_mag_abund_Art)/2, 3)) %>%
  
  dplyr::select(OTU,
                Phylum, Class, Order, Family, Genus, Species,
                mean_mag_abund_Art, mean_mag_abund_Nat, average, condition_higher_abund,
                Completeness, Contamination, Genome_Size, GC_Content, Total_Coding_Sequences
                ) %>% 
  
  dplyr::rename(
    "MAG ID" = OTU,
    "Average Relative Abundance"= average,
    "Genome Size" = Genome_Size,
    "GC Content" = GC_Content,
    "Number of coding sequences" = Total_Coding_Sequences,
     "Average Relative Abundance in Contructed"= mean_mag_abund_Art,
     "Average Relative Abundance in Natural"= mean_mag_abund_Nat,
     "Significantly more abundant in:"= condition_higher_abund
  ) -> mag_analysis_nice

fwrite(mag_analysis_nice, "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/final/supplementary/mag_analysis_nice_13jan25.csv",
       sep = ";", dec = ",")

```
## Preproc vOTUs (import vir_analysis)
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
# Import Virus information
vir_analysis<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/data/analysis/vir_analysis_03jan25.csv")

phold_description<-fread("/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/vir/pipeline3_29aug24/pharokka_phold/phold/phold_all_cds_functions.tsv") %>%
  filter(Contig %in% vir_analysis$votu)

phold_description_w<-phold_description %>%
pivot_wider(names_from = Description,  values_from = Count)

phold_annot<-fread("/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/vir/pipeline3_29aug24/pharokka_phold/phold/phold_per_cds_predictions.tsv") %>%
  filter(contig_id %in% vir_analysis$votu)
names(phold_annot)[1]="votu"

pharokka_description<-fread("/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/vir/pipeline3_29aug24/pharokka_phold/pharokka_cds_functions.tsv")

pharokka_annot<-fread("/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/vir/pipeline3_29aug24/pharokka_phold/pharokka_cds_final_merged_output.tsv")

bakta<-fread("/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/vir/pipeline3_29aug24/bakta/all_OTUs_no_doubleBB.tsv")
names(bakta)[1]="votu"
bakta %<>%
  filter(votu %in% vir_analysis$votu)

bakta_hypotheticals<-fread("/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/vir/pipeline3_29aug24/bakta/all_OTUs_no_doubleBB.hypotheticals.tsv")
names(bakta_hypotheticals)[1]="votu"
bakta_hypotheticals %<>%
  filter(votu %in% vir_analysis$votu)

```
# ~~~ MAGs ~~~
## Differential abundance of MAGs 
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=FALSE}
mag_analysis %>% 
  group_by(condition_higher_abund) %>%
  count(sort = T) %>%
  ungroup %>%
  mutate(prop=100*n/sum(n))
```

## Presence-absence of MAGs 
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=FALSE}
mag_analysis %>%
  mutate(
    presence_absence =
    case_when(
      (mean_mag_abund_Art == 0 & mean_mag_abund_Nat>0)
      |
      (mean_mag_abund_Nat == 0 & mean_mag_abund_Art>0) ~ T, .default = F)
  ) %>% group_by(condition_higher_abund, presence_absence) %>% count
```

## Phage-defence integrations
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
def_paths<-list.files(
  path="/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaG/coassemby_07feb24/mags_c70c10a99/defensefinder_11set",
  recursive = T,
  full.names = T,
  pattern = "systems.tsv"
  )

# Loop over the files, add MAG IDs
fread(def_paths[2])
foreach(i = def_paths)%do%{
  fread(i) %>% add_column(OTU=str_extract(i,"[:alpha:]+_SemiBin_[:digit:]+"), .before = "sys_id")
} %>% rbindlist()->def_sys

#filter the table based on a specific type
def_sys %>% filter(type == "SoFIC") %>% pull(OTU)->mags_sofi
```

#----------------------------------------------------------------------
# ~~~ vOTUs ~~~
## Size distribution of vOTUs
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show'}
vir_qual<-fread("/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/vir/pipeline3_29aug24/Rtables/genomad_virsorter_checkv_filtered.csv") %>%
  mutate(seq_name=str_replace(seq_name, "\\|\\|", "_")) %>%
  filter(seq_name %in% vir_analysis$votu)

vir_qual %>% summarise(
  mean=mean(checkv_contig_length)/1e3,
  sd=sd(checkv_contig_length)/1e3
)

# Genome size by Class
vir_analysis[Class == "Malgrandaviricetes", size_kb] %>% range
vir_analysis[Class == "Caudoviricetes", size_kb] %>% range
vir_analysis[Class == "Caudoviricetes", size_kb]  %>% hist
```

#---
# Virus functional annotation
# Import vocabularies 
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show'}
vfdb_vocabulary<-fread("~/rstudio-server/storage16TB/riccardo/tools/phold/vfdb_description_output.csv")
defencefinder_vocabulary<-fread("~/rstudio-server/storage16TB/riccardo/tools/phold/defensefinder_plddt_over_70_metadata.tsv")
```
## Condense phold categories
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show'}
# Count annotated
phold_description$Description %>% unique
phold_description %>%
  filter(!Description %in% c("CDS", "unknown function")) %>%
  group_by(Contig) %>%
  summarise(sum_known=sum(Count))-> vir_sum_fun_annot

# Add to phold_description
phold_description_w %<>%
  merge(vir_sum_fun_annot)

# Group categories
phold_description_w %>%
  mutate(
    annot_prop=round(sum_known/(sum_known +`unknown function`) ,3),
    structural= `head and packaging` + tail + connector,
    DNARNA=`DNA, RNA and nucleotide metabolism`,
    integration_excision = `integration and excision`,
    transcription = `transcription regulation`,
    moron= `moron, auxiliary metabolic gene and host takeover`,
    unk= `unknown function`) %>% 
  dplyr::rename(votu=Contig) %>%
  dplyr::select(votu, 
                structural, 
                DNARNA, integration_excision, lysis,
                transcription,
                moron,
                other,
                unk,
                annot_prop,
                CDS
                )-> phold_annot_summary

phold_annot_summary %>%
  dplyr::rename("Structural components" = structural,
                "DNA, RNA, Nucleotide metabolism" = DNARNA,
                "Integration and excision" = integration_excision,
                "Lysis" = lysis,
                "Transcription regulations" = transcription,
                "Auxiliary genes and host takeover" = moron,
                "Other" = other,
                "Unknown" = unk
                ) ->phold_annot_summary1
```

## Heatmap: distribution of phold categories across vOTU
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show'}
require(ComplexHeatmap)

#Remove rare classes
vir_analysis %>%
  group_by(Class) %>% count %>%
  filter(n>2) %>% pull(Class) -> not_rare_classes

vir_analysis %>%
  filter(Class %in% not_rare_classes) %>%
  pull(votu)->votu_not_rare_classes

#Preapre matrix
phold_annot_summary1 %>%
  dplyr::select(-c(annot_prop, CDS)) %>%
  filter(votu %in% votu_not_rare_classes) %>%
  column_to_rownames("votu") %>% as.matrix()-> phold_annot_summary_mat

# Transform into presence-absence
phold_annot_summary_mat[phold_annot_summary_mat>1]=1

# Define column properties
lev<-colnames(phold_annot_summary_mat)

set.seed(25012025)
column_col<-
  structure(
    sample(hcl.colors(length(unique(lev)), palette = "RdYlGn", alpha = 0.7)),
    names = unique(lev))
column_col-> phold_cat_colors
#colorspace::swatchplot(column_col)

col_annot<-
  columnAnnotation(
  "Functional \n Category" = factor(lev, levels = unique(lev)),
  show_legend = F,
  col = list("Functional \n Category"=column_col),
  annotation_name_gp = gpar(fontsize = 10),
  annotation_name_side = "left"
  )

lev2=vir_analysis[votu %in% hclust_sorted_votus, Class]
lev2[lev2 == ""] = "Unknown"
load(file = "/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/data/analysis/virus_class_colors.RData")
names(virus_class_colors)[names(virus_class_colors) == ""] = "Unknown"
virus_class_colors[names(virus_class_colors) == "Unknown"] = "grey80"

# Set the order of row splits
ord<-lev2 %>% table %>% sort(decreasing = F) %>% names
lev2 = factor(lev2,
              levels = (ord), # by frequency
              ordered = T
              )

# Sort the matrix within each Class
foreach(i = not_rare_classes) %do% {
  
  vir_analysis %>%
    filter(Class == i) %>% pull(votu) -> votu
  
  phold_annot_summary_mat[votu,] -> curr_mat
  
  curr_mat %>%
    dist %>% hclust %>% cutree(h=1) %>% sort %>%
    names -> sorted_votus
  phold_annot_summary_mat[sorted_votus,] %>% as.data.frame %>%
    rownames_to_column("votu")
  
} %>% rbindlist %>% column_to_rownames("votu") %>% as.matrix -> hclust_sorted_votus_mat
phold_annot_summary_mat<-hclust_sorted_votus_mat


row_col <- structure(
  virus_class_colors[ord],
  names = ord
)

# Create a row annotation object
row_annot <- rowAnnotation(
  "vOTU \n Class" = lev2,
  col=list("vOTU \n Class" = row_col),
  annotation_name_gp = gpar(fontsize = 10),
  show_legend = F
)

phold_annot_summary_mat %>% Heatmap(
          col=c("white", "lightblue"),
          show_row_names = FALSE,
          show_row_dend = FALSE,
          top_annotation = col_annot,
          right_annotation = row_annot,
          cluster_rows = F,       # Row clustering interferes with row_split order
          cluster_row_slices = F,
          cluster_columns = F, 
          column_split = factor(colnames(phold_annot_summary_mat), levels=colnames(phold_annot_summary_mat)),
          column_title_gp = gpar(fontsize = 12, fontface="bold"),
          column_title_rot = 45, 
          row_split = lev2,
          row_title_gp = gpar(fontsize = 0),
          row_title_side = "right",
          row_title_rot = 0,
          row_names_gp = gpar(fontsize=0),
          column_names_gp = gpar(fontsize=0),
          show_heatmap_legend = F,
          border=F,
          heatmap_width = unit(17,"cm"),
          heatmap_height = unit(13,"cm")
  ) -> hm_vir_phold_desc;

png(filename = "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/final/figures/heatmap_vir_phold_categories.png",
    width = 20, height = 21, res = 600, units = "cm")
draw(hm_vir_phold_desc, heatmap_legend_side = "top")
dev.off()
```
## Heatmap by phold category
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show'}
# Rename `function` column to a more handy name
phold_annot %>%
  dplyr::rename(func=`function`) -> phold_annot1

phold_annot1 %>%
  mutate(func= 
           case_when(
             func %in% c("head and packaging", "tail", "connector") ~ "Structural components",
             func == "moron, auxiliary metabolic gene and host takeover" ~ "Auxiliary genes and host takeover",
             func == "other" ~ "Other", .default = func)) -> phold_annot2

# Select a category
for(cur_fun in unique(phold_annot2$func)[c(2,4,5)]) {

# Pick a color
color=phold_cat_colors[cur_fun]

phold_annot2 %>%
  filter(func == cur_fun & votu %in% votu_not_rare_classes) %>%
  group_by(votu, product) %>%
  count %>%
  pivot_wider(names_from = product, values_from = n, values_fill = 0) %>%
  column_to_rownames("votu")-> phold_cat_sum_df

colSums(phold_cat_sum_df) %>% quantile(probs = seq(0,1,0.1)) -> decile


#filter out products with less than 10 hits
phold_cat_sum_df[,colSums(phold_cat_sum_df)>decile["80%"]] -> phold_cat_sum_df_f

#Transform to matrix
phold_cat_sum_df_f %>% as.matrix -> phold_cat_mat
phold_cat_mat[phold_cat_mat>0]=1

vir_analysis[votu %in% rownames(phold_cat_mat),Class] %>% unique -> classes
# Sort the matrix within each Class
foreach(i = classes) %do% {
  
  vir_analysis[votu %in% rownames(phold_cat_mat),] %>%
    filter(Class == i) %>% pull(votu) -> votu
  
  phold_cat_mat[votu,] -> curr_mat
  
  curr_mat %>%
    dist %>% hclust %>% cutree(h=1) %>% sort %>%
    names -> sorted_votus
  phold_cat_mat[sorted_votus,] %>% as.data.frame %>%
    rownames_to_column("votu")
  
} %>% rbindlist %>% column_to_rownames("votu") %>% as.matrix -> hclust_sorted_votus_mat
phold_cat_mat<-hclust_sorted_votus_mat

# Define column titles properties
lev<-rep(cur_fun, ncol(phold_cat_mat))

column_col<-
  structure(
    color,
    names = cur_fun)

col_annot1<-
  columnAnnotation(
  "Gene \n annotation" = factor(lev, levels = unique(cur_fun)),
  show_legend = F,
  col = list("Gene \n annotation" = column_col),
  annotation_name_gp = gpar(fontsize = 0),
  annotation_name_side = "left")

col_annot2<-
  columnAnnotation(
  "Gene count" = anno_barplot(
    -colSums(phold_cat_mat[,names(sort(colSums(phold_cat_mat), T))]),
    axis_param = list(
      at = pretty(-colSums(phold_cat_mat), n=2),  # Adjust axis ticks
      labels = rev(pretty(colSums(phold_cat_mat), n=2))  # Show positive labels
      ),
    ylim=range(pretty(-colSums(phold_cat_mat), n=2))
    ),
  show_legend = F)


lev2=vir_analysis[votu %in% rownames(phold_cat_mat), Class]
lev2[lev2 == ""] = "Unknown"
load(file = "/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/data/analysis/virus_class_colors.RData")
names(virus_class_colors)[names(virus_class_colors) == ""] = "Unknown"
virus_class_colors[names(virus_class_colors) == "Unknown"] = "grey80"

# Set the order of row splits
ord<-lev2 %>% table %>% sort(decreasing = F) %>% names

lev2 = factor(lev2,
              levels = ord, # by frequency
              ordered = T, labels = ord
              )
row_col<- 
  structure(
    virus_class_colors,
    names = names(virus_class_colors))

# Create a row annotation object
row_annot <- rowAnnotation(
  "vOTU \n Class" = lev2,
  col=list("vOTU \n Class"=row_col[levels(lev2)]),
  annotation_name_gp = gpar(fontsize = 10),
  annotation_name_side="top",
  annotation_name_rot = 45,
  show_legend = F
)

phold_cat_mat[,names(sort(colSums(phold_cat_mat), T))] %>% # order columns by abundace
  Heatmap(
          col=c("white", as.character(color)),
          show_row_names = FALSE,
          show_row_dend = FALSE,
          top_annotation = col_annot1,
          bottom_annotation = col_annot2,
          right_annotation = row_annot,
          cluster_rows = F,
          cluster_row_slices = T,
          cluster_columns = F,
          column_split = rep(cur_fun, ncol(phold_cat_mat)),
          rect_gp = gpar(col = "white", lwd = 0),
          column_title_gp = gpar(fontsize = 0),
          column_title_rot = 0, 
          column_title_side = "bottom",
          row_split = lev2,
          row_title_gp = gpar(fontsize = 0),
          row_title_side = "right",
          row_title_rot = 0,
          row_names_gp = gpar(fontsize=0),
          column_names_gp = gpar(fontsize=10),
          column_names_rot = 45,
          show_heatmap_legend = F,
          border=F,
          heatmap_width = unit(8+6,"cm"),
          heatmap_height = unit(18,"cm"),
          raster_quality = 10
  ) -> hm_vir_phold_fun

png(filename = 
      paste0(
      "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/final/figures/heatmap_vir_phold_", cur_fun, ".png"),
    width = 22, height = 20, res = 600, units = "cm", bg = "transparent")
draw(hm_vir_phold_fun)
dev.off()
}


```

# Phold stats
## General
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show'}
# Proportions of CDS assigned to phold categories
n_cds=sum(phold_description[Description == "CDS",Count])
phold_description %>%
  group_by(Description) %>%
  summarise(
    sum=sum(Count),
    prop=round((sum/n_cds)*100,2)     
            ) %>% arrange(desc(sum))


# Distribution of CDS
phold_annot_summary1$CDS %>% quantile
phold_annot_summary1$CDS %>% mean
phold_annot_summary1$CDS %>% sd
phold_annot_summary1$CDS %>% hist

# Distribution of annotated proportion
phold_annot_summary1$annot_prop %>% quantile
phold_annot_summary1$annot_prop %>% mean
phold_annot_summary1$annot_prop %>% sd
phold_annot_summary1$annot_pro %>% hist


# Distribution of categories across vOTUs
phold_description %>%
  group_by(Description) %>%
  summarise(
    n=n(),
    sum=sum(Count>0),
    prop=round((sum/n)*100, 2)
            ) %>% arrange(desc(prop))

# Distribution of products across vOTUs
phold_annot2 %>%
  group_by(product) %>%
  summarise(
    n=n(),
    prop=round((n/nrow(phold_annot2))*100, 2)
            ) %>% arrange(desc(prop)) %>% print(n = 50)

```

## Structural components
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show'}
phold_annot2 %>% 
  filter(func == "Structural components") -> struc

struc %>%
  count(product) %>% 
  mutate(percentile_group = ntile(n, 10)) %>%
  arrange(desc(n)) %>% view

phold_annot_summary1 %>%
  summarise(
    n=n(),
    sum = sum(`Structural components`>0),
    prop=round((sum/n)*100, 2)
            ) %>% arrange(desc(prop))
```

## DNARNA
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show'}
phold_annot2 %>% 
  filter(func == "DNA, RNA and nucleotide metabolism") -> dnarna

dnarna %>%
  count(product) %>% 
  mutate(percentile_group = ntile(n, 10)) %>%
  arrange(desc(n)) %>% view
```

## Moron and auxiliary
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show'}
phold_annot2 %>% 
  filter(func == "Auxiliary genes and host takeover") -> aux

aux %>%
  count(product) %>% 
  mutate(percentile_group = ntile(n, 10)) %>%
  arrange(desc(n)) %>% view

# Expand VFDB annotations
aux %>% 
  dplyr::select(votu, cds_id, product, tophit_protein) %>%
  filter(product == "VFDB virulence factor protein") %>%
  dplyr::rename(prot_id=tophit_protein) %>% merge(vfdb_vocabulary, by="prot_id") -> aux_vfdb

aux_vfdb %>% 
  group_by(vf_category, description) %>% 
  count -> aux_vfdb_count

# Expand DefenceFinder annotations
aux %>% 
  dplyr::select(votu, cds_id, product, tophit_protein) %>%
  filter(product == "DefenseFinder protein") %>%
  mutate(id = parse_number(str_extract(tophit_protein, "_[:digit:]+$"))) %>% 
  merge(defencefinder_vocabulary, by="id") -> aux_defencefinder


aux_defencefinder %>% 
  group_by(gene_name) %>% 
  count -> aux_defencefinder_count
```
## Other
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show'}
phold_annot2 %>% 
  filter(func == "Other") -> other

other %>%
  count(product) %>% 
  mutate(percentile_group = ntile(n, 10)) %>%
  arrange(desc(n)) %>% view
```

## Supplementary 5: Virus gene annotations
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show'}
require(openxlsx)
wb <- createWorkbook()
phold_annot_summary1
addWorksheet(wb, "Table 1")
writeData(wb, "Table 1", phold_annot_summary1)
"Gene Counts by vOTUs and Functional category"

funcs<-unique(phold_annot2$func)[-1]
for (i in 1:length(funcs)){
phold_annot2 %>%
  filter(func == funcs[i]) %>%
  group_by(func, product) %>% 
  count %>% arrange(desc(n)) %>% 
  dplyr::select(product, n) %>%
  dplyr::rename(
    "Function" = func,
    "Gene annotation" = product,
    "Gene count" = n
                ) -> gene_counts
  
  phold_annot2 %>%
  filter(func == funcs[i]) %>%
  group_by(func, product) %>% 
  summarise(n=sum(length(unique(votu)))) %>%   
    dplyr::rename(
      "Function" = func,
      "Gene annotation" = product,
      "Number of vOTU" = n) -> votu_counts
  
  sel<-merge(gene_counts, votu_counts, by = c("Gene annotation", "Function")) %>%
    relocate("Function", .before = "Gene annotation")
  
  addWorksheet(wb, paste0("Table ", 1+i))
  writeData(wb, paste0("Table ", 1+i), sel)
}

saveWorkbook(wb,
             "/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/final/supplementary/Supplementary File 8 - Virus gene annotation_v2.xlsx",
             overwrite = F)
```

# Utils
## Exctract gene sequences by name
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show'}
phold_annot2 %>% count(product) %>% arrange(desc(n)) %>% view
gene<-"VFDB virulence factor protein"
phold_annot2 %>% filter(product == gene) %>% 
  mutate(seq_id=paste0(votu, ":", cds_id))-> phold_selected_gene

#Select vOTUs
phold_selected_gene %>% pull(seq_id) %>% unique -> seq_ids

# Load auxiliary genes
Biostrings::readAAStringSet(
  filepath = "/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/vir/pipeline3_29aug24/pharokka_phold/phold/phold_moron_other_aa.fasta",format = "fasta") -> aux.fna

names(aux.fna)
aux.fna[names(aux.fna) %in% seq_ids] -> selected_votus_fna

selected_votus_fna[1] %>% as.character

```

## Select votus by auxiliary genes
```{r, echo = TRUE, warning=FALSE, message=FALSE, results='show'}
phold_annot2 %>% count(product) %>% arrange(desc(n)) %>% view
gene<-"VFDB virulence factor protein"
phold_annot2 %>% filter(product == gene) -> phold_selected_gene

#Select vOTUs
phold_selected_gene %>% pull(votu) %>% unique -> votus

#Pick the sequences of the viruses
Biostrings::readDNAStringSet(
  filepath = "/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/vir/pipeline3_29aug24/OTU/all_OTUs_no_doubleBB.fasta",format = "fasta") -> votus_fna
votus_fna[names(votus_fna) %in% votus] -> selected_votus_fna

dir.create("/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/vir/pipeline3_29aug24/OTU_morons")

Biostrings::writeXStringSet(
  selected_votus_fna,
  filepath = "/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/vir/pipeline3_29aug24/OTU_morons/votus_phosphoadenosine_phosphosulfate_reductase.fa",format = "fasta")

```

#-------------------------------------------------------------------------------
# ~~~ MAG - vOTUs ~~~
# Summarise iphop results
## Number of vOTUs associated to a host tax
```{r, echo = TRUE, warning=FALSE, message=FALSE, results='show'}
iphop_mags<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/vir/pipeline3_29aug24/iphop_mags_vOTUs/Host_prediction_to_genome_m90.csv") %>% rename(votu=Virus)

iphop_mags$`Host taxonomy`%>% str_split(";", simplify = T) %>%
  as.data.frame -> iphop_mags_tax
names(iphop_mags_tax) = c("Domain", "Phylum", "Class", "Order", "Family", "Genus")
cbind(iphop_mags, iphop_mags_tax[,-8]) -> iphop_mags2

# Remove vOTUs not in vir_analysis (unmapped)
iphop_mags2[votu %in% vir_analysis$votu] -> iphop_mags3

iphop_mags3 %>% group_by(Phylum, votu) %>% count -> iphop_phy
iphop_mags3 %>% group_by(Phylum, Class, votu) %>% count -> iphop_class
iphop_mags3 %>% group_by(Phylum, Family, votu) %>% count -> iphop_fam

iphop_fam %>%
  group_by(Phylum, Family) %>%
  count -> iphop_fam_count

iphop_class %>%
  mutate(Phylum = str_remove_all(Phylum, "p__"),
         Class = str_remove_all(Class, "c__")) -> iphop_phy_class_count

```
## Summarise number of vOTUs associated to a host tax
```{r, echo = TRUE, warning=FALSE, message=FALSE, results='show'}
iphop_phy_class_count %>%
  ungroup %>%
  mutate(prop=100*(n/sum(n)))

# Count the number of unique Phyla  
unique(iphop_phy_class_count$Phylum) %>% length

# Count the total number of vOTU  
iphop_phy_class_count$n %>% sum

# Info about the viruses
vir_analysis[votu %in% iphop_phy$votu] -> votu_hits

votu_hits %>%
  count(Class) %>% 
  ungroup %>% mutate(prop=n/sum(n)*100)

votu_hits %>%
  summarise(sum_vir = sum(str_detect((votu),"vir_")),
            sum_metag = sum(str_detect((votu),"metag_"))
            )

# Count the unique number of vOTU  
iphop_phy$votu %>% unique %>% length

# phyla predicted but not matched
setdiff(
  unique(iphop_phy_class_count$Phylum),
  str_remove(mag_analysis$Phylum, "p__")
  ) -> non_matched
length(non_matched)

# phyla predicted but and matched
intersect(
  str_remove(mag_analysis$Phylum, "p__"),
  unique(iphop_phy_class_count$Phylum)) -> matched
length(matched)

# Propotion of matched
100*(length(matched)/ (length(non_matched) + length(matched)))

```
#
# PLOT: Barplots of predicted Classes and Phyla
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show'}
#load colors
iphop_phy_class_count %>% filter(n>1) %>%
  group_by(Phylum, Class) %>%
  summarise(sum=sum(n)) %>%
  ggplot(aes(y=reorder(Class, sum), x=sum, fill = Phylum)) + 
  geom_bar(stat="identity", color="black") +
  scale_fill_manual(values = tree_colors, na.value = NA) +
  ylab("Predicted host Phylum") + xlab("n° of vOTUs") + 
  theme_bw() +
  theme(legend.position = "none",
        axis.text.y = element_text(size = 15),
        axis.text.x = element_text(size = 15),
        axis.title.y = element_text(size = 0),
        axis.title.x = element_text(size = 15),
        
        ) -> gg_iphop_phy_count;gg_iphop_phy_count

ggsave(plot = gg_iphop_phy_count,
       filename =  paste0("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/final/figures/iphop_phy_class_count_14apr25.png"),
         width =5.5, height = 6, units = "in", scale = 1)

mag_analysis %>%
  mutate(
    iphop_fam=
    case_when(
    Family %in% iphop_fam_count$Family ~ T, .default = F))
```

# Extract iphop associations
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show'}
vir_analysis %>%
  dplyr::select(votu, mag_interaction) %>%
  mutate(OTU=str_split(mag_interaction, ";")) %>% 
  unnest(cols = c(OTU)) %>% 
  filter(grepl("SemiBin", OTU)) %>%
  dplyr::select(votu, OTU) %>% rename(vir=votu, mag=OTU) -> mag_votu
```

# Association between MAG taxonomy and vOTU tax prediction
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show'}
# Correlation between number of vOTU in predicted MAG Class and n.o MAG in Class 
mag_analysis_nice[,.(`MAG ID`, Class)] %>%
  mutate(Class = str_remove(Class, "c__")) %>%
  group_by(Class) %>%
  summarise(sum_class = n()) %>%
  merge(
iphop_phy_class_count %>%
  group_by(Class) %>% 
  summarise(sum_votu =sum(n)), by = "Class") -> votu_pred_count_gem_abund

# Check distribution
shapiro.test(votu_pred_count_gem_abund$sum_class)
shapiro.test(votu_pred_count_gem_abund$sum_votu)

cor.test(votu_pred_count_gem_abund$sum_class, 
         votu_pred_count_gem_abund$sum_votu,
         method = "spearman"
        )

cor.test(votu_pred_count_gem_abund$sum_class, 
         votu_pred_count_gem_abund$sum_votu,
         method = "kendall"
        )

plot(votu_pred_count_gem_abund$sum_class, 
         votu_pred_count_gem_abund$sum_votu)

# Correlation between average abundance of vOTU from with Class is predicted and MAG Class abundance
mag_analysis_nice[,.(Class, `Average Relative Abundance`)] %>%
  mutate(Class = str_remove(Class, "c__")) %>%
  group_by(Class) %>%
  summarise(sum_mag_class = mean(`Average Relative Abundance`)) %>%
  merge(
iphop_phy_class_count %>%
  ungroup %>%
  select(c(votu, Class)) %>%
  unique %>%
  merge(vir_analysis[,.(votu, mean_vir_all)], by="votu") %>%
  group_by(Class) %>%
  summarise(sum_votu_class=mean(mean_vir_all))
  , by = "Class") -> 
  class_pred_votu_mag_abundance

# Double check with data shuffling
#class_pred_votu_mag_abundance %<>%
#  mutate(sum_mag_class=sample(sum_mag_class))
  
# Check distribution
shapiro.test(class_pred_votu_mag_abundance$sum_mag_class)
shapiro.test(class_pred_votu_mag_abundance$sum_votu_class)

cor.test(class_pred_votu_mag_abundance$sum_mag_class, 
         class_pred_votu_mag_abundance$sum_votu_class,
         method = "pearson"
        )

plot(class_pred_votu_mag_abundance$sum_mag_class, 
         class_pred_votu_mag_abundance$sum_votu_class)

#Check vOTU-GEM correlation
iphop_mags[grepl("SemiBin", `Host genome`),] -> votu_gem_association

load("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/data/analysis/mag_MPSE_norm_mags99c70c10a99.RData")
mag_MPSE_norm@assays@data@listData[["RelhellingerBySample"]] -> GEMs_bysample
colnames(GEMs_bysample)

load("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/data/analysis/vir_MPSE_norm_04Oct24.RData")
vir_MPSE_norm@assays@data@listData[["RelhellingerBySample"]] -> vOTUs_bysample

colnames(GEMs_bysample) %<>% toupper() %>% str_replace("ART", "CON") %>% str_remove("\\.1")
colnames(vOTUs_bysample) %<>% toupper() %>% str_replace("ART", "CON") %>% str_remove("\\.1")

colnames(GEMs_bysample)
colnames(vOTUs_bysample) 

intersect(colnames(vOTUs_bysample), colnames(GEMs_bysample)) -> shared_columns

rbind(
vOTUs_bysample[,shared_columns],
GEMs_bysample[,shared_columns]) %>% scale -> mat

mat[c(votu_gem_association$votu,votu_gem_association$`Host genome`),] %>% 
  t %>% cor -> cor_mat

cor_mat[
  grepl("vir|metag", rownames(cor_mat)),
  grepl("SemiBin", colnames(cor_mat))
  ] %>% reshape2::melt() %>% filter(abs(value)>0.95) -> cor_mat_f_l

plot(mat["art_SemiBin_1041",],
     mat["metag_nat_2436508",]
     )

plot(mat["art_SemiBin_244",],
     mat["metag_art_1946551_full",]
     )

fread(file = "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/data/analysis/vir_analysis_aux_7apr25.csv")-> vir_analysis_aux
vir_analysis_aux[votu %in% cor_mart_f_l$Var1] %>% view()
mag_analysis[OTU %in% cor_mart_f_l$Var2] %>% view()


```
# ----------------------------------------------------------------------------------
# AMGs + tax pred
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show'}
vir_analysis %>% filter(host_taxonomy != "") -> vir_analysis_host

phold_annot_summary %>%
  dplyr::select(-c(annot_prop, CDS, unk)) %>%
  column_to_rownames("votu") %>% as.matrix()-> phold_annot_summary_mat
phold_annot_summary_mat[phold_annot_summary_mat>1]=1

# Create a row annotation object
row_annot <- 
  rowAnnotation(
  Phylum = vir_analysis_host$Phylum,
  "Host taxonomy" = vir_analysis_host$host_taxonomy %>% str_extract("p__[:alpha:]+")
)

phold_annot_summary_mat[vir_analysis_host$votu,] %>%
  Heatmap(col=c("white", "red4"),
          show_row_names = FALSE,
          show_row_dend = FALSE,
          right_annotation = row_annot,
          cluster_rows = T,
          cluster_columns = F, row_split = vir_analysis_host$Phylum
          )

phold_annot %>% 
  filter(votu %in% vir_analysis_host$votu & grepl("other",`function`))
```

## Annotations of MAG - vOTU interactions
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show'}
vir_analysis2 %>% filter(grepl("art|nat", mag_interaction)) -> vir_an_mags
phold_description %>% filter(Contig %in% vir_an_mags$votu) %>%
  pivot_wider(names_from = Description, values_from = Count) %>% view
phold_annot %>% filter(contig_id %in% vir_an_mags$votu) %>% 
  group_by('function', product) %>% tally %>% view
```
# ----------------------------------------------------------------------------------
# Contig Taxonomy

## Relative abundance by superkingdom
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show'}
load("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaG/coassemby_07feb24/Rdata/TPM_Viruses_Eukaryota_Bacteria_Archaea_MPSE.Rdata")
se_hell<-TPM_Viruses_Eukaryota_Bacteria_Archaea

#Hellinger normalization
se_hell_norm <- 
  mp_decostand(se_hell, method = "hellinger", .abundance = Abundance)

#Calculate abundance ----
se_hell_norm %<>%
  mp_cal_abundance(
    .abundance = hellinger,
    force = T,
    action = "add"
  )
se_hell_norm %>%
  as.data.table %>%
  select(OTU, Sample, superkingdom, RelhellingerBySample) -> test

test %>%
  group_by(superkingdom, Sample) %>%
  summarise(sum=sum(RelhellingerBySample)) %>%
  ungroup() %>%
  group_by(superkingdom) %>%
  summarise(mean=mean(sum)) %>% view

se_hell_norm %>%
  group_by(superkindom) %>%
  summarise(sum=sum(RelhellingerBySample))

```

## Barplots
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show'}
#Import MPSE data ----
load("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaG/coassemby_07feb24/Rdata/TPM_Viruses_Eukaryota_Bacteria_Archaea_MPSE.Rdata")
load("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaG/coassemby_07feb24/Rdata/TPM_Eukaryota_MPSE.Rdata")
load("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaG/coassemby_07feb24/Rdata/TPM_Archaea_MPSE.Rdata")
load("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaG/coassemby_07feb24/Rdata/TPM_Bacteria_MPSE.Rdata")
load("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaG/coassemby_07feb24/Rdata/TPM_Eukaryota_MPSE.Rdata")

se_hell<-TPM_Eukaryota
se_hell<-TPM_Archaea
se_hell<-TPM_Bacteria
se_hell<-TPM_Viruses_Eukaryota_Bacteria_Archaea
se_hell$group %<>% toupper %>% str_replace("ART", "RES") %>% factor(levels = c("RES", "NAT"))
rownames(se_hell@colData) %<>% str_replace("Art", "Res")

#Hellinger normalization
se_hell_norm <- 
  mp_decostand(se_hell, method = "hellinger", .abundance = Abundance)

#Calculate abundance ----
se_hell_norm %<>%
  mp_cal_abundance(
    .abundance = hellinger,
    force = T,
    action = "add"
  )

#Plot abundance
se_hell_norm %>%
  mp_plot_abundance(
    .abundance=RelhellingerBySample,
    .group=group, 
    taxa.class = phylum, 
    topn = 24,
    relative = TRUE,
    force = T
  ) -> barplots;

barplots +
  theme(axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        legend.text = element_text(size = 6)
        ) -> b1

ggsave(
  plot = b1,
  filename = "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/final/supplementary/barplots_euk.png",
  device = "png",
  dpi = 600,
  units = "cm",
  height = 15,
  width = 30
  )

```

```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show'}


```