---
title: "bosc23_metab_fba"
author: "Riccardo Frizzo"
date: "2026-26-06"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

#Init
## Packages
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
require(cobrar); require(tidyverse); require(data.table); require(foreach); library(parallel); require(magrittr); require(ggpubr); require(vegan); require(caret);library(ggrepel);require(DT);require(rstatix); require(patchwork);require(scales)
```

## Variables
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
qprobs=seq(0,1,0.01)
load("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/data/tax/tree_colors_14apr25.RData")
tree_colors<-c(tree_colors, "gray10")
names(tree_colors)[15]="Thermoproteota" #Add the archaea :)
```
#---     
# Metabolites vocabulary
## Prepare
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=FALSE}
require(rcdk)
key_metabolites<-fread("/home/riccardo/rstudio-server/storage16TB/riccardo/tools/gapseq_feb25/gapseq/dat/seed_metabolites_edited.tsv") %>%
  mutate(has_C=str_detect(formula, "C[:digit:]|C[:upper:]"),
         has_N=str_detect(formula, "N"),
         has_S=str_detect(formula, "S"), 
         has_H=str_detect(formula, "H"),
         has_O=str_detect(formula, "O"),
         ) 
## Define chemical classes
cl<-makeCluster(25)
doParallel::registerDoParallel(cl)
foreach(met_row=1:nrow(key_metabolites[,]), .errorhandling = "remove")%dopar%{
  print(paste0(met_row, "/", nrow(key_metabolites)))
  mols <- rcdk::parse.smiles(as.character(key_metabolites[met_row,"smiles"]))
  if(!is.null(unlist(mols))){
    tibble::add_column(key_metabolites[met_row,],
               is_COH= rcdk::matches(query ="([CX4][OX2H])", mols),
               is_COOH= rcdk::matches(query ="C(=O)O", mols),
               carbonyl = rcdk::matches(query ="[CX3]=O", mols)
               #ketone =rcdk::matches(query = "[#6][CX3](=O)[#6]", mols),
               #aldehyde = rcdk::matches(query ="[CX3H1](=O)[#6]", mols),
               #carboxylic_acid = rcdk::matches(query ="[CX3](=O)[OX2H1]", mols),
               #carboxylate = rcdk::matches(query ="[CX3](=O)[O-]", mols)
               )
                    }else{
                      tibble::add_column(key_metabolites[met_row,],
                        is_COH=NA,
                        is_COOH=NA,
                        carbonyl = NA,
                        #ketone = NA,
                        #aldehyde = NA,
                        #carboxylic_acid = NA,
                        #carboxylate = NA
                                  
                        )
                    }
} %>% rbindlist -> key_metabolites
stopCluster(cl)

fwrite(key_metabolites, "/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/key_metabolites.tsv")
```
## Load
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
# Import classes
fread("/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/key_metabolites.tsv") %>%
  mutate(has_H=str_detect(formula, "H"),
         has_O=str_detect(formula, "O"),
         has_P=str_detect(formula, "P"),
        simple_category=case_when(
            has_S ~ "Sulfur compounds",
            has_C & !has_S & !has_N & carbonyl & as.numeric(mass)<100 ~ "Carbonyl compounds" ,
            has_C & !has_S & !has_N & is_COH & as.numeric(mass)>100 ~ "Carbohydrates" ,
            has_C & !has_S & has_N & carbonyl ~ "N-Carbonyl compounds",
            has_C & !has_S & has_N ~ "Nitrogen compounds",
            has_C & !has_S & has_P ~ "Phosporous compounds",
            !has_C | (has_C & !has_H) ~ "Inorganic compounds", .default = NA)) -> key_metabolites

carbonyl_cpp_str<-"Fumarate|Succinate|2-Oxoglutarate|Citrate|Itaconate|4-Hydroxybenzoate|Acetoacetate|Chorismate|Malate"
aa_cpp_str<-"Alanine|Glycine|Isoleucine|Leucine|Proline|Valine|Phenylalanine|Tryptophan|Aspartate|Glutamate|Arginine|Histidine|Lysine|Serine|Threonine|Cysteine|Methionine|Asparagine|Glutamine|Tyrosine|Ala-Ala"

key_metabolites %<>%
  mutate(simple_category=
           case_when(grepl(carbonyl_cpp_str, name) ~ "Carbonyl compounds",
                     grepl("Methanol|Ethanol|Glycerol|1,2-Propanediol", name) ~ "Alcohols",
                     grepl("Xylan|Inulin|starch", name) ~ "Polysaccharides",
                     grepl("Carbon monoxide", name) ~ "Inorganic compounds",
                     grepl("Dodecanoic acid|lipid", name) ~ "Fatty acids",
                     simple_category=="Carbohydrates" & grepl("ol", name) ~ "Sugar Alcohols",
                     grepl(aa_cpp_str, name) ~ "Amino acids",
                     grepl("chloroethene", name) ~ "Chlorocarbons",
                     grepl("Zn2+|Fe3+|Fe2+|Mn|Mg|K+|Cu2+|Zn2+|Ca2+|Cobalt", name) ~ "Metals",
                     grepl("Melibiose", name) ~ "Carbohydrates",
                     .default = simple_category
                     ))

fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/data/analysis/fba_summary_ox_curated_30dec24.tsv") %>%
  dplyr::select(id, subclass)-> hmdb_metabolite_classes
```

# Reactions vocabulary
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
#fread("/home/riccardo/rstudio-server/storage16TB/riccardo/tools/gapseq/dat/seed_reactions_corrected.tsv")-> key_seed_reactions
fread("/home/riccardo/rstudio-server/storage16TB/riccardo/tools/gapseq_feb25/gapseq/dat/seed_reactions_corrected.tsv")-> key_seed_reactions

# Watch out!! A single reaction name, matches several IDs
(count(key_seed_reactions, name)$n>1) %>% table
(count(key_seed_reactions, name)$n>5) %>% table

# IDs are unique
(count(key_seed_reactions, id)$n>1) %>% table

# Simplify reactions DB
names(key_seed_reactions)

key_seed_reactions %>% 
  dplyr::select(id, name, compound_ids,
                 definition, equation 
                ) %>%
  dplyr::rename(react_id2=id,
                react_name=name) -> key_reactions_simple
```

# Pathways vocabulary
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
# Import pathway hierarchy
gapseq_pwy_hierarchy<-
  rbind(fread("/home/riccardo/rstudio-server/storage16TB/riccardo/tools/gapseq_feb25/gapseq/dat/meta_pwy.tbl"),
        fread("/home/riccardo/rstudio-server/storage16TB/riccardo/tools/gapseq_feb25/gapseq/dat/custom_pwy.tbl"))
gapseq_pwy_hierarchy <- gapseq_pwy_hierarchy[!duplicated(id)]
gapseq_pwy_hierarchy %<>%
  mutate(sbs=str_remove_all(id, "\\|")) %>% 
  dplyr::select(sbs, name, hierarchy) %>% 
  dplyr::rename(Name=name) %>%
  unique

# read and prepare pathway DB
pwyDB <- rbind(fread("/home/riccardo/rstudio-server/storage16TB/riccardo/tools/gapseq_feb25/gapseq/dat/meta_pwy.tbl"),
               fread("/home/riccardo/rstudio-server/storage16TB/riccardo/tools/gapseq_feb25/gapseq/dat/custom_pwy.tbl"))

pwyDB <- pwyDB[!duplicated(id)]
pwyDB <- pwyDB[, .(id, name, spont, reaId)]
pwyDB[, reaNr := length(unlist(strsplit(reaId, ","))), by = id] # Count n° of reactions
pwyDB[, spontNr := length(unlist(strsplit(spont, ","))), by = id] # Count n° of spont reactions
pwyDB %<>% mutate(id=str_remove_all(id, "\\|"))

pwyDB %>% dplyr::select(id, name) %>% 
  dplyr::rename(sbs = id) %>% mutate(sbs=str_remove_all(sbs, "\\|")) -> key_pwy
rm(pwyDB)

gapseq_pwy_hierarchy$hierarchy %>%
  str_extract("\\|Pathways\\|[:graph:]+") %>%
  str_remove_all("\\|") %>%
  str_split(pattern = ",", simplify = T) %>% data.table -> sbs_hier
names(sbs_hier)=paste0("h", seq(1,8))

cbind(gapseq_pwy_hierarchy, sbs_hier) %>%
  dplyr::rename(sbs_ids=sbs) %>%
  select(sbs_ids, Name, h2, h3,h4,h5) -> sbs_hier1

```

# Import MAG analysis and prepare GEM analysis
```{r, echo = T, warning=FALSE, message=FALSE, results='show', eval=TRUE}
# Import MAG analysis for taxonomy and other parameters
mag_analysis<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/data/analysis/mag_analysis_co_de_24sep24.tsv") %>%
  rowwise %>%
  mutate(mean_mag_abund=
           mean(mean_mag_abund_Art, mean_mag_abund_Nat), .before = "mean_mag_abund_Art") %>%
  data.table %>%
  mutate(Phylum=str_remove(Phylum, "p__"),
         Order=str_remove(Order, "o__")
         )

mag_analysis$Order[mag_analysis$Order==""]="Unknown"
mag_analysis %>% pull(OTU)->mag_ids

# Make gem_analysis
gem_analysis <- 
  mag_analysis[,c("OTU", "Phylum", "Class", "Order", "Family","mean_mag_abund", "virus_presence")]
```

#----
# Prepare metaB data
## Import metaB IDs and assign a modelSEED id
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
# Import assigned metabolites 
fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaB/LCGC_metabolite_ids_subclasses_curated_classified_hmdb_24dic24.csv") %>%
  mutate(Subclass = str_remove(Subclass, "\t"))-> metab_ids_curated

# Use KEGG IDs and HMDB IDs to find the cpd id in gapseq (modelSEED) db
foreach(i = 1:nrow(metab_ids_curated))%do%{
  
  metab_ids_curated[i,]$HMDB->hmdbid
  metab_ids_curated[i,]$KEGG->keggid
  metab_ids_curated[i,]$name_orig->name_orig
  metab_ids_curated[i,]$Subclass->subclass
  
  #Find modelSEED ids
  key_metabolites %>% 
    filter(
        grepl(keggid, keggID) | 
        grepl(hmdbid, hmdbID)) %>% 
    dplyr::select(id, name, keggID, hmdbID) %>%
    add_column(name_orig=name_orig,
               subclass=subclass)->hits
  if(nrow(hits)==0){
    data.frame(id=NA,
               name=NA,
               keggID=NA,
               hmdbID=NA,
               name_orig=name_orig,
               subclass=subclass) # Create empty line in case of nohit
    }else(hits)
  } %>% rbindlist -> metab_ids_modelseed

metab_ids_modelseed %>% 
  filter(is.na(id)) -> not_matched

# Manual curation of non matches (go to ModelSEED Compunds and look at synonyms)
not_matched[1,]$id="cpd24232";not_matched[2,]$name="eicosane";
not_matched[3,]$id="cpd35300";not_matched[3,]$name="gamma-Glu-Leu";
not_matched[8,]$id="cpd33424";not_matched[8,]$name="Glu-Thr";
not_matched[9,]$id="cpd25630";not_matched[9,]$name="N-acetyl-(L)-arginine";
not_matched[12,]$id="cpd33751";not_matched[12,]$name="2-octanone";
not_matched[13,]$id="cpd33751";not_matched[13,]$name="2-nonanone";
not_matched[15,]$id="cpd15618";not_matched[15,]$name="nonanal";
not_matched[16,]$id="cpd15644";not_matched[16,]$name="undecanal";
not_matched[18,]$id="cpd02498";not_matched[18,]$name="2,3-Dihydroxy-isovalerate";
not_matched[19,]$id="cpd32753";not_matched[19,]$name="3-methylinosine";
not_matched[20,]$id="cpd32380";not_matched[20,]$name="5-methylcytidine";
not_matched[21,]$id="cpd00093";not_matched[21,]$name="Anthranilate";
not_matched[22,]$id="cpd10271";not_matched[22,]$name="2-Ethyltoluene";

not_matched %>% 
  filter(!is.na(id)) -> rematched

metab_ids_modelseed %>% 
  filter(!is.na(id)) -> matched

rbind(matched, rematched) %>% add_column(orig="metab")-> metab_react_matched
not_matched->metab_react_not_matched


rm(rematched, matched, not_matched, hits, metab_ids_modelseed, metab_ids_curated)
```

## SANITY CHECK: Metabolite merging
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
## Check ambiguos matches
metab_react_matched %>%
  group_by(name_orig) %>% count %>% 
  filter(n>1) %>%
  pull(name_orig) -> ambiguos

metab_react_matched[name_orig %in% ambiguos] 

# Check duplicates
metab_react_matched %>%
  group_by(id) %>% count(id) %>%
  filter(n>1) %>% pull(id) -> dups
metab_react_matched[id %in% dups] %>% arrange(id)

# Check unmatched
metab_react_not_matched %>% 
  filter(is.na(id)) 

# Final dataset
metab_react_matched %>% data.table->metab_react_matched_cpds
rm(metab_react_matched, ambiguos, dups)

fwrite(metab_react_matched_cpds, "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/metab_react_matched_cpds.csv")
```

# Check predicted growth media
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=T}
list.files("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaG/coassemby_07feb24/mags_c70c10a99/gapseq", pattern = "-medium.csv", full.names = T) -> medias

foreach(i=medias)%do%{
    fread(i) %>%
      add_column(OTU = str_extract(i, "[:alpha:]+_SemiBin_[:digit:]+"),
                 .before = "compounds"
                 )
} %>% rbindlist() -> all_medias

all_medias %>%
  select(OTU, compounds, maxFlux) %>%
  pivot_wider(names_from = compounds, values_from = maxFlux, values_fill = 0) -> all_medias_w


```

# Prepare growth media
## Growth medium by GEM
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
min_medium<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaG/coassemby_07feb24/mags_c70c10a99/gapseq/mean_min_medium_6sep24.csv")
metabolites<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/metab_react_matched_cpds.csv")

metabolites %>%
  add_column(maxFlux=1) %>%
  dplyr::rename(compounds=id) %>%
  select(name,compounds,maxFlux) %>%
  unique %>%
  filter(name != "")-> metab_medium

metab_medium %>%
  select(compounds, maxFlux) %>%
  unique %>%
  pivot_wider(names_from = compounds, values_from = maxFlux, values_fill = 0) -> metab_medium_w

foreach(i = 1:nrow(all_medias_w), .combine = "rbind")%do%{
  all_medias_w[i,] %>% colnames -> min_med
  min_metab<-c(min_med, setdiff(metab_medium$compounds, min_med))
  cbind(all_medias_w[i,], metab_medium_w[1,])[,min_metab]
} -> min_metab_med_df_w

min_metab_med_df_w %>%
  reshape2::melt() %>% data.table->min_metab_med_df

colnames(min_metab_med_df) = c("OTU", "compounds", "MaxFlux")

min_metab_med_df %>%
  left_join(key_metabolites[,c("id", "name")] %>%
              dplyr::rename(compounds=id)) %>% 
  select(compounds, MaxFlux, name, OTU)->min_metab_med_df1

dir.create("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaG/coassemby_07feb24/mags_c70c10a99/gapseq/gapseq_min_metab_maxflux1_bygem_22apr26")

mags<-unique(min_metab_med_df1$OTU)
for(i in mags){
  min_metab_med_df1[OTU == i] %>%
    filter(MaxFlux != 0) %>%
    select(compounds, name, MaxFlux) %>%
    fwrite(
      paste0("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaG/coassemby_07feb24/mags_c70c10a99/gapseq_min_metab_maxflux1_bygem_22apr26/", i, "-metab-medium.csv")
    )
}


```

## Growth medium by GEM, gapseq medium
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=T}
min_medium

gc_mean<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/GC_means.csv")

gc_mean %>% mutate(
  mean_scaled=rescale(mean,to=c(0.1, 10))) -> gc_mean_scaled


lc_mean<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/LC_means.csv")
lc_mean %>% mutate(
  mean_scaled=rescale(mean,to=c(0.1,10))) -> lc_mean_scaled

metab_react_matched_cpds<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/metab_react_matched_cpds.csv")

metab_react_matched_cpds %>%
  left_join(
    rbind(lc_mean_scaled, gc_mean_scaled),
    by="name_orig") %>%
  filter(!is.na(mean_scaled)) %>%
  mutate(mean_scaled=round(mean_scaled,3))-> metab_react_matched_cpds_mean_scaled #Tryptophan betaine is removed 

#Example:
"gapseq medium -m ../gapseq/art_SemiBin_423-draft.RDS -p ../gapseq/art_SemiBin_423-all-Pathways.tbl -c cpd00007:0;cpd00159:10"

paste0(
  metab_react_matched_cpds_mean_scaled$id,
  ":",
  metab_react_matched_cpds_mean_scaled$mean_scaled, collapse = ";")

# Copy and paste this string to the gapseq medium command or to a bash env var

# Import the media for the suppplementary
medium_files<-list.files(path = "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaG/coassemby_07feb24/mags_c70c10a99/gapseq_min_metab_maxflux1_bygem_22apr26/", pattern = "\\d-medium.csv", full.names = T)

foreach(i=medium_files)%do%{
    fread(i) %>%
      add_column(MAG = str_extract(i, "[:alpha:]+_SemiBin_[:digit:]+"),
                 .before = "compounds"
                 )
} %>% rbindlist() -> all_medias
all_medias %>%
  mutate(name=case_when(name == "" ~ compounds, .default = name)) %>%
  select(MAG, name, maxFlux) %>% 
  group_by(MAG,name) %>%
  summarise(max=max(maxFlux)) %>%
  pivot_wider(names_from = name, values_from = max) %>% 
  fwrite("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/metab_medium.csv", sep = "\t", dec = ",")




```
##Alt: Prepare common growth medium
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
min_medium<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaG/coassemby_07feb24/mags_c70c10a99/gapseq/mean_min_medium_6sep24.csv")
metabolites<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/metab_react_matched_cpds.csv")

metabolites %>%
  add_column(maxFlux=10) %>%
  dplyr::rename(compounds=id) %>%
  select(compounds, name,  maxFlux) -> metab_medium

metab_medium[!compounds %in% min_medium$compounds]
min_medium[compounds %in% metab_medium$compounds]

rbind(min_medium[!compounds %in% metab_medium$compounds],
      metab_medium) %>% unique->min_metab_medium

fwrite(min_metab_medium, "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/min_metab_medium_v4_all_metab10.csv")
```
Gapfill

# Check transporters
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
list.files("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaG/coassemby_07feb24/mags_c70c10a99/gapseq", pattern = "-Transporter.tbl", full.names = T) -> trans

foreach(i=trans)%do%{
    fread(i) %>%
      add_column(OTU = str_extract(i, "[:alpha:]+_SemiBin_[:digit:]+"),
                 .before = "id"
                 )
} %>% rbindlist() %>% data.table -> all_trans

all_trans %>%
  left_join(gem_analysis[, c("OTU", "Order")]) -> all_trans1

all_trans1 %>%
  select(OTU, sub) %>%
  unique %>% add_column(pres=1) %>%
  pivot_wider(names_from = sub,
              values_from = pres, values_fill = 0)-> all_trans_w

all_trans_w %>% column_to_rownames("OTU") %>%
  colSums() %>% sort

all_trans1[grepl("xylan", sub), Order] %>% table %>% sort

```
#-

#Import models
```{r eval=FALSE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
#job::job({
mag_fba_metab <- function(m) {
  model<-cobrar::readSybilmod (
    paste0("/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaG/coassemby_07feb24/mags_c70c10a99/gapseq_min_metab_maxflux1_bygem_22apr26/", m, ".RDS"))
  return(model)
}

list.files("/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaG/coassemby_07feb24/mags_c70c10a99/gapseq_min_metab_maxflux1_bygem_22apr26/", pattern = ".RDS") %>%
  str_extract("[:alpha:]+_SemiBin_[:digit:]+") -> aval
mag_ids<-aval
length(mag_ids)


#mag_fba_min <- function(m) {
#  model<-cobrar::readSBMLmod(
#    paste0("/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaG/coassemby_07feb24/mags_c70c10a99/gapseq/mean_min_medium_models_nocore_reac/all_models_xml/", m, ".xml"))
#  return(model)
#}

#cl<-makeCluster(30)
#parLapply(cl, mag_ids, mag_fba_min)->models_min
#stopCluster(cl);gc()
#names(models_min)=mag_ids

cl<-makeCluster(25)
parLapply(cl, mag_ids, mag_fba_metab)->models_metab
stopCluster(cl);gc()
names(models_metab)=mag_ids
#})


#cl<-makeCluster(30)
#parLapply(cl, mag_ids, mag_fba_min)->models_min
#stopCluster(cl);gc()
#names(models_metab)=mag_ids
#})

# GEM check
#models_metab[["nat_SemiBin_286"]] %>% fba()
#models_metab[["nat_SemiBin_286"]] %>% fva()


```

#Prepare sbs react (do)
```{r eval=FALSE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
foreach(mod=models_metab)%do%{
  sbs<-mod@subSys %>% as.matrix
  rownames(sbs)=mod@react_id
  sbs %>% reshape2::melt() %>% as.data.frame
} %>% rbindlist() %>% filter(value) %>%
  select(Var1, Var2) %>% unique %>%
  dplyr::rename(react_id=Var1, sbs=Var2)-> sbs_react

sbs_react %>% mutate(sbs=str_remove(sbs, "subsys_") %>%
                       str_replace("_", "-")) -> sbs_react1

sbs_react1 %>%
  left_join(key_pwy) -> sbs_react

fwrite(sbs_react, "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/sbs_react_27apr26.csv")
```
#Prepare sbs react (load)
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
sbs_react<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/sbs_react_27apr26.csv") %>%
  group_by(react_id) %>%
  summarise(
    sbs_ids=paste(sbs, collapse = "|"),
    sbs_names=paste(name, collapse = "|")) %>%
  mutate(react_id2=str_remove_all(react_id, "_c0"))
```


#Prepare react-gene df (do)
```{r eval=FALSE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
foreach(mod=models_metab)%do%{
tibble(
  OTU=mod@mod_id,
  react_id=mod@react_id,
  react_name=mod@react_name,
  gene=mod@genes) %>% rowwise() %>% 
    mutate(gene=paste(unlist(gene), collapse = ";"))
} %>% rbindlist -> gem_react_genes
fwrite(gem_react_genes, "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/gem_react_genes_27apr26.csv")
```

#Prepare react-gene df (load)
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
gem_react_genes<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/gem_react_genes_27apr26.csv")
```
#FBA: Get default reactions and exchanges fluxes
## Simulate growth (do)

```{r eval=FALSE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
length(models_metab[])

# FBA METAB ----
cl<-makeCluster(25)
doParallel::registerDoParallel(cl)

foreach(mod = models_metab[], .verbose = T) %dopar% {
  #mod <- cobrar::changeBounds(model = mod, react = "EX_cpd00007_e0", lb = -0.01, ub=0)  #O2
  # Simulate aerobic growth
  cobrar::fba(mod) -> sol
  #tibble::add_column(cobrar::fva(mod), 
  #                   OTU=mod@mod_id,
  #                   sim_O2=T, .before = "react") -> fva
  
  data.table::data.table(
    OTU = mod@mod_id,
    #sim_O2=T,
    react_id=mod@react_id,
    fba_flux=sol@fluxes
    #fva_min=fva$min.flux,
    #fva_max=fva$max.flux,
    #red_costs=sol@redCosts
  )-> fba_a;fba_a

  
  # Simulate anaerobic growth by setting O2 lower reaction bound to 0
#  mod <- cobrar::changeBounds(model = mod, react = "EX_cpd00007_e0", lb = 0)  #O2
#  cobrar::fba(mod) -> sol
#  tibble::add_column(cobrar::fva(mod), 
#                     OTU=mod@mod_id,
#                     sim_O2=F, .before = "react") -> fva
#  
#  data.table::data.table(
#    OTU = mod@mod_id,
#    sim_O2=F,
#    react_id=mod@react_id,
#    fba_flux=sol@fluxes,
#    fva_min=fva$min.flux,
#    fva_max=fva$max.flux,
#    red_costs=sol@redCosts
#  )-> fba_an
  
#  rbind(fba_a, fba_an)
} -> fba_comp
fba_comp %>% rbindlist() -> fba_comp_metab

#fwrite(fba_comp_metab, "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/fba_fva_min_gems_20nov25.csv", nThread = 20)
fwrite(fba_comp_metab, "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/fba_fva_metab_maxflux_conc 0.1-10_22apr26.csv")
stopCluster(cl)

gc()
fba_comp_metab$OTU %>% unique
```
## Simulate growth (load)
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
fba_comp_metab<- fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/fba_fva_metab_maxflux_conc 0.1-10_22apr26.csv")

#fba_comp_metab<- fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/fba_fva_metab_gems_27oct25.csv")

#fba_comp_metab<- fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/pfba_fva_metab_gems_02feb25.csv")



```
#Import subsystem/reaction table (do)
```{r eval=FALSE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
foreach(mod=models_metab)%do%{
  sbs<-mod@subSys %>% as.matrix
  rownames(sbs)=mod@react_id
  sbs %>% reshape2::melt() %>% as.data.frame
} %>% rbindlist() %>% filter(value) %>%
  select(Var1, Var2) %>% unique %>%
  dplyr::rename(react_id=Var1, sbs=Var2)-> sbs_react

sbs_react %>% mutate(sbs=str_remove(sbs, "subsys_") %>%
                       str_replace("_", "-")) -> sbs_react1

fwrite(sbs_react1, "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/sbs_react_27apr26.csv")

```

#Import subsystem/reaction table (load)
```{r eval=T, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
sbs_react1<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/sbs_react_27apr26.csv")
```
#---
# Analysis I

## Number of MAGs by phylum
```{r eval=FALSE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
gem_analysis %>%
  group_by(Phylum) %>%
  summarise(abund=sum(mean_mag_abund),
            n=n()) %>%
  arrange(desc(abund)) -> phylum_nmags
```
## Number of MAGs by order
```{r eval=T, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
gem_analysis %>%
  group_by(Order) %>%
  summarise(abund=sum(mean_mag_abund),
            n=n()) %>%
  arrange(desc(abund)) -> order_nmags


order_nmags %>%
  filter(n>=5) %>% pull(Order) %>% length

order_nmags %>%
  filter(n>=5) %>% pull(n) %>% sum
```

## Number of MAGs by family
```{r eval=T, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
gem_analysis %>%
  group_by(Family) %>%
  summarise(abund=sum(mean_mag_abund),
            n=n()) %>%
  arrange(desc(abund)) -> family_nmags


family_nmags %>%
  filter(n>=5) %>% pull(Family) %>% length

family_nmags %>%
  filter(n>=5) %>% pull(n) %>% sum -> testable_fam
family_nmags %>%
  filter(n>=5) %>% pull(Family) -> testable_fam
```
## Define who is growing with O2 and without (for analysis)
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
# Check distribution of biomass vector
fba_comp_metab[react_id=="EX_cpd11416_c0"] %>% pull(fba_flux) %>% sort-> growth_vec

qprobs<-seq(0,1,0.01)
quantile(growth_vec, qprobs)%>% round(10)

# Look for a big drop (several orders of magnitude) in the quantile distribution
# 0.1 is a suitable threshold

fba_comp_metab[react_id=="EX_cpd11416_c0"] %>%
   filter(fba_flux>0.1) -> OTU_growth

gem_analysis[OTU_growth[,c("OTU", "fba_flux")],
             growth := i.fba_flux, on = "OTU"]


```

# Define who is growing fast and slow
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
growth_vec<-gem_analysis$growth%>% sort
growth_vec %>% quantile()->growth_quant;growth_quant

#Define threshold at 50%
growth_threshold<-growth_quant[["50%"]]


# Define growth categories
gem_analysis[, growth_category := fcase(growth > growth_threshold, "High",
                                        growth <= growth_threshold, "Low",
                                        default = NA)]
 
gem_analysis %>%
  group_by(Order) %>%
  summarise(growth_range = paste0(round(min(growth, na.rm = T), 2),"-", round(max(growth, na.rm = T), 2)),
            growth_median = round(median(growth, na.rm = T),2)
            )


#gem_analysis %>%
#  ggplot(aes(y=reorder(Order,growth), x=growth)) +
#  geom_boxplot()
```

#---
# Tidy raw datasets
## Exchange fluxes (products and subs)
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
# Tidy, add info
# Filter low fluxes 
fba_comp_metab %>%
  filter(
    grepl("EX_", react_id),
    !grepl("DM|bio|EX_cpd11416_c0", react_id)) %>%
  mutate(id=str_remove_all(react_id, "EX_|_c0|_e0")) %>%
  left_join(key_metabolites[,
                            c("id", "name", "formula",
                              "has_C", "has_N","has_S",
                              "is_COH", "is_COOH","mass","simple_category")]) %>%
  left_join(
    gem_analysis[, 
                 c("OTU","Phylum", "Class", "Order","Family",
                   "growth_category")]) %>%
  select(OTU, react_id,
         #sim_O2,
         fba_flux,
         #fva_min, fva_max, red_costs,
         id, name, formula,
         has_C, has_N, has_S, , is_COH, is_COOH, mass, simple_category,
         Phylum, Class, Order, Family,
         growth_category) -> fba_comp_metab1
fba_comp_metab1$OTU %>% unique
  #filter(abs(fba_flux)>1e-3)-> fba_comp_metab1 #HERE FLUX FILTERING 

# Fix metaB categories
fba_comp_metab1[,c("name", "simple_category")] %>% unique -> metab_cat
metab_cat$simple_category %>% unique
metab_cat[is.na(simple_category)]

carbonyl_cpp_str<-"Fumarate|Succinate|2-Oxoglutarate|Citrate|Itaconate|4-Hydroxybenzoate|Acetoacetate|Chorismate|Malate"
aa_cpp_str<-"Alanine|Glycine|Isoleucine|Leucine|Proline|Valine|Phenylalanine|Tryptophan|Aspartate|Glutamate|Arginine|Histidine|Lysine|Serine|Threonine|Cysteine|Methionine|Asparagine|Glutamine"
lipid<-"Decanoate|Myristic acid|tetradecenoate|Palmitate|tetradecenoate|Dodecanoic acid|lipid"

fba_comp_metab1 %<>%
  mutate(simple_category=
           case_when(grepl(carbonyl_cpp_str, name) ~ "Carbonyl compounds",
                     grepl("Methanol|Ethanol|Glycerol|1,2-Propanediol|1,3-Propanediol", name) ~ "Alcohols",
                     grepl("Xylan|Inulin|starch", name) ~ "Polysaccharides",
                     grepl("Carbon monoxide|H2S|Sulfate|Sulfite", name) ~ "Inorganic compounds",
                     grepl(lipid, name) ~ "Fatty acids",
                     simple_category=="Carbohydrates" & grepl("ol", name) ~ "Sugar Alcohols",
                     grepl(aa_cpp_str, name) ~ "Amino acids",
                     grepl("chloroethene", name) ~ "Chlorocarbons",
                     grepl("Zn2+|Fe3+|Fe2+|Mn|Mg|K+|Cu2+|Zn2+|Ca2+|Cobalt", name) ~ "Metals",
                     .default = simple_category
                     ))


# Separate aero and anaero

tidy_exhanges<-fba_comp_metab1
#tidy_exhanges<-
#  rbind(
#  fba_comp_metab1[sim_O2==T & anaerobic==F],
#  fba_comp_metab1[sim_O2==F & anaerobic==T])


## Filter substrates
tidy_exhanges %>%
  filter(
    fba_flux<0,
    ) %>% mutate(fba_flux=-(fba_flux)) -> fba_substrates
tidy_exhanges %>%
  filter(
    fba_flux>0,
    ) %>% mutate(fba_flux=(fba_flux)) -> fba_products

fba_substrates %>%
  filter(name == "Betaine" & fba_flux<=0.1) %>%
  ggplot(aes(x=name, y=fba_flux, color=Order)) +
  geom_boxplot() + geom_jitter() 
  


# Export products and substrates 
tidy_exhanges %>%
  group_by(name) %>%
  summarise(
    nmag_substrate = sum(fba_flux<0),
    nmag_product = sum(fba_flux>0),
  ) %>% fwrite("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/fba_substrates_products_nmags.csv")

```

## Test exchange flux frequencies
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
# Fisher test on substrates
fba_substrates[] %>%
  #filter(fba_flux>=0.1) %>%
  group_by(name, Order, simple_category) %>%
  summarise(n_use=sum(fba_flux>0)) %>%
  left_join(order_nmags) %>%
  #filter(Order %in% sel_orders) %>%
  mutate(prop=100*(n_use/n)) %>% data.table -> subs

foreach(i=unique(subs$name), .errorhandling = "remove")%do%{
  fba_substrates[name==i] -> dt
  dt %>%
    group_by(Order) %>%
    summarise(
              n_use=n()
              ) %>%
    left_join(order_nmags) %>%
    filter(n>=5)-> dt1

total_use <- sum(dt1$n_use)
total_n   <- sum(dt1$n)

dt1 %>%
  rowwise() %>%
  mutate(
    # 2x2 contingency table per order
    ft = list(fisher.test(matrix(c(
      n_use,                          # in order, with pathway
      n - n_use,                      # in order, without pathway
      total_use - n_use,              # not in order, with pathway
      total_n - n - total_use + n_use # not in order, without pathway
    ), nrow = 2))),
    p_value  = ft$p.value,
    odds_ratio = ft$estimate,
    direction  = ifelse(n_use / n > total_use / total_n, "enriched", "depleted")
  ) %>%
  ungroup() %>%
  mutate(p_adj = p.adjust(p_value, method = "BH")) %>%
  select(Order, n_use, n, direction, odds_ratio, p_value, p_adj) %>%
  arrange(p_adj) %>% add_column(Name=i)
} %>% rbindlist -> sub_fish

sub_fish%>%
  filter(p_adj<=0.05 & direction == "enriched") -> sub_fish_sig


# Fisher test on products
fba_products[] %>%
  group_by(name, Order, simple_category) %>%
  summarise(n_use=sum(fba_flux>0)) %>%
  left_join(order_nmags) %>%
  #filter(Order %in% sel_orders) %>%
  mutate(prop=100*(n_use/n)) %>% data.table -> prods

foreach(i=unique(prods$name), .errorhandling = "remove")%do%{
  fba_products[name==i] -> dt
  dt %>%
    group_by(Order) %>%
    summarise(
              n_use=n()
              ) %>%
    left_join(order_nmags) %>%
    filter(n>=5)-> dt1

total_use <- sum(dt1$n_use)
total_n   <- sum(dt1$n)

dt1 %>%
  rowwise() %>%
  mutate(
    # 2x2 contingency table per order
    ft = list(fisher.test(matrix(c(
      n_use,                          # in order, with pathway
      n - n_use,                      # in order, without pathway
      total_use - n_use,              # not in order, with pathway
      total_n - n - total_use + n_use # not in order, without pathway
    ), nrow = 2))),
    p_value  = ft$p.value,
    odds_ratio = ft$estimate,
    direction  = ifelse(n_use / n > total_use / total_n, "enriched", "depleted")
  ) %>%
  ungroup() %>%
  mutate(p_adj = p.adjust(p_value, method = "BH")) %>%
  select(Order, n_use, n, direction, odds_ratio, p_value, p_adj) %>%
  arrange(p_adj) %>% add_column(Name=i)
} %>% rbindlist -> prod_fish

prod_fish%>%
  filter(p_adj<=0.05 & direction == "enriched") -> prod_fish_sig


# Results
prod_fish_sig
sub_fish_sig
```


## FBA exchanges PCA
```{r eval=F, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
c("Bacteroidales", "Flavobacteriales", "Cytophagales", "Desulfobulbales", "Desulfobacterales",
  "Rhodobacterales", "Pseudomonadales", "Chromatiales",
  "Rhizobiales", "Woeseiales"
             ) -> sel_orders


tidy_exhanges %>%
  filter(Order %in% sel_orders) %>%
  filter(fba_flux<0) %>%
  #filter(name %in% sub_fish_sig$Name) %>%
  mutate(fba_flux=abs(fba_flux)) %>%
  #filter(fba_flux>0.1) %>%
  select(OTU, name, fba_flux) %>%
  pivot_wider(names_from = name, values_from = fba_flux, values_fill = 0) %>% 
  column_to_rownames("OTU") %>% as.matrix() -> mat

order_vec<-gem_analysis[OTU %in% rownames(mat), Order]
names(order_vec)<-rownames(mat)
ord_cols<-hcl.colors(palette = "Zissou 1", n = length(unique(order_vec)))
names(ord_cols) = unique(order_vec)

#order_vec<-rownames(mat)
#ord_cols<-hcl.colors(palette = "Set3", n = length(order_vec))
#names(ord_cols) = (order_vec)

require(factoextra);require(FactoMineR)
pca_res <- PCA(mat,
           scale.unit = T,   # standardize variables
           graph = FALSE)

fviz_pca_ind(
  pca_res,
  axes = c(1, 2),
  geom = c("point"),
  repel = FALSE,
  habillage = "none",
  addEllipses = T,
  ellipse.type  = "confidence",    # or "convex" / "norm" / "t"

  col.ind       = order_vec[rownames(mat)],          # color by group
  palette       = ord_cols,
  fill.ind = "white",
  col.ind.sup = "blue",
  alpha.ind = 1,
  select.ind = list(name = NULL, cos2 = NULL, contrib = NULL)
)

fviz_pca_biplot(pca_res,
                # --- samples ---
                geom.ind      = "point",
                pointshape    = 19,
                col.ind       = order_vec,          # color by group
                palette       = ord_cols,
                addEllipses   = T,            # confidence ellipses
                ellipse.type  = "confidence",    # or "convex" / "norm" / "t"
                ellipse.level = 0.95,            # 95% CI
                label         = c("var"),           # label variables only,
                labelsize     = 3,  
                # --- variables (arrows) ---
                col.var       = "black",
                alpha.var =    0.1,
                repel         = TRUE,  
                title         = "PCA Biplot",
                select.var = list(name = NULL,
                                  cos2 = 50,
                                  contrib = NULL)

                )+
  theme_bw() +
  theme(axis.text = element_text(size=13))


# ── PCoA ──────────────────────────────────────────────────────────────────────
dist_matrix <- vegdist(scale((mat)), method = "euclidean", na.rm = T)
pcoa        <- ape::pcoa(dist_matrix)

# ── envfit arrows ─────────────────────────────────────────────────────────────
env <- envfit(pcoa$vectors, mat, permutations = 999)

# ── Variance explained ────────────────────────────────────────────────────────
var_exp <- round(pcoa$values$Relative_eig * 100, 1)

# ── plot with base vegan ───────────────────────────────────────────────────────
plot(pcoa$vectors,
     col = ord_cols[order_vec[rownames(mat)]],
     xlab = paste0("PCoA1 (", var_exp[1], "%)"),
     ylab = paste0("PCoA2 (", var_exp[2], "%)"),
     main = "PCoA", lwd = 2, pch=20)
     
plot(env, p.max = 0.05, col="black")       # only significant arrows

ordiellipse(pcoa$vectors[, 1:2],
            groups = order_vec[rownames(mat)],
            kind   = "se",                      # "se", "sd", or "ehull"
            conf   = 0.95,
            draw   = "polygon",
            col    = ord_cols,
            border = ord_cols,
            alpha  = 0.1,
            lwd    = 1.5) -> p

legend("topright",
       legend = names(ord_cols),
       col    = ord_cols,
       pt.bg  = ord_cols,
       pch    = 21,
       cex    = 0.9,
       bty    = "n")    


```

## Reaction fluxes
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
fba_comp_metab[grepl("rxn", react_id) &
               !grepl("DM|bio|EX_", react_id)] %>%
  mutate(react_id2=str_remove(react_id,"_c0|_e0")) %>%
   left_join(key_reactions_simple) %>%
  left_join(gem_analysis[,
                         c("OTU","growth_category")]) %>%
  left_join(gem_analysis[, c("OTU","Phylum", "Class", "Order")])->fba_reactions
fba_reactions$fba_flux %>% abs %>% quantile(seq(0,1,0.01)) %>%
  round(5)
# 1e-3 is a reasonable threshold 

fba_reactions[abs(fba_flux)>1e-3] %>%
  left_join(gem_analysis[, c("OTU","Phylum", "Class", "Order")]) -> fba_reactions_f
index_reactions<-fba_reactions_f[,c("react_name")] %>% unique
```

# 
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}

key_metabolites[grepl("h2s", name, ignore.case = T)]
fba_reactions[grepl("cpd00239", equation, ignore.case = T)] %>%
  pull(react_name) %>% table %>% sort



key_metabolites[grepl("methanethiol", name, ignore.case = T)]
fba_reactions[grepl("cpd00324", equation, ignore.case = T)] %>%
  pull(react_name) %>% table %>% sort


```

#----
# INTEGRATION WITH VIR
## Import vocabularies 
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
vfdb_vocabulary<-fread("~/rstudio-server/storage16TB/riccardo/tools/phold/vfdb_description_output.csv")
defencefinder_vocabulary<-fread("~/rstudio-server/storage16TB/riccardo/tools/phold/defensefinder_plddt_over_70_metadata.tsv")
card_vocabulary<-fread("~/rstudio-server/storage16TB/riccardo/tools/phold/card_plddt_over_70_metadata.tsv")
```

## Import and tidy data
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
vir_analysis<-fread("/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/data/analysis/vir_analysis_03jan25.csv")
mag_analysis<-fread("/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/data/analysis/mag_analsis_fba_sbs_16nov24.csv")
phold_annot<-fread("/home/riccardo/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/vir/pipeline3_29aug24/pharokka_phold/phold/phold_per_cds_predictions.tsv") %>%
  filter(contig_id %in% vir_analysis$votu)
names(phold_annot)[1]="votu"

phold_annot %>%
  dplyr::rename(func=`function`) -> phold_annot1

phold_annot1 %>%
  mutate(func= 
           case_when(
             func %in% c("head and packaging", "tail", "connector") ~ "Structural components",
             func == "moron, auxiliary metabolic gene and host takeover" ~ "Auxiliary genes and host takeover",
             func == "other" ~ "Other", .default = func)) -> phold_annot2


phold_annot2 %>% dplyr::select(cds_id, func, product) -> phold_annot_simple

blastp_out<-list.files(
  "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaG/coassemby_07feb24/mags_c70c10a99/blastp_mags_phold_morons_07jan25/",
  pattern = ".faa", full.names = T)

foreach(i = blastp_out[])%do%{
fread(i) %>%
    add_column(OTU=str_extract(i, "[:alpha:]+_SemiBin_[:digit:]+")) -> hits
  names(hits) = c("qseqid", "sseqid", "pident", "length", "mismatch", "gapopen", 
             "qstart", "qend", "sstart", "send", "evalue", "bitscore", 
             "qlen", "slen", "OTU")
  hits
  } %>% rbindlist -> blastp

columns <- c("qseqid", "sseqid", "pident", "length", "mismatch", "gapopen", 
             "qstart", "qend", "sstart", "send", "evalue", "bitscore", 
             "qlen", "slen")
react_sbs_metab_info<-fread(
   "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/final/supplementary/gem_reactions_17mar25.csv")
```

# Expand VFDB annotations
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
phold_annot2 %>% 
  filter(func == "Auxiliary genes and host takeover") -> aux
aux %>% 
  dplyr::select(votu, cds_id, product, tophit_protein) %>%
  filter(product == "VFDB virulence factor protein") %>%
  dplyr::rename(prot_id=tophit_protein) %>% merge(vfdb_vocabulary, by="prot_id") -> aux_vfdb

aux_vfdb %>% 
  group_by(description) %>% 
  count(sort = T) -> aux_vfdb_count
names(aux_vfdb_count) = c("Gene", "Count")
aux_vfdb_count %<>% add_column(Category = "VFDB", .before = "Gene")

```

## BLASTn: All votu cds vs MAGs proteins
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
list.files("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/vir/pipeline3_29aug24/OTU_morons2/blastn_aux_oth_mags/", full.names = T)->files_paths
list.files("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/vir/pipeline3_29aug24/OTU_morons2/blastn_aux_oth_mags/", full.names = F)->files_names

foreach(i=files_paths)%do%{
  fread(i)
    } %>% rbindlist->blastn
names(blastn) = c("cds_id", "sseqid", "pident", "length", "mismatch", "gapopen", 
             "qstart", "qend", "sstart", "send", "evalue", "bitscore", 
             "qlen", "slen")
blastn -> blastn_f

left_join(
  blastn_f,phold_annot_simple) %>%
  mutate(
  qperc_align=(length/qlen)*100,
  sperc_align=(length/slen)*100,
  ) %>%
  filter(pident>95 & qperc_align>95)-> blastn_all_votu_genes
blastn_all_votu_genes$product %>%
table %>% sort

blastn_all_votu_genes %>%
  group_by(func, product) %>% count %>% view
  
```
## Moron and auxiliary, and others
### Prepare
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval = FALSE}
# Select and summarise OTHER annot
phold_annot2 %>% 
  filter(func == "Auxiliary genes and host takeover") -> aux

aux %>%
  count(product) %>% 
  arrange(desc(n)) -> aux_count
names(aux_count) = c("Gene", "Count")
aux_count %<>% add_column(Category = "AUX", .before = "Gene")


# Expand VFDB annotations
aux %>% 
  dplyr::select(votu, cds_id, product, tophit_protein) %>%
  filter(product == "VFDB virulence factor protein") %>%
  dplyr::rename(prot_id=tophit_protein) %>% merge(vfdb_vocabulary, by="prot_id") -> aux_vfdb

aux_vfdb %>% 
  group_by(description) %>% 
  count(sort = T) -> aux_vfdb_count
names(aux_vfdb_count) = c("Gene", "Count")
aux_vfdb_count %<>% add_column(Category = "VFDB", .before = "Gene")


# Expand DefenceFinder annotations
aux %>% 
  dplyr::select(votu, cds_id, product, tophit_protein) %>%
  filter(product == "DefenseFinder protein") %>%
  mutate(id = parse_number(str_extract(tophit_protein, "_[:digit:]+$"))) %>% 
  merge(defencefinder_vocabulary, by="id") -> aux_defencefinder


aux_defencefinder %>% 
  group_by(gene_name) %>% 
  count(sort = T) -> aux_defencefinder_count
names(aux_defencefinder_count) = c("Gene", "Count")
aux_defencefinder_count %<>% add_column(Category = "DF", .before = "Gene")

# Expand CARD annotations
aux %>% 
  dplyr::select(votu, cds_id, product, tophit_protein) %>%
  filter(product == "CARD resistance protein") %>% 
  dplyr::rename("Protein Accession" = tophit_protein) %>%
  merge(card_vocabulary, by="Protein Accession") -> aux_card


# Select and summarise OTHER annot
phold_annot2 %>% 
  filter(func == "Other") -> oth

oth %>%
  count(product) %>% 
  arrange(desc(n)) -> oth_count
names(oth_count) = c("Gene", "Count")
oth_count %<>% add_column(Category = "OTH", .before = "Gene")


rbindlist(list(aux_count, aux_defencefinder_count, aux_vfdb_count, oth_count)) %>%
  arrange(desc(Count)) -> aux_oth_gene_count 
 
 fwrite(
    aux_oth_gene_count,
    file = "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/final/supplementary/vir_aux_other_counts_11mar25.csv")

```

## BLASTp: Morons and other vs MAGs proteins
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
# Check that all the MAGs have been processed
length(unique(blastp$OTU))

blastp %>%
  mutate(qseqid_gene = str_split(qseqid, "\\:", simplify = T)[,2]) %>%
  mutate(
  qperc_align=(length/qlen)*100,
  sperc_align=(length/slen)*100,
  )-> blastp1

blastp1[qperc_align>130]


phold_annot$`function` %>% table
unique(blastp1$qseqid) %>% length
unique(blastp1$sseqid) %>% length
mag_analysis$Total_Coding_Sequences %>% mean

blastp1 %>%
  dplyr::select(sseqid, OTU)%>%
  unique %>%
  group_by(OTU) %>% count()

# Select mag_analysis MAGs only
blastp1 %>%
  filter(OTU %in% mag_analysis$OTU) -> blastp2

#blastp2 %>% 
#  filter(pident>=75,
#         qperc_align>=95,
#         evalue<1e-10
#         )-> blastp_f

blastp2 %>% 
  filter(pident>=30,
         qperc_align>=70,
         evalue<1e-5
         )-> blastp_f

blastp_f %>%
  dplyr::select(OTU, qseqid, qseqid_gene,sseqid, qperc_align, sperc_align, evalue, length) %>%
  mutate(qseqid = str_split(qseqid, ":", simplify = T)[,1]) %>%
  dplyr::rename(cds_id = qseqid_gene, votu=qseqid, mag_gene=sseqid) %>%
  left_join(phold_annot_simple, by = "cds_id") %>% data.table -> blast_phold

blast_phold %>% 
  group_by(func, product) %>%
  summarise(n=n())-> blast_phold_summary

phold_annot2[votu %in%
blast_phold[grepl("S-adenosylmethionine decarboxylase", product), votu]] 

mag_analysis[OTU %in% blast_phold$OTU]
vir_analysis[votu %in% blast_phold$qseqid] 

blast_phold %>%
  group_by(func, product) %>%
  summarise(n_mags_hits=length(unique(OTU)),
            v_otus_hits=length(unique(votu))
            )  -> blastp_count

phold_annot2 %>%
  group_by(func,product) %>%
  summarise(n_votu=length(unique(votu))) %>%
  right_join(blastp_count, by=c("func","product")) %>% 
  arrange(desc(v_otus_hits)) %>%
  mutate(prop_mags=round(100*(n_mags_hits/311),2)) 
phold_annot2
```

# Subset CDS
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=F}
require(Biostrings)
readDNAStringSet("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/vir/pipeline3_29aug24/OTU_morons2/pharokka_phold/prodigal-gv.ffn")-> aux_cds.fa
names(aux_cds.fa) %>% str_extract(".*_[:digit:]+") -> cds_ids

selected_cds<-names(aux_cds.fa)[cds_ids%in%selected_aux[,cds_id]]
selected_sequences<-aux_cds.fa[selected_cds]
names(selected_sequences) %<>% 
  str_extract(".*_[:digit:]+") #%>% str_replace("\\|", "_")

names(selected_sequences)

writeXStringSet(
  selected_sequences,
  "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/vir/pipeline3_29aug24/contigs_versus_selected_AUX/selected_aux_cds.fa",format = "fasta")
```

# Check blastn hits on contigs
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=TRUE}
rbind(
fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/vir/pipeline3_29aug24/contigs_versus_selected_AUX/aux_cds_art_contigs.tsv"),
fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/vir/pipeline3_29aug24/contigs_versus_selected_AUX/aux_cds_nat_contigs.tsv"))-> blastn_aux_contigs

names(blastn_aux_contigs) = c("cds_id", "sseqid", "pident", "length", "mismatch", "gapopen", 
             "qstart", "qend", "sstart", "send", "evalue", "bitscore", 
             "qlen", "slen")

left_join(
  blastn_aux_contigs,phold_annot_simple) %>%
  mutate(
  qperc_align=(length/qlen)*100,
  sperc_align=(length/slen)*100
  ) %>%
  filter(pident>95 & qperc_align>95)-> blastn_all_votu_genes
```

# BlastP results
```{r, echo = FALSE, warning=FALSE, message=FALSE, results='show', eval=F}
blast_phold %>%
  left_join(aux_vfdb[,c("cds_id", "description")], by="cds_id") %>%
  mutate(
  product=case_when(cds_id %in% aux_vfdb$cds_id ~ description, .default=product)) -> blast_phold
  
blast_phold %>%
  filter(grepl(aux_oth_string, product, ignore.case=T)) %>%
  left_join(gem_analysis[,c("OTU", "Phylum", "Order")]) -> blast_phold_aux

blast_phold_aux %>% group_by(Order) %>% count(sort = T)

blast_phold_aux %>%
  group_by(product) %>% count %>% 
  fwrite("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/blasp_aux_product_annot.csv")

vir_analysis[votu %in% blast_phold_aux$votu, mean_vir_all] %>% sum
```
# Host prediction
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
iphop_mags<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/vir/pipeline3_29aug24/iphop_mags_vOTUs/Host_prediction_to_genome_m90.csv") %>% janitor::clean_names()

iphop_mags$host_taxonomy %>%
  str_split(';', simplify = T) %>%
  as.data.frame -> iphop_expanded_tax
names(iphop_expanded_tax) = c('Domain', 'Phylum', 'Class', 'Order', 'Family', 'Genus', 'Species')
iphop_expanded_tax$Order %>% table %>% sort

cbind(iphop_mags, iphop_expanded_tax) %>% data.table -> iphop_mags2

# How many virus per order (no mags)
iphop_mags2[!grepl("SemiBin", host_genome)] %>%
  group_by(Order) %>%
  summarise(n_votu=length(unique(virus))) %>%
  arrange(desc(n_votu)) -> votu_order_count

# How many virus per order (only mags)
iphop_mags2[grepl("SemiBin", host_genome)] %>%
  group_by(Order) %>%
  summarise(n_votu=length(unique(virus))) %>%
  arrange(desc(n_votu))

iphop_mags2[grepl("SemiBin", host_genome)] %>%
  summarise(sum=length(unique(virus)))

iphop_mags2[grepl("SemiBin", host_genome)] %>%
  summarise(sum=length(unique(host_genome)))
  
mag_analysis$Order %>% table %>% sort

iphop_mags2[grepl("SemiBin", host_genome)] %>% 
  fwrite("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/iphop_vOTU-MAGs.csv")

# How many virus per order (all)
iphop_mags2 %>%
  group_by(Order) %>%
  summarise(n_votu=length(unique(virus))) %>%
  arrange(desc(n_votu)) %>%
  mutate(Order=str_remove(Order, "o__"))-> votu_order_count_all
```

# Check association between number of predicted vOTU and number ov MAGs by order
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
left_join(
  order_nmags[,c("Order", "n")],
  votu_order_count_all) %>%
  column_to_rownames("Order") %>%
  filter(!is.na(n_votu))-> order_votu_mag_counts

# Distribution
order_votu_mag_counts$n %>% shapiro_test()
order_votu_mag_counts$n_votu %>% shapiro_test()

# Test

cor.test(order_votu_mag_counts$n, order_votu_mag_counts$n_votu, method = "spearman")
plot(order_votu_mag_counts$n, order_votu_mag_counts$n_votu)


```

# AUX: Find overlapping annotations between blastp_hits and gems genes (1)
```{r eval=FALSE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
mag_prot<-list.files("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaG/coassemby_07feb24/mags_c70c10a99/checkm2_mags/protein_files/", pattern = ".faa", full.names = T)

foreach(i=mag_prot[-1])%do%{
  data.table(
    OTU=str_extract(i, "//[:graph:]+.faa") %>% str_remove_all("//|.faa"),
    header=readAAStringSet(i) %>% names
  )
} %>% rbindlist -> mag_prot_regions

mag_prot_regions %>%
  mutate(mag_gene=str_extract(header, "megahit_[:digit:]+_[:digit:]+"),
         contig=str_remove(mag_gene, "_[:digit:]+$"),
         region_span=str_extract(header, "# [:digit:]+ # [:digit:]+ "),
         range=str_replace(region_span, " # ","_") %>% str_remove("# ") %>% str_remove_all(" "),
         gene=paste0(contig,"_", range),
         ) %>%
  select(OTU,mag_gene, gene) -> mag_prot_regions1   # load and tidy all CDS

mag_prot_regions1[mag_gene %in% blast_phold$mag_gene] %>% #select CDS that matched viral functions (blastp)
  rowwise %>%
  mutate(
         contig=str_remove(mag_gene, "_[:digit:]+$"),
         id=paste0(OTU,"-", contig) %>% str_remove_all(" "),
         range=str_extract(gene, "[:digit:]+_[:digit:]+$"),
          start=sort(parse_integer(unlist(strsplit(range,"_"))))[1],
          end=sort(parse_integer(unlist(strsplit(range,"_"))))[2]
             ) %>%
  select(id,mag_gene, contig, start, end) %>% unique -> blastp_mag_contig_regions

gem_react_genes %>%
  group_by(OTU, gene, react_id, react_name) %>%
  reframe(gene=unlist(strsplit(gene,";"))) -> gem_react_genes_l

gem_react_genes_l %>%
    mutate(
         contig=str_remove(gene, "_[:digit:]+_[:digit:]+$"),
         id=paste0(OTU,"-", contig) %>% str_remove_all(" "),
         range=str_extract(gene, "[:digit:]+_[:digit:]+$")) %>%
  filter(contig %in% blastp_mag_contig_regions$contig) %>%
  rowwise %>%
  mutate(
    start=sort(parse_integer(unlist(strsplit(range,"_"))))[1],
    end=sort(parse_integer(unlist(strsplit(range,"_"))))[2]
         ) %>%
  select(id, gene, start, end) -> gem_react_contig_regions

fwrite(blastp_mag_contig_regions %>% select(-c(mag_gene, contig)) %>% unique, "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/blastp_mag_contig_regions.bed",
col.names = F,
sep = "\t", dec = ",")

fwrite(gem_react_contig_regions %>% select(-gene) %>% unique, "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/gem_react_contig_regions.bed", col.names = F,
sep = "\t", dec = ",")

# do bedtools
"bedtools intersect -a blastp_mag_contig_regions.bed -b gem_react_contig_regions.bed -wa -wb > aux_gems"

fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/aux_gems") -> aux_gems
names(aux_gems)=c("aux_contig", "aux_start", "aux_end", "gem_contig", "gem_start", "gem_end")
names(gem_react_contig_regions)=c("gem_contig", "gem_gene", "gem_start", "gem_end")
blastp_mag_contig_regions %<>% select(-contig)
names(blastp_mag_contig_regions)=c("aux_contig", "aux_gene", "aux_start", "aux_end")

aux_gems %>%
  left_join(blastp_mag_contig_regions) %>%
  left_join(gem_react_contig_regions) %>%
  #select(aux_contig, aux_gene, gem_gene) %>%
  unique %>%
  left_join(
    blast_phold[,c("mag_gene","func", "product")] %>%
      dplyr::rename(aux_gene=mag_gene)
      ) %>%
  left_join(gem_react_genes_l[,c("gene", "react_id", "react_name")] %>% unique %>%
              dplyr::rename(gem_gene=gene)) %>% unique -> blastp_aux_gem_react




fwrite(blastp_aux_gem_react, "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/blastp_aux_gem_react.csv")


  
```

# Analyse AUX functions in GEMS
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
blastp_aux_gem_react<- fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/blastp_aux_gem_react.csv")

# select genes aux genes that matched in GEMs
blastp_aux_gem_react %>%
  filter(grepl("Other|Auxiliary", func)) %>%
    mutate(gem=str_split(gem_contig, "-", simplify = T)[,1],
           mag_gene=paste0(gem, "-", aux_gene)) %>%
  select(mag_gene, product, react_name, react_id) %>% 
  unique-> blastp_aux_gem_react_aux

# Link votu and MAGs to genes containing AUX
blast_phold %>%
  filter(grepl("Other|Auxiliary", func)) %>%
  mutate(mag_gene=paste0(OTU, "-", mag_gene)) %>%
  filter(mag_gene %in% blastp_aux_gem_react_aux$mag_gene) %>%
  select(mag_gene, OTU, votu) %>% unique-> mag_gene_votu

# Match gene annotation with MAG and vOTU identity
full_join(blastp_aux_gem_react_aux, mag_gene_votu, by = "mag_gene") %>% 
  unique %>%filter(!is.na(votu)|!is.na(OTU)) ->  gem_votu_matches

gem_votu_matches %>%
  select(mag_gene, product, react_name, react_id, OTU, votu) %>% unique -> gem_votu_matches2

gem_votu_matches2 %>%
  group_by(product, react_name, react_id) %>%
  filter(!grepl("transport|resistance|repressor", product),
         !grepl("ABC|transport", react_name)
         ) %>%
  summarise(
    n=n(),
    nmag=length(unique(OTU)),
    nvotu=length(unique(votu)),
    gems_id=paste(unique(OTU), collapse = "|"),
            votu_id=paste(unique(votu), collapse = "|")
            ) %>% 
  ungroup %>%
  group_by(product) %>%
  slice_max(order_by = n, n = 5, with_ties = FALSE)%>%
  left_join(sbs_react[, c("react_id", "sbs_names")]) -> top_aux_gem_react

fwrite(top_aux_gem_react, "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/putative_avg_in_gems_annot_09feb26_v2.csv", sep="\t", dec=",")

```

# Describe selection of putative AVGs
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}

top_aux_gem_react_classes<-
  fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/putative_avg_in_gems_annot_16feb26_v6.tsv") %>%
  unique

# How many initial genes is aux and oth?
phold_annot_simple %>%
  filter(grepl("Aux|Oth", func)) %>% pull(product) %>% unique %>% length

gem_votu_matches2 %>%
  group_by(product, react_name, react_id) %>%
  filter(grepl("transport|resistance|repressor", product),
         !grepl("ABC|transport", react_name)
         ) %>% pull(product) %>% unique

# Final selection of putative AVGs
top_aux_gem_react_classes %>% 
  filter(`Unlikely AMG`!="T") -> top_aux_gem_react_classes_f

top_aux_gem_react_classes_f$product
top_aux_gem_react_classes_f$react_name

top_aux_gem_react %>%
  select(product, react_name, votu_id, gems_id) %>% 
  right_join(top_aux_gem_react_classes_f, by = c("product", "react_name", "votu_id", "gems_id")) %>% 
  filter(!is.na(react_name)) %>% unique %>% arrange(nvotu) %>% data.table -> top_aux_gem_react_summary2

foreach(i=1:nrow(top_aux_gem_react_summary2))%do%{
  top_aux_gem_react_summary2[i,gems_id] %>% strsplit("\\|") %>% unlist %>% unique -> gem_ids
  top_aux_gem_react_summary2[i,react_name] -> react_nam
  
  gem_analysis[OTU %in% gem_ids, Order] %>% unique %>% paste(collapse = "|") -> order_vec
  fba_reactions_f[OTU %in% gem_ids & react_name %in% react_nam, fba_flux] -> fba_flux
  
  top_aux_gem_react_summary2[i,] %>% 
    add_column(median_flux=(median(abs(fba_flux))),
               gem_orders=order_vec
               )
} %>% rbindlist-> top_aux_gem_react_summary3

top_aux_gem_react_summary3 %>%
  select(product, react_name, Class, votu_id, gems_id) %>%
  fwrite("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/aux_in_gems.csv", sep = "\t", dec = ".")

top_aux_gem_react_summary3 %>%
  select(Class, react_name, median_flux, nvotu, sbs_names, nmag, react_id) %>%
  group_by(Class, react_name,react_id, sbs_names) %>%
  summarise(median_flux=mean(median_flux),
            nvotu=sum(nvotu),
            nmag=sum(nmag),
            ) -> top_aux_gem_react_summary4

top_aux_gem_react_summary4 %>%
  group_by(Class) %>% count(name = "ngenes", sort = T) -> ngenes
ngenes$ngenes %>% sum
ngenes %>%
  mutate(prop=100*(ngenes/52))

top_aux_gem_react_summary4 %>%
  group_by(Class) %>% summarise(nvotu=sum(nvotu)) -> nvotus

full_join(ngenes, nvotus) %>% arrange(nvotu)->n_genes_votus

n_genes_votus$Class=factor(n_genes_votus$Class, levels = n_genes_votus$Class)

top_aux_gem_react_summary4$Class=
  factor(top_aux_gem_react_summary4$Class, levels = n_genes_votus$Class)
top_aux_gem_react_summary4$Class %>% unique

  ggplot() + 
  geom_bar(data= top_aux_gem_react_summary4,
           aes(y=Class, x=nvotu, fill=median_flux),
           stat="identity", color="black", size=0.7) +
  geom_text(data = n_genes_votus,
            aes(x = nvotu+1, y=Class, label = paste0("", ngenes))) +
  theme_bw() +
    xlab("Number of vOTUs with AVG") +
    ylab("Microbial metabolism") +
  scale_fill_gradientn(na.value = "white",colours = 
                      c(
                      low = "cadetblue",
                      mid="orange",
                      high = "red")) +
    theme(legend.position = "bottom",axis.text.y = element_text(size = 12)) +
    labs(fill = expression(atop("Median reaction flux", mmol~g[DW]^{-1}~h^{-1}))) -> p;p
  
ggsave(plot = p,
    "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/barplot_putative_avg_in_gems.png",
    width = 10, height = 8, units = "in", dpi = 600)

file.show("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/barplot_putative_avg_in_gems.png")
```

# Boxplots AVGs in GEMs, vOTU counts coloured by viral mechanism
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
top_aux_gem_react_summary_class<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/aux_in_gems_v7.csv") %>% select(Class, Class1, mechanism) %>% unique


top_aux_gem_react_summary4 %>%
  left_join(top_aux_gem_react_summary_class, by="Class") %>%
  mutate(
    mechanism = 
      case_when(
        mechanism == "Functional maintenance" ~ "Functional host maintenance",
        .default = mechanism))-> top_aux_gem_react_summary5

top_aux_gem_react_summary5 %>%
  filter(mechanism != "") %>%
  group_by(Class1) %>%
  summarise(nvotu=sum(nvotu)) %>%
  arrange((nvotu)) %>% pull(Class1) -> sorted_class_vec

top_aux_gem_react_summary5 %>%
  filter(mechanism != "") %>%
  group_by(Class1,mechanism) %>%
  summarise(nvotu=sum(nvotu)) %>%
  mutate(Class1=factor(Class1, levels = sorted_class_vec)) %>%
    ggplot() + 
  geom_bar(aes(y=Class1, x=nvotu, fill=mechanism),
           stat="identity", color="black", size=0.7) +
  theme_bw() +
    xlab("") +
    ylab("AVG functional category") +
  scale_fill_manual(values =
                      c(
                      "Resource scavenging" = alpha("paleturquoise3",0.2),
                      "Functional host maintenance" = alpha("#F2B342",0.8))
                    ) +
  guides(fill=guide_legend("AVG-related strategy", nrow = 2),
         ) +
  theme(legend.position = "bottom", text = element_text(size=16)) -> p

ggsave(plot = p,
    "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/barplot_putative_avg_in_gems_count_only.png",
    width = 8.35, height = 6, units = "in", dpi = 600)

file.show("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/barplot_putative_avg_in_gems_count_only.png")

```

# Number of AVGs examined
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
top_aux_gem_react_summary5$react_name %>% unique

```

# Proportion of AVGs by functional category
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}

top_aux_gem_react_summary5 %>%
  filter(!is.na(median_flux)) %>%
  group_by(Class1) %>%
  summarise(
    min   = min(median_flux, na.rm = TRUE),
    max   = max(median_flux, na.rm = TRUE),
    range = max(median_flux, na.rm = TRUE) - min(median_flux, na.rm = TRUE),
    abs_median = median(abs(median_flux), na.rm = TRUE)
    
  )
```

# Median flux values by category of AVGs
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
length(unique(top_aux_gem_react_summary2$react_name))
top_aux_gem_react_summary3 %>% 
  select(react_name, Class) %>% unique %>%
  group_by(Class) %>%
  summarise(prop=100*(n()/50)) %>% arrange(prop)

```

# Verify react flux values and pathways
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
top_aux_gem_react_summary3 %>%
  filter(grepl("glutamine", react_name, ignore.case=T)) %>%
  pull(react_id) -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("glutamine", react_name)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags


tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )

"D-Glucosamine phosphate"
key_metabolites[name=="D-Glucosamine phosphate", c("id", "name")]
find_metab_react("cpd00288", tidy_reactions_f[]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable

####################

top_aux_gem_react_summary3 %>%
  filter(grepl("D-glycero-beta-d-manno-heptose", react_name, ignore.case=T)) %>%
  pull(react_id) -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("D-glycero-beta-d-manno-heptose", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags, definition] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )

####################

top_aux_gem_react_summary3 %>%
  filter(grepl("arabinose", react_name, ignore.case=T)) %>%
  pull(react_id) -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("arabinose", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags, definition] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )

key_metabolites[name=="D-Arabinose5-phosphate", c("id", "name")]
find_metab_react("cpd00817", tidy_reactions_f[]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable

####################

top_aux_gem_react_summary3 %>%
  filter(grepl("nicotin", react_name, ignore.case=T)) %>%
  pull(react_id) -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("nicotin", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

gem_analysis[OTU %in% mags]

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags, definition] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )

####################

top_aux_gem_react_summary3 %>%
  filter(grepl("cobal", react_name, ignore.case=T)) %>%
  pull(react_id) -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("cobal", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

gem_analysis[OTU %in% mags] %>% arrange(Order)

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags, definition] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )

####################

top_aux_gem_react_summary3 %>%
  filter(grepl("this", react_name, ignore.case=T)) %>%
  pull(react_id) -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("this", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

gem_analysis[OTU %in% mags] %>% arrange(Order)

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags, definition] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )

####################

top_aux_gem_react_summary3 %>%
  filter(grepl("S-adeno", react_name, ignore.case=T)) %>%
  pull(react_id) -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("S-adeno", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

gem_analysis[OTU %in% mags] %>% arrange(Order)

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags, definition] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )
####################

top_aux_gem_react_summary3 %>%
  filter(grepl("oxo-L", react_name, ignore.case=T)) %>%
  pull(react_id) -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("oxo-L", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

gem_analysis[OTU %in% mags] %>% arrange(Order)

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags, definition] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )

####################

top_aux_gem_react_summary3 %>%
  filter(grepl("glutamate", react_name, ignore.case=T)) %>%
  pull(react_id) -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("glutamate", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

gem_analysis[OTU %in% mags] %>% arrange(Order)

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags, definition] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )

####################

top_aux_gem_react_summary3 %>%
  filter(grepl("aspartate", react_name, ignore.case=T)) %>%
  pull(react_id) -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("aspartate", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

gem_analysis[OTU %in% mags] %>% arrange(Order)

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags, definition] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )
####################

top_aux_gem_react_summary3 %>%
  filter(grepl("photo", react_name, ignore.case=T)) %>%
  pull(react_id) -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("photo", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

gem_analysis[OTU %in% mags] %>% arrange(Order)

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags, c("react_name", "definition")] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )

####################

top_aux_gem_react_summary3 %>%
  filter(grepl("sulf", react_name, ignore.case=T)) %>%
  pull(react_id) %>% unique -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("sulf", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

gem_analysis[OTU %in% mags] %>% arrange(Order)

tidy_reactions[react_id %in% react_ids[] & OTU %in% mags, definition] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )

####################

top_aux_gem_react_summary3 %>%
  filter(grepl("ferre", react_name, ignore.case=T)) %>%
  pull(react_id) %>% unique -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("ferre", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

gem_analysis[OTU %in% mags] %>% arrange(Order)

tidy_reactions[react_id %in% react_ids[] & OTU %in% mags, definition] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )

####################

top_aux_gem_react_summary3 %>%
  filter(grepl("complex", react_name, ignore.case=T)) %>%
  pull(react_id) %>% unique -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("ferre", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

gem_analysis[OTU %in% mags] %>% arrange(Order)

tidy_reactions[react_id %in% react_ids[] & OTU %in% mags, definition] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )

####################

top_aux_gem_react_summary3 %>%
  filter(grepl("urea", react_name, ignore.case=T)) %>%
  pull(react_id) %>% unique -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("urea", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

gem_analysis[OTU %in% mags] %>% arrange(Order)

tidy_reactions[react_id %in% react_ids[] & OTU %in% mags, definition] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )

####################

top_aux_gem_react_summary3 %>%
  filter(grepl("acetate", react_name, ignore.case=T)) %>%
  pull(react_id) %>% unique -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("acetate", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

gem_analysis[OTU %in% mags] %>% arrange(Order)

tidy_reactions[react_id %in% react_ids[] & OTU %in% mags, definition] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )
####################

top_aux_gem_react_summary3 %>%
  filter(grepl(":pyruvate,phosphate phosphotransferase", react_name, ignore.case=T)) %>%
  pull(react_id) %>% unique -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl(":pyruvate,phosphate phosphotransferase", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

gem_analysis[OTU %in% mags] %>% arrange(Order)

tidy_reactions[react_id %in% react_ids[] & OTU %in% mags, definition] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )

key_metabolites[name=="Phosphoenolpyruvate"]
find_metab_react("cpd00061", tidy_reactions_f[OTU %in% mags]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable

####################

top_aux_gem_react_summary3 %>%
  filter(grepl("hexadecanal", react_name, ignore.case=T)) %>%
  pull(react_id) %>% unique -> react_ids

top_aux_gem_react_summary3 %>%
  filter(grepl("hexadecanal", react_name, ignore.case=T)) %>% 
  pull(gems_id) %>% strsplit("\\|") %>% unlist %>% unique -> mags

gem_analysis[OTU %in% mags] %>% arrange(Order)

tidy_reactions[react_id %in% react_ids[] & OTU %in% mags, definition] %>% unique

tidy_reactions[react_id %in% react_ids[1] & OTU %in% mags] %>% 
  summarise(pos=sum(fba_flux>0),
            neg=sum(fba_flux<0)
            )

#################
key_metabolites[grepl("phosphoadenyl",name)]
find_metab_react("cpd00044", tidy_reactions[]) %>%
  filter(utilisation == "consumed")-> res; res %>% datatable

```

# Verify gene location in genome
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
top_aux_gem_react_summary3 %>%
  #filter(grepl("5-oxo", react_name)) %>% 
  filter(Class == "Carbonyl-compound") %>%
  pull(votu_id) %>% strsplit("\\|") %>% unlist %>% unique -> votu_selection

{
vir_func_cat<-phold_annot2$func %>% unique
vir_func_cat_colors<-hcl.colors(n=9, palette = "Zissou 1")
names(vir_func_cat_colors)<-vir_func_cat
vir_func_cat_colors["unknown function"]="white"
vir_func_cat_colors["Other"]="orange"
vir_func_cat_colors["Auxiliary genes and host takeover"]="red"
vir_func_cat_colors["Structural components"]="lightblue"
vir_func_cat_colors["integration and excision"]="#C05763"
vir_func_cat_colors["lysis"]="#FDF7DC"
vir_func_cat_colors["transcription regulation"]="#4C9168"
vir_func_cat_colors["DNA, RNA and nucleotide metabolism"]="#B7D886"
}

phold_annot2[votu %in% votu_selection] %>%
  group_by(votu) %>%
  summarise(max(end)) %>% arrange(votu)


vir_analysis[votu %in% votu_selection]
phold_annot2[votu %in% votu_selection] %>%
  mutate(annot=
           case_when(func == "Structural components" ~ product,
                     func == "Auxiliary genes and host takeover" ~ product,
                     func == "Other" ~ product,
                     ))->toplot
toplot$votu %<>% factor(levels=votu_selection)
require(gggenes)
ggplot(toplot,
       aes(xmin = start, xmax = end, y = votu, fill = func)) +
    geom_gene_arrow() +
  geom_gene_label(aes(label=product),min.size = 0) +
  scale_fill_manual(values=vir_func_cat_colors, name = "Virus gene\nfunctional category") +
  theme_genes() +
  facet_wrap(~ votu, scales = "free", ncol = 1) +
  theme(legend.position = "bottom") -> p;p

ggsave(plot = p,
       filename= "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/select_aux_votu_test.png",
       width=20, height = 30, units = "in", dpi = 600)
file.show("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/select_aux_votu_test.png")

```
# How many glycosyltransferases?
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
phold_annot2 %>%
  filter(grepl("Aux|Oth", func)) %>%
  group_by(product) %>%
  summarise(nvotu=length(unique(votu))) %>%
  arrange(desc(nvotu)) %>% print(n=50)
```

# How many vOTUs with the selected genes?
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
unique(unlist(strsplit(top_aux_gem_react_summary3$votu_id, "\\|"))) -> aux_votus

100*(length(aux_votus)/nrow(vir_analysis))

vir_analysis[votu %in% aux_votus, mean_vir_all] %>% sum

phold_annot2[grepl("Aux|Oth", func)]

```


# Analyse AUX functions in GEMS
```{r eval=F, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}

"Galactose-3-O-Sulfotransferase"


phold_annot2%>%
  filter(grepl("Galactose-3-O-Sulfotransferase", product, ignore.case=T)) %>% pull(votu) %>% unique-> votus

votus %>% length
vir_analysis[votu %in% votus] %>% view

blast_phold[grepl("Galactose-3-O-sulfotransferase", product), OTU] %>% unique %>% length

blastp_aux_gem_react_count %>%
  filter(grepl("Galactose", react_name, ignore.case=T))-> matches;
matches %>% datatable

tidy_reactions[react_id %in% matches$react_id] %>% 
  group_by(react_id, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25)

##################################################
"bifunctional heptose"

blast_phold[grepl("bifunctional heptose", product), OTU] %>% unique %>% length

blastp_aux_gem_react_count %>%
  filter(grepl("bifunctional heptose", product, ignore.case=T))-> matches;matches

tidy_reactions_f[react_id %in%  c("rxn03916_c0", "rxn03918_c0")] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") %>%
  datatable

gapseq_pwy_hierarchy[sbs=="PWY0-1241"]

##################################################
"phosphoheptose" #no clear match

blast_phold[grepl("phosphoheptose", product), OTU] %>% unique %>% length

blastp_aux_gem_react_count %>%
  filter(grepl("phosphoheptose", product, ignore.case=T))-> matches;matches 

tidy_reactions[react_id %in%  c("rxn03916_c0", "rxn03918_c0")] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
  res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

#################################################
"glutamine-D-fructose"
phold_annot2%>%
  filter(grepl("glutamine-D-fructose", product, ignore.case=T)) %>% pull(votu) %>% unique-> votus

votus %>% length
vir_analysis[votu %in% votus]

blastp_aux_gem_react_count %>%
  filter(grepl("glutamine-D-fructose", product, ignore.case=T))-> matches;matches

tidy_reactions_f[react_id %in% c("rxn00790_c0", "rxn00555_c0", "rxn15064_c0")] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions_f[react_id %in% c("rxn00790_c0", "rxn00555_c0", "rxn15064_c0")] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view
#################################################
"gamma-glutamyl cyclotransferase" # matches several oxido reductases acting on alcohols

blastp_aux_gem_react_count %>%
  filter(grepl("gamma-glutamyl cyclotransferase", product, ignore.case=T))-> matches;matches

tidy_reactions[react_id %in% matches$react_id] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions_f[react_id %in% matches$react_id] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("glutamyl",definition)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable


#################################################
"cysteine dioxygenase" # matches several different reactions, no reference to cysteine or alanine

blastp_aux_gem_react_count %>%
  filter(grepl("cysteine dioxygenase", product, ignore.case=T))-> matches;matches

tidy_reactions[react_id %in% matches$react_id] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions_f[react_id %in% matches$react_id] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("cysteine",react_name)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable


#################################################
"2OG" # no matches in GEMs
blast_phold%>%
  filter(grepl("2OG", product, ignore.case=T))

blastp_aux_gem_react_count %>%
  filter(grepl("2OG", product, ignore.case=T))-> matches;matches

tidy_reactions[react_id %in% matches$react_id] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions_f[react_id %in% matches$react_id] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("2-Oxoglutarate",react_name)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable


#################################################
"Porphyrin"
blast_phold%>%
  filter(grepl("Porphyrin", product, ignore.case=T))

blastp_aux_gem_react_count %>%
  filter(grepl("Porphyrin", product, ignore.case=T))-> matches;matches

tidy_reactions[react_id %in% matches$react_id] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(mean_fba=mean(abs(fba_flux)),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions[react_id %in% matches$react_id] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view


gem_analysis[OTU %in% 
               unique(tidy_reactions[react_id %in% matches$react_id, OTU])] %>%
  group_by(Order) %>% count(sort = T)

#################################################
"arabinose" 
phold_annot2%>%
  filter(grepl("arabinose", product, ignore.case=T)) %>% pull(votu) %>% unique-> votus
votus %>% length()
vir_analysis[votu %in% votus]


blastp_aux_gem_react_count %>%
  filter(grepl("arabinose", product, ignore.case=T))-> matches;matches

tidy_reactions[react_id %in% "rxn01117_c0"] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(mean_fba=mean(abs(fba_flux)),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions_f[react_id %in% "rxn01117_c0"] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("",react_name)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

#################################################
"S-adenosyl" 
blast_phold%>%
  filter(grepl("S-adenosyl", product, ignore.case=T))

blastp_aux_gem_react_count %>%
  filter(grepl("S-adenosyl", product, ignore.case=T))-> matches;matches

tidy_reactions[react_id %in% matches$react_id] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(mean_fba=median((fba_flux)),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions_f[react_id %in% matches$react_id] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("",react_name)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gem_analysis[OTU %in% 
               unique(tidy_reactions[react_id %in% matches$react_id, OTU])] %>%
  group_by(Order) %>% count(sort = T)

# Methanol
key_metabolites[name=="Putrescine"]
find_metab_react("cpd00118", tidy_reactions[]) %>% datatable

#################################################
"legionaminate" 
blast_phold%>%
  filter(grepl("legiona", product, ignore.case=T))

blastp_aux_gem_react_count %>%
  filter(grepl("legiona", product, ignore.case=T))-> matches;matches

tidy_reactions_f[react_id %in% matches$react_id] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(mean_fba=mean(abs(fba_flux)),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions[react_id %in% matches$react_id] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") %>%
  filter(grepl("phosphoenolpyruvate", react_name))  -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("",react_name)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

#################################################
"ferredoxin" 
phold_annot2%>%
  filter(grepl("ferredoxin", product, ignore.case=T)) %>% pull(votu) %>% unique-> votus

votus %>% length
vir_analysis[votu %in% votus] 


blastp_aux_gem_react_count %>%
  filter(grepl("ferredoxin", product, ignore.case=T))-> matches;matches

tidy_reactions_f[react_id %in% "rxn05937_c0"] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions_f[react_id %in% "rxn05937_c0"] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

mag_analysis[OTU %in% 
               unique(tidy_reactions_f[react_id %in% matches$react_id, OTU])] %>%
  group_by(Order,Family) %>% count(sort = T)

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("",react_name)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

#################################################
"sulfurylase"
blast_phold%>%
  filter(grepl("sulfurylase", product, ignore.case=T))

blastp_aux_gem_react_count %>%
  filter(grepl("sulfurylase", product, ignore.case=T))-> matches;matches

tidy_reactions[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions_f[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("",react_name)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

#################################################
"glutamine amidotransferase" # no matches in GEMs
blast_phold%>%
  filter(grepl("glutamine amidotransferase", product, ignore.case=T)) 

blastp_aux_gem_react_count %>%
  filter(grepl("glutamine amidotransferase", product, ignore.case=T))-> matches;matches

votus %>% length()
vir_analysis[votu %in% votus]


tidy_reactions[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("",react_name)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

#################################################
"enoyl-coa" # no matches in GEMs
blast_phold%>%
  filter(grepl("enoyl-coa", product, ignore.case=T))

blastp_aux_gem_react_count %>%
  filter(grepl("enoyl-coa", product, ignore.case=T))-> matches;matches

tidy_reactions[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("",react_name)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

#################################################
"pnuc" # no matches in GEMs
blast_phold%>%
  filter(grepl("pnuc", product, ignore.case=T))

blastp_aux_gem_react_count %>%
  filter(grepl("pnuc", product, ignore.case=T))-> matches;matches

tidy_reactions[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("",react_name)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable
#######################################
"methylamine" # no matches in GEMs
blast_phold%>%
  filter(grepl("methylamine", product, ignore.case=T))

blastp_aux_gem_react_count %>%
  filter(grepl("methylamine", product, ignore.case=T))-> matches;matches

tidy_reactions[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("",react_name)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

####################################################
"phosphofructokinase" # no matches in GEMs
blast_phold%>%
  filter(grepl("phosphofructokinase", product, ignore.case=T))

blastp_aux_gem_react_count %>%
  filter(grepl("phosphofructokinase", product, ignore.case=T))-> matches;matches

tidy_reactions[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("",react_name)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

####################################################
"cobalto" # no matches in GEMs
blast_phold%>%
  filter(grepl("cobalto", product, ignore.case=T))

blastp_aux_gem_react_count %>%
  filter(grepl("cobalto", product, ignore.case=T))-> matches;matches

tidy_reactions[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("",react_name)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

####################################################
"Fe-S" # no matches in GEMs
blast_phold%>%
  filter(grepl("Fe-S", product, ignore.case=T))

blastp_aux_gem_react_count %>%
  filter(grepl("Fe-S", product, ignore.case=T))-> matches;matches

tidy_reactions[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("",react_name)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

####################################################
"phosphoadenosine p"

phold_annot2%>%
  filter(grepl("phosphoadenosine", product, ignore.case=T)) %>% pull(votu) %>% unique-> votus
votus %>% length()
vir_analysis[votu %in% votus]

blastp_aux_gem_react_count %>%
  filter(grepl("phosphoadenosine", product, ignore.case=T))-> matches;matches

tidy_reactions_f[react_id %in% matches$react_id] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25)

tidy_reactions_f[react_id=="rxn00379_c0"] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

mag_analysis[OTU %in% 
               unique(tidy_reactions_f[react_id %in% matches$react_id, OTU])] %>%
  group_by(Order,Family) %>% count(sort = T)

#################################################
"nicotinamide-nucleo" # seems to have a fairly different function
blast_phold%>%
  filter(grepl("nicotinamide-nucleo", product, ignore.case=T))

blastp_aux_gem_react_count %>%
  filter(grepl("nicotinamide-nucleo", product, ignore.case=T))-> matches;matches

tidy_reactions[react_id %in% c("rxn03372_c0", "rxn00361_c0")] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions[react_id %in% "rxn25839_c0"] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("",react_name)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

####################################################
"nicotinamide phospho" # seems to have a fairly different function
blast_phold%>%
  filter(grepl("nicotinamide phosp", product, ignore.case=T))

blastp_aux_gem_react_count %>%
  filter(grepl("nicotinamide phosp", product, ignore.case=T))-> matches;matches

tidy_reactions_f[react_id %in% matches$react_id] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux) %>% round(4),
            n=n()
            ) %>% print(n=25) %>% datatable

tidy_reactions[react_id %in% matches$react_id] %>% 
  group_by(react_id,react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

gapseq_pwy_hierarchy[sbs %in% unlist(strsplit(res$subsytems_ids, ";"))] %>% view

# Check functions involving the metabolite/functiona group
tidy_reactions[grepl("",react_name)] %>% 
  group_by(react_id, react_id2, react_name, definition) %>% 
  summarise(median_fba=median(fba_flux),
            n=n()
            ) %>%
  left_join(react_sbs_metab_info[,c("react_id2", "subsytems_ids", "subystems")], by="react_id2") -> res
res %>% datatable

#################################################

```



#----
# GATHER RESULTS

#AUX: Merge blastp aux hits with gems genes
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
gem_react_genes %>% 
  group_by(OTU, react_id, react_name) %>%
  reframe(gene=strsplit(gene, ";") %>% unlist) -> gem_react_genes_l
```

# AUX genes co-occurring
```{r, echo = TRUE, warning=FALSE, message=FALSE, results='show', eval=F}
# Genes together
selected_aux1 %>%
  group_by(votu) %>%
  summarise(
    n_aux=n(),
    aux_genes=paste(product, collapse = ";")
  ) 

selected_aux1 %>%
  select(votu, product) %>%
  unique %>%
  add_column(p=1) %>%
  pivot_wider(names_from = product, values_from = p, values_fill = 0) %>%
  column_to_rownames("votu") %>% as.matrix -> aux_votu_mat
  aux_votu_mat[rowSums(aux_votu_mat)>1,]-> aux_votu_mat1
  
  Heatmap(aux_votu_mat1,
          col = c("white", "orange"),
          rect_gp = gpar(col = "gray20", lwd = 0.4),
          row_names_gp = gpar(fontsize=0),
          column_names_gp = gpar(fontsize=8),
          column_names_rot = 45,
          heatmap_width = unit(10, "in"),
          heatmap_height = unit(10, "in")
          ) 
```

# AUX: Distribution of AVG across predicted vOTUs
```{r, echo = TRUE, warning=FALSE, message=FALSE, results='show', eval=F}
iphop_mags2[] %>%
  select(votu=virus, Order) %>% 
  unique %>%
  left_join(phold_annot3) %>%
  select(votu, Order, func, product) %>% unique %>%
    filter(grepl("Aux|Oth", func)) %>%
  ungroup %>%
  group_by(Order) %>%
  count(product) %>%
  view
```

# DEF: Classify metab use by order
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
foreach(metabolite=unique(tidy_exhanges$id), .errorhandling = "remove")%do%{
  tidy_exhanges[id==metabolite] -> dt
  metab_name<-unique(dt$name)
  dt %>%
    group_by(Order) %>%
    summarise(
              median_imp=median(fba_flux[fba_flux<0]),
              median_exp=median(fba_flux[fba_flux>0]),
              n_imp=sum(fba_flux<0),
              n_exp=sum(fba_flux>0),
              n_use=n()
              ) %>%
    left_join(order_nmags) %>%
    filter(n>=5) %>%
    rowwise() %>%
    mutate(
      fisher = list(
        fisher.test(
          matrix(c(n_imp, n - n_imp,
                   n_exp, n - n_exp),
                 nrow = 2, byrow = TRUE)
        )
      ),
      p_value   = fisher$p.value,
      odds_ratio = fisher$estimate) %>%
    ungroup() %>%
    mutate(p_adj = p.adjust(p_value, method = "fdr")) %>%
    filter(p_adj<=0.05) %>%
    add_column(id=metabolite, name=metab_name) %>%
    select(id, name, Order, median_imp, median_exp, n_imp, n_exp, n_use, n, odds_ratio, p_adj)
  } %>% rbindlist -> order_metab_fisher

order_metab_fisher %>%
  filter(odds_ratio>1) -> order_metab_fisher_imported
order_metab_fisher %>%
  filter(odds_ratio<1) -> order_metab_fisher_exported
  

order_metab_fisher_imported %>%
  group_by(Order) %>%
  summarise(name=paste0(name, collapse = "; ")) %>% view

order_metab_fisher_exported %>%
  group_by(Order) %>%
  summarise(name=paste0(name, collapse = "; ")) %>% view

# Number of orders classifiables
order_metab_fisher %>%
  pull(Order) %>% unique %>% length #20

# Average number of aminoacids classified
order_metab_fisher %>% 
  left_join(key_metabolites %>%
  select(name, simple_category) %>% unique)->order_metab_fisher1


# Check frequent substrates
order_metab_fisher1 %>%
  filter(odds_ratio>1) %>%
  group_by(simple_category, name) %>% 
  summarise(prop=100*(length(unique(Order))/20)) %>%
  filter(prop>=90) -> frequent_metab

frequent_metab %>%
  group_by(simple_category) %>% count(sort = T)


# how many frequent metabolites are detect?
frequent_metab %>%
  filter(!simple_category %in% c("Metals", "Inorganic compounds")) %>%
  pull(name) %>% length

intersect(
metab_react_matched_cpds3$name,
frequent_metab$name) %>% length

13/22

# What is their median fluxes
order_metab_fisher1[name %in% frequent_metab$name] %>%
    filter(odds_ratio>1) %>%
  group_by(name) %>%
    summarise(median=median(median_imp )) -> freq_median_f

table(freq_median_f$median<(-1))

```

# Top producers by Order
```{r eval=T, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
order_nmags %>%
  filter(n>=5) %>% pull(Order) -> teastable

foreach(i=unique(tidy_exhanges$name), .errorhandling = "remove")%do%{
  tidy_exhanges %>%
  #filter(abs(fba_flux) > 0.00) %>%
  filter(Order %in% teastable) %>%
  filter(name == i) -> dt
  g <- unique(dt$Order)
  do.call(rbind, lapply(g, function(gr) {
    x <- dt$fba_flux[dt$Order == gr]
    y <- dt$fba_flux[dt$Order != gr]
    data.frame(
      group = gr,
      n_group = length(x),
      n_rest = length(y),
      median_group = median(x, na.rm = TRUE),
      median_rest = median(y, na.rm = TRUE),
      p_value = wilcox.test(x, y, alternative = "greater", exact = FALSE)$p.value
    )
      }))  %>% 
    add_column(name=i) -> df
  df %>%
      add_column(p_adj=
        df$p_value %>% p.adjust(method = "fdr"))
  } %>% rbindlist() %>%
  filter(p_value<=0.05) %>% filter(median_group>0) %>%
  mutate(p_adj=round(p_adj, 5)) %>%
    mutate(FC=round((median_rest-median_group)/median_group, 3))-> top_prod

top_prod %>%
  group_by(group) %>%
  summarise(name=paste0(name, collapse = "; "))

top_prod %>%
  group_by(name) %>%
  summarise(n=n(),
            group=paste0(group, collapse = "; ")) %>%
  arrange(desc(n)) %>% print(n=Inf) -> top_prod_summary

top_prod_summary$name

```


# Top consumers by Order
```{r eval=T, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
order_nmags %>%
  filter(n>=5) %>% pull(Order) -> teastable

foreach(i=unique(tidy_exhanges$name), .errorhandling = "remove")%do%{
  tidy_exhanges %>%
  filter(Order %in% teastable) %>%
  #filter(fba_flux < 0.00) %>%
  filter(name == i) -> dt
  g <- unique(dt$Order)
  do.call(rbind, lapply(g, function(gr) {
    x <- dt$fba_flux[dt$Order == gr]
    y <- dt$fba_flux[dt$Order != gr]
    data.frame(
      group = gr,
      n_group = length(x),
      n_rest = length(y),
      median_group = median(x, na.rm = TRUE),
      median_rest = median(y, na.rm = TRUE),
      p_value = wilcox.test(x, y, alternative = "less", exact = FALSE)$p.value
    ) 
  })) %>% 
    add_column(name=i) -> df
  df %>%
      add_column(p_adj=
        df$p_value %>% p.adjust(method = "BH"))
  } %>% rbindlist() %>%
  filter(p_adj<=0.05) %>% filter(median_group<0) %>% 
  mutate(p_adj=round(p_adj, 5))%>%
    mutate(FC=round((median_rest-median_group)/median_group, 3))-> top_sub

top_sub %>%
  group_by(group) %>%
  summarise(name=paste0(name, collapse = "; "))

top_sub %>%
  group_by(name) %>%
  summarise(n=n(),
            group=paste0(group, collapse = "; ")) %>%
  arrange(desc(n)) %>% print(n=Inf) %>% view
```

# Pathway distribution 
```{r eval=FALSE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
files<-list.files(path = "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaG/coassemby_07feb24/mags_c70c10a99/gapseq/",
                  pattern ="-all-Pathways.tbl", full.names = T)
foreach(i = files)%do%{
fread(i) %>%
  filter(Prediction) %>% 
  mutate(sbs_ids = str_remove_all(ID, "\\|")) %>%
  left_join(sbs_hier1) %>%
    add_column(OTU=str_extract(i, "[:alpha:]{3}_SemiBin_[:digit:]+"))
  } %>% rbindlist -> pws

pws %>%
  select(OTU, Name, h2, h3, h4, h5, ReactionsFound) %>%
  left_join(gem_analysis[,c("OTU", "Order")]) %>%
  rowwise %>%
  mutate(n_react=length(unlist(strsplit(ReactionsFound, " ")))) %>%
  data.table -> pws1


pws1$h3 %>% unique

pws1 %>%
  select(h2, h3, h4, h5, Name) %>% unique -> pwy_unique

pws1 %>%
  group_by(h3, h4, h5, Name, Order) %>% 
    left_join(gem_analysis[,c("OTU", "mean_mag_abund")]) %>%
  summarise(n_mag=n(),
            sum_abund=round(sum(mean_mag_abund),2),
            #orders=paste0(unique(Order), collapse = ";")
            ) -> pws_summary_order

```

##Classify pathway by order
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
foreach(i=unique(pws1$Name), .errorhandling = "remove")%do%{
  pws1[Name==i] -> dt
  dt %>%
    group_by(Order) %>%
    summarise(
              n_pwy=n()
              ) %>%
    left_join(order_nmags) %>%
    filter(n>=5)-> dt1

total_pwy <- sum(dt1$n_pwy)
total_n   <- sum(dt1$n)

dt1 %>%
  rowwise() %>%
  mutate(
    # 2x2 contingency table per order
    ft = list(fisher.test(matrix(c(
      n_pwy,                          # in order, with pathway
      n - n_pwy,                      # in order, without pathway
      total_pwy - n_pwy,              # not in order, with pathway
      total_n - n - total_pwy + n_pwy # not in order, without pathway
    ), nrow = 2))),
    p_value  = ft$p.value,
    odds_ratio = ft$estimate,
    direction  = ifelse(n_pwy / n > total_pwy / total_n, "enriched", "depleted")
  ) %>%
  ungroup() %>%
  mutate(p_adj = p.adjust(p_value, method = "BH")) %>%
  select(Order, n_pwy, n, direction, odds_ratio, p_value, p_adj) %>%
  arrange(p_adj) %>% add_column(Name=i)
} %>% rbindlist -> pws_fish

pws_fish %>%
  filter(p_adj<0.05 & 
         direction == "enriched" &
          odds_ratio > 1 
           )-> pwy_fish_ehnr

pwy_fish_ehnr %>%
  left_join(sbs_hier1) -> pwy_fish_ehnr_h

pwy_fish_ehnr_h$Order %>% unique
    
```

#---
#FIGURES
### Produced plots
```{r eval=F, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
top_prod1<-rbind(top_prod,top_sub) %>% data.table()

iphop_mags2 %>% 
  group_by(Phylum, Order) %>%
  summarise(sum=length(unique(virus))) %>%
    arrange(sum) %>% 
  filter(sum>=5 & 
         Order %in% mag_analysis$Order) %>% unique %>%
  pull(Order) %>% str_remove_all("o__")-> sorted_votus2

order_nmags %>%
  filter(n>=5) %>% pull(Order)-> testable

foreach(i=c("Hydrogen", "CO2", "H2S",
            "Ammonium", "indol","Methanethiol",
            "Acetate", "Propionate", "Butyrate"
            ),.inorder = T)%do%{

tidy_exhanges %>%
  filter(Order %in% testable) %>%
  filter(name==i) %>%
  mutate(significance=case_when(Order %in% top_prod1[name==i, group]~ "s", .default = "ns")
          )-> df
              
df %>% group_by(Order) %>% 
    summarise(median=median(fba_flux)) %>%
  arrange((median)) -> sorted

ggplot(df %>%
        mutate(Order=factor(Order, levels=sorted$Order)),
       aes(x = fba_flux, y = Order, fill = significance)) +
  geom_boxplot(outliers = F) +
  geom_jitter(size=0.1, alpha = 0.3) +
  geom_vline(xintercept=median(df$fba_flux), color="red", linetype = "dashed", size=0.5) +
  theme_bw() +
  ggtitle(i) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") +
  ylab("") +
  xlab("") +
  scale_fill_manual(
    values = c(s="red", ns="white"), na.value = "white")

} -> pl

(pl[[1]]+pl[[2]]+pl[[3]] + coord_cartesian(xlim = c(-1, 2)))  / 
  (pl[[4]]+pl[[5]]+pl[[6]]) / 
  (pl[[7]]+pl[[8]]+pl[[9]]) ->pf

ggsave(plot = pf, filename = "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/boxplot_products_vir.png",
       height = 10, width = 16, units = "in", dpi = 600
       )
file.show("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/boxplot_products_vir.png")

```

##Heatmap: MetaB utilisation by order v2
```{r eval=FALSE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
tidy_exhanges %>%
  group_by(name,simple_category) %>%
  summarise(median_pos=median(fba_flux[fba_flux>0], na.rm = T),
            median_neg=median(fba_flux[fba_flux<0], na.rm = T),
            prop_pos=round(100*(sum(fba_flux[fba_flux>0], na.rm = T)/311),2),
            prop_neg=round(100*(sum(fba_flux[fba_flux<0], na.rm = T)/311),2)
            ) %>%
  arrange(desc(median_pos)) -> exchanges_summary

#####  HM O2
require(ComplexHeatmap)
tidy_exhanges<-fba_comp_metab1

# vOTU abundance by host order
votu_host_order_abund<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/votu_host_order_mean.csv")
votu_host_order_abund$Order = str_remove(votu_host_order_abund$Order,"o__")
votu_host_order_abund_vec<-votu_host_order_abund$mean
names(votu_host_order_abund_vec) = votu_host_order_abund$Order


# Select orders
# Select orders
c("Bacteroidales", "Rhizobiales", "Pseudomonadales",  "Rhodobacterales", "Chromatiales", "Flavobacteriales", "Desulfobacterales", "Desulfobulbales", "Cytophagales",  "Woeseiales", "JAAYZQ01", "SZUA-229"
             ) -> sel_orders


## Wider alternative
votu_order_count_all %>%
  filter(n_votu>=2) -> order_min2_votu

# Select which categories to plot
unique(tidy_exhanges$simple_category)
selected_classes<-c(
    "Polysaccharides", "Carbohydrates", "Sugar Alcohols",
    "Alcohols", "Carbonyl compounds",
    "Sulfur compounds",
    "N-Carbonyl compounds","Amino acids", "Nitrogen compounds", "Inorganic compounds")

tax_lev="Order"

tidy_exhanges[Order %in% sel_orders] %>%
  filter(abs(fba_flux)>0.001) %>%
  group_by(name, Order, simple_category) %>%
  summarise(n_flux=n()) %>%
  left_join(order_nmags[,c("Order", "n")]) %>%
  mutate(prop=n_flux/n) %>%
  filter(prop>0.5) -> freq_exc

tidy_exhanges[Order %in% sel_orders] %>%
  filter(abs(fba_flux)>0.001) %>%
  select(name, simple_category, !!sym(tax_lev), fba_flux) %>%
  filter(simple_category %in% selected_classes) %>%
  group_by(name, simple_category, !!sym(tax_lev)) %>%
  summarise(median=median(fba_flux)) %>%
  mutate(
    tax=as.factor(!!sym(tax_lev)),
    name=factor(name, levels=unique(tidy_exhanges$name)))-> medians

#semi_join(medians, freq_exc, by=c("name", "Order")) %>%
#  select(name, simple_category, Order, median, tax)-> medians

quant<-medians$median %>% quantile(seq(0,1,0.05));quant

medians %>%
  ungroup %>%
  mutate(f_cat=case_when(
    median>0 & median<=quant[["85%"]] ~ 0.5,
    median>quant[["85%"]] & median<=quant[["100%"]] ~ 1,
    median<0 & median>=quant[["35%"]] ~ -0.5,
    median>=quant[["0%"]] & median<quant[["35%"]] ~ -1
  )) %>% 
  dplyr::select(tax, name, f_cat) %>% data.table ->medians1 
  medians1$name %>% unique
  medians1[CJ(tax = unique(tax), name = levels(medians1$name)), on = .(tax, name)]->medians2
  medians2$name %>% unique
  medians2[is.na(f_cat), f_cat := 0]
  
  medians2 %>%
  pivot_wider(names_from = name, values_from = f_cat, values_fill = 0) %>%
  column_to_rownames("tax") %>%
  as.matrix  -> order_mat

tidy_exhanges[name %in% medians$name,c("name", "simple_category")] %>%
    filter(simple_category %in% selected_classes) %>%
  unique %>% filter(name %in% colnames(order_mat)) %>%
  mutate(simple_category=
           factor(
           simple_category,
         levels = 
          selected_classes)) %>% arrange(simple_category) -> metab_cat

metab_class_vec=metab_cat$simple_category
names(metab_class_vec)=metab_cat$name


# Define splitting vectors
gem_analysis$Phylum %>% as.factor->phy_vec
names(phy_vec)=gem_analysis %>% pull(!!sym(tax_lev))

# Define HM body colors
flux_colors=c(alpha("cadetblue",0.8),
              alpha("cadetblue",0.8),
              "white",
              alpha("red",0.8),
              alpha("red",0.8)
              )

# Transpose matrix
reorder_mat<-order_mat %>% t
reorder_mat<-reorder_mat[names(metab_class_vec),]

# Bottom annot

# Number of predicted host
votu_order_count_all %>%
  filter(Order %in% colnames(reorder_mat)) -> votu_hits
votu_vec<-votu_hits$n_votu
names(votu_vec) = votu_hits$Order


# Relative abundance by Order
abund_order_df<-gem_analysis %>%
  group_by(Order) %>%
  summarise(sum=sum(mean_mag_abund))
abund_order_df$sum->abund_order_df_vec
names(abund_order_df_vec)=abund_order_df$Order
abund_order_df_vec %>% sort

# nGEMs by order
ngems_order_df<-gem_analysis %>%
  group_by(Order) %>%
  summarise(sum=length(OTU))
ngems_order_df$sum->ngems_order_df_vec
names(ngems_order_df_vec)=ngems_order_df$Order
ngems_order_df_vec %>% sort


# Predicted biomass growth
biomass_order_df<-gem_analysis %>%
  group_by(Order) %>%
  summarise(mean=mean(growth))
biomass_order_df$mean->biomass_order_df_vec
names(biomass_order_df_vec)=biomass_order_df$Order


bottom_annot<-
columnAnnotation(
  
  
   "Number of vOTUs\n linked to order" = 
     anno_barplot(x = votu_vec[colnames(reorder_mat)],
                 gp=gpar(fill="gray80"),
                 axis_param = list(labels_rot = 0,
                                   gp = gpar(fontsize = 6),
                                   side = "right"
                                   #direction="reverse"
                                   ),
     ),
  

  "Avg. rel. abundance of vOTUs\n linked to order" = 
     anno_barplot(x = votu_host_order_abund_vec[colnames(reorder_mat)],
                 gp=gpar(fill="gray80"),
                 axis_param = list(labels_rot = 0,
                                   gp = gpar(fontsize = 6),
                                   side = "right"
                                   #direction="reverse"
                                   ),
     ),
  
    "MAG rel. abundance\n by order (%)" = 
    anno_barplot(x = abund_order_df_vec[colnames(reorder_mat)],
                 gp=gpar(fill="gray80"),
                 axis_param = list(labels_rot = 0,
                                   gp = gpar(fontsize = 6),
                                   side = "right"

                                   #direction="reverse"
                                   ),
                 ),
  
    spacer=anno_empty(border = F, height = unit(2, "mm")),
  show_legend = F,
  annotation_name_rot = 0,
  annotation_name_side = "right",
  annotation_name_gp = gpar(fontsize = 10),
  annotation_name_offset = unit(0.6, "cm"),
  gap=unit(2,"mm")
  )

# Top annot
HeatmapAnnotation(
  "MAG phylum" = phy_vec[colnames(reorder_mat)],
  which = "column",
  col = list("MAG phylum"=tree_colors),
  border = T,  show_legend = F,
  show_annotation_name = F,
    annotation_name_side = "right"
)-> top_annot

# Right annot
rownames(reorder_mat)%in%
metab_react_matched_cpds_mean_scaled[, name] %>% as.character->metaB_presence_vec
names(metaB_presence_vec)=rownames(reorder_mat)


right_annot<-
rowAnnotation(
  #spacer=anno_empty(border = F, height = unit(0.5, "mm")),
  "Detected metabolites"=metaB_presence_vec[rownames(reorder_mat)],
  annotation_name_side = "bottom",
  col=list("Detected metabolites"=c("FALSE"="white","TRUE"="gray80")),
  border=T,
  show_legend = F)

# left annot
fba_median_pos<-exchanges_summary$median_pos
names(fba_median_pos)<-exchanges_summary$name

fba_median_neg<-exchanges_summary$median_neg %>% abs
names(fba_median_neg)<-exchanges_summary$name

fba_median_neg[fba_median_neg>30]
fba_median_neg[fba_median_neg>30] = 30

left_annot<-rowAnnotation(
  "Median export flux\n[mmol/gDWh]" = 
    anno_barplot(x = fba_median_pos[rownames(reorder_mat)],
                 gp=gpar(fill="gray80"),
                 axis_param = list(labels_rot = 90,
                                   gp = gpar(fontsize = 6),
                                   direction="reverse"
                                   )),
    "Median import flux\n[mmol/gDWh]" = 
    anno_barplot(x = fba_median_neg[rownames(reorder_mat)],
                gp=gpar(fill="gray80"),
                axis_param = list(labels_rot = 90,
                                  gp = gpar(fontsize = 6),
                                  direction="reverse"
                                  )),
  spacer=anno_empty(border = F, width = unit(1, "mm")),

  show_legend = F,
  annotation_name_rot = 90,
  annotation_name_side = "top",
  annotation_name_gp = gpar(fontsize = 10),
  annotation_name_offset = unit(0.1, "cm"),
  gap=unit(2,"mm")
  )


#reorder_mat<-order_mat[,(metab_cat$name)]
dim(order_mat[, names(metab_class_vec)])
Heatmap(
  reorder_mat,
  col=flux_colors, na_col = "white",
  rect_gp = gpar(col = "gray20", lwd = 0.3),
  border_gp = gpar(col = "black", lwd = 0.7),
  row_gap= unit(3, "mm"),
  column_gap= unit(3, "mm"),
  row_split = metab_class_vec,
  column_title_rot = 45,
  column_title_gp = gpar(fontsize=13),
  column_names_rot = 45,
  column_names_gp  = gpar(fontsize=13),
  column_title_side = "bottom",
  column_split = phy_vec[colnames(reorder_mat)],
  column_names_side = "top",
  row_title_rot=0,
  row_names_gp = gpar(fontsize=10),
  row_title_gp  = gpar(fontsize=13),
  border = T,
  cluster_rows = T,
  cluster_columns = T,
  cluster_column_slices = F,
  cluster_row_slices = F,
  show_column_dend = F,
  show_row_dend = F,
  heatmap_height = unit(19, "in"),
  #heatmap_width = unit(14, "in"),
  show_heatmap_legend = F,
  right_annotation = right_annot,
  #top_annotation = bottom_annot,
  #bottom_annotation = top_annot
  #left_annotation = left_annot
) ->hm_ex
png(filename = 
      paste0(
      "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/hm_exchanges_test_v4_metab_med_0.1-10.png"),
    width = 10, height = 20, units = "in", res = 300)
draw(hm_ex, heatmap_legend_side = "bottom", annotation_legend_side = "bottom")
dev.off()

file.show("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/hm_exchanges_test_v4_metab_med_0.1-10.png")

```




##Heatmap: MetaB frequency utilisation by order v2
```{r eval=FALSE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
tidy_exhanges %>%
  group_by(name,simple_category) %>%
  summarise(median_pos=median(fba_flux[fba_flux>0], na.rm = T),
            median_neg=median(fba_flux[fba_flux<0], na.rm = T),
            prop_pos=round(100*(sum(fba_flux[fba_flux>0], na.rm = T)/311),2),
            prop_neg=round(100*(sum(fba_flux[fba_flux<0], na.rm = T)/311),2)
            ) %>%
  arrange(desc(median_pos)) -> exchanges_summary

#####  HM O2
require(ComplexHeatmap)
tidy_exhanges<-fba_comp_metab1

# vOTU abundance by host order
votu_host_order_abund<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/votu_host_order_mean.csv")
votu_host_order_abund$Order = str_remove(votu_host_order_abund$Order,"o__")
votu_host_order_abund_vec<-votu_host_order_abund$mean
names(votu_host_order_abund_vec) = votu_host_order_abund$Order


# Select orders
# Select orders
c("Bacteroidales", "Rhizobiales", "Pseudomonadales",  "Rhodobacterales", "Chromatiales", "Flavobacteriales", "Desulfobacterales", "Desulfobulbales", "Cytophagales",  "Woeseiales", "JAAYZQ01", "SZUA-229", "Caldilineales"
             ) -> sel_orders


## Wider alternative
votu_order_count_all %>%
  filter(n_votu>=2) -> order_min2_votu

# Select which categories to plot
unique(tidy_exhanges$simple_category)
selected_classes<-c(
    "Polysaccharides", "Carbohydrates", "Sugar Alcohols",
    "Alcohols", "Carbonyl compounds",
    "Sulfur compounds",
    "N-Carbonyl compounds","Amino acids", "Nitrogen compounds", "Inorganic compounds")

# Heatmap for substrates
tax_lev="Order"
tidy_exhanges[Order %in% sel_orders] %>%
  filter((fba_flux)<(-0.001)) %>%
  group_by(name, Order, simple_category) %>%
  summarise(n_flux=n()) %>%
  left_join(order_nmags[,c("Order", "n")]) %>%
  mutate(prop=n_flux/n) %>%
  filter(prop>0.50) -> freq_exc

  freq_exc %>%
    select(name, prop) %>%
  pivot_wider(names_from = name, values_from = prop, values_fill = 0) %>%
  column_to_rownames("Order") %>%
  as.matrix  -> order_mat

tidy_exhanges[name %in% freq_exc$name,c("name", "simple_category")] %>%
    filter(simple_category %in% selected_classes) %>%
  unique %>% filter(name %in% colnames(order_mat)) %>%
  mutate(simple_category=
           factor(
           simple_category,
         levels = 
          selected_classes)) %>% arrange(simple_category) -> metab_cat

metab_class_vec=metab_cat$simple_category
names(metab_class_vec)=metab_cat$name


# Define splitting vectors
gem_analysis$Phylum %>% as.factor->phy_vec
names(phy_vec)=gem_analysis %>% pull(!!sym(tax_lev))

# Transpose matrix
reorder_mat<-order_mat %>% t
reorder_mat<-reorder_mat[names(metab_class_vec),]

# Right annot
rownames(reorder_mat)%in%
metab_react_matched_cpds_mean_scaled[, name] %>% as.character->metaB_presence_vec
names(metaB_presence_vec)=rownames(reorder_mat)


right_annot<-
rowAnnotation(
  #spacer=anno_empty(border = F, height = unit(0.5, "mm")),
  " "=metaB_presence_vec[rownames(reorder_mat)],
  annotation_name_side = "bottom",
  col=list(" "=c("FALSE"="white","TRUE"="gray80")),
  border=T,
  show_legend = F)

#reorder_mat<-order_mat[,(metab_cat$name)]
dim(order_mat[, names(metab_class_vec)])
Heatmap(
  reorder_mat,
  col=c("white","white", "skyblue", "cadetblue"), na_col = "white",
  rect_gp = gpar(col = "gray20", lwd = 0.3),
  border_gp = gpar(col = "black", lwd = 0.7),
  row_gap= unit(3, "mm"),
  column_gap= unit(3, "mm"),
  row_split = metab_class_vec,
  column_title_rot = 45,
  column_title_gp = gpar(fontsize=0),
  column_names_rot = 45,
  column_names_gp  = gpar(fontsize=13),
  column_title_side = "bottom",
  column_split = phy_vec[colnames(reorder_mat)],
  column_names_side = "top",
  row_title_rot=0,
  row_names_gp = gpar(fontsize=10),
  row_title_gp  = gpar(fontsize=13),
  border = T,
  cluster_rows = T,
  cluster_columns = F,
  cluster_column_slices = F,
  cluster_row_slices = F,
  show_column_dend = F,
  show_row_dend = F,
  #heatmap_height = unit(10, "in"),
  #heatmap_width = unit(14, "in"),
  show_heatmap_legend = T,
  right_annotation = right_annot,
  width           = ncol(reorder_mat) * 1,   # fixed: same cols in every slice
  height          = nrow(reorder_mat) * 1,
  heatmap_legend_param =list(title = "% of GEMs\n with exchange",
                             direction = "horizontal",
                             border=T,
                             title_position = "topcenter",
                             legend_height = unit(8, "cm"),
                             grid_width    = unit(16, "mm"),
                             title_gp      = gpar(fontsize = 14, fontface = "bold"),
                             labels_gp     = gpar(fontsize = 12)
                             
                             )
) ->hm_ex_sub


# Heatmap for products
tax_lev="Order"
tidy_exhanges[Order %in% sel_orders] %>%
  filter((fba_flux)>(0.001)) %>%
  group_by(name, Order, simple_category) %>%
  summarise(n_flux=n()) %>%
  left_join(order_nmags[,c("Order", "n")]) %>%
  mutate(prop=n_flux/n) %>%
  filter(prop>0.50) -> freq_exc

  freq_exc %>%
    select(name, prop) %>%
  pivot_wider(names_from = name, values_from = prop, values_fill = 0) %>%
  column_to_rownames("Order") %>%
  as.matrix  -> order_mat

tidy_exhanges[name %in% freq_exc$name,c("name", "simple_category")] %>%
    filter(simple_category %in% selected_classes) %>%
  unique %>% filter(name %in% colnames(order_mat)) %>%
  mutate(simple_category=
           factor(
           simple_category,
         levels = 
          selected_classes)) %>% arrange(simple_category) -> metab_cat

metab_class_vec=metab_cat$simple_category
names(metab_class_vec)=metab_cat$name


# Define splitting vectors
gem_analysis$Phylum %>% as.factor->phy_vec
names(phy_vec)=gem_analysis %>% pull(!!sym(tax_lev))

# Transpose matrix
reorder_mat<-order_mat %>% t
reorder_mat<-reorder_mat[names(metab_class_vec),]

# Right annot
rownames(reorder_mat)%in%
metab_react_matched_cpds_mean_scaled[, name] %>% as.character->metaB_presence_vec
names(metaB_presence_vec)=rownames(reorder_mat)


right_annot<-
rowAnnotation(
  #spacer=anno_empty(border = F, height = unit(0.5, "mm")),
  "Detected metabolites"=metaB_presence_vec[rownames(reorder_mat)],
  annotation_name_side = "bottom",
  col=list("Detected metabolites"=c("FALSE"="white","TRUE"="gray80")),
  border=T,
  show_legend = F,
  annotation_name_rot  = 45)

#reorder_mat<-order_mat[,(metab_cat$name)]
dim(order_mat[, names(metab_class_vec)])
Heatmap(
  reorder_mat,
  col=c("white", "white","tan1", "red"), na_col = "white",
  rect_gp = gpar(col = "gray20", lwd = 0.3),
  border_gp = gpar(col = "black", lwd = 0.7),
  row_gap= unit(3, "mm"),
  column_gap= unit(3, "mm"),
  row_split = metab_class_vec,
  column_title_rot = 45,
  column_title_gp = gpar(fontsize=13),
  column_names_rot = 45,
  column_names_gp  = gpar(fontsize=13),
  column_title_side = "bottom",
  column_split = phy_vec[colnames(reorder_mat)],
  column_names_side = "top",
  row_title_rot=0,
  row_names_gp = gpar(fontsize=10),
  row_title_gp  = gpar(fontsize=13),
  border = T,
  cluster_rows = T,
  cluster_columns = F,
  cluster_column_slices = F,
  cluster_row_slices = F,
  show_column_dend = F,
  show_row_dend = F,
  #heatmap_height = unit(10, "in"),
  #heatmap_width = unit(14, "in"),
  show_heatmap_legend = F,
  right_annotation = right_annot,
  width           = ncol(reorder_mat) * 1,   # fixed: same cols in every slice
  height          = nrow(reorder_mat) * 1
) ->hm_ex_prod




png(filename = 
      paste0(
      "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/hm_exchanges_test_v4_metab_med_0.1-10.png"),
    width = 8, height = 15, units = "in", res = 600)
draw(hm_ex_sub %v% hm_ex_prod,
     heatmap_legend_side = "bottom",
     padding = unit(c(15, 5, 10, 5), "mm")
     )
dev.off()

file.show("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/hm_exchanges_test_v4_metab_med_0.1-10.png")

```


## PWYs by key categories
```{r eval=FALSE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
# Import pwy data
files<-list.files(path = "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaG/coassemby_07feb24/mags_c70c10a99/gapseq/",
                  pattern ="-all-Pathways.tbl", full.names = T)
foreach(i = files)%do%{
fread(i) %>%
  filter(Prediction) %>% 
  mutate(sbs_ids = str_remove_all(ID, "\\|")) %>%
  left_join(sbs_hier1) %>%
    add_column(OTU=str_extract(i, "[:alpha:]{3}_SemiBin_[:digit:]+"))
  } %>% rbindlist -> pws

# Number of unique pws
pws$ID %>% unique %>% length()
tidy_exhanges$name %>% unique %>% length

pws %>%
  select(OTU, Name, h2, h3, h4, h5, ReactionsFound) %>%
  left_join(gem_analysis[,c("OTU", "Order")]) %>%
  rowwise %>%
  mutate(n_react=length(unlist(strsplit(ReactionsFound, " ")))) %>%
  data.table -> pws1


# vOTU abundance by host order
votu_host_order_abund<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/votu_host_order_mean.csv")
votu_host_order_abund$Order = str_remove(votu_host_order_abund$Order,"o__")
votu_host_order_abund_vec<-votu_host_order_abund$mean
names(votu_host_order_abund_vec) = votu_host_order_abund$Order

# Select thos with at least two votus
votu_order_count_all %>%
  filter(n_votu>=2) %>% arrange(n_votu) -> order_min2_votu


# Select relevant pwys and summarise
pws1$h4 %>% unique
pws1[h3 == "CARBOXYLATES-DEG", Name] %>% unique %>% view
pws1[grepl("Xylan|Inulin|Pectin|Cellulose|Starch", Name, ignore.case = T) &
     !grepl("Biosynthesis", Name, ignore.case = T)] %>% add_column(type="Polysaccharide \ndegradation") -> df1

pws1[grepl("Calvin|Wood|Reductive TCA|3-hydroxypropanoate cycle", Name)] %>% add_column(type="CO2 fixation") -> df2

pws1[grepl("lactate ferment|pyruvate fermentation to acetate|ethanol|fermentation to butanoate|	
propanoate fermentation|acetate conversion", Name, ignore.case = F) & 
       h4 == c("Fermentation-to-Acids")] %>% add_column(type="Fermentation") -> df3

pws1[grepl("glyoxylate cycle", Name, ignore.case = F)] %>% add_column(type="Energy metabolism") -> df4

pws1[grepl("Tyrosine|Arginine|histidine", Name, ignore.case = T) & 
     !grepl("arginine decarboxylase", Name, ignore.case = T) &
     h4 %in% c("Proteinogenic-Amino-Acids-Degradation") | 
     h4 == "4-Aminobutyraye-Degradation"] %>% 
  add_column(type="Proteinogenic \namino acid \ndegradation") -> df5

pws1[h4 %in% c("Purine-Degradation", "Pyrimidine-Degradation") &
     !grepl("L-arginine degradation IV", Name, ignore.case = T)] %>% add_column(type="Nucleotide degradation")-> df6

pws1[grepl("trehalose|xylose|fucose|melibiose|lactose|cellobiose", Name, ignore.case = T) & 
      h4 %in% c("Sugars-And-Polysaccharides-Degradation") &
     !grepl("periplasmic|cytosolic|trehalose degradation IV", Name, ignore.case = T)
     ] %>% 
  add_column(type="Carbohydrate degradation")-> df7

rbind(df1, df2) %>%
  rbind(df3) %>%
  rbind(df4) %>%
  rbind(df5) %>%
  rbind(df6) %>%
  rbind(df7) %>%
    filter(Order %in% order_min2_votu$Order)-> df

df$type %>% unique

df %>%
  group_by(Order, Name, type) %>% 
  summarise(nmag = length(unique(OTU))) %>% 
  left_join(order_nmags, by= "Order") %>%
  mutate(p=100*(nmag/n)) %>% select(Order, Name, type) %>%
  unique %>% 
  mutate(p=case_when(type=="Polysaccharide \ndegradation"~ 1,
                     type=="CO2 fixation"~ 2,
                     type=="Fermentation"~ 3,
                     type=="Energy metabolism"~4,
                     type=="Nucleotide degradation"~ 5,
                     type=="Proteinogenic \namino acid \ndegradation"~ 6,
                     type=="Carbohydrate degradation" ~ 7
                     )) %>% select(Order, Name, p) -> pwy_count
  

pwy_count %>% pivot_wider(names_from = Order, values_from = p, values_fill = 0) %>% 
  column_to_rownames("Name")  %>% as.matrix -> mat

# Define splitting vectors columns
gem_analysis$Phylum %>% as.factor->phy_vec
names(phy_vec)=gem_analysis %>% pull(Order)

# Define row splits
df %>% select(Name, type) %>% unique -> rsplit
rsplit$type %<>% factor(levels = c(
                         "Polysaccharide \ndegradation",
                         "Carbohydrate degradation",
                         "CO2 fixation",
                         "Fermentation",
                         "Energy metabolism",
                         "Nucleotide degradation",
                         "Proteinogenic \namino acid \ndegradation"
                       ))
rsplit$type -> rsplit_vec
names(rsplit_vec)=rsplit %>% pull(Name)

# Top annot
HeatmapAnnotation(
  "MAG phylum" = phy_vec[colnames(mat)],
  which = "column",
  col = list("MAG phylum"=tree_colors),
  border = T,  show_legend = F,
  show_annotation_name = F,
    annotation_name_side = "right"
)-> top_annot



require(ComplexHeatmap)

library(circlize)
col_fun <- colorRamp2(
  breaks = c(0, 1, 2, 3, 4, 5, 6, 7),
  colors = c(
    "white",                  
    "green4",
    "gray50",
    "orange",
    "lightgoldenrod1",
    "lightblue",
    "lightcoral",
    "olivedrab3"
  ),transparency = 0.3
)

Heatmap(
  mat[rsplit$Name,],
  name="xxx",
  col = col_fun,
  rect_gp = gpar(col = "gray20", lwd = 0.5),
  border_gp = gpar(col = "black", lwd = 0.9),
  row_gap= unit(3, "mm"),
  column_gap= unit(3, "mm"),
  row_split = rsplit_vec[],
  column_title_rot = 45,
  column_names_rot = 45,
  #column_names_gp  = gpar(fontsize=2),
  column_split = phy_vec[colnames(mat)],
  column_title_side = "bottom",
  #column_title_gp = gpar(fontsize=1)
  #row_split = aux_gem_annot_vec[rownames(mag_vir_hm)],
  row_title_rot=0, 
  #row_title_gp = gpar(fontsize=1),
  row_names_gp = gpar(fontsize=9),
  column_names_side = "top",
  border = T,
  cluster_rows = T,
  cluster_columns = T,
  cluster_column_slices = F,
  cluster_row_slices = F,
  show_column_dend = F,
  show_row_dend = F,
  #heatmap_height = unit(3, "in"),
  heatmap_width = unit(11, "in"),
  show_heatmap_legend = F,
  top_annotation = top_annot
) -> hm;hm

png(filename = 
      paste0(
      "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/hm_pws.png"),
    width = 16, height = 12, units = "in", res = 600)
draw(hm)
dev.off()

file.show("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/hm_pws.png")

cairo_pdf(file = 
      paste0(
      "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/hm_pws.pdf"),
    width = 16, height = 12)
draw(hm)
dev.off()

file.show("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/hm_pws.pdf")
```


## Heatmap: PWYs by key categories v2
```{r eval=FALSE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
# Import pwy data
pwy_processes<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/pathway_process_heatmap_long_reduced.csv") %>% data.table()

pwy_processes[, c("pathway", "process")] %>% unique -> pwy_processes_u

# Select orders
c("Bacteroidales", "Flavobacteriales", "Cytophagales", "Desulfobulbales", "Desulfobacterales",
  "Rhodobacterales", "Pseudomonadales", "Chromatiales", "Caldilineales",
  "Rhizobiales", "Woeseiales", "JAAYZQ01", "SZUA-229"
             ) -> sel_orders

## Wider alternative
votu_order_count_all %>%
  filter(n_votu>=2) -> order_min2_votu


files<-list.files(path = "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/metaG/coassemby_07feb24/mags_c70c10a99/gapseq/",
                  pattern ="-all-Pathways.tbl", full.names = T)
foreach(i = files)%do%{
fread(i) %>%
  filter(Prediction) %>% 
  mutate(sbs_ids = str_remove_all(ID, "\\|")) %>%
  left_join(sbs_hier1) %>%
    add_column(OTU=str_extract(i, "[:alpha:]{3}_SemiBin_[:digit:]+"))
  } %>% rbindlist -> pws

pws %>%
  select(OTU, Name, h2, h3, h4, h5, ReactionsFound) %>%
  left_join(gem_analysis[,c("OTU", "Order")]) %>%
  rowwise %>%
  mutate(n_react=length(unlist(strsplit(ReactionsFound, " ")))) %>%
  data.table -> pws1


# vOTU abundance by host order
votu_host_order_abund<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/votu_host_order_mean.csv")
votu_host_order_abund$Order = str_remove(votu_host_order_abund$Order,"o__")
votu_host_order_abund_vec<-votu_host_order_abund$mean
names(votu_host_order_abund_vec) = votu_host_order_abund$Order

# Impoort manually curated pwys
selected_pwys_cat<-fread("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/tables/pwy_selectio_hm_18mag26 - pathway_classifiation (3).csv") %>% data.table

selected_pwys_cat$cat%>%unique
selected_pwys_cat$Name%>%unique

# Define splitting vectors columns
gem_analysis$Phylum %>% as.factor->phy_vec
names(phy_vec)=gem_analysis %>% pull(Order)


# Check pwy distribution by categories
pws1 %>%
  filter(Order %in% sel_orders) %>%
  filter(Name %in% pwy_fish_ehnr_h$Name) %>% 
  group_by(Order, h2, Name) %>% 
  summarise(nmag = length(unique(OTU))) -> enh_pwy_summary

pws1 %>%
  filter(Order %in% sel_orders) %>%
  filter(Name %in% selected_pwys_cat$Name) %>% 
  group_by(Order, h2, Name) %>% 
  summarise(nmag = length(unique(OTU))) -> sel_pwy_summary

enh_pwy_summary %>%
  group_by(h2, Name) %>%
  summarise(nmags=sum(nmag)) %>% 
  mutate(prop=
           100*(nmags/nrow(gem_analysis[Order %in% sel_orders]))
         ) -> enh_pwy_summary1
enh_pwy_summary1$h2 %>% table

sel_pwy_summary %>%
  group_by(h2, Name) %>%
  summarise(nmags=sum(nmag)) %>% 
  mutate(
    prop=100*(nmags/nrow(gem_analysis[Order %in% sel_orders]))
         )-> sel_pwy_summary1


# Sulphur hm --------------------------------------------------------

selected_pwys_cat[cat == "Sulphur cycling"]-> rsplit_sel

# Define row splits
rsplit_sel$type %<>% factor(levels = unique(rsplit_sel$type))
rsplit_sel$type -> rsplit_vec
names(rsplit_vec)=rsplit_sel %>% pull(Name)


# Prepare frequency matrix

pws1 %>%
  filter(Order %in% sel_orders) %>%
  filter(Name %in% rsplit_sel$Name) %>% 
  group_by(Order, Name) %>% 
  summarise(nmag = length(unique(OTU))) %>%
  ungroup() %>%
  complete(Order, Name, fill = list(nmag = 0)) %>%   # keeps all orders
  left_join(order_nmags, by = "Order") %>%
  mutate(p = 100 * (nmag / n)) %>%
  select(Order, Name, p) %>%
  unique() %>% 
  mutate(p = case_when(p < 25 ~ 0, .default = p)) -> pwy_count
  


pwy_count %>% pivot_wider(names_from = Order, values_from = p, values_fill = 0) %>% 
  column_to_rownames("Name")  %>% as.matrix -> mat

length(rsplit_vec)
nrow(mat)

# Top annot
HeatmapAnnotation(
  "MAG phylum" = phy_vec[colnames(mat)],
  which = "column",
  col = list("MAG phylum"=tree_colors),
  border = T,  show_legend = F,
  show_annotation_name = F,
    annotation_name_side = "right"
)-> top_annot

c("white",                  
    "green4",
    "gray50",
    "orange",
    "lightgoldenrod1",
    "lightblue",
    "lightcoral",
    "olivedrab3")


require(ComplexHeatmap)

Heatmap(
  mat[],
  name="xxx",
  col = c("white", "white", "tan1", "orange"),
  rect_gp = gpar(col = "gray20", lwd = 0.5),
  border_gp = gpar(col = "black", lwd = 0.9),
  row_gap= unit(3, "mm"),
  column_gap= unit(3, "mm"),
  row_split = rsplit_vec[rownames(mat)],
  column_title_rot = 45,
  column_names_rot = 45,
  #column_names_gp  = gpar(fontsize=2),
  column_split = phy_vec[colnames(mat)],
  column_title_side = "bottom",
  column_title_gp = gpar(fontsize=0),
  #row_split = aux_gem_annot_vec[rownames(mag_vir_hm)],
  row_title_rot=0, 
  #row_title_gp = gpar(fontsize=1),
  row_names_gp = gpar(fontsize=12),
  column_names_side = "top",
  border = T,
  cluster_rows = T,
  cluster_columns = F,
  cluster_column_slices = F,
  cluster_row_slices = F,
  show_column_dend = F,
  show_row_dend = F,
  #heatmap_height = unit(3, "in"),
  heatmap_width = unit(10, "in"),
  show_heatmap_legend = T,
  heatmap_legend_param =list(title = "% of GEMs\nwith pathway",
                             direction = "horizontal",
                             border=T,
                             title_position = "topcenter",
                             legend_height = unit(8, "cm"),
                             grid_width    = unit(16, "mm"),
                             title_gp      = gpar(fontsize = 14, fontface = "bold"),
                             labels_gp     = gpar(fontsize = 12)
                             
                             ),
  top_annotation = NULL,
  width           = ncol(mat) * 1,   # fixed: same cols in every slice
  height          = nrow(mat) * 1  
) -> hm1;hm1


# Nitrogen hm --------------------------------------------------------
selected_pwys_cat[cat == "Nitrogen mineralisation"]-> rsplit_sel

# Define row splits
rsplit_sel$type %<>% factor(levels = unique(rsplit_sel$type))
rsplit_sel$type -> rsplit_vec
names(rsplit_vec)=rsplit_sel %>% pull(Name)


# Prepare frequency matrix

pws1 %>%
  filter(Order %in% sel_orders) %>%
  filter(Name %in% rsplit_sel$Name) %>%
  group_by(Order, Name) %>% 
  summarise(nmag = length(unique(OTU))) %>%
  ungroup() %>%
  complete(Order, Name, fill = list(nmag = 0)) %>%   # keeps all orders
  left_join(order_nmags, by = "Order") %>%
  mutate(p = 100 * (nmag / n)) %>%
  select(Order, Name, p) %>%
  unique() %>%
  mutate(p = case_when(p < 25 ~ 0, .default = p)) -> pwy_count


pwy_count %>% pivot_wider(names_from = Order, values_from = p, values_fill = 0) %>% 
  column_to_rownames("Name")  %>% as.matrix -> mat

length(rsplit_vec)
nrow(mat)

# Top annot
HeatmapAnnotation(
  "MAG phylum" = phy_vec[colnames(mat)],
  which = "column",
  col = list("MAG phylum"=tree_colors),
  border = T,  show_legend = F,
  show_annotation_name = F,
    annotation_name_side = "right"
)-> top_annot

require(ComplexHeatmap)

Heatmap(
  mat[],
  name="xxx",
  col = c("white", "white", "salmon", "red"),
  rect_gp = gpar(col = "gray20", lwd = 0.5),
  border_gp = gpar(col = "black", lwd = 0.9),
  row_gap= unit(3, "mm"),
  column_gap= unit(3, "mm"),
  row_split = rsplit_vec[rownames(mat)],
  column_title_rot = 45,
  column_names_rot = 45,
  #column_names_gp  = gpar(fontsize=2),
  column_split = phy_vec[colnames(mat)],
  column_title_side = "bottom",
  column_title_gp = gpar(fontsize=0),
  #row_split = aux_gem_annot_vec[rownames(mag_vir_hm)],
  row_title_rot=0, 
  #row_title_gp = gpar(fontsize=1),
  row_names_gp = gpar(fontsize=12),
  column_names_side = "top",
  border = T,
  cluster_rows = T,
  cluster_columns = F,
  cluster_column_slices = F,
  cluster_row_slices = F,
  show_column_dend = F,
  show_row_dend = F,
  #heatmap_height = unit(3, "in"),
  heatmap_width = unit(10, "in"),
  show_heatmap_legend = F,
  heatmap_legend_param =list(title = "% of GEMs\nwith pathway",
                             direction = "horizontal",
                             border=T,
                             title_position = "topcenter",
                             legend_height = unit(8, "cm"),
                             grid_width    = unit(16, "mm"),
                             title_gp      = gpar(fontsize = 14, fontface = "bold"),
                             labels_gp     = gpar(fontsize = 12)
                             
                             ),
  top_annotation = NULL,
  width           = ncol(mat) * 1,   # fixed: same cols in every slice
  height          = nrow(mat) * 1  
) -> hm2;hm2

# Polysach and carbo hm --------------------------------------------------------
selected_pwys_cat[cat == "Carbohydrates degradation"]-> rsplit_sel

# Define row splits
rsplit_sel$type %<>% factor(levels = unique(rsplit_sel$type))
rsplit_sel$type -> rsplit_vec
names(rsplit_vec)=rsplit_sel %>% pull(Name)


# Prepare frequency matrix
pws1 %>%
  filter(Order %in% sel_orders) %>%
  filter(Name %in% rsplit_sel$Name) %>%
  group_by(Order, Name) %>% 
  summarise(nmag = length(unique(OTU))) %>%
  ungroup() %>%
  complete(Order, Name, fill = list(nmag = 0)) %>%   # keeps all orders
  left_join(order_nmags, by = "Order") %>%
  mutate(p = 100 * (nmag / n)) %>%
  select(Order, Name, p) %>%
  unique() %>%
  mutate(p = case_when(p < 25 ~ 0, .default = p)) -> pwy_count
  


pwy_count %>% pivot_wider(names_from = Order, values_from = p, values_fill = 0) %>% 
  column_to_rownames("Name")  %>% as.matrix -> mat

length(rsplit_vec)
nrow(mat)

# Top annot
HeatmapAnnotation(
  "MAG phylum" = phy_vec[colnames(mat)],
  which = "column",
  col = list("MAG phylum"=tree_colors),
  border = T,  show_legend = F,
  show_annotation_name = F,
    annotation_name_side = "right"
)-> top_annot

require(ComplexHeatmap)

Heatmap(
  mat[],
  name="xxx",
  col = c("white", "white", "olivedrab3", "darkgreen"),
  rect_gp = gpar(col = "gray20", lwd = 0.5),
  border_gp = gpar(col = "black", lwd = 0.9),
  row_gap= unit(3, "mm"),
  column_gap= unit(3, "mm"),
  row_split = rsplit_vec[rownames(mat)],
  column_title_rot = 45,
  column_names_rot = 45,
  #column_names_gp  = gpar(fontsize=2),
  column_split = phy_vec[colnames(mat)],
  column_title_side = "bottom",
  column_title_gp = gpar(fontsize=0),
  #row_split = aux_gem_annot_vec[rownames(mag_vir_hm)],
  row_title_rot=0, 
  #row_title_gp = gpar(fontsize=1),
  row_names_gp = gpar(fontsize=12),
  column_names_side = "top",
  border = T,
  cluster_rows = T,
  cluster_columns = F,
  cluster_column_slices = F,
  cluster_row_slices = F,
  show_column_dend = F,
  show_row_dend = F,
  #heatmap_height = unit(3, "in"),
  heatmap_width = unit(10, "in"),
  show_heatmap_legend = F,
  heatmap_legend_param =list(title = "% of GEMs\nwith pathway",
                             direction = "horizontal",
                             border=T,
                             title_position = "topcenter",
                             legend_height = unit(8, "cm"),
                             grid_width    = unit(16, "mm"),
                             title_gp      = gpar(fontsize = 14, fontface = "bold"),
                             labels_gp     = gpar(fontsize = 12)
                             
                             ),
  top_annotation = NULL,
  width           = ncol(mat) * 1,   # fixed: same cols in every slice
  height          = nrow(mat) * 1  
) -> hm3;hm3

# Fermentation --------------------------------------------------------
selected_pwys_cat[cat %in% c("Fermentation", "Carboxylate degradation")]-> rsplit_sel


# Define row splits
rsplit_sel$type %<>% factor(levels = unique(rsplit_sel$type))
rsplit_sel$type -> rsplit_vec
names(rsplit_vec)=rsplit_sel %>% pull(Name)


# Prepare frequency matrix
pws1 %>%
  filter(Order %in% sel_orders) %>%
  filter(Name %in% rsplit_sel$Name) %>%
  group_by(Order, Name) %>% 
  summarise(nmag = length(unique(OTU))) %>%
  ungroup() %>%
  complete(Order, Name, fill = list(nmag = 0)) %>%   # keeps all orders
  left_join(order_nmags, by = "Order") %>%
  mutate(p = 100 * (nmag / n)) %>%
  select(Order, Name, p) %>%
  unique() %>%
  mutate(p = case_when(p < 25 ~ 0, .default = p)) -> pwy_count
  


pwy_count %>% pivot_wider(names_from = Order, values_from = p, values_fill = 0) %>% 
  column_to_rownames("Name")  %>% as.matrix -> mat

length(rsplit_vec)
nrow(mat)

# Top annot
HeatmapAnnotation(
  "MAG phylum" = phy_vec[colnames(mat)],
  which = "column",
  col = list("MAG phylum"=tree_colors),
  border = T,  show_legend = F,
  show_annotation_name = F,
    annotation_name_side = "right"
)-> top_annot

require(ComplexHeatmap)

Heatmap(
  mat[],
  name="xxx",
  col = c("white", "white", "skyblue", "turquoise4"),
  rect_gp = gpar(col = "gray20", lwd = 0.5),
  border_gp = gpar(col = "black", lwd = 0.9),
  row_gap= unit(3, "mm"),
  column_gap= unit(3, "mm"),
  row_split = rsplit_vec[rownames(mat)],
  column_title_rot = 45,
  column_names_rot = 45,
  #column_names_gp  = gpar(fontsize=2),
  column_split = phy_vec[colnames(mat)],
  column_title_side = "bottom",
  column_title_gp = gpar(fontsize=12),
  #row_split = aux_gem_annot_vec[rownames(mag_vir_hm)],
  row_title_rot=0, 
  #row_title_gp = gpar(fontsize=1),
  row_names_gp = gpar(fontsize=12),
  column_names_side = "top",
  border = T,
  cluster_rows = T,
  cluster_columns = F,
  cluster_column_slices = F,
  cluster_row_slices = F,
  show_column_dend = F,
  show_row_dend = F,
  #heatmap_height = unit(3, "in"),
  heatmap_width = unit(10, "in"),
  show_heatmap_legend = F,
  heatmap_legend_param =list(title = "% of GEMs\nwith pathway",
                             direction = "horizontal",
                             border=T,
                             title_position = "topcenter",
                             legend_height = unit(8, "cm"),
                             grid_width    = unit(16, "mm"),
                             title_gp      = gpar(fontsize = 14, fontface = "bold"),
                             labels_gp     = gpar(fontsize = 12)
                             
                             ),
  top_annotation = NULL,
  width           = ncol(mat) * 1,   # fixed: same cols in every slice
  height          = nrow(mat) * 1   
) -> hm4;hm4




png(filename = 
      paste0(
      "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/hm_pws_pca.png"),
    width = 16, height = 15, units = "in"
    , res = 300)

draw(hm3 %v% hm2 %v% hm4 %v% hm1,
     heatmap_legend_side = "bottom",
     padding = unit(c(2, 2, 15, 2), "mm"))

dev.off()

file.show("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/hm_pws_pca.png")

```

## PWY PCoA
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
# Select orders
c("Bacteroidales", "Rhizobiales", "Pseudomonadales",  "Rhodobacterales", "Chromatiales", "Flavobacteriales", "Desulfobacterales", "Desulfobulbales", "Cytophagales",  "Woeseiales", "JAAYZQ01", "SZUA-229", "Caldilineales"
             ) -> sel_orders

# Select orders
#c("Bacteroidales", "Rhizobiales", "Pseudomonadales",  "Rhodobacterales", "Chromatiales", "Flavobacteriales", #"Desulfobacterales","Desulfobulbales",  "Woeseiales", "SZUA-229"
#             ) -> sel_orders

## Wider alternative
votu_order_count_all %>%
  filter(n_votu>=2)-> order_min2_votu

# Select pathways
selected_pwys_cat$Name %>% unique -> pws_categories

pws1 %>% 
    group_by(OTU, Name) %>%
    filter(Order %in% sel_orders) %>%
    filter(Name %in% pwy_fish_ehnr$Name) %>%
    count() %>%
  pivot_wider(names_from = OTU, values_from = n, values_fill = 0) %>% 
  column_to_rownames("Name") %>% as.matrix()  %>% t -> mat

# Simiplify names
gsub("dimethylsulfoniopropanoate ", "DMSP ", colnames(mat)) -> colnames(mat)
gsub("dimethylsulfoniopropanoate ", "DMSP ", pws_categories) -> pws_categories1
gsub("degradation", "deg.", colnames(mat)) -> colnames(mat)
gsub("degradation", "deg.", pws_categories1) -> pws_categories1
mat[mat>1]=1

order_vec<-gem_analysis[OTU %in% rownames(mat), Order]
phy_vec<-gem_analysis[OTU %in% rownames(mat), Phylum]

names(order_vec)<-rownames(mat)
ord_cols<-hcl.colors(palette = "Zissou 1", n = length(unique(order_vec)))
names(ord_cols) = unique(order_vec)

ord_cols <- c(
  "Bacteroidales"     = "green4",
   "Flavobacteriales"  = "aquamarine4",
  "Cytophagales"      = "springgreen",

  "Chromatiales"      = "#40ABA5",
  "SZUA-229"          = "cyan1",
  
  "JAAYZQ01"          = "black",
  "Caldilineales"     = "gray",

  "Desulfobacterales" = "gold",
  "Desulfobulbales"   = "olivedrab2",
  
  "Rhodobacterales"   = "mediumvioletred",
  "Woeseiales"        = "orange",
  "Rhizobiales"       = "#F5191C",
  "Pseudomonadales"   = "lightsalmon"
)



require(factoextra);require(FactoMineR)

library(vegan)
library(ape)

# 1. Jaccard distance on binary matrix
jac_dist <- vegdist(mat, method = "jaccard")

# 2. PCoA
pcoa_res <- pcoa(jac_dist)

ncomp <- 10  # axes to retain

# 3. Sample scores
scores <- pcoa_res$vectors[, 1:ncomp]

# 4. Project variables onto PCoA axes via correlation
loadings <- cor(mat, scores)
loadings %>% view
# 5. Positive eigenvalues only
eig_all_pos <- pcoa_res$values$Eigenvalues[pcoa_res$values$Eigenvalues > 0]
eig_retained <- eig_all_pos[1:ncomp]

# 6. Build prcomp-compatible object
pca_like <- list(
  x        = scores,
  rotation = loadings,
  sdev     = sqrt(eig_all_pos),   # all positive — fixes denominator
  center   = colMeans(mat),
  scale    = FALSE
)
class(pca_like) <- "prcomp"


fviz_pca_ind(pca_like,
                # --- samples ---
                geom.ind      = "point",
                pointshape    = 19,
                col.ind       = order_vec,          # color by group
                palette       = ord_cols,
                addEllipses   = T,            # confidence ellipses
                ellipse.type  = "convex",    # or "convex" / "norm" / "t"
                ellipse.level = 0.95)  +
  theme_bw() +
  guides(color = guide_legend(ncol = 2),
         fill  = guide_legend(ncol = 2)) +
  labs(color="MAG Order", fill = "MAG Order") +
  theme(axis.text = element_text(size=20),
        text = element_text(size=20),
        legend.position = "none",
        legend.position.inside = c(0.81, 0.9),
        panel.border = element_rect(color="black", linewidth = 1)
        )  +
        ylab("PCoA2 (11.5%)") + xlab("PCoA1 (14.5%)") + ggtitle("") -> pwy_pca 



fviz_pca_biplot(pca_like,
                # --- samples ---
                geom.ind      = "point",
                pointshape    = 19,
                col.ind       = order_vec,          # color by group
                palette       = ord_cols,
                addEllipses   = T,            # confidence ellipses
                ellipse.type  = "convex",    # or "convex" / "norm" / "t"
                ellipse.level = 0.95,            # 95% CI
                label         = c("var"),           # label variables only,
                labelsize     = 4,  
                # --- variables (arrows) ---
                col.var       = "black",
                alpha.var =    0.2,
                repel         = TRUE,  
                title         = "",
                select.var = list(name = pws_categories1,
                                  cos2 = 0.4,
                                  contrib = NULL)
                ) +
  theme_bw() +
  guides(color = guide_legend(ncol = 2),
         fill  = guide_legend(ncol = 2)) +
  labs(color="GEM Order", fill = "GEM Order") +
  theme(axis.text = element_text(size=20),
        text = element_text(size=20),
        legend.position = "inside",
        legend.position.inside = c(0.81, 0.9))  +
        ylab("PCoA2 (11.5%)") + xlab("PCoA1 (14.5%)")-> pwy_pca_bi;pwy_pca_bi

 

png(filename = 
      paste0(
      "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/hm_pws_biplot.png"),
    width = 16, height = 13, units = "in", res = 600)

pwy_pca_bi

dev.off()
file.show("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/hm_pws_biplot.png")

png(filename = 
      paste0(
      "~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/hm_pws_pca.png"),
    width = 16, height = 13, units = "in", res = 600)

pwy_pca

dev.off()
file.show("~/rstudio-server/nvme_4tb/riccardo/analysis/boschettona23/Results/man_v2/figures/hm_pws_pca.png")

```
## PWY PCoA check
```{r eval=TRUE, echo = FALSE, warning = FALSE, message=FALSE, results='hide'}
# ── PCoA ──────────────────────────────────────────────────────────────────────
dist_matrix <- vegdist(mat, method = "jaccard")
pcoa_res        <- ape::pcoa(dist_matrix)
# ── envfit arrows ─────────────────────────────────────────────────────────────
env <- envfit(pcoa_res$vectors, mat, permutations = 999)

# ── Variance explained ────────────────────────────────────────────────────────
var_exp <- round(pcoa_res$values$Relative_eig * 100, 1)

# ── plot with base vegan ───────────────────────────────────────────────────────
plot(pcoa$vectors,
     col = ord_cols[order_vec[rownames(mat)]],
     xlab = paste0("PCoA1 (", var_exp[1], "%)"),
     ylab = paste0("PCoA2 (", var_exp[2], "%)"),
     main = "PCoA", lwd = 2, pch=20)
     
plot(env, p.max = 0.001, col="black")       # only significant arrows

ordiellipse(pcoa$vectors[, 1:2],
            groups = order_vec[rownames(mat)],
            kind   = "se",                      # "se", "sd", or "ehull"
            conf   = 0.95,
            draw   = "polygon",
            col    = ord_cols[order_vec[rownames(mat)]],
            border = ord_cols[order_vec[rownames(mat)]],
            alpha  = 0.1,
            lwd    = 1.5) -> p

legend("topright",
       legend = names(ord_cols),
       col    = ord_cols,
       pt.bg  = ord_cols,
       pch    = 21,
       cex    = 0.9,
       bty    = "n")    

```
